---
title: "Analysis Log: 2020 Ocean Sciences High Density CCommunities "
author: "Robert.McGuinn@NOAA.gov; 843-460-9696"
date: '`r Sys.Date()`'
output: word_document
always_allow_html: yes
---

```{r import_data_and_fix, echo=FALSE, cache=TRUE, eval=T}

##### add NDB ##### 
setwd("C:/rworking/deepseatools/indata")
indata<-read.csv("DSCRTP_NatDB_20191217-0.csv", header = T)
filt <- indata %>%
  filter(Flag == "0")

##### add data from Tom Hourigan from 2019 ISDSC7 meeting##### 
x <- "20190719_0_CAPSTONE_AnalysisHiDensity_THourigan.xlsx"
setwd("C:/rworking/deepseatools/indata")
y <- read.xlsx(x, sheet = 1)
y <- y %>% filter(
  EventID != 'D2-EX1605L1-13',
  is.na(ScientificName) == F)

##### strip of the Density measurements for later merge step
yd <- y %>% dplyr::select(EventID, Density)
names(yd) <- c('EventID','Density2')
yd <- yd %>% group_by(EventID, Density2) %>% summarize(n=n()) 

##### extract the records from the NDB that have matching EventIDs with Tom's file ##### 
a <- filt %>% filter(EventID %in% yd$EventID)%>% dplyr::select(names(y))
 
##### merge the Density values from Tom's original file ##### 
b <- merge(a,yd, all.x = T)
b$Density <- b$Density2
b <- b %>% dplyr::select(-Density2, -n)

##### add four more dives from most recent DB ##### 
z <- filt %>% filter(EventID %in% c("NA101-H1715","NA101-H1717","NA101-H1720","NA101-H1722"), 
                      DatasetID == "OET_NA101") %>% dplyr::select(names(y))

d <- rbind(b,z) 

# Notes from Tom Hourigan on how to set Densities for the added 4 dives
# NA101-H1715 - High Density, >3000/km
# NA101-H1717 - Very High Density, >10000/km
# NA101-H1720 - High Density, >3000/km
# NA101-H1722 - High Density, >3000/km

##### add Density values to new dives issues #####
d$Density <- ifelse(test = d$EventID %in%  c('NA101-H1715','NA101-H1720','NA101-H1722'), 
                    yes = ">3000/km", 
                    no = d$Density)

d$Density <- ifelse(test = d$EventID %in%  c('NA101-H1717'), 
                    yes = ">10000/km", 
                    no = d$Density)

##### export data ##### 
setwd("C:/rworking/deepseatools/reports/2020_Ocean_Sciences_Hourigan/data")
write.xlsx(d, '20200131-0_Hawaii_CAPSTONE_plus_THourigan_RMcGuinn.xlsx')  

```

```{r setup, include=FALSE}
opts_chunk$set(fig.height = 5,
               fig.width = 7, 
               dpi = 300)
```

# Analysis Log

* Google Drive project folder location: [2020_Ocean_Sciences_Hourigan](https://drive.google.com/open?id=1NqMKNQYSGMTGR_KcSh8NdTV6ZQgsmMkA)
* Local path - Robert McGuinn's computer ("C:\rworking\deepseatools\reports\2020_Ocean_Sciences_Hourigan")

## Dataset: Coral & sponge records from all of the CAPSTONE with high density dives

* Dataset used in this analysis provided by Tom Hourigan + 4 new dives added from the National Database version 20191217-0.  The dataset is in the '~2020_Ocean_Sciences_Hourigan/data' folder: *20200131-0_Hawaii_CAPSTONE_plus_THourigan_RMcGuinn.xlsx* 

The added dives:

** NA101-H1715 - High Density, >3000/km
** NA101-H1717 - Very High Density, >10000/km
** NA101-H1720 - High Density, >3000/km
** NA101-H1722 - High Density, >3000/km)

The density ranks are per 'EventID' (or dive)

* High-Density (>3000/km in the 'Density' variable)
* Very High Density (>10000/km in the 'Density' varable)

## Summary tables for taxa, grouped by EventID

#### Table: Coral and sponge summary by individual tax(limit: taxonrank is below genus)

```{r table_coral_and_sponge_summary_by_EventID, echo=FALSE} 
x <- d %>%
  filter(
    # TaxonRank == 'genus' |
    #   TaxonRank == 'species' |
    #   TaxonRank == 'subspecies',
    #Phylum == 'Cnidaria'
  ) %>%
  arrange(ScientificName) %>%
  group_by(EventID, ScientificName) %>%
  summarize(
    TaxonRank = paste(unique(TaxonRank), collapse= ' | '),
    Phylum = paste(unique(Phylum), collapse= ' | '),
    Taxa = paste(unique(ScientificName), collapse= ' | '),
    IndividualCount_sum = sum(IndividualCount),
    Densities = paste(unique(Density), collapse = ' | '),
    #Richness = length(unique(ScientificName)),
    Locality = paste(unique(Locality), collapse= ' | '),
    MinimumDepth = min(DepthInMeters),
    MaximumDepth = max(DepthInMeters), 
    NumberRecords = n()

  )

x <- x %>% arrange(EventID, desc(IndividualCount_sum))


##### write it to excel ##### 
filename = "20200203-1_Summary_of_sorals_and_sponges_by_EventID_ScientificName_not_limited_by_TaxonRank_RPMcGuinn.xlsx"
setwd("C:/rworking/deepseatools/reports/2020_Ocean_Sciences_Hourigan/tables")
write.xlsx(x, filename)  
```

#### Table: Coral summary by EventID (limit: taxonrank is below genus)

```{r table_coral_summary_by_EventID, echo=FALSE} 
x <- d %>%
  filter(
    # TaxonRank == 'genus' |
    #   TaxonRank == 'species' |
    #   TaxonRank == 'subspecies',
    Phylum == 'Cnidaria'
  ) %>%
  arrange(ScientificName) %>%
  group_by(EventID, ScientificName) %>%
  summarize(
    Densities = paste(unique(Density), collapse = ' | '),
    Taxa = paste(unique(ScientificName), collapse= ' | '),
    Richness = length(unique(ScientificName)),
    Locality = paste(unique(Locality), collapse= ' | '),
    MinimumDepth = min(DepthInMeters),
    MaximumDepth = max(DepthInMeters), 
    NumberRecords = n(),
    IndividualCount_sum = sum(IndividualCount)
  )

x <- x %>% arrange(EventID, desc(IndividualCount_sum))


##### write it to excel ##### 
filename = "20200203-0_Summary_of_Corals_by_EventID_not_limited_by_TaxonRank_RPMcGuinn.xlsx"
setwd("C:/rworking/deepseatools/reports/2020_Ocean_Sciences_Hourigan/tables")
write.xlsx(x, filename)  
```

*Coral* occurrences per EventID (dive) where TaxonRank is limited to 'genus, species, or subspecies'. The table is arranged in descending order of the sum of the IndividualCount. 

For an Excel version of the table, see the '~2020_Ocean_Sciences_Hourigan/tables' folder
Filename: *`r filename`*

#### Table: Sponge summary by EventID (limit: taxonrank is below genus)

```{r table_sponge_summary_by_EventID, echo=FALSE} 
x <- d %>%
  filter(
    TaxonRank == 'genus' |
      TaxonRank == 'species' |
      TaxonRank == 'subspecies',
    Phylum == 'Porifera'
  ) %>%
  arrange(ScientificName) %>%
  group_by(EventID) %>%
  summarize(
    Densities = paste(unique(Density), collapse = ' | '),
    Taxa = paste(unique(ScientificName), collapse= ' | '),
    Richness = length(unique(ScientificName)),
    Locality = paste(unique(Locality), collapse= ' | '),
    MinimumDepth = min(DepthInMeters),
    MaximumDepth = max(DepthInMeters), 
    NumberRecords = n(),
    IndividualCount_sum = sum(IndividualCount)
  )

x <- x %>% arrange(desc(IndividualCount_sum))


##### write it to excel ##### 
filename = "20200131-0_Summary_of_Sponges_by_EventID.xlsx"
setwd("C:/rworking/deepseatools/reports/2020_Ocean_Sciences_Hourigan/tables")
write.xlsx(x, filename)  

```

*Sponge* occurrences per EventID (dive) where TaxonRank is limited to 'genus, species, or subspecies'. The table is arranged in descending order of the sum of the IndividualCount. 

For an Excel version of the table, see the '~2020_Ocean_Sciences_Hourigan/tables' folder
Filename: *`r filename`*

## Coral summary by 4 distinct zones defined by region and depth. 

1. North - Shallow 
2. North - Deep
3. South - Shallow
4. South - Deep

North: > 10 degrees latitude  
South: < 10 degrees latitude  
Shallow: 0-1000 m  
Deep: > 1000 m

#### Figure: Coral community (North - Shallow) 

``` {r north_shallow, echo=FALSE, cache=FALSE, dpi=300, warnings = FALSE}

latlimit <- 10
depthlimit <- 1000

x <- d %>% filter(
  #is.na(Genus) == T,
  # Order != "Alcyonacea",
  Phylum == "Cnidaria", 
  as.numeric(Latitude) > latlimit,
  as.numeric(DepthInMeters) < depthlimit,
  # grepl('Wake', Locality)
  )

# limit x by count
nlimit <- 100
counts <- x %>% group_by(ScientificName) %>% summarize(IndividualCount_sum = sum(IndividualCount))
counts <- counts %>% filter(IndividualCount_sum > nlimit)
counts$abundance_cat <- NA
counts$abundance_cat[counts$IndividualCount_sum > 0 & counts$IndividualCount_sum <= 1000] <- "0-1,000 Individuals"
counts$abundance_cat[counts$IndividualCount_sum > 1000 & counts$IndividualCount_sum <= 2000] <- "1,001-2,000 Individuals"
counts$abundance_cat[counts$IndividualCount_sum > 2000 & counts$IndividualCount_sum <= 3000] <- "2,001-3,000 Individuals"
counts$abundance_cat[counts$IndividualCount_sum > 3000 & counts$IndividualCount_sum <= 4000] <- "3,001-4,000 Individuals"
counts$abundance_cat[counts$IndividualCount_sum > 4000 & counts$IndividualCount_sum <= 5000] <- "4,001-5,000 Individuals"
counts$abundance_cat[counts$IndividualCount_sum > 5000 & counts$IndividualCount_sum <= 6000] <- "5,001-6,000 Individuals"
counts$abundance_cat[counts$IndividualCount_sum > 6000 & counts$IndividualCount_sum <= 7000] <- "6,001-7,000 Individuals"
counts$abundance_cat[counts$IndividualCount_sum > 7000 & counts$IndividualCount_sum <= 10000] <- "7,001-10,000 Individuals"
counts$abundance_cat[counts$IndividualCount_sum > 10000 & counts$IndividualCount_sum <= 20000] <- "10,001- 20,000 Individuals"
counts$abundance_cat[counts$IndividualCount_sum > 20000] <- "> 20,000 Individuals"
counts$abundance_cat <- factor(counts$abundance_cat, levels = c("0-1,000 Individuals", "1,001-2,000 Individuals", "2,001-3,000 Individuals", "3,001-4,000 Individuals", "4,001-5,000 Individuals", "5,001-6,000 Individuals", "6,001-7,000 Individuals", "7,001-10,000 Individuals", "10,001- 20,000 Individuals", "> 20,000 Individuals"))

x <- x %>% filter(ScientificName %in% counts$ScientificName)
x <- merge(x,counts)
# table(x$Genus)

# plot
g <- ggplot(x, aes(reorder(ScientificName, as.numeric(DepthInMeters), FUN=median), as.numeric(DepthInMeters),fill=abundance_cat)) + 
  geom_boxplot() +
  scale_y_reverse() +
  # ylim(2000,1500) +
  ylab("Depth (meters)") + 
  xlab("Taxon") +
  labs(fill = "Abundance Classes") +
  theme_bw(base_size = 22, base_family = "Cambria") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0, face = 'italic'))#  +
  #geom_hline(yintercept = 300, col = 'grey')

g + scale_fill_manual(values = rev(brewer.pal(10,'Spectral')))
# set.seed(8)
# g + scale_fill_manual(values = sample(big_palette))
# g + scale_color_manual(values = brewer.pal(12, "Paired")[c(10,9,8,7,6,5,4,3,2,1)])
```

#### Figure: Coral community (North - Deep) 

``` {r north_deep, echo=FALSE, cache=FALSE, dpi=300, warnings = FALSE}

latlimit <- 10
depthlimit <- 1000

x <- d %>% filter(
  #is.na(Genus) == T,
  # Order != "Alcyonacea",
  Phylum == "Cnidaria", 
  as.numeric(Latitude) > latlimit,
  as.numeric(DepthInMeters) > depthlimit,
  # grepl('Wake', Locality)
  )

# limit x by count
nlimit <- 100
counts <- x %>% group_by(ScientificName) %>% summarize(IndividualCount_sum = sum(IndividualCount))
counts <- counts %>% filter(IndividualCount_sum > nlimit)
counts$abundance_cat <- NA
counts$abundance_cat[counts$IndividualCount_sum > 0 & counts$IndividualCount_sum <= 1000] <- "0-1,000 Individuals"
counts$abundance_cat[counts$IndividualCount_sum > 1000 & counts$IndividualCount_sum <= 2000] <- "1,001-2,000 Individuals"
counts$abundance_cat[counts$IndividualCount_sum > 2000 & counts$IndividualCount_sum <= 3000] <- "2,001-3,000 Individuals"
counts$abundance_cat[counts$IndividualCount_sum > 3000 & counts$IndividualCount_sum <= 4000] <- "3,001-4,000 Individuals"
counts$abundance_cat[counts$IndividualCount_sum > 4000 & counts$IndividualCount_sum <= 5000] <- "4,001-5,000 Individuals"
counts$abundance_cat[counts$IndividualCount_sum > 5000 & counts$IndividualCount_sum <= 6000] <- "5,001-6,000 Individuals"
counts$abundance_cat[counts$IndividualCount_sum > 6000 & counts$IndividualCount_sum <= 7000] <- "6,001-7,000 Individuals"
counts$abundance_cat[counts$IndividualCount_sum > 7000 & counts$IndividualCount_sum <= 10000] <- "7,001-10,000 Individuals"
counts$abundance_cat[counts$IndividualCount_sum > 10000 & counts$IndividualCount_sum <= 20000] <- "10,001- 20,000 Individuals"
counts$abundance_cat[counts$IndividualCount_sum > 20000] <- "> 20,000 Individuals"
counts$abundance_cat <- factor(counts$abundance_cat, levels = c("0-1,000 Individuals", "1,001-2,000 Individuals", "2,001-3,000 Individuals", "3,001-4,000 Individuals", "4,001-5,000 Individuals", "5,001-6,000 Individuals", "6,001-7,000 Individuals", "7,001-10,000 Individuals", "10,001- 20,000 Individuals", "> 20,000 Individuals"))

x <- x %>% filter(ScientificName %in% counts$ScientificName)
x <- merge(x,counts)
# table(x$Genus)

# plot
g <- ggplot(x, aes(reorder(ScientificName, as.numeric(DepthInMeters), FUN=median), as.numeric(DepthInMeters),fill=abundance_cat)) + 
  geom_boxplot() +
  scale_y_reverse() +
  # ylim(2000,1500) +
  ylab("Depth (meters)") + 
  xlab("Taxon") +
  labs(fill = "Abundance Classes") +
  theme_bw(base_size = 22, base_family = "Cambria") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0, face = 'italic'))#  +
  #geom_hline(yintercept = 300, col = 'grey')

g + scale_fill_manual(values = rev(brewer.pal(10,'Spectral')))
# set.seed(8)
# g + scale_fill_manual(values = sample(big_palette))
# g + scale_color_manual(values = brewer.pal(12, "Paired")[c(10,9,8,7,6,5,4,3,2,1)])
```


#### Figure: Coral community (South - Shallow) 

``` {r south_shallow, echo=FALSE, cache=FALSE, dpi=300, warnings = FALSE}

latlimit <- 10
depthlimit <- 1000

x <- d %>% filter(
  #is.na(Genus) == T,
  # Order != "Alcyonacea",
  Phylum == "Cnidaria", 
  as.numeric(Latitude) < latlimit,
  as.numeric(DepthInMeters) < depthlimit,
  # grepl('Wake', Locality)
  )

# limit x by count
nlimit <- 10
counts <- x %>% group_by(ScientificName) %>% summarize(IndividualCount_sum = sum(IndividualCount))
counts <- counts %>% filter(IndividualCount_sum > nlimit)
counts$abundance_cat <- NA
counts$abundance_cat[counts$IndividualCount_sum > 0 & counts$IndividualCount_sum <= 1000] <- "0-1,000 Individuals"
counts$abundance_cat[counts$IndividualCount_sum > 1000 & counts$IndividualCount_sum <= 2000] <- "1,001-2,000 Individuals"
counts$abundance_cat[counts$IndividualCount_sum > 2000 & counts$IndividualCount_sum <= 3000] <- "2,001-3,000 Individuals"
counts$abundance_cat[counts$IndividualCount_sum > 3000 & counts$IndividualCount_sum <= 4000] <- "3,001-4,000 Individuals"
counts$abundance_cat[counts$IndividualCount_sum > 4000 & counts$IndividualCount_sum <= 5000] <- "4,001-5,000 Individuals"
counts$abundance_cat[counts$IndividualCount_sum > 5000 & counts$IndividualCount_sum <= 6000] <- "5,001-6,000 Individuals"
counts$abundance_cat[counts$IndividualCount_sum > 6000 & counts$IndividualCount_sum <= 7000] <- "6,001-7,000 Individuals"
counts$abundance_cat[counts$IndividualCount_sum > 7000 & counts$IndividualCount_sum <= 10000] <- "7,001-10,000 Individuals"
counts$abundance_cat[counts$IndividualCount_sum > 10000 & counts$IndividualCount_sum <= 20000] <- "10,001- 20,000 Individuals"
counts$abundance_cat[counts$IndividualCount_sum > 20000] <- "> 20,000 Individuals"
counts$abundance_cat <- factor(counts$abundance_cat, levels = c("0-1,000 Individuals", "1,001-2,000 Individuals", "2,001-3,000 Individuals", "3,001-4,000 Individuals", "4,001-5,000 Individuals", "5,001-6,000 Individuals", "6,001-7,000 Individuals", "7,001-10,000 Individuals", "10,001- 20,000 Individuals", "> 20,000 Individuals"))

x <- x %>% filter(ScientificName %in% counts$ScientificName)
x <- merge(x,counts)
# table(x$Genus)

# plot
g <- ggplot(x, aes(reorder(ScientificName, as.numeric(DepthInMeters), FUN=median), as.numeric(DepthInMeters),fill=abundance_cat)) + 
  geom_boxplot() +
  scale_y_reverse() +
  # ylim(2000,1500) +
  ylab("Depth (meters)") + 
  xlab("Taxon") +
  labs(fill = "Abundance Classes") +
  theme_bw(base_size = 22, base_family = "Cambria") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0, face = 'italic'))#  +
  #geom_hline(yintercept = 300, col = 'grey')

g + scale_fill_manual(values = rev(brewer.pal(10,'Spectral')))
# set.seed(8)
# g + scale_fill_manual(values = sample(big_palette))
# g + scale_color_manual(values = brewer.pal(12, "Paired")[c(10,9,8,7,6,5,4,3,2,1)])
```

#### Figure: Coral community (South - Deep) 

``` {r south_deep, echo=FALSE, cache=FALSE, dpi=300, warnings = FALSE}

latlimit <- 10
depthlimit <- 1000

x <- d %>% filter(
  # is.na(Genus) == T,
  # Order != "Alcyonacea",
  Phylum == "Cnidaria", 
  as.numeric(Latitude) < latlimit,
  as.numeric(DepthInMeters) > depthlimit,
  # grepl('Wake', Locality)
  )

# limit x by count
nlimit <- 10
counts <- x %>% group_by(ScientificName) %>% summarize(IndividualCount_sum = sum(IndividualCount))
counts <- counts %>% filter(IndividualCount_sum > nlimit)
counts$abundance_cat <- NA
counts$abundance_cat[counts$IndividualCount_sum > 0 & counts$IndividualCount_sum <= 1000] <- "0-1,000 Individuals"
counts$abundance_cat[counts$IndividualCount_sum > 1000 & counts$IndividualCount_sum <= 2000] <- "1,001-2,000 Individuals"
counts$abundance_cat[counts$IndividualCount_sum > 2000 & counts$IndividualCount_sum <= 3000] <- "2,001-3,000 Individuals"
counts$abundance_cat[counts$IndividualCount_sum > 3000 & counts$IndividualCount_sum <= 4000] <- "3,001-4,000 Individuals"
counts$abundance_cat[counts$IndividualCount_sum > 4000 & counts$IndividualCount_sum <= 5000] <- "4,001-5,000 Individuals"
counts$abundance_cat[counts$IndividualCount_sum > 5000 & counts$IndividualCount_sum <= 6000] <- "5,001-6,000 Individuals"
counts$abundance_cat[counts$IndividualCount_sum > 6000 & counts$IndividualCount_sum <= 7000] <- "6,001-7,000 Individuals"
counts$abundance_cat[counts$IndividualCount_sum > 7000 & counts$IndividualCount_sum <= 10000] <- "7,001-10,000 Individuals"
counts$abundance_cat[counts$IndividualCount_sum > 10000 & counts$IndividualCount_sum <= 20000] <- "10,001- 20,000 Individuals"
counts$abundance_cat[counts$IndividualCount_sum > 20000] <- "> 20,000 Individuals"
counts$abundance_cat <- factor(counts$abundance_cat, levels = c("0-1,000 Individuals", "1,001-2,000 Individuals", "2,001-3,000 Individuals", "3,001-4,000 Individuals", "4,001-5,000 Individuals", "5,001-6,000 Individuals", "6,001-7,000 Individuals", "7,001-10,000 Individuals", "10,001- 20,000 Individuals", "> 20,000 Individuals"))

x <- x %>% filter(ScientificName %in% counts$ScientificName)
x <- merge(x,counts)
# table(x$Genus)

# plot
g <- ggplot(x, aes(reorder(ScientificName, as.numeric(DepthInMeters), FUN=median), as.numeric(DepthInMeters),fill=abundance_cat)) + 
  geom_boxplot() +
  scale_y_reverse() +
  # ylim(2000,1500) +
  ylab("Depth (meters)") + 
  xlab("Taxon") +
  labs(fill = "Abundance Classes") +
  theme_bw(base_size = 22, base_family = "Cambria") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0, face = 'italic'))#  +
  #geom_hline(yintercept = 300, col = 'grey')

g + scale_fill_manual(values = rev(brewer.pal(10,'Spectral')))
# set.seed(8)
# g + scale_fill_manual(values = sample(big_palette))
# g + scale_color_manual(values = brewer.pal(12, "Paired")[c(10,9,8,7,6,5,4,3,2,1)])
```


```{r setdepthclass, echo=FALSE, cache = FALSE} 

##### setting depth categories ##### 
d$DepthCat[as.numeric(d$DepthInMeters) <= 1000] <- "shallow"
d$DepthCat[as.numeric(d$DepthInMeters) > 1000] <- "deep"
d$DepthCat <- factor(d$DepthCat, levels = c("shallow", "deep"))

##### setting location categories ##### 
d$loc[as.numeric(d$Latitude) > 10] <- "north"
d$loc[as.numeric(d$Latitude) < 10] <- "south"

##### setting temperature categories #####
d$temp_cat[as.numeric(d$Temperature) > 5] <- "warm: > 5"
d$temp_cat[as.numeric(d$Temperature) >= 3 & as.numeric(d$Temperature) <= 5] <- "cold: 3 to 5"
d$temp_cat[as.numeric(d$Temperature) < 3] <- "colder: less than 3"
d$temp_cat <- factor(d$temp_cat, levels = c("warm: > 5", "cold: 3 to 5", "colder: less than 3"))

##### setting oxygen categories #####
# qplot(as.numeric(d$Oxygen))
d$ox_cat[as.numeric(d$Oxygen) >= 4] <- "rich"
d$ox_cat[as.numeric(d$Oxygen) < 4] <- "poor"

##### longitude cat #####
d$longloc[as.numeric(d$Longitude) > 90] <- "west"
d$longloc[as.numeric(d$Longitude) < 90] <- "east"

```


``` {r NMDS, echo=FALSE, cache = FALSE, dpi=300, warning=FALSE, message=FALSE, results='hide', cache=T}

x <- d %>% 
  filter(
    #DepthCat == 'deep',
    #EventID != "D2-EX1605L1-13"
    #Phylum == "Porifera"
  )

##### create site X species matrix #####
library(vegan)
site.sp <- dcast(x, EventID ~ ScientificName, value.var = "IndividualCount", fun.aggregate = sum)

# setwd("C:/rworking/digs/outdata")
# write.csv(site.sp, 'site.sp.csv')

# creating a site variable
site.sp$site <- site.sp$EventID

# getting rid of non-needed variables
site.sp <- site.sp %>%
  dplyr::select(-EventID)

# # moving the site variable to the beginning
# col_idx <- grep("site", names(site.sp))
# site.sp <- site.sp[, c(col_idx, (1:ncol(site.sp))[-col_idx])]
# # names(site.sp)

# set the site as the row.names
row.names(site.sp) <- site.sp$site

# remove the site variable
site.sp <- site.sp %>%
  dplyr::select(-site)

# making it a matrix
site.sp <- as.matrix(site.sp)

##### running NMDS starting with site X species matrix #####
#install.packages("vegan")
library(vegan)
NMDS <- metaMDS(site.sp, distance = "jaccard", binary = F, k=3, trymax = 30)

##### extracting the site and species scores for use with ggplot2 ##### 
#Using the scores function from vegan to extract the site scoresand convert to a data.frame
site.scores <- as.data.frame(scores(NMDS))  
site.scores$site <- rownames(site.scores)  # create a column of site names, from the rownames of data.scores
# head(site.scores)  #look at the data
# merge information from d (to get categories)
site.scores <- merge(site.scores, x, by.x = 'site', by.y = 'EventID')
site.scores <- site.scores %>% group_by(site) %>% summarise(NMDS1 = NMDS1[1],
                                                            NMDS2 = NMDS2[1],
                                                            DepthCat = DepthCat[1],
                                                            temp_cat = temp_cat[1],
                                                            loc = loc[1])


# extracting species scores
species.scores <- as.data.frame(scores(NMDS, "species"))  #Using the scores function from vegan to extract the species scores and convert to a data.frame
species.scores$species <- rownames(species.scores)  # create a column of species, from the rownames of species.scores
# head(species.scores)  #look at the data
species.scores$spec_code <- 1:(length(species.scores$species))

NMDS

##### plotting NMDS ##### 
ggplot() +
  geom_point(data=site.scores,aes(x=NMDS2,y=NMDS1, color=DepthCat), size=5) + # add the point markers
  geom_text(data=site.scores,aes(x=NMDS2,y=NMDS1, label = site), size=4, position = position_nudge(y = .0001)) + # add the text
  #geom_point(data=species.scores,aes(x=NMDS1,y=NMDS2), shape = 3, size=4) + 
  coord_equal() +
  theme_bw(base_size = 15, base_family = "Cambria") +
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank())


ggplot() +
  geom_point(data=site.scores,aes(x=NMDS2,y=NMDS1, color=temp_cat), size=5) + # add the point markers
  geom_text(data=site.scores,aes(x=NMDS2,y=NMDS1, label = site), size=4, position = position_nudge(y = .0001)) + # add the text
  #geom_point(data=species.scores,aes(x=NMDS1,y=NMDS2), shape = 3, size=4) + 
  coord_equal() +
  theme_bw(base_size = 15, base_family = "Cambria") +
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank())

ggplot() +
  geom_point(data=site.scores,aes(x=NMDS2,y=NMDS1, color=loc), size=5) + # add the point markers
  geom_text(data=site.scores,aes(x=NMDS2,y=NMDS1, label = site), size=4, position = position_nudge(y = .0001)) + # add the text
  #geom_point(data=species.scores,aes(x=NMDS1,y=NMDS2), shape = 3, size=4) + 
  coord_equal() +
  theme_bw(base_size = 15, base_family = "Cambria") +
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank())


ggplot() +
  #geom_point(data=site.scores,aes(x=NMDS2,y=NMDS1, color=loc), size=5) + # add the point markers
  geom_text(data=species.scores,aes(x=NMDS2,y=NMDS1, label = spec_code), size=4, position = position_nudge(y = .0001)) + # add the text
  geom_point(data=species.scores,aes(x=NMDS2,y=NMDS1), shape = 3, size=4) + 
  coord_equal() +
  theme_bw(base_size = 15, base_family = "Cambria") +
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank())

print(NMDS)
```

Table of species codes (key to the lables in the species score NMDS plot above)

```{r spec_codes, echo=F}
x <- species.scores %>%  dplyr::select(spec_code)

kable(x)

```

## Diversity estimates.

##### Table: Shannon diversity per EventID
All taxa (both coral and sponges at all taxonomic ranks included in estimates): [LINK](https://drive.google.com/open?id=1PWfh5Tc6Fk8mM4jHeNe31XesLXDzW1np)

```{r diversity, echo=F}

plots <- diversity(site.sp, index = "shannon") #makes an object
#that the diversity values #are written to.
# summary(plots) #gives summary statistics for the plots
# median(plots) #gives the median
# mean(plots) #gives the mean
# class(plots)
# max(plots[1:30])
# # View(plots)

# creating plots
plots_df <- data.frame(plots)
# View(plots_df)
plots_df$EventID <- row.names(plots_df)
# names(d)
d <- merge(d, plots_df, all.x = T)
# length(d$EventID)
# length(unique(d$EventID))
x <- d %>% group_by(EventID, plots) %>% summarize(n=n(), 
                                                 Latitude = paste(mean(as.numeric(Latitude)), collapse = ' | '), 
                                                 Longitude = paste(mean(as.numeric(Longitude)), collapse = ' | '))
#View(x)

setwd("C:/rworking/deepseatools/reports")
write.xlsx(x, '20190725-0_Summary_of_Corals_and_Sponges_Diversity_by_EventID.xlsx') 

## Calculating various Diversity Indices

# H <- diversity(site.sp)
# simp <- diversity(site.sp, "simpson")
# invsimp <- diversity(site.sp, "inv")
# ## Unbiased Simpson (Hurlbert 1971, eq. 5) with rarefy:
# unbias.simp <- rarefy(site.sp, 2) - 1
# ## Fisher alpha
# alpha <- fisher.alpha(site.sp)
# ## Plot all
# pairs(cbind(H, simp, invsimp, unbias.simp, alpha, J), pch="+", col="blue")
# ## Species richness (S) and Pielou's evenness (J):
# S <- specnumber(site.sp) ## rowSums(site.sp > 0) does the same...
# J <- H/log(S)

```

```{r mapping_diversity , echo=FALSE, eval=F}

x$Latitude <- as.numeric(x$Latitude)
x$Longitude <- as.numeric(x$Longitude)

pal <- colorNumeric(
  palette = "Reds",
  domain = x$plots)


m <- leaflet()
m <- addProviderTiles(m, "Esri.OceanBasemap") #Esri.OceanBasemap, "CartoDB.DarkMatter"
m <- addCircleMarkers(m, data=x,
                      radius=5,
                      weight=0,
                      #fillColor= "green",
                      color = ~pal(plots),
                      fillOpacity=1,
                      popup = paste("EventID:", x$EventID, "<br>"))
                                   
m

```

## Summary of community composition at each EventID.  

#### Figure: 56 figures follow, one per each EventID. (To find a specific EventID use ctrl-F in Word.)

Figures are arranged by ascending order of EventID. Each figure represents 'top 10' taxa in each EventID. The list (legend) on the side of the figure is in rank order and the color blocks are also in rank order by sum of IndividualCount as a proportion of total.Colors between figures m represent different taxa, because there are differing Top 10 taxa. All taxa, sponges and corals, were used regardless of TaxonRank. To find a specific EventID use ctrl-F in Word.

```{r communities_by_eventID, echo=F, eval=T, warning=F, message=F}

latlimit <- 10
depthlimit <- 1000

##### bring in the original data (d) and filter according to EventIDs of interest
x <- d %>% filter(
  TaxonRank == 'genus' |
    TaxonRank == 'species',
  #Density == '>10000/km', 
  #Phylum == "Cnidaria"#,
  #as.numeric(Latitude) > latlimit,
  #as.numeric(DepthInMeters) < depthlimit
  ) %>%
  group_by(ScientificName, EventID) %>%
  summarize(sum = sum(IndividualCount)) %>%
  arrange(desc(sum))

##### getting the top ten species/eventID from the combined list when ordered by 
# sum of IndividualCount
x <- x[1:55,]

##### filter
z <- d %>% filter(
  ScientificName %in% x$ScientificName
  #EventID %in% x$EventID
  )

# checking
length(unique(x$EventID))
length(unique(z$EventID))
length(unique(d$EventID))

length(unique(x$ScientificName))
length(unique(z$ScientificName))
length(unique(d$ScientificName))

##### make some colors #####
palette1 <- brewer.pal(9,"Set1")
palette2 <- brewer.pal(8,"Set2")
palette3 <- brewer.pal(9,"Set3")
palette4 <- brewer.pal(8,"Dark2")
palette5 <- brewer.pal(8,"Accent")
big_palette <- c(palette1,palette2,palette3, palette4, palette5)

##### set colors to rep specific taxa #####
# mycols <- c("Keratoisidinae" = "red",
#             "Acanella weberi" = "blue",
#             "Hemicorallium" = "darkgreen",
#             "Madrepora oculata" = "black",
#             "Paramuricea" = "steelblue3",
#             "Paragorgia" = "gray64",
#             "Paracalyptrophora hawaiiensis" = "violet",
#             "Plexauridae" = "tan", 
#             "Chrysogorgia" = "lightsalmon",
#             "Iridogorgia" = "gray25",
#             "Parisis" = "green"                   
#             )

mycols <- big_palette
names(mycols) <- levels(factor(x$ScientificName))

##### run the loop ##### 

for (id in unique(z$EventID)){

# get the right EventID
yo <- z %>% filter(EventID == id)

# create the proper oridering
ord <- aggregate(yo$IndividualCount, by=list(yo$ScientificName), FUN=sum) 
ord <- ord %>% arrange(x)
yo$ScientificName <- factor(yo$ScientificName, levels=ord$Group.1)
yo <- yo[order(yo$ScientificName),]


# make the plot
p <- ggplot(yo, aes(x = EventID, y = IndividualCount, fill = ScientificName)) +
  geom_bar(stat = "identity", position = "fill" ) +
  coord_flip() +
  scale_colour_manual(values = mycols, 
                      aesthetics = c("colour", "fill"),
                      drop = FALSE, guide=guide_legend(reverse=TRUE)
                      ) +
  theme_minimal() + 
  theme(legend.position = "bottom", 
        legend.title = element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y = element_blank()
        ) +
  coord_polar("y", start=0) + 
  labs(title = NULL, y = NULL, x = NULL)
print(id)
print(p)
}



```


## Notes from Tom Hourigan on 20200117

I wanted to get started on this, based mostly on the work you already did for the DSC symposium. I think we could also add results from 4 additional Dives in data set OET_NA101

EventID:
NA101-H1715 - High Density
NA101-H1717 - Very High Density
NA101-H1720 - High Density
NA101-H1722 - High Density   


## Notes from Tom Hourigan 20200131

Elements of the analysis for the Ocean Sciences 2020 poster.  

* Large map showing all of the Hi-Density and Very High Density sites for the region (Don't worry about the map for now - but it will be similar to this with the four additional OET NWHI sites from 2018)

* NMDS plot of all the  Hi-Density and Very High Density sites.  I like the one you prepared with temperature categories.  Wonder if we could do a Sorensen’s index of similarity.to see if the different temperature sites really separated out significantly?

* Inset map focused on the Hawaiian Archipelago/Johnston/Musician Seamounts. 

* representative picture of Very High Density communities with a smaller picture of the bathymetry and a pie chart of the abundance of the top ten taxa.  What I want to show is if different taxa predominate in different communities (as I think they do).  To show this most clearly, the same taxa should have the same color on different charts. (ALternatively, instead of a pie chart, a bar chart would work, but it might be more work - see below).  I am focusing on the Hawaiian Archipelago, because it should represent a relatively cohesive biogeographic unit.

* I will probably work up an additional chart showing the total # of CAPSTONE & OET dives (or hours) in each depth strata along with the ones having Hi-Density and Very High Density communities in order to give a measure of survey effort (lower priority for you - I can probably manage this)


