---
title: "Analysis Log: 2020 Ocean Sciences High Density Communities "
author: "Robert.McGuinn@NOAA.gov; 843-460-9696"
date: 'Last ran on: `r Sys.Date()`'
output: word_document
always_allow_html: yes
editor_options: 
  chunk_output_type: console
---

## Install
```{r install, warning=F, message=F, echo=FALSE}
library(tidyverse)
library(RColorBrewer)
library(rmarkdown)
library(knitr)
library(openxlsx)
library(magrittr)
library(mime)
library(scales)
library(vegan)
devtools::install_github("slowkow/ggrepel")

```

## Setup 'knitr' options
```{r setup, include=FALSE}
opts_chunk$set(fig.height = 8,
               fig.width = 10, 
               dpi = 300)
```

## Load

```{r import_data, echo=F, cache=TRUE, eval=T}
##### add NDB ##### 
# setwd("C:/rworking/deepseatools/indata")
# indata<-read.csv("DSCRTP_NatDB_20191217-0.csv", header = T)
# filt <- indata %>%
#   filter(Flag == "0")

##### add data from Tom Hourigan from 2019 ISDSC7 meeting##### 
x <- "20190719_0_CAPSTONE_AnalysisHiDensity_THourigan.xlsx"
setwd("C:/rworking/deepseatools/indata")
y <- read.xlsx(x, sheet = 1)
y <- y %>% filter(
  EventID != 'D2-EX1605L1-13',
  is.na(ScientificName) == F)

###### cleanup ##### 
rm(x)
```

## Fix

```{r fix_data, echo=F, cache=F, eval=T}
##### strip off the Density measurements for later merge step
yd <- y %>% dplyr::select(EventID, Density)
names(yd) <- c('EventID','Density2')
yd <- yd %>% group_by(EventID, Density2) %>% summarize(n=n()) 


##### extract the records from the NDB that have matching EventIDs with Tom's file ##### 
a <- filt %>% filter(EventID %in% yd$EventID)%>% dplyr::select(names(y))
 

##### merge the Density values from Tom's original file ##### 
b <- merge(a,yd, all.x = T)
b$Density <- b$Density2
b <- b %>% dplyr::select(-Density2, -n)

##### add four more dives from most recent DB ##### 
z <- filt %>% filter(EventID %in% c("NA101-H1715","NA101-H1717","NA101-H1720","NA101-H1722"), 
                      DatasetID == "OET_NA101") %>% dplyr::select(names(y))

d <- rbind(b,z) 

# Notes from Tom Hourigan on how to set Densities for the added 4 dives
# NA101-H1715 - High Density, >3000/km
# NA101-H1717 - Very High Density, >10000/km
# NA101-H1720 - High Density, >3000/km
# NA101-H1722 - High Density, >3000/km

##### add Density values to new dives issues #####
d$Density <- ifelse(test = d$EventID %in%  c('NA101-H1715','NA101-H1720','NA101-H1722'), 
                    yes = ">3000/km", 
                    no = d$Density)

d$Density <- ifelse(test = d$EventID %in%  c('NA101-H1717'), 
                    yes = ">10000/km", 
                    no = d$Density)

##### work on 2 NA values in dataset ##### 

# set the NA
d <- d %>%  mutate(ScientificName = ifelse(is.na(ScientificName),
                                           yes = "Narella calamus",
                                           no = as.character(ScientificName)))

# d %>% filter(ScientificName == 'Narella calamus') %>% dplyr::select(ScientificName, EventID)
# x %>% filter(ScientificName == 'Narella calamus') %>% dplyr::select(ScientificName, EventID)
# d %>% filter(is.na(ScientificName) == T) %>% 
#   dplyr::select(ScientificName, EventID)


##### export data ##### 
# setwd("C:/rworking/deepseatools/reports/2020_Ocean_Sciences_Hourigan/data")
# write.xlsx(d, '20200131-0_Hawaii_CAPSTONE_plus_THourigan_RMcGuinn.xlsx')  

##### cleanup #####
rm(a)
rm(b)
rm(yd)
rm(z)
```

## Project folder on Google Drive

Google Drive project folder location: [2020_Ocean_Sciences_Hourigan](https://drive.google.com/open?id=1NqMKNQYSGMTGR_KcSh8NdTV6ZQgsmMkA)

## Pie charts by EventID

The figures below are arranged by ascending alphabetical order by EventID.  Each figure represents the 'top 10' taxa in each EventID.  The legend on the bottom of the figure is in rank order from left to right, as is the arrangement of the taxa within the pie chart (starting a top going around clockwise from most to least abundant).  The color wedges are also in rank order by sum of IndividualCount as a proportion of the total for the top ten taxa. *Colors for each taxa are consistent across all figures.* All taxa, sponges and corals, were used regardless of TaxonRank (no other taxonomic filters applied). To find a specific EventID use *ctrl-F* to find in Word.  The EventID label is included *above* each figure as well as within the image of each figure (this can be left off if needed.) 

## Summarize by EventID

```{r summarize_by_EventID, echo=F, eval=T, warning=F, message=F}

# set limts
# latlimit <- 10
# depthlimit <- 1000

##### bring in the original data (d) and summarize according to EventIDs of interest #####
x <- d %>% filter(
  #TaxonRank == 'genus' |
  #TaxonRank == 'species',
  #Density == '>10000/km', 
  #Phylum == "Cnidaria"#,
  #as.numeric(Latitude) > latlimit,
  #as.numeric(DepthInMeters) < depthlimit
  ) %>%
  group_by(ScientificName, EventID) %>%
  summarize(sum = sum(IndividualCount)) %>%
  arrange(desc(sum))
```

## Make color map

```{r make_colors, echo=F, eval=T, warning=F, message=F}

topten <- d %>%
 group_by(EventID, ScientificName)%>%
 summarize(individual_count_sum = sum(IndividualCount))%>%
 top_n(10, individual_count_sum) %>% arrange(EventID, desc(individual_count_sum))

##### make a random color list that matches length of top ten #####

cl <- colors(distinct = TRUE)
set.seed(15887) # to set random generator seed
mycols <- sample(cl, 138)
names(mycols) <- unique(topten$ScientificName)
#mycols

##### writing to xlsx #####

# setwd("C:/rworking/deepseatools/reports/2020_Ocean_Sciences_Hourigan/data")
# write.xlsx(topten, '20200205-0_top_ten_summary_THourigan_RMcGuinn.xlsx')  

##### checking ##### 

# table(topten$EventID)
# length(unique(topten$ScientificName))
# 
# # checking
# length(unique(x$EventID))
# length(unique(z$EventID))
# length(unique(d$EventID))
# 
# length(unique(x$ScientificName))
# length(unique(z$ScientificName))
# length(unique(d$ScientificName))
# length(unique(topten$ScientificName))

```

## Filter to only species appearing in the top ten list  

```{r filter_to_top_ten, echo=F, eval=T, warning=F, message=F}
##### filter d down to just the species appearing in the top ten ##### 
z <- d %>% filter(
  ScientificName %in% topten$ScientificName
# EventID %in% c("NA101-H1715","NA101-H1720", "D2-EX1504L2-17")
  )
```

## Create pie charts by EventID

```{r create_community_charts, echo=F, eval=f, warning=F, message=F}
##### run the loop ##### 
for (id in unique(z$EventID)){

# get the right EventID
yo <- z %>% filter(EventID == id)

##### create the proper oridering for the factor levels #####
ord <- aggregate(yo$IndividualCount, by=list(yo$ScientificName), FUN=sum) 
ord <- ord %>% arrange(desc(x))
ord <- ord[1:10,]
ord$EventID <- id
ord$Group.1 <- factor(ord$Group.1, levels = rev(ord$Group.1))

##### make the pie chart ##### 
p <- ggplot(ord, aes(x = EventID, y = x, fill = Group.1)) +
  geom_bar(stat = "identity", position = "fill" ) +
  coord_flip() +
  scale_colour_manual(values = mycols, 
                      aesthetics = c("colour", "fill"),
                      drop = FALSE, guide=guide_legend(reverse=TRUE, nrow=5)) +
  theme_minimal() + 
  theme(legend.position = "bottom", 
        legend.title = element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y = element_blank(),
        text = element_text(size=30), 
        legend.text=element_text(size=24)) +
  coord_polar("y", start=0) + 
  #guides(fill = guide_legend(nrow = 2)) +
  labs(title = NULL, y = NULL, x = NULL)
print(id)
print(p)

##### save each pie chat to a folder #####

setwd("C:/rworking/deepseatools/reports/2020_Ocean_Sciences_Hourigan/figures")
ggsave(paste(id,'.png',sep=''))

}
###### clean up everything except core objects ###### 
rm(list=setdiff(ls(), c("filt", "indata", "d", "z")))
```

## NMDS   
### Setting up depth, temp, and location categories

```{r setdepthclass, echo=FALSE, cache = FALSE} 

##### setting depth categories ##### 
z$DepthCat[as.numeric(z$DepthInMeters) <= 1000] <- "shallow"
z$DepthCat[as.numeric(z$DepthInMeters) > 1000] <- "deep"
z$DepthCat <- factor(z$DepthCat, levels = c("shallow", "deep"))

##### setting location categories ##### 
z$loc[as.numeric(z$Latitude) > 10] <- "north"
z$loc[as.numeric(z$Latitude) < 10] <- "south"

##### setting temperature categories #####
z$temp_cat[as.numeric(z$Temperature) > 5] <- "warm: > 5"
z$temp_cat[as.numeric(z$Temperature) >= 3 & as.numeric(z$Temperature) <= 5] <- "cold: 3 to 5"
z$temp_cat[as.numeric(z$Temperature) < 3] <- "colder: less than 3"
z$temp_cat <- factor(z$temp_cat, levels = c("warm: > 5", "cold: 3 to 5", "colder: less than 3"))

##### setting oxygen categories #####
# qplot(as.numeric(z$Oxygen))
z$ox_cat[as.numeric(z$Oxygen) >= 4] <- "rich"
z$ox_cat[as.numeric(z$Oxygen) < 4] <- "poor"

##### longitude cat #####
z$longloc[as.numeric(z$Longitude) > 90] <- "west"
z$longloc[as.numeric(z$Longitude) < 90] <- "east"

```

### Filter dataset 

```{r}
x <- z %>% 
  filter(
    # DepthCat == 'deep',
    # EventID %in% c("D2-EX1504L2-11", "D2-EX1504L2-07", 
    #                "D2-EX1504L2-13", "NA101-H1722", "D2-EX1708-12",
    #                "D2-EX1708-07", "D2-EX1708-17", "D2-EX1708-03")
    #Phylum == "Porifera"
  )

```

### Create site by species matrix.

```{r}

##### create site X species matrix #####

site.sp <- dcast(x, EventID ~ ScientificName, value.var = "IndividualCount", fun.aggregate = sum)

# setwd("C:/rworking/digs/outdata")
# write.csv(site.sp, 'site.sp.csv')

# creating a site variable
site.sp$site <- site.sp$EventID

# getting rid of non-needed variables
site.sp <- site.sp %>%
  dplyr::select(-EventID)

# # moving the site variable to the beginning
# col_idx <- grep("site", names(site.sp))
# site.sp <- site.sp[, c(col_idx, (1:ncol(site.sp))[-col_idx])]
# # names(site.sp)

# set the site as the row.names
row.names(site.sp) <- site.sp$site

# remove the site variable
site.sp <- site.sp %>%
  dplyr::select(-site)

# making it a matrix
site.sp <- as.matrix(site.sp)

```

### Calculating diversity metrics 

```{r}

alpha = mean(specnumber(site.sp))
beta  = ncol(site.sp)/alpha - 1
alpha
beta

```

### Calculating dissimilarity matrices 

```{r}
## Bray-Curtis as binary (presence/absence) of species per equation above
sp_bray_binary = vegdist(site.sp, method='bray', binary=T)

## Bray-Curtis as amount (density) of species
sp_bray = vegdist(site.sp, method='bray')

## transformed Bray-Curtis as amount (density) of species
sp_bray_transformed = vegdist(wisconsin(sqrt(site.sp)), method='bray')

## write to data folder
# setwd("C:/rworking/deepseatools/reports/2020_Ocean_Sciences_Hourigan/tables")
# write.csv(as.matrix(sp_bray_binary),'sp_bray_binary.csv')
# write.csv(as.matrix(sp_bray), 'sp_bray.csv')
# write.csv(as.matrix(sp_bray_transformed), 'sp_bray_transformed.csv')

```

### Look at histgrams of ecological distance

```{r}

i = lower.tri(as.matrix(sp_bray_binary))
hist(as.matrix(sp_bray_binary)[i]     , 
     xlab='ecological distance', main='sp_bray_binary')

hist(as.matrix(sp_bray)[i]            ,  
     xlab='ecological distance', main='sp_bray')

hist(as.matrix(sp_bray_transformed)[i],  
     xlab='ecological distance', main='sp_bray_transformed')

##### checking ##### 
## look at individual values
m = as.matrix(sp_bray_transformed)
diag(m) = NA
i_max = which(m == max(m, na.rm=T), arr.ind=T)
i_min = which(m == min(m, na.rm=T), arr.ind=T)

View(m)

##### checking ##### 
## look at pairs of sites.
# y <- z %>% filter(EventID == 'D2-EX1708-13')
# a <- z %>% filter(EventID == 'D2-EX1504L2-10')
# b <- z %>% filter(EventID == 'D2-EX1605L3-06')
# 
# length()
# intersect(y$ScientificName, a$ScientificName)
# intersect(b$ScientificName, a$ScientificName)

```


### Run NMDS on site X species matrix #####

```{r}

NMDS <- metaMDS(site.sp, distance = "bray", binary = F, k=3, trymax = 30)

```

### Extracting the site and species scores for use with ggplot2 ##### 

```{r}

#Using the scores function from vegan to extract the site scoresand convert to a data.frame
site.scores <- as.data.frame(scores(NMDS))  
site.scores$site <- rownames(site.scores)  # create a column of site names, from the rownames of data.scores
# head(site.scores)  #look at the data
# merge information from d (to get categories)
site.scores <- merge(site.scores, z, by.x = 'site', by.y = 'EventID')

#summarize by site 
site.scores <- site.scores %>% group_by(site) %>% summarise(NMDS1 = NMDS1[1],
                                                            NMDS2 = NMDS2[1],
                                                            DepthCat = DepthCat[1],
                                                            temp_cat = temp_cat[1],
                                                            loc = loc[1])

# extracting species scores
species.scores <- as.data.frame(scores(NMDS, "species"))  #Using the scores function from vegan to extract the species scores and convert to a data.frame
species.scores$species <- rownames(species.scores)  # create a column of species, from the rownames of species.scores
# head(species.scores)  #look at the data
species.scores$spec_code <- 1:(length(species.scores$species))

#NMDS

```

### Plotting NMDS results

```{r}

ggplot() +
  geom_point(data=site.scores, aes(x=NMDS2,y=NMDS1, color= DepthCat), size=5) + # add the point markers
  geom_text(data=site.scores,aes(x=NMDS2,y=NMDS1, label = site), size=4, position = position_nudge(y = .0001)) + # add the text
  #geom_point(data=species.scores,aes(x=NMDS1,y=NMDS2), shape = 3, size=4) +
  coord_equal() +
  theme_bw(base_size = 15, base_family = "Cambria") +
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        panel.background = element_blank(),
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank())

ggplot() +
  geom_point(data=site.scores,aes(x=NMDS2,y=NMDS1, color = temp_cat), size=5) + # add the point markers
  geom_text(data=site.scores,aes(x=NMDS2,y=NMDS1, label = site), size=4, position = position_nudge(y = .0001)) + # add the text
  #geom_point(data=species.scores,aes(x=NMDS1,y=NMDS2), shape = 3, size=4) + 
  coord_equal() +
  theme_bw(base_size = 15, base_family = "Cambria") +
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank())

# ggplot() +
#   geom_point(data=site.scores,aes(x=NMDS2,y=NMDS1, color=loc), size=5) + # add the point markers
#   geom_text(data=site.scores,aes(x=NMDS2,y=NMDS1, label = site), size=4, position = position_nudge(y = .0001)) + # add the text
#   #geom_point(data=species.scores,aes(x=NMDS1,y=NMDS2), shape = 3, size=4) + 
#   coord_equal() +
#   theme_bw(base_size = 15, base_family = "Cambria") +
#   theme(axis.text.x = element_blank(),  # remove x-axis text
#         axis.text.y = element_blank(), # remove y-axis text
#         axis.ticks = element_blank(),  # remove axis ticks
#         axis.title.x = element_text(size=18), # remove x-axis labels
#         axis.title.y = element_text(size=18), # remove y-axis labels
#         panel.background = element_blank(), 
#         panel.grid.major = element_blank(),  #remove major-grid labels
#         panel.grid.minor = element_blank(),  #remove minor-grid labels
#         plot.background = element_blank())
# 
# 
ggplot() +
  #geom_point(data=site.scores,aes(x=NMDS2,y=NMDS1, color=loc), size=5) + # add the point markers
  geom_text(data=species.scores,aes(x=NMDS2,y=NMDS1, label = species), size=4, position = position_nudge(y = .0001)) + # add the text
  geom_point(data=site.scores,aes(x=NMDS2,y=NMDS1, color = temp_cat), shape = 3, size=4) +
  coord_equal() +
  theme_bw(base_size = 15, base_family = "Cambria") +
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        panel.background = element_blank(),
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank())

print(NMDS)

```

### Looking at clusters 

```{r}
y <- vegdist(site.sp, method = "bray", binary = F)
#heatmap(as.matrix(y))

# h = hclust(y)
# plot(h, main = "Clustering cars using Bray Curtis method", sub = "", xlab="", axes = FALSE, hang = -1)
# lines(x = c(0,0), y = c(0,200), type = "n") # force extension of y axis
# axis(side = 2, at = seq(0,100,10), labels = seq(100,0,-10))

fit <- hclust(sp_bray_transformed)
plot(fit, main = "Clustering cars using Bray Curtis method", sub = "", xlab="", axes = FALSE) # display dendogram
groups <- cutree(fit, k=4) # cut tree into 5 clusters
# draw dendogram with red borders around the 5 clusters
rect.hclust(fit, k=4, border="red")

```

### Table of species codes 

```{r spec_codes, echo=F}
x <- species.scores %>%  dplyr::select(spec_code)

kable(x)

```

### Diversity estimates.

#### Table: Shannon diversity per EventID
All taxa (both coral and sponges at all taxonomic ranks included in estimates).

```{r diversity, echo=F}

plots <- diversity(site.sp, index = "shannon") #makes an object
#that the diversity values #are written to.
# summary(plots) #gives summary statistics for the plots
# median(plots) #gives the median
# mean(plots) #gives the mean
# class(plots)
# max(plots[1:30])
# # View(plots)

# creating plots
plots_df <- data.frame(plots)
# View(plots_df)
plots_df$EventID <- row.names(plots_df)
# names(d)
d <- merge(d, plots_df, all.x = T)
# length(d$EventID)
# length(unique(d$EventID))
x <- d %>% group_by(EventID, plots) %>% summarize(n=n(), 
                                                 Latitude = paste(mean(as.numeric(Latitude)), collapse = ' | '), 
                                                 Longitude = paste(mean(as.numeric(Longitude)), collapse = ' | '))
#View(x)

setwd("C:/rworking/deepseatools/reports")
write.xlsx(x, '20190725-0_Summary_of_Corals_and_Sponges_Diversity_by_EventID.xlsx') 

## Calculating various Diversity Indices

# H <- diversity(site.sp)
# simp <- diversity(site.sp, "simpson")
# invsimp <- diversity(site.sp, "inv")
# ## Unbiased Simpson (Hurlbert 1971, eq. 5) with rarefy:
# unbias.simp <- rarefy(site.sp, 2) - 1
# ## Fisher alpha
# alpha <- fisher.alpha(site.sp)
# ## Plot all
# pairs(cbind(H, simp, invsimp, unbias.simp, alpha, J), pch="+", col="blue")
# ## Species richness (S) and Pielou's evenness (J):
# S <- specnumber(site.sp) ## rowSums(site.sp > 0) does the same...
# J <- H/log(S)

```
