---
title: "2020_OceanSciences_NMDS"
author: "Robert McGuinn"
date: "2/11/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

## NMDS   
### Setting up depth, temp, and location categories

```{r setdepthclass, echo=FALSE, cache = FALSE} 

##### setting depth categories ##### 
z$DepthCat[as.numeric(z$DepthInMeters) <= 1000] <- "shallow"
z$DepthCat[as.numeric(z$DepthInMeters) > 1000] <- "deep"
z$DepthCat <- factor(z$DepthCat, levels = c("shallow", "deep"))

##### setting location categories ##### 
z$loc[as.numeric(z$Latitude) > 10] <- "north"
z$loc[as.numeric(z$Latitude) < 10] <- "south"

##### setting temperature categories #####
z$temp_cat[as.numeric(z$Temperature) > 5] <- "warm: > 5"
z$temp_cat[as.numeric(z$Temperature) >= 3 & as.numeric(z$Temperature) <= 5] <- "cold: 3 to 5"
z$temp_cat[as.numeric(z$Temperature) < 3] <- "colder: less than 3"
z$temp_cat <- factor(z$temp_cat, levels = c("warm: > 5", "cold: 3 to 5", "colder: less than 3"))

##### setting oxygen categories #####
# qplot(as.numeric(z$Oxygen))
z$ox_cat[as.numeric(z$Oxygen) >= 4] <- "rich"
z$ox_cat[as.numeric(z$Oxygen) < 4] <- "poor"

##### longitude cat #####
z$longloc[as.numeric(z$Longitude) > 90] <- "west"
z$longloc[as.numeric(z$Longitude) < 90] <- "east"

```

### Filter dataset 

```{r}
x <- z %>% 
  filter(
    # DepthCat == 'deep',
    # EventID %in% c("D2-EX1504L2-11", "D2-EX1504L2-07", 
    #                "D2-EX1504L2-13", "NA101-H1722", "D2-EX1708-12",
    #                "D2-EX1708-07", "D2-EX1708-17", "D2-EX1708-03")
    #Phylum == "Porifera"
  )

```

### Create site by species matrix.

```{r}

##### create site X species matrix #####

site.sp <- dcast(x, EventID ~ ScientificName, value.var = "IndividualCount", fun.aggregate = sum)

# setwd("C:/rworking/digs/outdata")
# write.csv(site.sp, 'site.sp.csv')

# creating a site variable
site.sp$site <- site.sp$EventID

# getting rid of non-needed variables
site.sp <- site.sp %>%
  dplyr::select(-EventID)

# # moving the site variable to the beginning
# col_idx <- grep("site", names(site.sp))
# site.sp <- site.sp[, c(col_idx, (1:ncol(site.sp))[-col_idx])]
# # names(site.sp)

# set the site as the row.names
row.names(site.sp) <- site.sp$site

# remove the site variable
site.sp <- site.sp %>%
  dplyr::select(-site)

# making it a matrix
site.sp <- as.matrix(site.sp)

```

### Calculating diversity metrics 

```{r}

alpha = mean(specnumber(site.sp))
beta  = ncol(site.sp)/alpha - 1
alpha
beta

```

### Calculating dissimilarity matrices 

```{r}
## Bray-Curtis as binary (presence/absence) of species per equation above
sp_bray_binary = vegdist(site.sp, method='bray', binary=T)

## Bray-Curtis as amount (density) of species
sp_bray = vegdist(site.sp, method='bray')

## transformed Bray-Curtis as amount (density) of species
sp_bray_transformed = vegdist(wisconsin(sqrt(site.sp)), method='bray')

## write to data folder
# setwd("C:/rworking/deepseatools/reports/2020_Ocean_Sciences_Hourigan/tables")
# write.csv(as.matrix(sp_bray_binary),'sp_bray_binary.csv')
# write.csv(as.matrix(sp_bray), 'sp_bray.csv')
# write.csv(as.matrix(sp_bray_transformed), 'sp_bray_transformed.csv')

```

### Look at histgrams of ecological distance

```{r}

i = lower.tri(as.matrix(sp_bray_binary))
hist(as.matrix(sp_bray_binary)[i]     , 
     xlab='ecological distance', main='sp_bray_binary')

hist(as.matrix(sp_bray)[i]            ,  
     xlab='ecological distance', main='sp_bray')

hist(as.matrix(sp_bray_transformed)[i],  
     xlab='ecological distance', main='sp_bray_transformed')

##### checking ##### 
## look at individual values
m = as.matrix(sp_bray_transformed)
diag(m) = NA
i_max = which(m == max(m, na.rm=T), arr.ind=T)
i_min = which(m == min(m, na.rm=T), arr.ind=T)

View(m)

##### checking ##### 
## look at pairs of sites.
# y <- z %>% filter(EventID == 'D2-EX1708-13')
# a <- z %>% filter(EventID == 'D2-EX1504L2-10')
# b <- z %>% filter(EventID == 'D2-EX1605L3-06')
# 
# length()
# intersect(y$ScientificName, a$ScientificName)
# intersect(b$ScientificName, a$ScientificName)

```


### Run NMDS on site X species matrix #####

```{r}

NMDS <- metaMDS(site.sp, distance = "bray", binary = T, k=3, trymax = 30)

```

### Extracting the site and species scores for use with ggplot2 ##### 

```{r}

#Using the scores function from vegan to extract the site scoresand convert to a data.frame
site.scores <- as.data.frame(scores(NMDS))  
site.scores$site <- rownames(site.scores)  # create a column of site names, from the rownames of data.scores
# head(site.scores)  #look at the data
# merge information from d (to get categories)
site.scores <- merge(site.scores, z, by.x = 'site', by.y = 'EventID')

#summarize by site 
site.scores <- site.scores %>% group_by(site) %>% summarise(NMDS1 = NMDS1[1],
                                                            NMDS2 = NMDS2[1],
                                                            DepthCat = DepthCat[1],
                                                            temp_cat = temp_cat[1],
                                                            ox_cat = ox_cat[1],
                                                            longloc = longloc[1],
                                                            loc = loc[1])

# extracting species scores
species.scores <- as.data.frame(scores(NMDS, "species"))  #Using the scores function from vegan to extract the species scores and convert to a data.frame
species.scores$species <- rownames(species.scores)  # create a column of species, from the rownames of species.scores
# head(species.scores)  #look at the data
species.scores$spec_code <- 1:(length(species.scores$species))

#NMDS

```

### Plotting NMDS results

```{r}

# ggplot() +
#   geom_point(data=site.scores, aes(x=NMDS2,y=NMDS1, color= longloc), size=5) + # add the point markers
#   geom_text_repel(data=site.scores,aes(x=NMDS2,y=NMDS1, label = site), size=4, position = position_nudge(y = .0001)) + # add the text
#   #geom_point(data=species.scores,aes(x=NMDS1,y=NMDS2), shape = 3, size=4) +
#   coord_equal() +
#   theme_bw(base_size = 15, base_family = "Cambria") +
#   theme(axis.text.x = element_blank(),  # remove x-axis text
#         axis.text.y = element_blank(), # remove y-axis text
#         axis.ticks = element_blank(),  # remove axis ticks
#         axis.title.x = element_text(size=18), # remove x-axis labels
#         axis.title.y = element_text(size=18), # remove y-axis labels
#         panel.background = element_blank(),
#         panel.grid.major = element_blank(),  #remove major-grid labels
#         panel.grid.minor = element_blank(),  #remove minor-grid labels
#         plot.background = element_blank())

ggplot() +
  geom_point(data=site.scores,aes(x=NMDS2,y=NMDS1, color = temp_cat), size=5) + # add the point markers
  geom_text_repel(data=site.scores,aes(x=NMDS2,y=NMDS1, label = site), size=4, position = position_nudge(y = .0001)) + # add the text
  #geom_point(data=species.scores,aes(x=NMDS1,y=NMDS2), shape = 3, size=4) + 
  coord_equal() +
  theme_bw(base_size = 15, base_family = "Cambria") +
  labs(color = "Temperature Category (Celcius)") +
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank())

# ggplot() +
#   geom_point(data=site.scores,aes(x=NMDS2,y=NMDS1, color=loc), size=5) + # add the point markers
#   geom_text(data=site.scores,aes(x=NMDS2,y=NMDS1, label = site), size=4, position = position_nudge(y = .0001)) + # add the text
#   #geom_point(data=species.scores,aes(x=NMDS1,y=NMDS2), shape = 3, size=4) + 
#   coord_equal() +
#   theme_bw(base_size = 15, base_family = "Cambria") +
#   theme(axis.text.x = element_blank(),  # remove x-axis text
#         axis.text.y = element_blank(), # remove y-axis text
#         axis.ticks = element_blank(),  # remove axis ticks
#         axis.title.x = element_text(size=18), # remove x-axis labels
#         axis.title.y = element_text(size=18), # remove y-axis labels
#         panel.background = element_blank(), 
#         panel.grid.major = element_blank(),  #remove major-grid labels
#         panel.grid.minor = element_blank(),  #remove minor-grid labels
#         plot.background = element_blank())
# 
# 
ggplot() +
  #geom_point(data=site.scores,aes(x=NMDS2,y=NMDS1, color=loc), size=5) + # add the point markers
  geom_point(data=site.scores,aes(x=NMDS2,y=NMDS1, color = temp_cat), size=5) +
  geom_text(data=species.scores,aes(x=NMDS2,y=NMDS1, label = species), size=4, position = position_nudge(y = .0001)) + # add the text
  
  coord_equal() +
  theme_bw(base_size = 15, base_family = "Cambria") +
  labs(color = "Temperature Category (Celcius)") +
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        panel.background = element_blank(),
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank())


print(NMDS)

```

### Looking at clusters 

```{r}
## Bray-Curtis as binary (presence/absence) of species per equation above
sp_bray_binary = vegdist(site.sp, method='bray', binary=T)

## Bray-Curtis as amount (density) of species
sp_bray = vegdist(site.sp, method='bray')

## transformed Bray-Curtis as amount (density) of species
sp_bray_transformed = vegdist(wisconsin(sqrt(site.sp)), method='bray')


# h = hclust(y)
# plot(h, main = "Clustering cars using Bray Curtis method", sub = "", xlab="", axes = FALSE, hang = -1)
# lines(x = c(0,0), y = c(0,200), type = "n") # force extension of y axis
# axis(side = 2, at = seq(0,100,10), labels = seq(100,0,-10))

fit <- hclust(sp_bray_transformed)
plot(fit, main = "", sub = "", xlab="Sites", axes = T, ylab = "Ecological Distance (Bray-Curtis - Wisconsin Transformed)", hang = -1) # display dendogram
groups <- cutree(fit, k=5) # cut tree into 5 clusters
# draw dendogram with red borders around the 5 clusters
rect.hclust(fit, k=5, border="red")

fit <- hclust(sp_bray_binary)
plot(fit, main = "", sub = "", xlab="Sites", axes = T, ylab = "Ecological Distance (Sorensen)", hang = -1) # display dendogram
groups <- cutree(fit, k=5) # cut tree into 5 clusters
# draw dendogram with red borders around the 5 clusters
rect.hclust(fit, k=5, border="red")

fit <- hclust(sp_bray)
plot(fit, main = "", sub = "", xlab="Sites", axes = T, ylab = "Ecological Distance (Bray-Curtis)", hang = -1) # display dendogram
groups <- cutree(fit, k=5) # cut tree into 5 clusters
# draw dendogram with red borders around the 5 clusters
rect.hclust(fit, k=5, border="red")


plot(NMDS, display='sites')
ordihull(NMDS, groups, lty=2, col='red', label=T)

```

### Table of species codes 

```{r spec_codes, echo=F}
x <- species.scores %>%  dplyr::select(spec_code)

kable(x)

```

### Diversity estimates.

#### Table: Shannon diversity per EventID
All taxa (both coral and sponges at all taxonomic ranks included in estimates).

```{r diversity, echo=F}

plots <- diversity(site.sp, index = "shannon") #makes an object
#that the diversity values #are written to.
# summary(plots) #gives summary statistics for the plots
# median(plots) #gives the median
# mean(plots) #gives the mean
# class(plots)
# max(plots[1:30])
# # View(plots)

# creating plots
plots_df <- data.frame(plots)
# View(plots_df)
plots_df$EventID <- row.names(plots_df)
# names(d)
d <- merge(d, plots_df, all.x = T)
# length(d$EventID)
# length(unique(d$EventID))
x <- d %>% group_by(EventID, plots) %>% summarize(n=n(), 
                                                  Latitude = paste(mean(as.numeric(Latitude)), collapse = ' | '), 
                                                  Longitude = paste(mean(as.numeric(Longitude)), collapse = ' | '))
#View(x)

setwd("C:/rworking/deepseatools/reports")
write.xlsx(x, '20190725-0_Summary_of_Corals_and_Sponges_Diversity_by_EventID.xlsx') 

## Calculating various Diversity Indices

# H <- diversity(site.sp)
# simp <- diversity(site.sp, "simpson")
# invsimp <- diversity(site.sp, "inv")
# ## Unbiased Simpson (Hurlbert 1971, eq. 5) with rarefy:
# unbias.simp <- rarefy(site.sp, 2) - 1
# ## Fisher alpha
# alpha <- fisher.alpha(site.sp)
# ## Plot all
# pairs(cbind(H, simp, invsimp, unbias.simp, alpha, J), pch="+", col="blue")
# ## Species richness (S) and Pielou's evenness (J):
# S <- specnumber(site.sp) ## rowSums(site.sp > 0) does the same...
# J <- H/log(S)

```

```{r}
varespec <- site.sp
vare.mds <- metaMDS(varespec, binary = F, try = 50)  #using all the defaults

data.scores <- as.data.frame(scores(vare.mds))  #Using the scores function from vegan to extract the site scores and convert to a data.frame
data.scores$site <- rownames(data.scores)  # create a column of site names, from the rownames of data.scores
data.scores <- merge(data.scores, z, by.x = 'site', by.y = 'EventID')
names(data.scores)

##### put results from hierarchical groupings ##### 

groups <- data.frame(groups)
groups$site<- row.names(groups)
data.scores <-merge(data.scores, groups)

##### extract species scores #####
species.scores <- as.data.frame(scores(vare.mds, "species"))  #Using the scores function from vegan to extract the species scores and convert to a data.frame
species.scores$species <- rownames(species.scores)  # create a column of species, from the rownames of species.scores
head(species.scores)  #look at the data

##### get hulls of hclust groups onto plot ##### 
grp.a <- data.scores[data.scores$groups == 1, ][chull(data.scores[data.scores$groups == 1, c("NMDS1", "NMDS2")]), ]  # hull values for grp A

grp.b <- data.scores[data.scores$groups == 2, ][chull(data.scores[data.scores$groups == 2, c("NMDS1", "NMDS2")]), ]  # hull values for grp B

grp.c <- data.scores[data.scores$groups == 3, ][chull(data.scores[data.scores$groups == 3, c("NMDS1", "NMDS2")]), ]  # hull values for grp B

grp.d <- data.scores[data.scores$groups == 4, ][chull(data.scores[data.scores$groups == 4, c("NMDS1", "NMDS2")]), ]  # hull values for grp B

grp.e <- data.scores[data.scores$groups == 5, ][chull(data.scores[data.scores$groups == 5, c("NMDS1", "NMDS2")]), ]  # hull values for grp B

hull.data <- rbind(grp.a, grp.b, grp.c, grp.d, grp.e)  #combine grp.a and grp.b
hull.data

##### get hulls of temperature onto plot ##### 
grp.a <- data.scores[data.scores$temp_cat == "warm: > 5", ][chull(data.scores[data.scores$temp_cat =="warm: > 5", c("NMDS1", "NMDS2")]), ]  # hull values for grp A

grp.b <- data.scores[data.scores$temp_cat == "cold: 3 to 5", ][chull(data.scores[data.scores$temp_cat =="cold: 3 to 5", c("NMDS1", "NMDS2")]), ]  # hull values for grp B

grp.c <- data.scores[data.scores$temp_cat == "colder: less than 3", ][chull(data.scores[data.scores$temp_cat =="colder: less than 3", c("NMDS1", "NMDS2")]), ]  # hull values for grp B

hull.data <- rbind(grp.a, grp.b, grp.c)  #combine grp.a and grp.b
hull.data$groups <- as.character(hull.data$groups)

##### plot with hulls ##### 
ggplot() + 
  geom_polygon(data=hull.data, aes(x=NMDS1,y=NMDS2,fill=groups,group=groups),alpha=0.30) + # add the convex hulls
  #geom_text_repel(data=species.scores, aes(x=NMDS1,y=NMDS2,label=species),alpha=0.5) +  # add the species labels
    geom_text_repel(data=data.scores, aes(x=NMDS1,y=NMDS2,label=site),alpha=0.5) +  # add the species labels
  geom_point(data=data.scores,aes(x=NMDS1,y=NMDS2, color = temp_cat), size=4) + # add the point markers
  scale_colour_manual(values=c("warm: > 5" = "red",
                               "cold: 3 to 5" = "yellow",
                               "colder: less than 3" = "blue")) +
  scale_fill_manual(values = c("1" = "red", 
                               "2" = "green", 
                               "3" = "blue", 
                               "4" = "pink", 
                               "5" = "gray")) +
  coord_flip() +
  theme_bw() + 
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank())


```


