---
title: "NOAA National Database for Deep-sea Corals and Sponges: Area of Interest / Query Summary"
author: "NOAA-NFMS-OHC-DSCRTP"
date: 'Robert.McGuinn@NOAA.gov; 843-460-9696. Report last ran: `r Sys.Date()`'
output: word_document
---

```{r packages, echo=F, warning=F, message=F}
#install.packages(tidyverse)
#install.packages('openxlsx')
library(tidyverse)
library(openxlsx)
library(sf)
library(rgdal)
library(RColorBrewer)
library(raster)
library(marmap)
library(leaflet)
library(extrafont)
library(RColorBrewer)
library(rmarkdown)
library(knitr)
library(maps)
library(rgdal)
library(raster)
library(marmap)
library(httr)
library(jsonlite)
library(rnaturalearth)
library(rnaturalearthdata)
library(openxlsx)
library(rgeos)

```

```{r load_NDB, eval=T, echo=F, message=FALSE, warning=FALSE, cache=TRUE}
##### input: latest version of NDB #####
setwd("C:/rworking/deepseatools/indata")
indata <- read_csv("DSCRTP_NatDB_20201021-0.csv")
filt <- indata %>%
  filter(Flag == "0", is.na(Phylum) == F)
rm(indata)
```

```{r filter_query_OR, echo=F, warning=F, message=F, eval=T}
##### get all the DatasetIDs in SEDCI region #####
ecogig <- c("NOAA_EX-12-02-L2",
            "BOEM_Lophelia_II",
            "WHOI_AT-18",
            "SOI_FK006B",
            "NOAA_HC-11-10",
            "OET_NA028",
            "OET_NA043",
            "OET_NA057",
            "OET_NA058",
            "ECOGIG_OI-16-10",
            "ECOGIG_O2-17-06")

ecogig <- unique(ecogig)

## DatasetIDs in the SEDCI rescue project (as of 2020-10-19)
sedci_rescue <- c("HBOI_SJ-10-07", "HBOI_SJ-10-07", "NOAA_SJ-09-08", "NOAA_SJ-09-08","HBOI_SJ-08-09","NOAA_SJ-07-05", "NOAA_SJ-07-05", "NOAA_SJ-07-06", "HBOI_SJ-06-02", "HBOI_SJ-06-05-L1", "HBOI_SJ-06-05-L3", "HBOI_SJ-06-02", "NOAA_SJ-05-04", "NOAA_SJ-05-08", "NOAA_SJ-05-10", "NOAA_SJ-05-11", "NOAA_SJ-04-05", "NOAA_SJ-02-08", "HBOI_EL-99-08", "HBOI_SJ-88-11", "USGS_TM-97-10", "NOAA_RB-03-07", "NOAA_SJ-05-11",
                  "NOAA_NF-09-01", "NOAA_NF-08-04", "NOAA_NF-07-06", "NOAA_NF-05-05", "NOAA_SJ-07-06", "HBOI_SJ-08-09", "NOAA_SJ-09-08", "NOAA_PC-10-02",
                  "NOAA_NF-14-01", "NOAA_NF-13-02", "NOAA_NF-12-01", "NOAA_NF-11-09", "NOAA_NF-11-09-L3", "NOAA_PC-11-05-L1", "NOAA_NF-11-09", "NOAA_NF-11-09-L3", "Sharuga_SM_2014", "NOAA_PC-12-03", "NOAA_NF-13-02", "NOAA_PC-13-03", "NOAA_NF-14-08","NOAA_PC-15-02", "NOAA_PC-16-02", "OET_NA058", "OET_NA028", "NOAA_PC-17-02", "HBOI_WS-17-125")

## unique DatasetIDs
sedci_rescue <- unique(sedci_rescue)

## sedci_rescue add ecogig
sedci_rescue <- union(sedci_rescue, ecogig)

## datasetIDs in the SEDCI rescue project (as of 2020-11-25)
sedci_new <- c("NOAA_DFH-30", "NOAA_NF-17-08",
               "NOAA_NF-17-08", "NOAA_DFH-32-33",
               "NOAA_DFH-32-33", "NOAA_EX-17-11",
               "NOAA_EX-18-03", "NOAA_EX-18-06",
               "NOAA_NF-18-04","NOAA_FGBNMS_DFH-35",
               "NOAA_FGBNMS_DFH-37", "NOAA_RESTORE_MT18")

sedci_new <- unique(sedci_new)

##### subset data #####
x <- filt %>% filter(DatasetID %in% sedci_new | DatasetID %in% sedci_rescue) 
##### build a variable for new vs. rescue datasets #####
x$sedci <-
  case_when(
    x$DatasetID %in% sedci_rescue ~ "rescue" ,
    x$DatasetID %in% sedci_new ~ "new",
  )

## define not in
`%notin%` <- Negate(`%in%`)

# length(x$Flag)
# table(x$sedci, useNA = 'always')

x$sedci2 <- 
  case_when(
  x$DatasetID %in% sedci_rescue ~ "Post SEDCI" ,
  x$DatasetID %in% sedci_new ~ "Post SEDCI",
  x$DatasetID %notin% sedci_rescue ~ "Pre SEDCI",
  x$DatasetID %notin% sedci_new ~ "Pre SEDCI"
)

x$sedci2 <- factor(x$sedci2, levels = c('Post SEDCI', 'Pre SEDCI'))
## check
# x %>% pull(CatalogNumber) %>% length()

# 
# filt %>%
#   filter(grepl("EX", DatasetID),
#          is.na(ImageURL) == F#,
#          #ObservationDate == '-999'
#          ) %>%
#   pull(DatasetID) %>% 
#   unique() %>% 
#   length()

```

```{r bbox_from_query, echo=F, warning=F, message=F, eval=T}
## set bounding box variables with query results.
minlat <- min(x$Latitude)
maxlat <- max(x$Latitude)
minlon <- min(x$Longitude)
maxlon <- max(x$Longitude)
```

```{r export_GIS, echo=FALSE, echo = FALSE, message = FALSE, warning=FALSE, eval=T}
##### load packages #####
library(arcgisbinding)
arc.check_product()

x_geo <- x

##### create spdf #####

coordinates(x_geo) <- c("Longitude", "Latitude")
proj4string(x_geo) <- "+proj=longlat +ellps=WGS84 +datum=WGS84"

##### create feature-class #####

# fgdb_path <- 'C:/rworking/sf/sf.gdb'
# arc.write(file.path(fgdb_path, 'x_geo'), data=x_geo, overwrite = TRUE)

```

```{r, echo=F}
x <- x %>%
  mutate(
    EcoRegion = case_when(
      Latitude >= 32 ~ "Northern Carolinian",
      Latitude < 32 & Latitude >= 28.489 ~ "Southern Carolinian", 
      Latitude < 28.489 ~ "Floridian"
    )
  )

x$EcoRegion <- factor(x$EcoRegion, levels = c('Northern Carolinian', 'Southern Carolinian', 'Floridian'))

class(x$EcoRegion)
table(x$EcoRegion)

```

```{r write_data, echo=F, warning=F, message=F, eval=F}
## write data to Excel
# setwd("C:/rworking/deepseatools/indata")
# write.xlsx(x,"20201019-2_subset_SEDCI_from_DSCRTP_NatDBNatDB_20200710-2_RPMcGuinn.xlsx",
#            row.names = F, quote = T)
```

```{r erddap_link, echo = FALSE}
x$erddap <- paste("https://www.ncei.noaa.gov/erddap/tabledap/deep_sea_corals.csv?ShallowFlag%2CDatabaseVersion%2CDatasetID%2CCatalogNumber%2CSampleID%2CTrackingID%2CImageURL%2CHighlightImageURL%2CCitation%2CRepository%2CScientificName%2CVerbatimScientificName%2CVernacularNameCategory%2CVernacularName%2CTaxonRank%2CAphiaID%2CLifeScienceIdentifier%2CPhylum%2CClass%2CSubclass%2COrder%2CSuborder%2CFamily%2CSubfamily%2CGenus%2CSubgenus%2CSpecies%2CSubspecies%2CScientificNameAuthorship%2CTypeStatus%2COperationalTaxonomicUnit%2CMorphospecies%2CCombinedNameID%2CSynonyms%2CIdentificationComments%2CIdentifiedBy%2CIdentificationDate%2CIdentificationQualifier%2CIdentificationVerificationStatus%2CAssociatedSequences%2COcean%2CLargeMarineEcosystem%2CCountry%2CFishCouncilRegion%2CLocality%2Clatitude%2Clongitude%2CDepthInMeters%2CDepthMethod%2CMinimumDepthInMeters%2CMaximumDepthInMeters%2CLocationComments%2CObservationDate%2CObservationYear%2CObservationTime%2CSurveyID%2CVessel%2CPI%2CPIAffiliation%2CPurpose%2CSurveyComments%2CStation%2CEventID%2CSamplingEquipment%2CVehicleName%2CSampleAreaInSquareMeters%2CfootprintWKT%2CfootprintSRS%2CIndividualCount%2CCategoricalAbundance%2CDensity%2CCover%2CVerbatimSize%2CMinimumSize%2CMaximumSize%2CWeightInKg%2CCondition%2CAssociatedTaxa%2COccurrenceComments%2CStartLatitude%2CStartLongitude%2CEndLatitude%2CEndLongitude%2CVerbatimLatitude%2CVerbatimLongitude%2CLocationAccuracy%2CNavType%2COtherData%2CHabitat%2CSubstrate%2CCMECSGeoForm%2CCMECSSubstrate%2CCMECSBiotic%2CTemperature%2CSalinity%2COxygen%2CpH%2CpHscale%2CpCO2%2CTA%2CDIC%2CRecordType%2CDataProvider%2CDataContact%2CModified%2CWebSite%2CEntryDate%2CReporter%2CReporterEmail%2CReporterComments%2CgisCRMDepth%2CgisGEBCODepth%2CgisEtopoDepth%2CgisMEOW%2CSynonymSearchProxy", "&latitude%3E=",
                  minlat, "&latitude%3C=", 
                  maxlat, "&longitude%3E=", 
                  minlon, "&longitude%3C=",
                  maxlon, sep='')

y <- unique(x$erddap)

```

# Download data
## Download comma delimited (CSV) file from ERDDAP: 
[LINK](`r y`)

Download data dictionary: [LINK](https://deepseacoraldata.noaa.gov/internal-documents/program-guidance/science-team-guidance-for-data-management/20170707.xlsx/at_download/file)

**Note:** When importing CSV files into Excel, use the data import wizard (under the 'data' tab) and set all column types to 'text' to avoid data tranformation issues. This step will prevent the automated alteration of dates in Excel.

## Database Version
`r paste('DSCRTP_NatDB_', x$DatabaseVersion[1], sep = '')`

# Area of Interest (AOI) Map

```{r zoom, echo=F}

# create a spatial points data frame
spdf<-x
coordinates(spdf) <- c("Longitude", "Latitude", "DepthInMeters")
proj4string(spdf) <- "+proj=longlat +ellps=WGS84 +datum=WGS84"
bbox<-bbox(spdf)

library(rnaturalearth)
library(rnaturalearthdata)

world <- ne_countries(scale = "medium", returnclass = "sf")
#class(world)

zoom <- 5
g <- ggplot(world) +
    geom_sf() +
    coord_sf(xlim = c(minlon-zoom, maxlon+zoom), ylim = c(minlat-zoom, maxlat+zoom), expand = FALSE)

# add sampling locations

g <- g + geom_point(aes(x=Longitude, y=Latitude), data=x, alpha=0.5, color = 'red', size = 2)
g

```

# 3D Bounding Box

```{r bounding_box, echo=FALSE, echo = FALSE, message = FALSE, warning=FALSE, dpi=300, fig.width=8, fig.height=4.1, eval=TRUE, fig.align='center'}
options(digits = 5)
spdf<-x
coordinates(spdf) <- c("Longitude", "Latitude", "DepthInMeters")
proj4string(spdf) <- "+proj=longlat +ellps=WGS84 +datum=WGS84"
bbox<-bbox(spdf)
bbox
```

# Summary of this Area of Interest by the Numbers

*Note:* The numbers below only reflect published records. Additional records may be retained by the Program for further review.

* **Number of records:** `r prettyNum(length(x$CatalogNumber), big.mark = ",")`
* **Number of coral records:** `r prettyNum(length(x$CatalogNumber[x$Phylum == "Cnidaria"]), big.mark = ",")` 
* **Number of sponge records:** `r prettyNum(length(x$CatalogNumber[x$Phylum == "Porifera"]), big.mark = ",")`
* **Records with images:** `r prettyNum(length(x$ImageURL[is.na(x$ImageURL) == FALSE]), big.mark = ",")`
* **Record type(s):** `r toString(unique(x$RecordType))`
* **Minimum depth (meters):** `r prettyNum(min(x[x$DepthInMeters != "-999",]$DepthInMeters), big.mark = ",")`
* **Maximum depth (meters):** `r prettyNum(max(x[x$DepthInMeters != "-999",]$DepthInMeters), big.mark = ",")`
* **Time frame:** `r min(x[x$ObservationYear != "-999",]$ObservationYear)` to `r max(x[x$ObservationYear != "-999",]$ObservationYear)`

# Where?

``` {r Ocean, echo=FALSE}
x <- x %>% mutate(Country = ifelse(is.na(Country) == T, 'International Waters', as.character(Country)))

sum_tbl <-
  x %>%
  #filter(Phylum == "Porifera") %>%
  group_by(Country) %>%
  summarize(Ocean = toString(unique(Ocean)), 
            FishCouncilRegion = toString(unique(FishCouncilRegion)),
            n = n()) %>%
  arrange(desc(n))
sum_tbl <- kable(sum_tbl, row.names = F, digits = 2, format.args = list(big.mark = ","))
sum_tbl

```

# Corals vs. sponges

```{r corals_sponges, echo=FALSE, cache = FALSE, dpi=300, fig.height=3, fig.width=8}
#library(extrafont)
#loadfonts(device = "win")

g <- ggplot(x, aes(Phylum, fill = Phylum)) +
  geom_bar() + 
 # coord_flip() + 
  theme(text = element_text(size=20)) + 
  ylab("Number of Records") + 
  theme_bw(base_size = 15, base_family = "Cambria")

#display.brewer.all(colorblindFriendly=TRUE)
g + scale_fill_manual(values = brewer.pal(12, "Paired")[c(8,7)])

```

# Distribution of Taxa by Order

```{r orders, echo=FALSE, cache = FALSE, dpi=300, fig.height=6, fig.width=8, message=FALSE, warning=FALSE}
##### building big pallette #####
# to get the colors from a palette:
palette1 <- brewer.pal(9,"Set1")
palette2 <- brewer.pal(8,"Set2")
palette3 <- brewer.pal(9,"Set3")
palette4 <- brewer.pal(8,"Dark2")
palette5 <- brewer.pal(8,"Accent")

# We can just stick a few color palettes together
big_palette <- c(palette1,palette2,palette3, palette4, palette5)

x <- within(x, 
             Order <- factor(Order, 
                                  levels=names(sort(table(Order), decreasing=TRUE))))

##### graphing the orders #####
g <- ggplot(x, aes(Order)) +
  geom_bar() + 
 # coord_flip() + 
  #theme(text = element_text(size=10)) + 
  ylab("Number of Records") + 
  facet_wrap(~Phylum, scales="free") +
  theme_bw(base_size = 12, base_family = "Cambria")

set.seed(7)
g + scale_fill_manual(values = sample(big_palette)) + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.05))

```

# Who Contributed to this Selection?

``` {r DataProvider, echo=FALSE}
# # Data Provider corrections
# x$DataProvider <- 
#   plyr::revalue(x$DataProvider,
#                 c('Kelley, Christopher (CKelley@hawaii.edu)' = 'NOAA, Deep Sea Coral Research & Technology Program and Office of Ocean Exploration and Research'))
# 
# x$DataContact <- 
#   plyr::revalue(x$DataContact, 
#                 c('2017-12-18' =  'Kelley, Chris: ckelley@hawaii.edu','Pante, Eric (eric.pante@univ-lr.fr)' = 'Pante, Eric: eric.pante@univ-lr.fr', "Battista, Tim; Tim.Battista@noaa.gov" = "Battista, Tim: Tim.Battista@noaa.gov", "Thoma, Jana" = "Thoma, Jana: jana.thoma@louisiana.edu"))
                  
# table(factor(x$DataProvider))
# table(factor(x$PI))
# table(factor(x$DataContact))

x$url <- paste0('https://deepseacoraldata.noaa.gov/Dataset%20Summaries/',
                      x$DatasetID, 
                      '.html', 
                      sep = '')

x$DashBoard <- paste0("[", x$DatasetID, "](", x$url, ")")

sum_tbl <-
  x %>%
  group_by(DataProvider) %>%
  summarize(DatasetID = toString(unique(DashBoard)), 
            PI = toString(unique(unlist(unique(strsplit(as.character(PI), '; '))))),
            DataContact = toString(unique(DataContact)),
            Reporter = toString(unique(Reporter)),
            n = n()) %>%
  arrange(desc(DatasetID))
sum_tbl <- kable(sum_tbl, row.names = F, digits = 2, format.args = list(big.mark = ","))
sum_tbl
``` 

# Expedition Details

``` {r exped_details, echo=FALSE}
sum_tbl <-
  x %>%
  group_by(DatasetID, SurveyID) %>% 
  summarize(Vessel = toString(unique(Vessel)),
            VehicleName = toString(unique(VehicleName)),
            PI = toString(unique(PI)),
            SamplingEquipment = toString(unique(SamplingEquipment)),
            RecordType = toString(unique(RecordType)),
            #BeginYear= min(as.numeric(ObservationYear)),
            EndYear= max(as.numeric(ObservationYear)),
            n = prettyNum(n(),big.mark = ',')) %>% 
  arrange(desc(EndYear))

View(sum_tbl)

sum_tbl$url <- paste0('https://deepseacoraldata.noaa.gov/Dataset%20Summaries/',
                      sum_tbl$DatasetID, 
                      '.html', 
                      sep = '')

sum_tbl$DashBoard <- paste0("[", "Dashboard Link", "](", sum_tbl$url, ")")
sum_tbl <- sum_tbl %>% dplyr::select(-url)
sum_tbl <- kable(sum_tbl, row.names = F, digits = 2)
sum_tbl
```

# Growth of Database Through Time within this Area of Interest

``` {r GrowthThroughTime_charts, echo=FALSE, dpi=300, eval = T}
## set options to remove scientific notation
options(scipen=10000, digits = 0)


subx <- x %>% filter (Phylum == "Cnidaria") %>%
  mutate(EntryDate = replace_na(as.character(EntryDate), '2019-12-18')) %>% 
  mutate(EntryDate = as.factor(EntryDate)) %>% 
  mutate(as.Date(EntryDate))

suby <- x %>% filter (Phylum == "Porifera") %>%
  mutate(EntryDate = replace_na(as.character(EntryDate), '2019-12-18')) %>% 
  mutate(EntryDate = as.factor(EntryDate)) %>% 
  mutate(as.Date(EntryDate))
 
ggplot(subx, aes(x=as.Date(EntryDate), color=Phylum)) +
  stat_bin(data = subx,
           aes(y=cumsum(..count..) / 1000), 
           geom="line", 
           binwidth = 2, 
           size = 1) +
  stat_bin(data = suby,
           aes(y=cumsum(..count..) / 1000),
           geom="line", 
           binwidth = 2, 
           size = 1) +
  ylab("Number of Records (thousands)") + 
  xlab("Year Record Added") + 
  xlim(as.Date('2012-01-01'), as.Date('2020-10-19')) +
  theme(axis.text.x=element_text(hjust= .2)) +
  guides(fill=FALSE) + 
  theme_bw() +
  theme(text=element_text(size=20,  family="Cambria")) +
  #theme(axis.text.x = element_text(angle = 90)) +
  theme(legend.title = element_blank()) +
  theme(axis.title.x = element_text(vjust=-2.5)) +
  theme(axis.title.y = element_text(vjust= 5)) +
  theme(plot.margin = unit(c(1,1,1,1), "cm"))+
  scale_color_manual(values = brewer.pal(12, "Paired")[c(10,9)]) +
  scale_y_continuous(label=scales::comma) +
  scale_y_continuous(labels = scales::number_format(accuracy = 1))+
  geom_vline(xintercept = as.Date("2016-01-01"), color = 'blue')# +
## ggtitle(paste("Database Version:", unique(filt$DatabaseVersion)))

## save the plot
ggsave(paste("c:/rworking/deepseatools/images/", 
             "20201019-3_SouthAtlantic_subset", 
             "NatDB_", 
             unique(x$DatabaseVersion), 
             ".png", 
             sep = ''),
       width = 8,
       height = 5.57,
       units = "in")

```



