---
title: "SEDCI Final Report Contribution"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=F, message=F)
```

# packages 
```{r packages, echo=F, warning=F, message=F}
#install.packages(tidyverse)
#install.packages('openxlsx')
#install.packages('here')
library(here)
library(tidyverse)
library(openxlsx)
library(sf)
library(rgdal)
library(RColorBrewer)
library(raster)
library(marmap)
library(leaflet)
library(extrafont)
library(RColorBrewer)
library(rmarkdown)
library(knitr)
library(maps)
library(rgdal)
library(raster)
library(marmap)
library(httr)
library(jsonlite)
library(rnaturalearth)
library(rnaturalearthdata)
library(openxlsx)
library(rgeos)
library(arcgisbinding)
arc.check_product()

```

# load NDB

```{r, cache=T, message=F, warning=F}
##### input: latest version of NDB #####
setwd("C:/rworking/deepseatools/indata")
indata<-read.csv("DSCRTP_NatDB_20210414-0.csv", header = T)
filt <- indata %>%
  filter(Flag == "0")

## clean
rm(indata)

```

# (create 'sub') subset the data and make into sp layer 

```{r}
##### filter the NDB to your liking #####
sub <- filt #%>% filter()

##### load from data subset #####

##  set project crs
projcrs <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"

geosub <- st_as_sf(x = sub,                         
           coords = c("Longitude", "Latitude"),
           crs = projcrs)

## cleanup 
#rm(filt)

```

# load geodata

```{r}

##### load protected areas from geodatabase and shapefiles #####

## CaribbeanFMC
setwd("C:/Users/Robert.McGuinn/Documents/ArcGIS/Packages/RTC2020_0d88f3/p20/")
grammanik <- sf::st_read(dsn = "reporttocongress6.gdb", layer = 'Grammanik_Bank_po_Project')
hind <- sf::st_read(dsn = "reporttocongress6.gdb", layer = 'Hind_Bank_Marine_Conservation_District')
mutton <- sf::st_read(dsn = "reporttocongress6.gdb", layer = 'mutton_snapper_spawn_po_Proj')
redhind <- sf::st_read(dsn = "reporttocongress6.gdb", layer = 'red_hind_spawn_po_Project')
navassa <- sf::st_read(dsn = "reporttocongress6.gdb", layer = 'Navassa')
bajodesico <- sf::st_read(dsn = "reporttocongress6.gdb", layer = 'Bajo_de_Sico_po_Project')

## SAFMC
stetson_miami <- sf::st_read("C:/Users/Robert.McGuinn/Documents/ArcGIS/Packages/RTC2020_0d88f3/commondata/stetson_update/StetsonMiami_po.shp")
oculinabankoeca <- sf::st_read(dsn = "C:/Users/Robert.McGuinn/Documents/ArcGIS/Packages/RTC2020_0d88f3/p20/reporttocongress1.gdb", layer = "OculinaBank_oeca_po_Project")
mpaIMS <- sf::st_read(dsn = "C:/Users/Robert.McGuinn/Documents/ArcGIS/Packages/RTC2020_0d88f3/p20/reporttocongress1.gdb", layer = "MPA_IMS_Project")
oculinabankold <- sf::st_read(dsn = "C:/Users/Robert.McGuinn/Documents/ArcGIS/Packages/RTC2020_0d88f3/p20/reporttocongress1.gdb", layer = "OculinaBank_po_old_Project1")
oculinabank <- sf::st_read(dsn = "C:/Users/Robert.McGuinn/Documents/ArcGIS/Packages/RTC2020_0d88f3/p20/reporttocongress1.gdb", layer = "OculinaBank_po_Project")
pourtales_minus_crab <- sf::st_read(dsn = "C:/Users/Robert.McGuinn/Documents/ArcGIS/Packages/RTC2020_0d88f3/p20/reporttocongress1.gdb", layer = "Pourtales_minus_crab")
capefear <- sf::st_read(dsn = "C:/Users/Robert.McGuinn/Documents/ArcGIS/Packages/RTC2020_0d88f3/p20/reporttocongress1.gdb", layer = "CapeFear_po_Project")
capelookout <- sf::st_read(dsn = "C:/Users/Robert.McGuinn/Documents/ArcGIS/Packages/RTC2020_0d88f3/p20/reporttocongress1.gdb", layer = "CapeLookout_po_Project")
capelookout_dissolve <- sf::st_read(dsn = "C:/Users/Robert.McGuinn/Documents/ArcGIS/Packages/RTC2020_0d88f3/p20/reporttocongress1.gdb", layer = "CL_merge_Dissolve")

## GOMEXFMC
hapc_c9nr <- sf::st_read("C:/rworking/sf/indata/geo/HAPCshapefiles/shpFinal2/Coral9NoRegs.shp")
hapc_c9r <- sf::st_read("C:/rworking/sf/indata/geo/HAPCshapefiles/shpFinal2/Coral9Regs.shp")
hapc_ewor <- sf::st_read("C:/rworking/sf/indata/geo/HAPCshapefiles/shpFinal2/ExistingWithOutRegs.shp")
hapc_ewr <- sf::st_read("C:/rworking/sf/indata/geo/HAPCshapefiles/shpFinal2/ExistingWithRegs.shp")

```

# align all geo files to project level crs

```{r}
##### set project crs #####
projcrs <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"

## caribbean
grammanik <- st_transform(grammanik, crs = projcrs)
hind <- st_transform(hind, crs = projcrs)
mutton <- st_transform(mutton, crs = projcrs)
redhind <- st_transform(redhind, crs = projcrs)
navassa <- st_transform(navassa, crs = projcrs)
bajodesico <- st_transform(bajodesico, crs = projcrs)

## southeast
stetson_miami <- st_transform(stetson_miami, crs = projcrs)
oculinabankoeca <- st_transform(oculinabankoeca, crs = projcrs)
mpaIMS <- st_transform(mpaIMS, crs = projcrs)
oculinabankold <- st_transform(oculinabankold, crs = projcrs)
oculinabank <- st_transform(oculinabank, crs = projcrs)
pourtales_minus_crab <- st_transform(pourtales_minus_crab, crs = projcrs)
capefear <- st_transform(capefear, crs = projcrs)
capelookout <- st_transform(capelookout, crs = projcrs)
capelookout_dissolve <- st_transform(capelookout_dissolve, crs = projcrs)

## hapc
hapc_c9nr <- st_transform(hapc_c9nr, crs = projcrs)
hapc_c9r <- st_transform(hapc_c9r, crs = projcrs)
hapc_ewor <- st_transform(hapc_ewor, crs = projcrs)
hapc_ewr <- st_transform(hapc_ewr, crs = projcrs)

```

# write all shapefiles to folder
```{r}

fgdb_path <- 'C:/rworking/sf/sf.gdb'
arc.write(file.path(fgdb_path, 'x_geo'), data=x_geo, overwrite = TRUE)


```




# do all interesections

```{r}
grammanik_points <- st_intersection(geosub, grammanik)
hind_points <- st_intersection(geosub, hind)
mutton_points <- st_intersection(geosub, mutton)
redhind_points <- st_intersection(geosub, redhind)
navassa_points <- st_intersection(geosub, navassa)
bajodesico_points <- st_intersection(geosub, bajodesico)

## southeast
stetson_miami_points <-  st_intersection(geosub, stetson_miami)
oculinabankoeca_points <- st_intersection(geosub, oculinabankoeca)
mpaIMS_points <- st_intersection(geosub, mpaIMS)
oculinabankold_points <- st_intersection(geosub, oculinabankold) 
oculinabank_points <- st_intersection(geosub, oculinabank) 
pourtales_minus_crab_points <- st_intersection(geosub, pourtales_minus_crab)
capefear_points <- st_intersection(geosub, capefear)
capelookout_points <- st_intersection(geosub, capelookout) 
capelookout_dissolve_points <- st_intersection(geosub, capelookout_dissolve)

## gomex
hapc_c9nr_points <- st_intersection(geosub, hapc_c9nr)
hapc_c9r_points <- st_intersection(geosub, hapc_c9r)
hapc_ewor_points <- st_intersection(geosub, hapc_ewor)
hapc_ewr_points <- st_intersection(geosub, hapc_ewr)
```

# assigning inside vs. outside fishing restricted area

```{r}
## create a vector of all CatalogNumbers 
yo <- 
  c(grammanik_points$CatalogNumber, 
    hind_points$CatalogNumber,
    mutton_points$CatalogNumber,
    redhind_points$CatalogNumber,
    navassa_points$CatalogNumber,
    bajodesico_points$CatalogNumber,
    stetson_miami_points$CatalogNumber,
    oculinabankoeca_points$CatalogNumber,
    mpaIMS_points$CatalogNumber,
    oculinabankold_points$CatalogNumber,
    oculinabank_points$CatalogNumber,
    pourtales_minus_crab_points$CatalogNumber,
    capefear_points$CatalogNumber,
    capelookout_points$CatalogNumber,
    capelookout_dissolve_points$CatalogNumber,
    # hapc_c9nr_points$CatalogNumber, # take out because of no regs
    hapc_c9r_points$CatalogNumber,
    # hapc_ewor_points$CatalogNumber, # take out because of no regs
    hapc_ewr_points$CatalogNumber)

## define not in
`%notin%` <- Negate(`%in%`)

## assign management status back to original subset## 
sub$HAPC <- 
  case_when(
  sub$CatalogNumber %in% yo ~ "Inside Managed Area",
  sub$CatalogNumber %notin% yo ~ "Outside Managed Area"
)

## checking
# length(mpas_cats$CatalogNumber)
# length(oculina_cats$CatalogNumber)
# length(oculinaexp_cats$CatalogNumber)
# length(hapc_cats$CatalogNumber)
# names(points_oculina)
# z <- points_oculina %>% filter(is.na(AREA_NAME) == F)
# plot(z["AREA_NAME"])
# names(points_mpas)
# z <- points_mpas %>% filter(is.na(NAME) == F)
# plot(z["NAME"])
# names(points_mpas)
# z <- points_oculinaexp %>% filter(is.na(AREA_NAME) == F)
# plot(z["AREA_NAME"])
```

# new variable: 'sedci', with values 'new' vs. 'rescue' 

```{r}
##### get all the DatasetIDs in SEDCI region #####
ecogig <- c("NOAA_EX-12-02-L2",
            "BOEM_Lophelia_II",
            "WHOI_AT-18",
            "SOI_FK006B",
            "NOAA_HC-11-10",
            "OET_NA028",
            "OET_NA043",
            "OET_NA057",
            "OET_NA058",
            "ECOGIG_OI-16-10",
            "ECOGIG_O2-17-06")

ecogig <- unique(ecogig)

## DatasetIDs in the SEDCI rescue project (as of 2020-10-19)
sedci_rescue <- c("HBOI_SJ-10-07", "HBOI_SJ-10-07", "NOAA_SJ-09-08", "NOAA_SJ-09-08","HBOI_SJ-08-09","NOAA_SJ-07-05", "NOAA_SJ-07-05", "NOAA_SJ-07-06", "HBOI_SJ-06-02", "HBOI_SJ-06-05-L1", "HBOI_SJ-06-05-L3", "HBOI_SJ-06-02", "NOAA_SJ-05-04", "NOAA_SJ-05-08", "NOAA_SJ-05-10", "NOAA_SJ-05-11", "NOAA_SJ-04-05", "NOAA_SJ-02-08", "HBOI_EL-99-08", "HBOI_SJ-88-11", "USGS_TM-97-10", "NOAA_RB-03-07", "NOAA_SJ-05-11", "NOAA_NF-09-01", "NOAA_NF-08-04", "NOAA_NF-07-06", "NOAA_NF-05-05", "NOAA_SJ-07-06", "HBOI_SJ-08-09", "NOAA_SJ-09-08", "NOAA_PC-10-02", "NOAA_NF-14-01", "NOAA_NF-13-02", "NOAA_NF-12-01", "NOAA_NF-11-09", "NOAA_NF-11-09-L3", "NOAA_PC-11-05-L1", "NOAA_NF-11-09", "NOAA_NF-11-09-L3", "Sharuga_SM_2014", "NOAA_PC-12-03", "NOAA_NF-13-02", "NOAA_PC-13-03", "NOAA_NF-14-08","NOAA_PC-15-02", "NOAA_PC-16-02", "OET_NA058", "OET_NA028", "NOAA_PC-17-02", "HBOI_WS-17-125")

## unique DatasetIDs
sedci_rescue <- unique(sedci_rescue)

## sedci_rescue add ecogig
sedci_rescue <- union(sedci_rescue, ecogig)

## datasetIDs in the SEDCI rescue project (as of 2020-11-25)
sedci_new <- c("NOAA_DFH-30", "NOAA_NF-17-08",
               "NOAA_NF-17-08", "NOAA_DFH-32-33",
               "NOAA_DFH-32-33", "NOAA_EX-17-11",
               "NOAA_EX-18-03", "NOAA_EX-18-06",
               "NOAA_NF-18-04","NOAA_FGBNMS_DFH-35",
               "NOAA_FGBNMS_DFH-37", "NOAA_RESTORE_MT18", "NOAA_EX-15-02-L3",
               "NOAA_EX-18-11", "NOAA_NF-19-01", "NOAA_EX-19-03-L2", "NOAA_EX-19-07")

sedci_new <- unique(sedci_new)

##### build a variable for new vs. rescue datasets #####
sub$sedci <-
  case_when(
    sub$DatasetID %in% sedci_rescue ~ "rescue" ,
    sub$DatasetID %in% sedci_new ~ "new",
  )

#length(x$Flag)
#table(sub$sedci, useNA = 'always')
```

# summary stats for SEDCI final report

```{r}
length(sedci_new)
length(sedci_rescue)
table(sub$sedci, useNA = 'always')

x <- sub %>% filter(sedci == "new" | sedci == "rescue")

table(x$sedci, useNA = 'always')
table(sub$sedci, useNA = 'always')

```
# Expedition Details

``` {r exped_details, echo=FALSE}
sum_tbl <-
  sub %>% 
  filter(sedci == "new" | sedci == "rescue") %>% 
  group_by(DatasetID) %>% 
  summarize(SurveyID = toString(unique(SurveyID)),
            Vessel = toString(unique(Vessel)),
            VehicleName = toString(unique(VehicleName)),
            PI = toString(unique(PI)),
            SamplingEquipment = toString(unique(SamplingEquipment)),
            RecordType = toString(unique(RecordType)),
            #BeginYear= min(as.numeric(ObservationYear)),
            EndYear= max(as.numeric(ObservationYear)),
            n = prettyNum(n(),big.mark = ',')) %>% 
  arrange(desc(EndYear))

View(sum_tbl)

sum_tbl$url <- paste0('https://deepseacoraldata.noaa.gov/Dataset%20Summaries/',
                      sum_tbl$DatasetID, 
                      '.html', 
                      sep = '')

sum_tbl$DashBoard <- paste0("[", "Dashboard Link", "](", sum_tbl$url, ")")
sum_tbl <- sum_tbl %>% dplyr::select(-url)
write.csv(sum_tbl, "c:/rworking/deepseatools/reports/20201207_SEDCI_Dataset_Summary_RPMcGuinn.csv")
sum_tbl <- kable(sum_tbl, row.names = F, digits = 2)
sum_tbl
```

# new variable: 'sedci2' with values 'Post SEDCI' and 'Pre SEDCI'

```{r}
##### build a variable for pre-SEDCI and post-SEDCI
sub$sedci2 <-
  case_when(
    sub$DatasetID %in% sedci_rescue ~ "Post SEDCI" ,
    sub$DatasetID %in% sedci_new ~ "Post SEDCI",
    sub$DatasetID %notin% sedci_rescue ~ "Pre SEDCI",
    sub$DatasetID %notin% sedci_new ~ "Pre SEDCI"
  )

##### order the levels properly to fit figure requirements #####
sub$sedci2 <- factor(sub$sedci2, levels = c('Post SEDCI', 'Pre SEDCI'))
```

# assigning depth bins

``` {r DepthClassTable_1, echo=FALSE, cache = FALSE} 
# The Bathyal Zone is generally defined as extending from approximately]
# 200 to 3500 m, with an Upper Bathyal (200-800) and Lower Bathyal
# (800-3500)

hi <- 800
y <- 200
z <- 50

sub$DepthCat[sub$DepthInMeters > hi] <- "very deep"
sub$DepthCat[sub$DepthInMeters > y & sub$DepthInMeters <= hi] <- "deep"
sub$DepthCat[sub$DepthInMeters >= z & sub$DepthInMeters <= y] <- "mesophotic"
sub$DepthCat[sub$DepthInMeters < z] <- "very shallow"
sub$DepthCat <- factor(sub$DepthCat, levels = c('very shallow','mesophotic', 'deep', 'very deep'))

```

# making the protection by depth figures

## Figure: Protection by DepthCat in Caribbean

```{r}
##### building big pallette #####
# to get the colors from a palette:
palette1 <- brewer.pal(9,"Set1")
palette2 <- brewer.pal(8,"Set2")
palette3 <- brewer.pal(9,"Set3")
palette4 <- brewer.pal(8,"Dark2")
palette5 <- brewer.pal(8,"Accent")

# We can just stick a few color palettes together
big_palette <- c(palette1,palette2,palette3, palette4, palette5)

##### making the figure #####
options(scipen=10000)

g <- sub %>% filter(FishCouncilRegion == "Caribbean"
  ) %>% 
  ggplot(aes(DepthCat, fill = sedci2)) +
  geom_bar() + 
  coord_flip() + 
  facet_wrap(~ HAPC) +
  theme(text = element_text(size=20)) + 
  ylab("Number of Records") + 
    xlab("Depth Zone") + 
    scale_x_discrete(limits = rev(levels(sub$DepthCat))) + 
  theme_bw(base_size = 15, base_family = "Cambria") +
  theme(legend.title = element_blank())

#display.brewer.all(colorblindFriendly=TRUE)
set.seed(10)
#g + scale_fill_manual(values = sample(big_palette)) +
g + scale_fill_manual(values = brewer.pal(9,"Greys")[c(4,7)]) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

## save the plot
ggsave(paste("c:/rworking/deepseatools/images/", 
             "20210623-0_Caribbean_Protection_by_Depth_SEDCI_", 
             "NatDB_", 
             unique(sub$DatabaseVersion), 
             ".png", 
             sep = ''),
       width = 8,
       height = 5.57,
       units = "in")


```

## Figure: Protection by DepthCat in South Atlantic

```{r}
##### building big pallette #####
# to get the colors from a palette:
palette1 <- brewer.pal(9,"Set1")
palette2 <- brewer.pal(8,"Set2")
palette3 <- brewer.pal(9,"Set3")
palette4 <- brewer.pal(8,"Dark2")
palette5 <- brewer.pal(8,"Accent")

# We can just stick a few color palettes together
big_palette <- c(palette1,palette2,palette3, palette4, palette5)

##### making the figure #####
options(scipen=10000)

g <- sub %>% filter(FishCouncilRegion == "South Atlantic"
  ) %>% 
  ggplot(aes(DepthCat, fill = sedci2)) +
  geom_bar() + 
  coord_flip() + 
  facet_wrap(~ HAPC) +
  theme(text = element_text(size=20)) + 
  ylab("Number of Records") + 
    xlab("Depth Zone") + 
    scale_x_discrete(limits = rev(levels(sub$DepthCat))) + 
  theme_bw(base_size = 15, base_family = "Cambria") +
  theme(legend.title = element_blank())

#display.brewer.all(colorblindFriendly=TRUE)
set.seed(10)
# g + scale_fill_manual(values = sample(big_palette)) + 
  g + scale_fill_manual(values = brewer.pal(9,"Greys")[c(4,7)]) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

## save the plot
ggsave(paste("c:/rworking/deepseatools/images/", 
             "20210623-0_South_Atlantic_Protection_by_Depth_SEDCI_", 
             "NatDB_", 
             unique(sub$DatabaseVersion), 
             ".png", 
             sep = ''),
       width = 8,
       height = 5.57,
       units = "in")


```

## Figure: Protection by DepthCat in Gulf of Mexico

```{r}
##### building big pallette #####
# to get the colors from a palette:
palette1 <- brewer.pal(9,"Set1")
palette2 <- brewer.pal(8,"Set2")
palette3 <- brewer.pal(9,"Set3")
palette4 <- brewer.pal(8,"Dark2")
palette5 <- brewer.pal(8,"Accent")

# We can just stick a few color palettes together
big_palette <- c(palette1,palette2,palette3, palette4, palette5)

##### making the figure #####
options(scipen=10000)

g <- sub %>% filter(FishCouncilRegion == "Gulf of Mexico"
  ) %>% 
  ggplot(aes(DepthCat, fill = sedci2)) +
  geom_bar() + 
  coord_flip() + 
  facet_wrap(~ HAPC) +
  theme(text = element_text(size=20)) + 
  ylab("Number of Records") + 
    xlab("Depth Zone") + 
    scale_x_discrete(limits = rev(levels(sub$DepthCat))) + 
  theme_bw(base_size = 15, base_family = "Cambria") +
  theme(legend.title = element_blank())

#display.brewer.all(colorblindFriendly=TRUE)
set.seed(10)
# g + scale_fill_manual(values = sample(big_palette)) + 
  g + scale_fill_manual(values = brewer.pal(9,"Greys")[c(4,7)]) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

## save the plot
ggsave(paste("c:/rworking/deepseatools/images/", 
             "20210623-0_GOMEX_Protection_by_Depth_SEDCI_", 
             "NatDB_", 
             unique(sub$DatabaseVersion), 
             ".png", 
             sep = ''),
       width = 8,
       height = 5.57,
       units = "in")


```

# prep data for special GOMEX figure

```{r}
## subset data to only that within GOMEX
sub <- filt %>% filter(FishCouncilRegion == 'Gulf of Mexico')

## create a vector of all CatalogNumbers 
gomex_designation_all <- 
  c(hapc_c9nr_points$CatalogNumber,
    hapc_c9r_points$CatalogNumber,
    hapc_ewor_points$CatalogNumber,
    hapc_ewr_points$CatalogNumber)

gomex_designation_2020 <-
  c(hapc_c9nr_points$CatalogNumber,
    hapc_c9r_points$CatalogNumber)

gomex_designation_existing <-
  c(hapc_ewor_points$CatalogNumber,
    hapc_ewr_points$CatalogNumber)

gomex_restricted_existing <- c(hapc_ewr_points$CatalogNumber)
gomex_restricted_2020 <- c(hapc_c9r_points$CatalogNumber)
gomex_unrestricted_existing <- c(hapc_ewor_points$CatalogNumber)
gomex_unrestricted_2020 <- c(hapc_c9nr_points$CatalogNumber)

## define not in
`%notin%` <- Negate(`%in%`)

## assign management status back to original subset## 
sub$designation <- 
  case_when(
  sub$CatalogNumber %in% gomex_designation_all ~ "Inside Designated Area",
  sub$CatalogNumber %notin% gomex_designation_all ~ "Outside Designated Area"
)

sub$fishing_restriction <- 
  case_when(
  sub$CatalogNumber %in% gomex_restricted_existing ~ "Pre-2016 Fishing-Restricted",
  sub$CatalogNumber %in% gomex_restricted_2020 ~ "2020 Fishing-Restricted",
  sub$CatalogNumber %in% gomex_unrestricted_existing ~ "Pre-2016 Fishing-Unrestricted",
  sub$CatalogNumber %in% gomex_unrestricted_2020 ~ "2020 Fishing-Unrestricted",
  sub$CatalogNumber %notin% gomex_designation_all ~ "Outside Managed Area"
  )

##### order the levels properly to fit figure requirements #####
sub$fishing_restriction <- factor(sub$fishing_restriction, levels = c('Pre-2016 Fishing-Restricted', '2020 Fishing-Restricted','Outside Managed Area','Pre-2016 Fishing-Unrestricted', '2020 Fishing-Unrestricted' ))


```

## new variable: 'sedci2' with values 'Post SEDCI' and 'Pre SEDCI'

```{r}
##### build a variable for pre-SEDCI and post-SEDCI
sub$sedci2 <-
  case_when(
    sub$DatasetID %in% sedci_rescue ~ "Post SEDCI" ,
    sub$DatasetID %in% sedci_new ~ "Post SEDCI",
    sub$DatasetID %notin% sedci_rescue ~ "Pre SEDCI",
    sub$DatasetID %notin% sedci_new ~ "Pre SEDCI"
  )

##### order the levels properly to fit figure requirements #####
sub$sedci2 <- factor(sub$sedci2, levels = c('Post SEDCI', 'Pre SEDCI'))
```

## new variable: assigning depth bins

``` {r DepthClassTable2, echo=FALSE, cache = FALSE} 
hi <- 800
y <- 200
z <- 50

sub$DepthCat[sub$DepthInMeters > hi] <- "very deep"
sub$DepthCat[sub$DepthInMeters > y & sub$DepthInMeters <= hi] <- "deep"
sub$DepthCat[sub$DepthInMeters >= z & sub$DepthInMeters <= y] <- "mesophotic"
sub$DepthCat[sub$DepthInMeters < z] <- "very shallow"
sub$DepthCat <- factor(sub$DepthCat, levels = c('very shallow','mesophotic', 'deep', 'very deep'))

```

## Figure: GOMEX Restricted fishing analysis

```{r}
##### building big pallette #####
# to get the colors from a palette:
palette1 <- brewer.pal(9,"Set1")
palette2 <- brewer.pal(8,"Set2")
palette3 <- brewer.pal(9,"Set3")
palette4 <- brewer.pal(8,"Dark2")
palette5 <- brewer.pal(8,"Accent")

# We can just stick a few color palettes together
big_palette <- c(palette1,palette2,palette3, palette4, palette5)

##### making the figure #####
options(scipen=10000)

g <- sub %>% #filter(
  # FishCouncilRegion == "Gulf of Mexico"
  # ) %>% 
  ggplot(aes(DepthCat, fill = sedci2)) +
  geom_bar() + 
  coord_flip() + 
  facet_wrap(~ fishing_restriction) +
  theme(text = element_text(size=10)) + 
  ylab("Number of Records") + 
    xlab("Depth Zone") + 
    scale_x_discrete(limits = rev(levels(sub$DepthCat))) + 
  theme_bw(base_size = 12, base_family = "Cambria") +
  theme(legend.title = element_blank()) +
  theme(legend.position = c(0.8, 0.2))

#display.brewer.all(colorblindFriendly=TRUE)
set.seed(10)
# g + scale_fill_manual(values = sample(big_palette)) + 
  g + scale_fill_manual(values = brewer.pal(9,"Greys")[c(4,7)]) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

## save the plot
ggsave(paste("c:/rworking/deepseatools/images/",
             "20210623-0_GOMEX_Fish_Resctriction_by_Depth_SEDCI_",
             "NatDB_",
             unique(sub$DatabaseVersion),
             ".png",
             sep = ''),
       width = 8,
       height = 5.57,
       units = "in")

```


  











<!-- # create tables for export -->
<!-- ```{r} -->
<!-- #### making frequency tables for each region -->
<!-- hapc_c9nr_tab <- table(geosub_hapc_c9nr_cats$Name, useNA = 'always') -->
<!-- hapc_c9r_tab <- table(geosub_hapc_c9r_cats$Name, useNA = 'always') -->
<!-- hapc_ewor_tab <- table(geosub_hapc_ewor_cats$Name, useNA = 'always') -->
<!-- hapc_ewr_tab <- table(geosub_hapc_ewr_cats$Name, useNA = 'always') -->

<!-- hapc_c9nr_tab_df <- as.data.frame(hapc_c9nr_tab) -->
<!-- hapc_c9r_tab_df <- as.data.frame(hapc_c9r_tab) -->
<!-- hapc_ewor_tab_df <- as.data.frame(hapc_ewor_tab) -->
<!-- hapc_ewr_tab_df <- as.data.frame(hapc_ewr_tab) -->

<!-- hapc_c9nr_tab_df$designation <- "c9_no_reg" -->
<!-- hapc_c9r_tab_df$designation <- "c9_reg" -->
<!-- hapc_ewor_tab_df$designation <- "existing_no_reg" -->
<!-- hapc_ewr_tab_df$designation <- "existing_reg" -->

<!-- hapc_all_tab_df <- rbind(hapc_c9nr_tab_df, -->
<!--                          hapc_c9r_tab_df, -->
<!--                          hapc_ewor_tab_df, -->
<!--                          hapc_ewr_tab_df) -->

<!-- ##### rename variable ##### -->
<!-- #hapc_all_tab_df2 <- hapc_all_tab_df %>% rename(Name = Var1) -->

<!-- ##### here I started over with a different selection from the top ##### -->
<!-- hapc_all_tab_df2_2013 <- hapc_all_tab_df %>% rename(Name = Var1) -->

<!-- ##### get rid of NAs before merging -->
<!-- table(hapc_all_tab_df$Var1, useNA = 'always') -->


<!-- hapc_all_tab_df2_2013_merge <- left_join(hapc_all_tab_df2, -->
<!--                                          hapc_all_tab_df2_2013, -->
<!--                                          by = "Name" -->
<!--                                          ) -->
<!-- ##### change the names ##### -->
<!-- names(hapc_all_tab_df2_2013_merge) <- c("Name", "Freq_up_to_2013", "designation","Freq_2014_and_up", "designation2") -->
<!-- names(hapc_all_tab_df2_2013_merge) -->

<!-- ##### get rid of NA ##### -->
<!-- hapc_all_tab_df2_2013_merge_cl <-  hapc_all_tab_df2_2013_merge %>% filter(is.na(Name) == F) -->

<!-- z <- c(hapc_c9nr$Name, hapc_c9r$Name, hapc_ewor$Name, hapc_ewr$Name) -->

<!-- x <- as.data.frame(x) -->
<!-- x <- x %>% rename(Name = x) -->
<!-- names(x) -->

<!-- table <- left_join(x, hapc_all_tab_df2_2013_merge_cl, -->
<!--                                          by = "Name") -->

<!-- ##### export table ##### -->
<!-- setwd("C:/rworking/deepseatools/indata") -->
<!-- table %>% -->
<!--   write.csv(paste("table.csv", sep = ''), row.names = FALSE) -->


<!-- ``` -->

<!-- # histogram of 'ObservationDate' by month -->
<!-- ```{r} -->
<!-- sub %>%  pull(ObservationDate) %>% as.Date() %>% hist("years") -->

<!-- ``` -->

<!-- # export to GIS -->
<!-- ```{r gis_export, eval=F} -->
<!-- ##### load packages ##### -->
<!-- library(arcgisbinding) -->
<!-- arc.check_product() -->

<!-- #### filter out data without coordinates -->
<!-- x <- sub %>% filter(Latitude != -999 | Longitude != -999) -->
<!-- x_geo <- x -->

<!-- ##### create spdf ##### -->
<!-- coordinates(x_geo) <- c("Longitude", "Latitude") -->
<!-- proj4string(x_geo) <- "+proj=longlat +ellps=WGS84 +datum=WGS84" -->

<!-- ##### create feature-class ##### -->
<!-- fgdb_path <- 'C:/rworking/sf/sf.gdb' -->
<!-- arc.write(file.path(fgdb_path, 'x_geo'), data=x_geo, overwrite = TRUE) -->
<!-- ``` -->

