---
title: "Analytical Task for Peter Etnoyer"
output: html_notebook
---

# intro

Original question posed by Peter Etnoyer on 2021-02-25:

"I need to know the 'non-null' average temperature, st dev, and temperature ranges of Paramuricea observations in the depth range 1000-2000 m in the Gulf of Mexico.

We assume based on past experience, that for our lab studies, the temperature range will be 5-10 C. This seemed like a good thing to check with DATA, and possible through our Nat DB. What do you think?"

# load packages

```{r}
library(tidyverse)
library(dplyr)
library(rmarkdown)
library(knitr)
library(flextable)
library(googlesheets4)
library(googlesheets)
library(googledrive)
library(openxlsx)
library(scales)
library(extrafont)
```

# load data

```{r}
# load the national database
setwd("C:/rworking/deepseatools/indata")
indata<-read_csv("DSCRTP_NatDB_20201021-0.csv")
filt <- indata %>%
  filter(Flag == "0", is.na(Phylum) == F)
rm(indata)

# load the other global climatology files
# https://www.ncei.noaa.gov/access/gulf-of-mexico-climate/bin/gomregcl.pl?parameter=t
setwd("C:/rworking/deepseatools/indata")
winter <-read.csv('gom_decav_t13mn01.csv', skip = 1)
spring <-read.csv('gom_decav_t14mn01.csv', skip = 1 )
summer <-read.csv('gom_decav_t15mn01.csv', skip = 1)
fall <-read.csv('gom_decav_t16mn01.csv', skip = 1)


round(mean(winter$X1000, na.rm = T), digits = 3)
round(mean(spring$X1000, na.rm = T), digits = 3)
round(mean(summer$X1000, na.rm = T), digits = 3)
round(mean(fall$X1000, na.rm = T), digits = 3)

round(mean(winter$X1500, na.rm = T), digits = 3)
round(mean(spring$X1500, na.rm = T), digits = 3)
round(mean(summer$X1500, na.rm = T), digits = 3)
round(mean(fall$X1500, na.rm = T), digits = 3)

round(mean(winter$X2000, na.rm = T), digits = 3)
round(mean(spring$X2000, na.rm = T), digits = 3)
round(mean(summer$X2000, na.rm = T), digits = 3)
round(mean(fall$X2000, na.rm = T), digits = 3)






```

# create summary table and histogram

```{r}
x <- filt %>% filter(
  Genus == "Paramuricea",
  as.numeric(DepthInMeters) > 1000,
  as.numeric(DepthInMeters) < 2000,
  Temperature != "-999", 
  FishCouncilRegion == "Gulf of Mexico"
  ) %>% 
  group_by(Genus) %>% 
  summarize(n=n(), 
            meanTemp = mean(Temperature), 
            sdTemp = sd(Temperature),
            minTemp = min(Temperature), 
            maxTemp = max(Temperature)
            ) %>% 
  View()

# create histogram
# hist(as.numeric(x$Temperature), breaks = 40)
qplot(x$Temperature)

```


