---
title: "Accession QA Report"
author: "NOAA-NFMS-OHC-DSCRTP: Robert.McGuinn@NOAA.gov"
date: "Date of report: `r Sys.Date()`"
output: 
  word_document:
     reference_docx: template.docx
always_allow_html: yes
editor_options: 
  chunk_output_type: console
---
# -
```{r load_packs, echo=F, message=FALSE, warning=F}
library(rmarkdown)
library(xfun)
library(tidyverse)
library(knitr)
library(flextable)
library(sp)
library(rgdal)
library(extrafont)
library(googlesheets4)
library(googledrive)
library(leaflet)
library(RColorBrewer)
library(terra)
library(sf)
library(maps)

```

```{r setup, include=FALSE}
opts_chunk$set(
  fig.width=7, 
  fig.height=3.5, 
  dpi = 300, 
  fig.align = 'left',
  message=FALSE, 
  warning=FALSE,
  cache.lazy = FALSE
)

```

```{r load_NDB, echo=F, message=FALSE, warning=FALSE, cache=T, eval=F}
##### manual input: latest version of NDB #####
## the data must be in the folder below
## get the latest from google drive and place in your local folder
setwd("C:/rworking/deepseatools/indata")
indata <- read.csv("DSCRTP_NatDB_20230620-0.csv", header = T) 
filt <- indata %>%
  filter(Flag == "0", 
         is.na(Phylum) == F)
## get the flagged records alone
# flagged <- sub %>%  filter(Flag == "1")
```

```{r OR_data_and_subset_local, echo=FALSE, message=FALSE, warning=FALSE, eval=F}
## manual: set direct path to file 
path <- "C:/rworking/deepseatools/indata/"
string <- paste(path,filename,".txt",sep = '')
sub <- read_csv(string) #or read_tsv as needed
flagged <- sub %>%  filter(Flag == "1")

```

```{r OR_data_and_subset_gdrive,echo=FALSE, message=FALSE, warning=FALSE, eval=T}
##### ##### 
## find the file in google drive by name [filename] is from runner
x <- drive_find(q = paste("name contains ", "'", filename,".zip", "'", sep = ''))
## getting the id as a character string
y <- x$id

## download the zip file
dl <- drive_download(as_id(y), path = "C:/rworking/deepseatools/indata/file.zip", overwrite = TRUE)

## extract just the file of interest from the zip file
sub <- read.csv(unz(dl$local_path, 
                    paste(filename, 
                          ".csv",
                          sep = '')),
                fileEncoding="latin9", header = T, stringsAsFactors = F)

## get the flagged records alone
flagged <- sub %>%  filter(Flag == "1")

```

```{r view_map, echo=F, eval=F}
##### gis inspection and export #####

##### create x from sub #####
x <- sub

##### filter data #####

# get rid of any missing Latitudes or Longitudes
x <- x %>% filter(Latitude != -999 | Longitude != -999)
# make copy to turn into spatial points data frame.
x_geo <- x

##### create simple features object data frame #####
projcrs <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"
x_geo_sf <- st_as_sf(x = x_geo,                         
           coords = c("Longitude", "Latitude"),
           crs = projcrs)

##### thin down the variables #####
x_geo_export <- x_geo_sf %>% dplyr::select(CatalogNumber, Phylum, ScientificName, SampleID, DepthInMeters, SurveyID, EventID)


##### make a leaflet map #####
library(leaflet)
m <- leaflet() %>% setView(lng = 5,lat = 38, zoom = 1)
m <- addProviderTiles(m, "Esri.OceanBasemap",  options = providerTileOptions(noWrap = TRUE))
m <- addCircleMarkers(m, data=x_geo_export, 
                        radius=5, 
                        weight=0, 
                        fillColor= "blue", 
                        fillOpacity=1,
                        popup = paste(
                          "<b><em>","Catalog Number:","</b></em>", sub$CatalogNumber, "<br>",
                          "<b><em>","Sample ID:","</b></em>", sub$SampleID, "<br>",
                          "<b><em>","Scientific Name:","</b></em>", sub$ScientificName, "<br>",
                          "<b><em>","Depth (meters):","</b></em>", sub$DepthInMeters, "<br>",
                          "<b><em>","Survey ID:","</b></em>", sub$SurveyID, "<br>",
                          "<b><em>","Event ID:","</b></em>", sub$EventID, "<br>",
                          "<b><em>","Image:","</b></em>",sub$ImageURL)
                                  
  )
m


##### Optional: export data it will guess the driver automatically based on the .shp extension #####
# st_write(x_geo_export, "C:/Users/Robert.Mcguinn/Documents/ArcGIS/Projects/fromr/shapefiles/x_geo_export.shp", delete_dsn = T)

```

```{r schema, echo=FALSE, cache = FALSE, warning=FALSE, message=FALSE, eval=FALSE}
##### download Google Sheet version of schema for use in R  documents ##### 
## run this command once per session to authorize googlesheets4 to your account
# gs4_auth()

# Register and download Google Sheet using googlesheets4::read_sheet
s <- read_sheet('1YDskzxY8OF-34Q8aI04tZvlRbhGZqBSysuie39kYHoI')

## checking 
# s %>% filter(FieldName == 'IdentificationVerificationStatus') %>% pull(FieldDescription)
# s %>% filter(FieldName == 'IdentificationVerificationStatus') %>% pull(ValidValues)
```

```{r mapit, include = F, eval=F}
sub2 <- sub %>%  
  filter(FlagReason == 'Horizontal position or depth is questionable')

m <- leaflet()
m <- addProviderTiles(m, "Esri.OceanBasemap")
m <- addCircleMarkers(m, data=sub,
                      radius=2,
                      weight=0,
                      fillColor= "green",
                      fillOpacity=.5,
                      popup = paste(
                        "<b><em>","Flag:","</b></em>", sub$Flag, "<br>",
                        "<b><em>","Catalog Number:","</b></em>", sub$CatalogNumber, "<br>",
                        "<b><em>","Record Type:","</b></em>", sub$RecordType, "<br>",
                        "<b><em>","DatasetID:","</b></em>", sub$DatasetID, "<br>",
                        "<b><em>","AccessionID:","</b></em>", sub$AccessionID, "<br>",
                        "<b><em>","DataProvider:","</b></em>", sub$DataProvider, "<br>",
                        "<b><em>","ObservationYear:","</b></em>", sub$ObservationYear, "<br>",
                        "<b><em>","Vessel:","</b></em>", sub$Vessel, "<br>",
                        "<b><em>","Locality:","</b></em>", sub$Locality, "<br>",
                        "<b><em>","Scientific Name:","</b></em>", sub$ScientificName, "<br>",
                        "<b><em>","Depth (meters):","</b></em>", sub$DepthInMeters, "<br>",
                        "<b><em>","Survey ID:","</b></em>", sub$SurveyID, "<br>",
                        "<b><em>","Event ID:","</b></em>", sub$EventID, "<br>",
                        "<b><em>","Latitude:","</b></em>", sub$Latitude, "<br>",
                        "<b><em>","Longitude:","</b></em>", sub$Longitude, "<br>",
                        "<b><em>","Image:","</b></em>",sub$ImageURL))
m <- addCircleMarkers(m, data=sub2,
                      radius=2,
                      weight=0,
                      fillColor= "black",
                      fillOpacity=.2,
                      popup = paste(
                        "<b><em>","Flag:","</b></em>", sub$Flag, "<br>",
                        "<b><em>","Catalog Number:","</b></em>", sub$CatalogNumber, "<br>",
                        "<b><em>","Record Type:","</b></em>", sub$RecordType, "<br>",
                        "<b><em>","DatasetID:","</b></em>", sub$DatasetID, "<br>",
                        "<b><em>","AccessionID:","</b></em>", sub$AccessionID, "<br>",
                        "<b><em>","DataProvider:","</b></em>", sub$DataProvider, "<br>",
                        "<b><em>","ObservationYear:","</b></em>", sub$ObservationYear, "<br>",
                        "<b><em>","Vessel:","</b></em>", sub$Vessel, "<br>",
                        "<b><em>","Locality:","</b></em>", sub$Locality, "<br>",
                        "<b><em>","Scientific Name:","</b></em>", sub$ScientificName, "<br>",
                        "<b><em>","Depth (meters):","</b></em>", sub$DepthInMeters, "<br>",
                        "<b><em>","Survey ID:","</b></em>", sub$SurveyID, "<br>",
                        "<b><em>","Event ID:","</b></em>", sub$EventID, "<br>",
                        "<b><em>","Latitude:","</b></em>", sub$Latitude, "<br>",
                        "<b><em>","Longitude:","</b></em>", sub$Longitude, "<br>",
                        "<b><em>","Image:","</b></em>",sub$ImageURL))

m

```

```{r echo = F, eval=T}
##### fix implicit NA #####
sub$Family <- as.character(sub$Family)
sub$Family <- forcats::fct_explicit_na(sub$Family)
```

```{r fit_width_table, echo = F}
# define ft width function 
FitFlextableToPage <- function(ft, pgwidth = 6){

  ft_out <- ft %>% autofit()

  ft_out <- width(ft_out, width = dim(ft_out)$widths*pgwidth /(flextable_dim(ft_out)$widths))
  return(ft_out)
}
```

# File Version Used in this Report
*`r substring(filename, 1, 10)`*

# Target AccessionID: 
*`r substring(filename,12)`*

AccessionID is currently: `r unique(sub$AccessionID)`

# DatasetID: *`r unique(factor(sub$DatasetID))`*

Purpose: `r unique(factor(sub$Purpose))`

SurveyComments: `r unique(factor(sub$SurveyComments))`

* Number of records: *`r length(sub$CatalogNumber)`*
* Number of records flagged: *`r length(flagged$CatalogNumber)`*
* Minimum depth (unflagged): *`r min(sub[sub$DepthInMeters != "-999" & sub$Flag == "0",]$DepthInMeters)`*
* Maximum depth (unflagged): *`r max(sub[sub$DepthInMeters != "-999" & sub$Flag == "0",]$DepthInMeters)`*
* Number of dives or tows recorded: *`r length(unique(sub[sub$EventID != "NA" | is.na(sub$EventID) == F,]$EventID))`* 
* Time frame: *`r min(as.numeric(min(sub$ObservationYear)))` to `r min(as.numeric(max(sub$ObservationYear)))`*
* Number of ImageURL populated: `r sub %>% filter(is.na(ImageURL) == F) %>% pull(ImageURL) %>% length()`
* Number of ImageFilePath Populated: `r sub %>% filter(is.na(ImageFilePath) == F) %>% pull(ImageURL) %>% length()`   

# Citation
``` {r Citation, echo=FALSE}
sum_tbl <-
  sub %>%
  group_by(Citation) %>%
  dplyr::summarize(Records=n())
sum_tbl <- kable(sum_tbl, row.names = F, digits = 2)
sum_tbl
```

# Citation Maker

``` {r CitationMaker, echo = FALSE, eval = T}

sub$CitationMaker <- paste(sub$DataProvider,'. ','Observation date range: ', 
                         min(sub$ObservationYear[sub$ObservationYear != "-999"]),' to ',
                         max(sub$ObservationYear[sub$ObservationYear != "-999"]),'. ',
                         'Coral or sponge occurrence observations submitted to the NOAA National Database for Deep Sea Corals and Sponges (www.deepseacoraldata.noaa.gov)', '. ',
                         #'DSCRTP Accession ID: ',sub$AccessionID, '. ',
                         #'Record type: ', sub$RecordType, '. ',
                         #'Vessel(s): ', sub$Vessel,'. ',
                         #'Sampling vehicle: ', sub$VehicleName,'. ',
                         #'Survey ID: ', sub$SurveyID,'. ',
                         'DSCRTP Dataset ID: ', sub$DatasetID, '. ',
                         'Principal investigator: ', sub$PI,'. ',
                         'Reporter: ', sub$Reporter, '. ',
                         'Database version: ', unique(filt$DatabaseVersion), '. ',
                         #'Data contact: ', sub$DataContact,'. ',
                         #'Reporter: ', sub$Reporter,'. ', 
                         #'Repository: ', sub$Repository,'. ',
                         # 'Web site [last accessed on YYYY-MM-DD]: ', sub$WebSite,'.', 
                         sep = '')
unique(sub$CitationMaker)
```

# Web site

``` {r Website, echo=FALSE}
sum_tbl <-
  sub %>%
  group_by(WebSite) %>%
  dplyr::summarize(Records=n())
sum_tbl <- kable(sum_tbl, row.names = F, digits = 2)
sum_tbl
```

# Data Origin

``` {r SourceSummary, echo=FALSE, warnings = F}
sum_tbl <-
  sub %>%
  group_by(DatasetID) %>%
  dplyr::summarize(
    SurveyID = paste(unique(SurveyID), collapse=" | "),
    DataProvider = paste(unique(DataProvider), collapse=" | "),
    PI = paste(unique(PI), collapse=" | "),
    Reporter = paste(unique(Reporter), collapse=" | "),
    ReporterEmail = paste(unique(ReporterEmail), collapse=" | "),
    DataContact = paste(unique(DataContact), collapse=" | "),
    Records = n()
  )


ft <- flextable(sum_tbl) %>% 
  theme_zebra() %>% 
  colformat_num(j = 7, 
                digits = 0, 
                big.mark = ",", 
                na_str = "NA") %>% 
  font(fontname = "Cambria", part = "all")  %>% 
  fontsize(size = 7, part = "all") # %>%
# align(j = 1:6, align = 'left', part = 'all')
 
ft <- set_table_properties(ft, width = 1, layout = "autofit")
ft
```

# Dive Table 

```{r, echo=F}

sum_tbl <-
  sub %>%
  group_by(SurveyID,EventID) %>%
  dplyr::summarize(
    Station = paste(unique(Station), collapse=" | "),
    Locality = paste(unique(Locality), collapse=" | "),
    Records = n()
  )


ft <- flextable(sum_tbl) %>% 
  theme_zebra() %>% 
  colformat_num(
    digits = 0, 
    big.mark = ",", 
    na_str = "NA"
  ) %>% 
  font(fontname = "Cambria", part = "all")  %>% 
  fontsize(size = 7, part = "all") # %>%
# align(j = 1:6, align = 'left', part = 'all')

ft <- set_table_properties(ft, width = 1, layout = "autofit")
ft

```


# Static Maps
## Locator Map

```{r, map, echo = FALSE, message = FALSE, warning=FALSE, dpi=300, fig.width=8, fig.height=4.1, eval=TRUE, fig.align='center'}

# # ##### Using ggmap libary #####
# #getting rid of  missing lat/long data
#library(ggmap)
sub2 <- sub %>%
  filter(Latitude != '-999' , Longitude != '-999')
# 
# # use this when not crossing dateline
# xpos <- sub2$Longitude
# # also use this additional calculation to transforming longitudes to 0-360 if crossing dateline as in Western AK and HI
# xpos <- ifelse(xpos < 0, (180 + xpos) + 180, xpos)
# 
# myLocation <- c(lon = mean(xpos), lat = mean(sub2$Latitude))
# #manually set center longitude if in an area that crosses dateline
# myLocation <- c(lon = 180, lat = mean(sub2$Latitude))
# 
# p <- get_map(location=myLocation, maptype = "watercolor", source = 'stamen', crop=FALSE, zoom=2)
# # # put transLong below if
# ggmap(p)+
# geom_point(aes(x = xpos, y = Latitude), data = sub2,
#  alpha = .5, color="darkred", size = 3)
# 
# 
# #> converting bounding box to center/zoom specification. (experimental)
# #> Map from URL : http://maps.googleapis.com/maps/api/staticmap?center=34.75309,-119.751995&zoom=16&size=640x640&scale=2&maptype=satellite&language=en-EN&sensor=false
# ggmap(sq_map) + geom_point(data = sub2, mapping = aes(x = Longitude, y = Latitude), color = "red")
# #> Warning: Removed 3 rows containing missing values (geom_point).


###### this will make a simple maps if you have problems with the method above #####
# library(dplyr)
# library(ggplot2)
gis<-sub2
coordinates(gis) <- c("Longitude", "Latitude", "DepthInMeters")
proj4string(gis) <- "+proj=longlat +ellps=WGS84 +datum=WGS84"
x<-bbox(gis)

m <- map_data("world") #, xlim = c(x[1,1]-1,x[1,2]+1), ylim = c(x[2,1]-1,x[2,2]+1))

z <- ggplot(m, aes(x = long, y = lat, group = group)) +
  geom_polygon()

z + geom_point(data=sub2,
               color = "red",
               size = 2,
               aes(x=Longitude, y = Latitude), inherit.aes = FALSE)
```
## 3D Bounding Box

```{r bounding box, echo=FALSE, echo = FALSE, message = FALSE, warning=FALSE, dpi=300, fig.width=8, fig.height=4.1, eval=TRUE, fig.align='center'}
spdf<-sub2
coordinates(spdf) <- c("Longitude", "Latitude", "DepthInMeters")
proj4string(spdf) <- "+proj=longlat +ellps=WGS84 +datum=WGS84"
x<-bbox(spdf)
x
```

# Expedition Details

``` {r source, echo=FALSE}
sum_tbl <-
  sub %>%
  group_by(SurveyID) %>%
  dplyr::summarize(
            Vessel = paste(unique(Vessel), collapse=" | "),
            VehicleName = paste(unique(VehicleName), collapse=" | "),
            SamplingEquipment = paste(unique(SamplingEquipment), collapse=" | "),
            NavType = paste(unique(NavType), collapse=" | "),
            LocationAccuracy = paste(unique(LocationAccuracy), collapse=" | "),
            ObservationDates = paste(unique(ObservationDate), collapse=" | "),
            Records = n()
            )


colkeys <- c("Records")
ft <- flextable(sum_tbl) %>% 
  theme_zebra() %>% #booktabs, vanilla, box, tron, tron_legacy, vader
  #set_header_labels(Genus_n = "# of Genera", n = "# of Records") %>% 
  colformat_num(j = 8, digits = 0, big.mark = ",", na_str = "NA") %>% 
  font(fontname = "Cambria", part = "all")  %>% #header, body, all
  fontsize(size = 7, part = "all") #%>%
  # width(j = 6, 1.5) %>% 
  # width(j = 7, 1.5)
 

ft <- set_table_properties(ft, width = 1, layout = "autofit")
ft


```

# Taxonomy Details
``` {r TaxDetails, echo=FALSE }
sum_tbl <-
  sub %>%
  # filter(RecordType == 'specimen') %>% 
  group_by(IdentifiedBy, RecordType) %>%
  dplyr::summarize(
    IdentificationQualifier = paste(unique(IdentificationQualifier), collapse=" | "),
    IdentificationVerificationStatus = paste(unique(IdentificationVerificationStatus), collapse=" | "),
    #Identification_Date = toString(unique(IdentificationDate)),
    Records = n()) %>%
  arrange(desc(Records))

ft <- flextable(sum_tbl) %>% 
  theme_zebra() %>% #booktabs, vanilla, box, tron, tron_legacy, vader
  #set_header_labels(Genus_n = "# of Genera", n = "# of Records") %>% 
  colformat_num(j=4, digits = 0, big.mark = ",", na_str = "NA") %>% 
  font(fontname = "Cambria", part = "all")  %>% #header, body, all
  fontsize(size = 7, part = "all")

ft <- set_table_properties(ft, width = 1, layout = "autofit")
ft
```

# List of Flagged Taxa
``` {r FlaggedScientificName, echo=FALSE }
sum_tbl <-
  sub %>% 
  arrange(ScientificName) %>% 
  filter(is.na(Phylum == T) | 
           Flag == "1") %>% 
  group_by(FlagReason) %>%
  dplyr::summarize(
    ScientificName = paste(unique(ScientificName), collapse=" | "),
    TaxonRank = paste(unique(TaxonRank), collapse=" | "),
    UnFlagged = length(Flag[Flag == "0"]), 
    Flagged = length(Flag[Flag == "1"]),
    FlagReason = paste(unique(FlagReason), collapse=" | "),
    Records = n())


ft <- flextable(sum_tbl) %>% 
  theme_zebra() %>% #booktabs, vanilla, box, tron, tron_legacy, vader
  colformat_num(digits = 0, big.mark = ",", na_str = "NA") %>%
  colformat_num(digits = 0, big.mark = ",", na_str = "NA") %>%
  font(fontname = "Cambria", part = "all")  %>% #header, body, all
  fontsize(size = 7, part = "all")# %>%
  #autofit() #%>% 
  #width(j = 2, 2)
 
ft <- set_table_properties(ft, width = 1, layout = "autofit")
ft
```


# List of ScientificNames (Coral) - Grouped by Family
``` {r ByScientificName, echo=FALSE }
sum_tbl <-
  sub %>% 
  arrange(ScientificName) %>% 
  filter(Phylum == 'Cnidaria') %>% 
  group_by(Family) %>%
  dplyr::summarize(
    ScientificName = paste(unique(ScientificName), collapse=" | "),
    TaxonRank = paste(unique(TaxonRank), collapse=" | "),
    UnFlagged = length(Flag[Flag == "0"]), 
    Flagged = length(Flag[Flag == "1"]),
    FlagReason = paste(unique(FlagReason), collapse=" | "),
    Records = n())


ft <- flextable(sum_tbl) %>% 
  theme_zebra() %>% #booktabs, vanilla, box, tron, tron_legacy, vader
  colformat_num(j = 4, digits = 0, big.mark = ",", na_str = "NA") %>%
  colformat_num(j = 7, digits = 0, big.mark = ",", na_str = "NA") %>%
  font(fontname = "Cambria", part = "all")  %>% #header, body, all
  fontsize(size = 7, part = "all")# %>%
  #autofit() #%>% 
  #width(j = 2, 2)
 
ft <- set_table_properties(ft, width = 1, layout = "autofit")
ft
```

# List of ScientificNames (Sponges) - Grouped by Family
``` {r ByScientificNameSponges, echo=FALSE }
sum_tbl <-
  sub %>% 
  arrange(ScientificName) %>% 
  filter(Phylum == 'Porifera') %>% 
  group_by(Family) %>%
  dplyr::summarize(
    ScientificName = paste(unique(ScientificName), collapse=" | "),
    TaxonRank = paste(unique(TaxonRank), collapse=" | "),
    UnFlagged = length(Flag[Flag == "0"]), 
    Flagged = length(Flag[Flag == "1"]),
    FlagReason = paste(unique(FlagReason), collapse=" | "),
    Records = n())

colkeys <- names(sum_tbl[,-1])
ft <- flextable(sum_tbl) %>% 
  theme_zebra() %>% #booktabs, vanilla, box, tron, tron_legacy, vader
  colformat_num(j = 4, digits = 0, big.mark = ",", na_str = "NA") %>%
  colformat_num(j = 7, digits = 0, big.mark = ",", na_str = "NA") %>%
  font(fontname = "Cambria", part = "all")  %>% #header, body, all
  fontsize(size = 7, part = "all") %>%
  autofit() # %>% 
  #width(j = 2, 2)
 

ft <- set_table_properties(ft, width = 1, layout = "autofit")
ft

```

# List of ScientificNames (Fish) - Grouped by Family
``` {r ByScientificNameFish, echo=FALSE }
sum_tbl <-
  sub %>% 
  arrange(ScientificName) %>% 
  filter(Phylum == 'Chordata') %>% 
  group_by(Family) %>%
  dplyr::summarize(
    ScientificName = paste(unique(ScientificName), collapse=" | "),
    TaxonRank = paste(unique(TaxonRank), collapse=" | "),
    UnFlagged = length(Flag[Flag == "0"]), 
    Flagged = length(Flag[Flag == "1"]),
    FlagReason = paste(unique(FlagReason), collapse=" | "),
    Records = n())

colkeys <- names(sum_tbl[,-1])
ft <- flextable(sum_tbl) %>% 
  theme_zebra() %>% #booktabs, vanilla, box, tron, tron_legacy, vader
  colformat_num(j = 4, digits = 0, big.mark = ",", na_str = "NA") %>%
  colformat_num(j = 7, digits = 0, big.mark = ",", na_str = "NA") %>%
  font(fontname = "Cambria", part = "all")  %>% #header, body, all
  fontsize(size = 7, part = "all") %>%
  autofit() # %>% 
  #width(j = 2, 2)
 

ft <- set_table_properties(ft, width = 1, layout = "autofit")
ft

```

# Location
## Localities and EventIDs (Dives/Tows/Hauls)
``` {r Locality, echo=FALSE}
sum_tbl <-
  sub %>%
  arrange(Locality, EventID) %>% 
  #filter(Phylum == "Porifera") %>%
  group_by(Locality) %>%
  dplyr::summarize(
    #min.Depth = min(as.numeric(MinimumDepthInMeters)),
    #max.Depth = max(as.numeric(MinimumDepthInMeters)), 
    #mean.Depth = mean(as.numeric(MinimumDepthInMeters)),
    #mean.DepthGIS = mean(as.numeric(gisDepth)),
    EventID = paste(unique(EventID), collapse=" | "),
    n = n()) %>%
  arrange(desc(n))


ft <- flextable(sum_tbl) %>% 
  theme_zebra() %>% 
  colformat_num(
    digits = 0, 
    big.mark = ",", 
    na_str = "NA"
  ) %>% 
  font(fontname = "Cambria", part = "all")  %>% 
  fontsize(size = 7, part = "all")

ft <- set_table_properties(ft, width = 1, layout = "autofit")
ft


```

## NavType
``` {r NavType, echo=FALSE}
sum_tbl <-
  sub %>%
  #filter(Phylum == "Porifera") %>%
  group_by(NavType) %>%
  dplyr::summarize(
    #min.Depth = min(as.numeric(MinimumDepthInMeters)),
    #max.Depth = max(as.numeric(MinimumDepthInMeters)), 
    #mean.Depth = mean(as.numeric(MinimumDepthInMeters)),
    #mean.DepthGIS = mean(as.numeric(gisDepth)),
    n = n()) %>%
  arrange(desc(n))


ft <- flextable(sum_tbl) %>% 
  theme_zebra() %>% 
  colformat_num(
    digits = 0, 
    big.mark = ",", 
    na_str = "NA"
  ) %>% 
  font(fontname = "Cambria", part = "all")  %>% 
  fontsize(size = 7, part = "all")

ft <- set_table_properties(ft, width = 1, layout = "autofit")
ft
```

## LocationAccuracy
``` {r LocationAccuracy, echo=FALSE}
sum_tbl <-
  sub %>%
  #filter(Phylum == "Porifera") %>%
  group_by(LocationAccuracy) %>%
  dplyr::summarize(
    #min.Depth = min(as.numeric(MinimumDepthInMeters)),
    #max.Depth = max(as.numeric(MinimumDepthInMeters)), 
    #mean.Depth = mean(as.numeric(MinimumDepthInMeters)),
    #mean.DepthGIS = mean(as.numeric(gisDepth)),
    n = n()) %>%
  arrange(desc(n))

ft <- flextable(sum_tbl) %>% 
  theme_zebra() %>% 
  colformat_num(
    digits = 0, 
    big.mark = ",", 
    na_str = "NA"
  ) %>% 
  font(fontname = "Cambria", part = "all")  %>% 
  fontsize(size = 7, part = "all")

ft <- set_table_properties(ft, width = 1, layout = "autofit")
ft
```

## Oceans
``` {r Ocean, echo=FALSE}
sum_tbl <-
  sub %>%
  #filter(Phylum == "Porifera") %>%
  group_by(Ocean) %>%
  dplyr::summarize(
    #min.Depth = min(as.numeric(MinimumDepthInMeters)),
    #max.Depth = max(as.numeric(MinimumDepthInMeters)), 
    #mean.Depth = mean(as.numeric(MinimumDepthInMeters)),
    #mean.DepthGIS = mean(as.numeric(gisDepth)),
    n = n()) %>%
  arrange(desc(n))

ft <- flextable(sum_tbl) %>% 
  theme_zebra() %>% 
  colformat_num(
    digits = 0, 
    big.mark = ",", 
    na_str = "NA"
  ) %>% 
  font(fontname = "Cambria", part = "all")  %>% 
  fontsize(size = 7, part = "all")

ft <- set_table_properties(ft, width = 1, layout = "autofit")
ft
```

## Countries
``` {r Country, echo=FALSE}

sum_tbl <-
  sub %>%
  #filter(Phylum == "Porifera") %>%
  group_by(Country) %>%
  dplyr::summarize(
    #min.Depth = min(as.numeric(MinimumDepthInMeters)),
    #max.Depth = max(as.numeric(MinimumDepthInMeters)), 
    #mean.Depth = mean(as.numeric(MinimumDepthInMeters)),
    #mean.DepthGIS = mean(as.numeric(gisDepth)),
    n = n()) %>%
  arrange(desc(n))


ft <- flextable(sum_tbl) %>% 
  theme_zebra() %>% 
  colformat_num(
    digits = 0, 
    big.mark = ",", 
    na_str = "NA"
  ) %>% 
  font(fontname = "Cambria", part = "all")  %>% 
  fontsize(size = 7, part = "all")

ft <- set_table_properties(ft, width = 1, layout = "autofit")
ft
```

## FishCouncilRegion(s)
``` {r FishCouncilRegions, echo=FALSE}
sum_tbl <-
  sub %>%
  #filter(Phylum == "Porifera") %>%
  group_by(FishCouncilRegion) %>%
  dplyr::summarize(
    #min.Depth = min(as.numeric(MinimumDepthInMeters)),
    #max.Depth = max(as.numeric(MinimumDepthInMeters)), 
    #mean.Depth = mean(as.numeric(MinimumDepthInMeters)),
    #mean.DepthGIS = mean(as.numeric(gisDepth)),
    n = n()) %>%
  arrange(desc(n))

ft <- flextable(sum_tbl) %>% 
  theme_zebra() %>% 
  colformat_num(
    digits = 0, 
    big.mark = ",", 
    na_str = "NA"
  ) %>% 
  font(fontname = "Cambria", part = "all")  %>% 
  fontsize(size = 7, part = "all")

ft <- set_table_properties(ft, width = 1, layout = "autofit")
ft
```

## Marine Eco-Regions of the World
``` {r MEOW, echo=FALSE}
sum_tbl <-
  sub %>%
  #filter(Phylum == "Porifera") %>%
  group_by(gisMEOW) %>%
  dplyr::summarize(
    #min.Depth = min(as.numeric(MinimumDepthInMeters)),
    #max.Depth = max(as.numeric(MinimumDepthInMeters)), 
    #mean.Depth = mean(as.numeric(MinimumDepthInMeters)),
    #mean.DepthGIS = mean(as.numeric(gisDepth)),
    n = n()) %>%
  arrange(desc(n))

ft <- flextable(sum_tbl) %>% 
  theme_zebra() %>% 
  colformat_num(
    digits = 0, 
    big.mark = ",", 
    na_str = "NA"
  ) %>% 
  font(fontname = "Cambria", part = "all")  %>% 
  fontsize(size = 7, part = "all")

ft <- set_table_properties(ft, width = 1, layout = "autofit")
ft
```


## Large Marine Ecosystem
``` {r LME, echo=FALSE}
sum_tbl <-
  sub %>%
  #filter(Phylum == "Porifera") %>%
  group_by(gisLME) %>%
  dplyr::summarize(
    #min.Depth = min(as.numeric(MinimumDepthInMeters)),
    #max.Depth = max(as.numeric(MinimumDepthInMeters)), 
    #mean.Depth = mean(as.numeric(MinimumDepthInMeters)),
    #mean.DepthGIS = mean(as.numeric(gisDepth)),
    n = n()) %>%
  arrange(desc(n))

ft <- flextable(sum_tbl) %>% 
  theme_zebra() %>% 
  colformat_num(
    digits = 0, 
    big.mark = ",", 
    na_str = "NA"
  ) %>% 
  font(fontname = "Cambria", part = "all")  %>% 
  fontsize(size = 7, part = "all")

ft <- set_table_properties(ft, width = 1, layout = "autofit")
ft
```


# Size, Abundance, and Condition
## Individual Count Summary
Nothing will appear here if no 'IndividualCount' values are present

``` {r IndividualCount, echo=FALSE}


sub2 <- sub %>% 
  filter(
    # is.na(Temperature) == F|
    as.character(IndividualCount) != '-999' 
    # as.character(Temperature) != 'NA' |
    # as.numeric(Temperature) > 0
  )

options(scipen=10000)

g <- ggplot(sub2, aes(IndividualCount,
                      #fill = Phylum
                      )) +
  geom_histogram(binwidth = 1) + 
  ylab("Number of Records")  +
  theme_bw(base_size = 15, base_family = "Cambria")

#display.brewer.all(colorblindFriendly=TRUE)
g + scale_fill_manual(values = brewer.pal(12, "Paired")[c(10,9)])


```


### IndividualCount
`r options(digits = 3)`
* min: `r min(sub$IndividualCount, digits = 3)`
* median: `r median(sub$IndividualCount, digits = 3)`
* max: `r max(sub$IndividualCount)`

`r options(digits = 5)`

## Categorical Abundance Values

Nothing will appear here if no 'CategoricalAbundance' values are present

``` {r CatAbundance, echo=FALSE}
sum_tbl <-
  sub %>%
  #filter(Phylum == "Porifera") %>%
  group_by(CategoricalAbundance) %>%
  dplyr::summarize(
    n = n()) %>%
  arrange(desc(CategoricalAbundance))

ft <- flextable(sum_tbl) %>% 
  theme_zebra() %>% 
  colformat_num(
    digits = 0, 
    big.mark = ",", 
    na_str = "NA"
  ) %>% 
  font(fontname = "Cambria", part = "all")  %>% 
  fontsize(size = 7, part = "all")

ft <- set_table_properties(ft, width = 1, layout = "autofit")
ft
```

## Density Values (per square meter)

Nothing will appear here if 'Density' values are not present

``` {r Density, echo=FALSE}

sub %>% filter(Density != -999) %>% ggplot(aes(x=Density)) +
  geom_histogram(binwidth = .2) +
  ylab("Number of Records")  +
  theme_bw(base_size = 15, base_family = "Cambria")

```

## Cover Values (%)

Arranged in descending order of median 'Cover' values. Nothing will appear here if no 'Cover' values are present

``` {r PercentCover, echo=FALSE, message=F, warning = F}
sum_tbl <-
  sub %>%
  filter(
    Cover != "-999" |
      is.na(Cover) == F
  ) %>%
  group_by(ScientificName) %>%
  dplyr::summarize(
    minCover = min(Cover),
    maxCover = max(Cover),
    medianCover = median(Cover),
    n = n()) %>%
  arrange(desc(medianCover))

ft <- flextable(sum_tbl) %>% 
  theme_zebra() %>% 
  colformat_num(
    digits = 0, 
    big.mark = ",", 
    na_str = "NA"
  ) %>% 
  font(fontname = "Cambria", part = "all")  %>% 
  fontsize(size = 7, part = "all")

ft <- set_table_properties(ft, width = 1, layout = "autofit")
ft

```

## Minimum and Maximum Size Values by Genus

This table will be empty if no size values present.

``` {r SizeSummary, echo=FALSE, warning = F}
sum_tbl <-
  sub %>%
  filter(
    VerbatimSize != '-999'
    ) %>%
  group_by(Genus) %>%
  dplyr::summarize(
    VerbatimSize_values = toString(unique(VerbatimSize)),
    min.Size = min(as.numeric(MinimumSize)),
    max.Size = max(as.numeric(MaximumSize)), 
    #mean.Depth = mean(as.numeric(MinimumDepthInMeters)),
    #mean.DepthGIS = mean(as.numeric(gisDepth)),
    n = n()) %>%
  arrange(desc(Genus))

ft <- flextable(sum_tbl) %>% 
  theme_zebra() %>% 
  colformat_num(
    digits = 0, 
    big.mark = ",", 
    na_str = "NA"
  ) %>% 
  font(fontname = "Cambria", part = "all")  %>% 
  fontsize(size = 7, part = "all")

ft <- set_table_properties(ft, width = 1, layout = "autofit")
ft
```

## Weight Values 

Table will be empty if no weight values are given.

``` {r WeightInKg, echo=FALSE,warning = F}
sum_tbl <-
  sub %>%
  filter(WeightInKg != "-999") %>%
  group_by(WeightInKg) %>%
  dplyr::summarize(
    #min.Depth = min(as.numeric(MinimumDepthInMeters)),
    #max.Depth = max(as.numeric(MinimumDepthInMeters)), 
    #mean.Depth = mean(as.numeric(MinimumDepthInMeters)),
    #mean.DepthGIS = mean(as.numeric(gisDepth)),
    n = n()) %>%
  arrange(desc(WeightInKg))

ft <- flextable(sum_tbl) %>% 
  theme_zebra() %>% 
  colformat_num(
    digits = 0, 
    big.mark = ",", 
    na_str = "NA"
  ) %>% 
  font(fontname = "Cambria", part = "all")  %>% 
  fontsize(size = 7, part = "all")

ft <- set_table_properties(ft, width = 1, layout = "autofit")
ft
```

## Condition Values 
``` {r Condition, echo=FALSE}
sum_tbl <-
  sub %>%
  #filter(is.na(Condition) == F) %>%
  group_by(Condition) %>%
  dplyr::summarize(
    #min.Depth = min(as.numeric(MinimumDepthInMeters)),
    #max.Depth = max(as.numeric(MinimumDepthInMeters)), 
    #mean.Depth = mean(as.numeric(MinimumDepthInMeters)),
    #mean.DepthGIS = mean(as.numeric(gisDepth)),
    n = n()) %>%
  arrange(desc(Condition))

ft <- flextable(sum_tbl) %>% 
  theme_zebra() %>% 
  colformat_num(
    digits = 0, 
    big.mark = ",", 
    na_str = "NA"
  ) %>% 
  font(fontname = "Cambria", part = "all")  %>% 
  fontsize(size = 7, part = "all")

ft <- set_table_properties(ft, width = 1, layout = "autofit")
ft
```

# Environment

## Habitat and Substrate

``` {r HabSub, echo=FALSE, warning = F}
sum_tbl <-
  sub %>%
  #filter(Phylum == "Porifera") %>%
  group_by(Habitat, Substrate) %>%
  dplyr::summarize(
    #Species = toString(unique(ScientificName)),
    # min.Depth = min(as.numeric(DepthInMeters)),
    # mean.Depth = mean(as.numeric(DepthInMeters)),
    # max.Depth = max(as.numeric(DepthInMeters)), 
    n = n()) %>%
  arrange(desc(n))

sum_tbl <- sum_tbl %>% filter(n > 100) 
  
ft <- flextable(sum_tbl) %>% 
  theme_zebra() %>% 
  colformat_num(
    digits = 0, 
    big.mark = ",", 
    na_str = "NA"
  ) %>% 
  font(fontname = "Cambria", part = "all")  %>% 
  fontsize(size = 7, part = "all")

ft <- set_table_properties(ft, width = 1, layout = "autofit")
ft


```

## Temperature, Salinity, Oxygen

Histogram of Temperature Values (Celcius) (empty if no Temperature values)

``` {r TemperatureHisto, echo=FALSE}
# ordering factors
sub2 <- sub %>% 
  filter(
   # is.na(Temperature) == F|
      as.character(Temperature) != '-999' 
     # as.character(Temperature) != 'NA' |
     # as.numeric(Temperature) > 0
    )

options(scipen=10000)

g <- ggplot(sub2, aes(Temperature,
                      #fill = Phylum
                      )) +
  geom_histogram(binwidth = .1) + 
  ylab("Number of Records")  +
  theme_bw(base_size = 15, base_family = "Cambria")

#display.brewer.all(colorblindFriendly=TRUE)
g + scale_fill_manual(values = brewer.pal(12, "Paired")[c(10,9)])


```

Histogram of Salinity Values (PSU) (empty if no Salinity values)

``` {r SalinityHisto, echo=FALSE}
# ordering factors
sub2 <- sub %>% 
  filter(
    is.na(Salinity) == F,
    Salinity != '-999'
    )

options(scipen=10000)

g <- ggplot(sub2, aes(Salinity 
                      #fill = Phylum
                      )) +
  geom_histogram(binwidth = .01) + 
  ylab("Number of Records")  +
  theme_bw(base_size = 15, base_family = "Cambria")

#display.brewer.all(colorblindFriendly=TRUE)
g + scale_fill_manual(values = brewer.pal(12, "Paired")[c(10,9)])


```



Histogram of Oxygen Values (ml/L) (empty if no Oxygen values)

``` {r OxygenHisto, echo=FALSE}
# ordering factors
sub2 <- sub %>% 
  filter(
    is.na(Oxygen) == F,
    Oxygen != '-999'
    )

options(scipen=10000)

g <- ggplot(sub2, aes(Oxygen, fill = Phylum)) +
  geom_histogram(binwidth = .1) + 
  ylab("Number of Records")  +
  theme_bw(base_size = 15, base_family = "Cambria")

#display.brewer.all(colorblindFriendly=TRUE)
g + scale_fill_manual(values = brewer.pal(12, "Paired"))#[c(10,9)])


```



## pH Values
``` {r pH, echo=FALSE}
sum_tbl <- sub %>% 
  #filter(Phylum == "Porifera") %>%
  group_by(pH) %>%
  dplyr::summarize(
    #min.Depth = min(as.numeric(MinimumDepthInMeters)),
    #max.Depth = max(as.numeric(MinimumDepthInMeters)), 
    #mean.Depth = mean(as.numeric(MinimumDepthInMeters)),
    #mean.DepthGIS = mean(as.numeric(gisDepth)),
    n = n()) %>%
  arrange(desc(pH))

ft <- flextable(sum_tbl) %>% 
  theme_zebra() %>% 
  colformat_num(
    digits = 0, 
    big.mark = ",", 
    na_str = "NA"
  ) %>% 
  font(fontname = "Cambria", part = "all")  %>% 
  fontsize(size = 7, part = "all")

ft <- set_table_properties(ft, width = 1, layout = "autofit")
ft
```

## pHscale Values
``` {r pHscale, echo=FALSE}
sum_tbl <- sub %>% 
  #filter(Phylum == "Porifera") %>%
  group_by(pHscale) %>%
  dplyr::summarize(
    #min.Depth = min(as.numeric(MinimumDepthInMeters)),
    #max.Depth = max(as.numeric(MinimumDepthInMeters)), 
    #mean.Depth = mean(as.numeric(MinimumDepthInMeters)),
    #mean.DepthGIS = mean(as.numeric(gisDepth)),
    n = n()) %>%
  arrange(desc(pHscale))

ft <- flextable(sum_tbl) %>% 
  theme_zebra() %>% 
  colformat_num(
    digits = 0, 
    big.mark = ",", 
    na_str = "NA"
  ) %>% 
  font(fontname = "Cambria", part = "all")  %>% 
  fontsize(size = 7, part = "all")

ft <- set_table_properties(ft, width = 1, layout = "autofit")
ft

```

## pCO2 Values
``` {r pHCO2, echo=FALSE}
sum_tbl <- sub %>% 
  #filter(Phylum == "Porifera") %>%
  group_by(pCO2) %>%
  dplyr::summarize(
    #min.Depth = min(as.numeric(MinimumDepthInMeters)),
    #max.Depth = max(as.numeric(MinimumDepthInMeters)), 
    #mean.Depth = mean(as.numeric(MinimumDepthInMeters)),
    #mean.DepthGIS = mean(as.numeric(gisDepth)),
    n = n()) %>%
  arrange(desc(pCO2))

ft <- flextable(sum_tbl) %>% 
  theme_zebra() %>% 
  colformat_num(
    digits = 0, 
    big.mark = ",", 
    na_str = "NA"
  ) %>% 
  font(fontname = "Cambria", part = "all")  %>% 
  fontsize(size = 7, part = "all")

ft <- set_table_properties(ft, width = 1, layout = "autofit")
ft

```

## TA Values
``` {r TA, echo=FALSE}
sum_tbl <- sub %>% 
  #filter(Phylum == "Porifera") %>%
  group_by(TA) %>%
  dplyr::summarize(
    #min.Depth = min(as.numeric(MinimumDepthInMeters)),
    #max.Depth = max(as.numeric(MinimumDepthInMeters)), 
    #mean.Depth = mean(as.numeric(MinimumDepthInMeters)),
    #mean.DepthGIS = mean(as.numeric(gisDepth)),
    n = n()) %>%
  arrange(desc(TA))

ft <- flextable(sum_tbl) %>% 
  theme_zebra() %>% 
  colformat_num(
    digits = 0, 
    big.mark = ",", 
    na_str = "NA"
  ) %>% 
  font(fontname = "Cambria", part = "all")  %>% 
  fontsize(size = 7, part = "all")

ft <- set_table_properties(ft, width = 1, layout = "autofit")
ft

```

## DIC Values
``` {r DIC, echo=FALSE}
sum_tbl <- sub %>% 
  #filter(Phylum == "Porifera") %>%
  group_by(DIC) %>%
  dplyr::summarize(
    #min.Depth = min(as.numeric(MinimumDepthInMeters)),
    #max.Depth = max(as.numeric(MinimumDepthInMeters)), 
    #mean.Depth = mean(as.numeric(MinimumDepthInMeters)),
    #mean.DepthGIS = mean(as.numeric(gisDepth)),
    n = n()) %>%
  arrange(desc(DIC))

ft <- flextable(sum_tbl) %>% 
  theme_zebra() %>% 
  colformat_num(
    digits = 0, 
    big.mark = ",", 
    na_str = "NA"
  ) %>% 
  font(fontname = "Cambria", part = "all")  %>% 
  fontsize(size = 7, part = "all")

ft <- set_table_properties(ft, width = 1, layout = "autofit")
ft

```

# Temporal  
## Observation Time Frame  
`r min(as.numeric(sub$ObservationYear[sub$ObservationYear != "-999"]), na.rm=TRUE)` to `r max(sub$ObservationYear, na.rm=TRUE)`

## Observation Dates 

### Number of ObservationDate values recorded:

`r x <- sub %>%
filter(
is.na(ObservationDate) == F |
ObservationDate != '-999'
) %>%
group_by(DatasetID) %>%
dplyr::summarize(
n = n()
) 

sum(x$n)`

### ObservationDate values grouped by SurveyID: 

```{r, echo=F}

x <- sub %>%
filter(
is.na(ObservationDate) == F |
ObservationDate != '-999'
) %>%
group_by(SurveyID) %>%
dplyr::summarize(ObservationDates = paste(unique(ObservationDate), collapse = ' | '),
n = n()
)

ft <- flextable(x) %>% 
  theme_zebra() %>% 
  colformat_num(
    digits = 0, 
    big.mark = ",", 
    na_str = "NA"
  ) %>% 
  font(fontname = "Cambria", part = "all")  %>% 
  fontsize(size = 7, part = "all")

ft <- set_table_properties(ft, width = 1, layout = "autofit")
ft


```


## Observation Time 

`r x <- sub %>%
filter(
is.na(ObservationTime) == F |
ObservationTime != '-999'
) %>%
group_by(DatasetID) %>%
dplyr::summarize(
n = n()
)`

### Number of ObservationTime values recorded: `r sum(x$n)`

View of the first 10 ObservationTime values: 

```{r, echo=F}
x <- sub %>%
filter(
is.na(ObservationTime) == F |
ObservationTime != '-999'
) %>%
group_by(ObservationTime) %>%
dplyr::summarize(
n = n()
) 


ft <- flextable(head(x, n=10)) %>% 
  theme_zebra() %>% 
  colformat_num(
    digits = 0, 
    big.mark = ",", 
    na_str = "NA"
  ) %>% 
  font(fontname = "Cambria", part = "all")  %>% 
  fontsize(size = 7, part = "all")

ft <- set_table_properties(ft, width = 1, layout = "autofit")
ft

```

## Modified Dates 
``` {r Modified, echo=FALSE}
sum_tbl <-
  sub %>%
  #filter(Phylum == "Porifera") %>%
  group_by(Modified) %>%
  dplyr::summarize(
    #min.Depth = min(as.numeric(MinimumDepthInMeters)),
    #max.Depth = max(as.numeric(MinimumDepthInMeters)), 
    #mean.Depth = mean(as.numeric(MinimumDepthInMeters)),
    #mean.DepthGIS = mean(as.numeric(gisDepth)),
    n = n()) %>%
  arrange(desc(Modified))

ft <- flextable(sum_tbl) %>% 
  theme_zebra() %>% 
  colformat_num(
    digits = 0, 
    big.mark = ",", 
    na_str = "NA"
  ) %>% 
  font(fontname = "Cambria", part = "all")  %>% 
  fontsize(size = 7, part = "all")

ft <- set_table_properties(ft, width = 1, layout = "autofit")
ft

```

## Entry Dates 
``` {r EntryDate, echo=FALSE}
sum_tbl <-
  sub %>%
  #filter(Phylum == "Porifera") %>%
  group_by(EntryDate) %>%
  dplyr::summarize(
    #min.Depth = min(as.numeric(MinimumDepthInMeters)),
    #max.Depth = max(as.numeric(MinimumDepthInMeters)), 
    #mean.Depth = mean(as.numeric(MinimumDepthInMeters)),
    #mean.DepthGIS = mean(as.numeric(gisDepth)),
    n = n()) %>%
  arrange(desc(EntryDate))

ft <- flextable(sum_tbl) %>% 
  theme_zebra() %>% 
  colformat_num(
    digits = 0, 
    big.mark = ",", 
    na_str = "NA"
  ) %>% 
  font(fontname = "Cambria", part = "all")  %>% 
  fontsize(size = 7, part = "all")

ft <- set_table_properties(ft, width = 1, layout = "autofit")
ft

```

## Entry Update 

``` {r EntryUpdate, echo=FALSE}
sum_tbl <-
  sub %>%
  #filter(Phylum == "Porifera") %>%
  group_by(EntryUpdate) %>%
  dplyr::summarize(
    #min.Depth = min(as.numeric(MinimumDepthInMeters)),
    #max.Depth = max(as.numeric(MinimumDepthInMeters)), 
    #mean.Depth = mean(as.numeric(MinimumDepthInMeters)),
    #mean.DepthGIS = mean(as.numeric(gisDepth)),
    n = n()) %>%
  arrange(desc(EntryUpdate))

ft <- flextable(sum_tbl) %>% 
  theme_zebra() %>% 
  colformat_num(
    digits = 0, 
    big.mark = ",", 
    na_str = "NA"
  ) %>% 
  font(fontname = "Cambria", part = "all")  %>% 
  fontsize(size = 7, part = "all")

ft <- set_table_properties(ft, width = 1, layout = "autofit")
ft

```

# Flagged Records
``` {r Flags_and_CatalogNumbers, echo=FALSE, warning=FALSE}
sum_tbl <-
  sub %>%
  filter(Flag == "1") %>%
  group_by(FlagReason) %>%
  dplyr::summarize(
    #CatalogNumbers = toString(CatalogNumber),
    taxa = paste(unique(ScientificName), collapse=" | "),
    n = n()
    ) %>%
  arrange(desc(n))

ft <- flextable(sum_tbl) %>% 
  theme_zebra() %>% 
  colformat_num(
    digits = 0, 
    big.mark = ",", 
    na_str = "NA"
  ) %>% 
  font(fontname = "Cambria", part = "all")  %>% 
  fontsize(size = 7, part = "all")

ft <- set_table_properties(ft, width = 1, layout = "autofit")
ft
```

# Depth
* Minimum Depth: `r min(sub[sub$DepthInMeters != '-999',]$DepthInMeters)`  
* Median Depth: `r median(sub[sub$DepthInMeters != '-999',]$DepthInMeters)`  
* Maximum Depth: `r max(sub[sub$DepthInMeters != '-999',]$DepthInMeters)`  

## Histogram of Depth Distribution
``` {r Coral_Depth_Summary_Boxplots, echo=FALSE, message=FALSE, warning=FALSE,fig.width=7, fig.height=3.5, dpi=300, cache=FALSE}

sub3 <- sub %>% filter(Flag == '0')

g <- ggplot(sub3, aes(DepthInMeters, fill = Phylum)) +
  geom_histogram(alpha = 0.5) + 
  xlim(0,7000) +
  ylab("Number of Records") + 
  xlab('DepthInMeters') + 
  scale_x_reverse() +   
  coord_flip() +
  theme_bw(base_size = 15, base_family = "Cambria")

#display.brewer.all(colorblindFriendly=TRUE)
g + scale_fill_manual(values = brewer.pal(12, "Paired")[c(10,9,8,7,6,5,4,3,2,1)])

# library(ggplot2)
# g <- ggplot(sub, aes(DepthInMeters)) +
#   geom_histogram(binwidth = 1) + 
#   theme(text = element_text(size=20)) +
#   scale_x_reverse() +
#   coord_flip() +
#   ylab("Number of Records") +
#   xlab("DepthInMeters") + 
#   theme_bw(base_size = 15, base_family = "Cambria")
# 
# g

```

# DepthCheck

``` {r Depth_Check, echo=FALSE, message=FALSE, warning=FALSE,fig.width=7, fig.height=3.5, dpi=300, eval = TRUE, cache=FALSE}

##### setting thresholds for difference and creating flags##### 
t <- 1000
z <- 500
percrm <- 30
perother <- 60

##### building selection conditions for each depth value #####
#finding depth difference for CRM
sub$CRMDepthDiff <- abs(sub$DepthInMeters - sub$gisCRMDepth)
sub$CRMDepthDiff <- (sub$CRMDepthDiff/sub$gisCRMDepth) * 100
#crm depth flags
crm <- sub %>%
  filter(CRMDepthDiff > percrm & gisCRMDepth != "-999") %>%
  dplyr::select(CatalogNumber, DatasetID, SampleID, TrackingID, ScientificName, Locality, DepthInMeters, gisEtopoDepth,
                gisGEBCODepth, gisCRMDepth, MinimumDepthInMeters, MaximumDepthInMeters,
                NavType, LocationAccuracy, LocationComments, Flag, FlagReason, SurveyID,
                Vessel, VehicleName, ObservationYear, Latitude, Longitude)
#finding depth difference for Etopo
sub$EtopoDepthDiff <- abs(sub$DepthInMeters - sub$gisEtopoDepth)
sub$EtopoDepthDiff <- (sub$EtopoDepthDiff/sub$gisEtopoDepth) * 100
#etopo depth flags
etopo <- sub %>%
  filter(EtopoDepthDiff > perother & gisEtopoDepth != "-999") %>%
  dplyr::select(CatalogNumber, DatasetID, SampleID, TrackingID, ScientificName, Locality, DepthInMeters, gisEtopoDepth,
                gisGEBCODepth, gisCRMDepth, MinimumDepthInMeters, MaximumDepthInMeters,
                NavType, LocationAccuracy, LocationComments, Flag, FlagReason, SurveyID,
                Vessel, VehicleName, ObservationYear, Latitude, Longitude)
#finding depth difference for GEBCO
sub$GEBCODepthDiff <- abs(sub$DepthInMeters - sub$gisGEBCODepth)
sub$GEBCODepthDiff <- (sub$GEBCODepthDiff/sub$gisGEBCODepth) * 100

#GEBCO depth flags
gebco <- sub %>%
  filter(GEBCODepthDiff > perother & gisGEBCODepth != "-999") %>%
  dplyr::select(CatalogNumber, DatasetID, SampleID, TrackingID, ScientificName, Locality, DepthInMeters, gisEtopoDepth,
                gisGEBCODepth, gisCRMDepth, MinimumDepthInMeters, MaximumDepthInMeters,
                NavType, LocationAccuracy, LocationComments, Flag, FlagReason, SurveyID,
                Vessel, VehicleName, ObservationYear, Latitude, Longitude)
#depth flags from reported values
depth <- sub %>%
  filter(as.numeric(DepthInMeters) == -999 | as.numeric(DepthInMeters) < 30) %>%
  dplyr::select(CatalogNumber, DatasetID, SampleID, TrackingID, ScientificName, Locality, DepthInMeters, gisEtopoDepth,
                gisGEBCODepth, gisCRMDepth, MinimumDepthInMeters, MaximumDepthInMeters,
                NavType, LocationAccuracy, LocationComments, Flag, FlagReason, SurveyID,
                Vessel, VehicleName, ObservationYear, Latitude, Longitude)


##### building joint file of all depth probs and de-dup
dx <- rbind(crm,etopo,gebco,depth)

##### testing for duplicated records #####
dx <- dx[!duplicated(dx), ]

```

## Depth Summary by ScientificName (Corals)

Table is sorted by median depth in ascending order.

``` {r DepthSumByScientificNameCorals, echo=FALSE}
sum_tbl <-
  sub %>%
  filter(Phylum == "Cnidaria") %>%
  group_by(ScientificName) %>%
  dplyr::summarize(
    min.Depth = min(as.numeric(DepthInMeters)),
    median.Depth = median(as.numeric(DepthInMeters)),
    max.Depth = max(as.numeric(DepthInMeters)), 
    n = n()) %>%
  arrange(median.Depth)

ft <- flextable(sum_tbl) %>% 
  theme_zebra() %>% 
  colformat_num(
    digits = 0, 
    big.mark = ",", 
    na_str = "NA"
  ) %>% 
  font(fontname = "Cambria", part = "all")  %>% 
  fontsize(size = 7, part = "all")

ft <- set_table_properties(ft, width = 1, layout = "autofit")
ft
```

## Depth Summary by ScientificName (Sponges)

Table is sorted by median depth in ascending order.

``` {r DepthSumByScientificNameSponges, echo=FALSE}
sum_tbl <-
  sub %>%
  filter(Phylum == "Porifera") %>%
  group_by(ScientificName) %>%
  dplyr::summarize(
    min.Depth = min(as.numeric(DepthInMeters)),
    median.Depth = median(as.numeric(DepthInMeters)),
    max.Depth = max(as.numeric(DepthInMeters)), 
    n = n()) %>%
  arrange(median.Depth)

ft <- flextable(sum_tbl) %>% 
  theme_zebra() %>% 
  colformat_num(
    digits = 0, 
    big.mark = ",", 
    na_str = "NA"
  ) %>% 
  font(fontname = "Cambria", part = "all")  %>% 
  fontsize(size = 7, part = "all")

ft <- set_table_properties(ft, width = 1, layout = "autofit")
ft
```

## Points > `r percrm` percent different than Coastal Relief Model
Figure will not show points if none meet the criteria.

``` {r Depth_CRM, echo=FALSE, message=FALSE, warning=FALSE,fig.width=7, fig.height=3.5, dpi=300, eval = TRUE, cache=FALSE}

p <- ggplot(crm, aes(DepthInMeters, gisCRMDepth))
p + 
  geom_point(aes(shape = factor(Flag), size = as.numeric(Flag)), size = 2) +
  scale_shape_manual(values=c(1,4)) +
  geom_vline(aes(xintercept = 50)) +
  geom_hline(aes(yintercept = 50)) +
  geom_abline(col = "gray60") +
  theme_bw(base_size = 15, base_family = "Cambria")

```  


``` {r crmtable, echo=FALSE, eval = F}

# ## Table of Points > `r percrm` percent different than Coastal Relief Model
# Nothing will show here if criteria is not met

sum_tbl <-
  crm %>%
  #filter(Flag == "1") %>%
  group_by(CatalogNumber) %>%
  dplyr::summarize(
    Scientific_Name = ScientificName,
    Depth = DepthInMeters,
    Etopo = gisEtopoDepth,
    GEBCO = gisGEBCODepth,
    CRM = gisCRMDepth, 
    Flagged = unique(Flag)) %>%
  arrange(desc(Flagged))
sum_tbl <- kable(sum_tbl, row.names = F, digits = 2)
sum_tbl

```


## Points > `r perother` percent different than ETOPO1 
Figure will not show points if none meet the criteria.

``` {r Depth_ETOPO, echo=FALSE, message=FALSE, warning=FALSE,fig.width=7, fig.height=3.5, dpi=300, eval = TRUE, cache=FALSE}
p <- ggplot(etopo, aes(DepthInMeters, gisEtopoDepth))
p + 
  geom_point(aes(shape = factor(Flag), size = as.numeric(Flag)), size = 2) +
  scale_shape_manual(values=c(1,4)) +
  geom_vline(aes(xintercept = 50)) +
  geom_hline(aes(yintercept = 50)) +
  geom_abline(col = "gray60") +
  theme_bw(base_size = 15, base_family = "Cambria")

```



``` {r etopotable, echo=FALSE, eval = F}

# ## Table of Points > `r perother` percent different than ETOPO
# Nothing will show here if criteria is not met

sum_tbl <-
  etopo %>%
  #filter(Phylum == "Porifera") %>%
  group_by(CatalogNumber) %>%
  dplyr::summarize(
    Scientific_Name = ScientificName,
    Depth = DepthInMeters,
    Etopo = gisEtopoDepth,
    GEBCO = gisGEBCODepth,
    CRM = gisCRMDepth,
   Flagged = unique(Flag)) %>%
  arrange(desc(Flagged))
sum_tbl <- kable(sum_tbl, row.names = F, digits = 2)
sum_tbl

```

## Points > `r perother` percent different than GEBCO
Figure will not show points if none meet the criteria.

``` {r Depth_GEBCO, echo=FALSE, message=FALSE, warning=FALSE,fig.width=7, fig.height=3.5, dpi=300, eval = TRUE, cache=FALSE}
p <- ggplot(gebco, aes(DepthInMeters, gisGEBCODepth))
p + 
  geom_point(aes(shape = factor(Flag), size = as.numeric(Flag)), size = 2) +
  scale_shape_manual(values=c(1,4)) +
  geom_vline(aes(xintercept = 50)) +
  geom_hline(aes(yintercept = 50)) +
  geom_abline(col = "gray60") +
  theme_bw(base_size = 15, base_family = "Cambria")

```


``` {r gebcotable, echo=FALSE, eval = F}

# ## Table of Points > `r perother` percent different than GEBCO
# Nothing will show here if criteria is not met## Table of Points 

sum_tbl <-
  gebco %>%
  #filter(Phylum == "Porifera") %>%
  group_by(CatalogNumber) %>%
  dplyr::summarize(
    Scientific_Name = ScientificName,
    Depth = DepthInMeters,
    Etopo = gisEtopoDepth,
    GEBCO = gisGEBCODepth,
    CRM = gisCRMDepth, 
    Flagged = unique(Flag)) %>%
  arrange(desc(Flagged))
sum_tbl <- kable(sum_tbl, row.names = F, digits = 2)
sum_tbl

```

## Points < 30m depth
Figure will not show points if none meet the criteria.

``` {r Depth_Shallow, echo=FALSE, message=FALSE, warning=FALSE,fig.width=7, fig.height=3.5, dpi=300, eval = TRUE, cache=FALSE}
p <- ggplot(depth, aes(DepthInMeters, gisGEBCODepth))
p + 
  geom_point(aes(shape = factor(Flag), size = as.numeric(Flag)), size = 2) +
  scale_shape_manual(values=c(1,4)) +
  geom_vline(aes(xintercept = 30)) +
  geom_hline(aes(yintercept = 30)) +
  geom_abline(col = "gray60") +
  theme_bw(base_size = 15, base_family = "Cambria")

```

## Table of Points < 30m depth
Nothing will show here if criteria is not met
``` {r shallowtable, echo=FALSE}
sum_tbl <-
  depth %>%
  #filter(Phylum == "Porifera") %>%
  group_by(CatalogNumber) %>%
  dplyr::summarize(
    Scientific_Name = ScientificName,
    Depth = DepthInMeters,
    Etopo = gisEtopoDepth,
    GEBCO = gisGEBCODepth,
    CRM = gisCRMDepth) %>%
  arrange(desc(Depth))
sum_tbl <- kable(sum_tbl, row.names = F, digits = 2)
sum_tbl

```

# Standardized Variable Checks

``` {r standardized_table_checks, echo=FALSE}
Vessel <- setdiff(unique(sub$Vessel), unique(filt$Vessel))
VehicleName <- setdiff(unique(sub$VehicleName), unique(filt$VehicleName))
PI <- setdiff(unique(sub$PI), unique(filt$PI))
PIAffiliation <- setdiff(unique(sub$PIAffiliation), unique(filt$PIAffiliation))
Repository <- setdiff(unique(sub$Repository), unique(filt$Repository))
IdentifiedBy <- setdiff(unique(sub$IdentifiedBy), unique(filt$IdentifiedBy))
IdentificationQualifier <- setdiff(unique(sub$IdentificationQualifier), unique(filt$IdentificationQualifier))
DataProvider <- setdiff(unique(sub$DataProvider), unique(filt$DataProvider))
DatasetID <- setdiff(unique(sub$DatasetID), unique(filt$DatasetID))
SurveyID <- setdiff(unique(sub$SurveyID), unique(filt$SurveyID))
```

If name appears below after variable, then it does not match the standard name. That means it needs to be changed to match, or it is actually a new name that we have not encountered before.

* Vessel: **`r Vessel`**  
* VehicleName: **`r VehicleName`**  
* PI: **`r PI`**  
* PIAffiliation: **`r PIAffiliation`**  
* Repository: **`r Repository`**  
* IdentifiedBy: **`r IdentifiedBy`**  
* IdentificationQualifier: **`r IdentificationQualifier`**  
* DataProvider: **`r DataProvider`**  
* DatasetID: **`r DatasetID`**  
* SurveyID: **`r SurveyID`**  

# Compare our current holdings for the same 'Vessel'

``` {r standardized_SurveyID_check, echo=FALSE, cache = FALSE, warning = F}

SurveyIDPattern <- filt %>% 
  filter(Vessel %in% unique(sub$Vessel)) %>% 
  group_by(Vessel) %>% 
  dplyr::summarise(
    VehicleName = paste(unique(VehicleName), collapse = ' | '),
    SurveyIDs = paste(unique(SurveyID), collapse = ' | '),
    DatasetIDs = paste(unique(DatasetID), collapse = ' | '),
    DataProviders = paste(unique(DataProvider), collapse = ' | '),
    Repository = paste(unique(Repository), collapse = ' | '),
    #ObservationDates = paste(unique(ObservationDate), collapse = ' | '),
    PIs = paste(unique(PI), collapse = ' | '), 
    DataContacts = paste(unique(DataContact), collapse = ' | '),
    #Reporters = paste(unique(Reporter), collapse = ' | ')
  )
  
ft <- flextable(SurveyIDPattern) %>% 
  theme_zebra() %>% 
  colformat_num(
    digits = 0, 
    big.mark = ",", 
    na_str = "NA"
  ) %>% 
  font(fontname = "Cambria", part = "all")  %>% 
  fontsize(size = 7, part = "all")

ft <- set_table_properties(ft, width = 1, layout = "autofit")
ft

```




