---
title: "Area of Interest Report - SFAA near Oculina Bank HAPC"
author: "Robert P. McGuinn"
date: "December 09, 2024"
output: word_document
---

```{r packages, eval=T, echo=F, message=FALSE, warning=FALSE, cache=FALSE}
## load packages
#install.packages(tidyverse)
#install.packages('openxlsx')
library(tidyverse)
library(openxlsx)
library(sf)
library(RColorBrewer)
library(raster)
library(leaflet)
library(extrafont)
library(RColorBrewer)
library(rmarkdown)
library(knitr)
library(maps)
library(raster)
library(httr)
library(jsonlite)
library(rnaturalearth)
library(rnaturalearthdata)
library(openxlsx)
library(googledrive)
library(spocc)


```

```{r load_NDB, echo=F, message = FALSE, warning = FALSE, cache = TRUE, eval = F}
## load NDB
source('c:/rworking/deepseatools/code/mod_load_current_NDB.R')

```

```{r set_ndb_version, echo=F}
## create needed variables
db_version <- unique(filt$DatabaseVersion)

```

```{r filter_geo_OR_query, echo=F, warning=F, message=F, eval=T}
##### filter occurrences by query or box or both ##### 

## set bounding box variables
minlat <- 25
maxlat <- 35
minlon <- -82
maxlon <- -71

## subset data by coordinates
x <- subset(filt, as.numeric(Latitude) > minlat &
                   as.numeric(Latitude) < maxlat &
                   as.numeric(Longitude) > minlon &
                   as.numeric(Longitude) < maxlon)
                
## also subset data by other variables by other criteria
# x <- x %>% filter(ScientificName == "")

## checking
# dim(x)

```

```{r create_geo_from_points, echo=FALSE, echo = FALSE, message = FALSE, warning=FALSE, eval=T}
## create spatial points data frame from a subset of points
x_geo <- x

## create spdf
coordinates(x_geo) <- c("Longitude", "Latitude")
proj4string(x_geo) <- "+proj=longlat +ellps=WGS84 +datum=WGS84"

## OPTIONAL: write out featureclass to geodatabase from sp object 
# fgdb_path <- "C:/data/aprx/aoi/aoi.gdb"
# arc.write(file.path(fgdb_path, 'x_geo'), data=x_geo, overwrite = TRUE)

```

```{r create_sf_from_points, message = FALSE, warning = FALSE, echo=F}
## transform coral and sponge occurrence points to 'sf' objects
##### transform coral and sponge points to sf #####
points <- st_as_sf(x_geo, wkt = "geom")

## OPTIONAL: write out featureclass from sf object
# fgdb_path <- "C:/data/aprx/aoi/aoi.gdb"
# arc.write(file.path(fgdb_path, 'points'), data=points, overwrite = TRUE)
rm(x)

```

```{r import_aoi, echo=FALSE, echo = F, message = FALSE, warning=FALSE, results = "hide", eval=T, }
shapefile_path <- "C:/rworking/deepseatools/indata/sfaa.shp"
sfaa <- st_read(shapefile_path)

##### get the Oculina HAPC ##### 
shapefile_path <- "C:/rworking/deepseatools/indata/20221104_protected_areas_update_HColeman/20221104_protected_areas.shp"
protected_areas <- st_read(shapefile_path)
oculina <- protected_areas %>% filter(Sitename == 'Oculina Bank HAPC')

##### check #####
# plot(oculina)
# plot(sfaa)

```

```{r buffering, echo=F}
##### buffer operations ##### 
## create 5000 meter buffer
sfaa_buf_5000 <- st_buffer(sfaa, dist=5000) # map units are in meters

##create 20000 meter buffer
sfaa_buf_20000 <- st_buffer(sfaa, dist=20000) # map units are in meters


```

```{r intersect, echo=F}
##### intersection operations ##### 
## transform CRS for coral and sponge points projection to match polygons
points_transform <- st_transform(points, crs = st_crs(sfaa))

## checking
# st_crs(points_transform)

## find points within the polygons
points_in_0 <- st_join(points_transform, sfaa, join = st_within) # no buffer
points_in_5000 <- st_join(points_transform, sfaa_buf_5000, join = st_within) # 5000 meter buffer
points_in_20000 <- st_join(points_transform, sfaa_buf_20000, join = st_within) # 20000 meter buffer

## checking
# dim(points_in_0)
# dim(points_in_5000)
# dim(points_in_20000)
# table(points_in_0$Shape_Area, useNA = 'always')
# table(points_in_5000$Shape_Area, useNA = 'always')
# table(points_in_20000$Shape_Area, useNA = 'always')

## select intersecting points only and create separate data frames
points_in_0_select <- points_in_0 %>% filter(is.na(Shape_Area) == F)
points_in_5000_select <- points_in_5000 %>% filter(is.na(Shape_Area) == F)
points_in_20000_select <- points_in_20000 %>% filter(is.na(Shape_Area) == F)

## checking 
# dim(points_in_0_select)
# dim(points_in_5000_select)
# dim(points_in_20000_select)


```

```{r load_images_to_local_folder, echo =F, eval=F}
## load any images to a local folder 
#### WARNING: This chunk takes a long time.  Go get a coffee #####
###### download all images and put them in a folder #####
z <- points_in_0_select %>% filter(is.na(ImageURL) == F)

## check 
length(z$ImageFilePath) # if this is zero, then the code below will return nothing

## manual: change directory to send images
dir.create('C:/rworking/deepseatools/indata/imageset_0_pt2', recursive = TRUE)
setwd("C:/rworking/deepseatools/indata/imageset_0_pt2")

## loop to download images 
for(i in 1:length(z$CatalogNumber)){
  download.file(as.character(z$ImageURL[i]),
                destfile = paste("DSCRTP",
                                 z$CatalogNumber[i],
                                 z$ScientificName[i],
                                 z$DepthInMeters[i],
                                 basename(as.character(z$ImageURL[i])),
                                 sep = '_'),
                mode = "wb")
}

```

```{r load_images_to_google_drive, echo=F, eval=F}
# load the images to a Google Drive folder
## WARNING: This chunk could take a longish time. Go get a coffee

##### Loading files from a local folder to Google Drive ##### 
## MANUAL CHANGE "folderurl" to the desired drive folder ID
folderurl <- "https://drive.google.com/drive/folders/1oOiCVRWJUc4fQQ1dec9q3APYfk1Ea4nj"

## get the list of files from the local folder
files <- list.files(path="C:/rworking/deepseatools/indata/imageset_5000", full.names=TRUE, recursive=FALSE)

## loop upload images to Google Drive
for(i in files){
drive_upload(i,
             path = as_id(folderurl),
             overwrite = T)
}

```

```{r get_obis_data, cache=T, echo=F, eval=T} 
##### getting OBIS data #####
## first get the original projection back (this CRS is what OBIS and GBIF expect)
## ID["EPSG",6326]]
sfaa_transform <- st_transform(sfaa, crs = st_crs(x_geo))
sfaa_transform <- st_make_valid(sfaa_transform)
sfaa_transform_wkt <- sf::st_as_text(sf::st_geometry(sfaa_transform))

## using bounding box variables defined above in another chunk
library(spocc)
out <- occ(geometry = sfaa_transform_wkt, from = 'obis', limit = 10000)
obis <- as.data.frame(out$obis$data)


## Convert to sf object using coordinates
obis_sf <- st_as_sf(
  obis,
  coords = c("longitude", "latitude"), # specify coordinate columns
  crs = 4326                         # set CRS to WGS84 (EPSG:4326)
)

## check 
# dim(obis)
# View(obis)
# names(obis)
# table(obis$catalogNumber)
# table(obis$id)
# table(obis$aphiaID)
# plot(obis_sf)

```

```{r get_obis_data_in_5000m_buffer, cache=T, echo=F, eval=T}
##### first transform back to the the original projection back (this is what OBIS and GBIF expect) #####
sfaa_buf_5000_transform <- st_transform(sfaa_buf_5000, crs = st_crs(x_geo))
sfaa_buf_5000_transform <- st_make_valid(sfaa_buf_5000_transform)
sfaa_buf_5000_transform_wkt <- sf::st_as_text(sf::st_geometry(sfaa_buf_5000_transform))

##### getting OBIS data #####
# using bounding box variables defined above
library(spocc)
out <- occ(geometry = sfaa_buf_5000_transform_wkt, from = 'obis', limit = 10000)
obisbuffer5000 <- as.data.frame(out$obis$data)

## Convert to sf object using coordinates
obisbuffer5000_sf <- st_as_sf(
  obisbuffer5000,
  coords = c("longitude", "latitude"), # specify coordinate columns
  crs = 4326                         # set CRS to WGS84 (EPSG:4326)
)

## check 
# dim(obisbuffer5000)
# View(obisbufrer5000)

```

```{r get_obis_data_in_20000m_buffer, cache=T, echo=F, eval=T}
##### first transform back to the the original projection back (this is what OBIS and GBIF expect) #####
sfaa_buf_20000_transform <- st_transform(sfaa_buf_20000, crs = st_crs(x_geo))
sfaa_buf_20000_transform <- st_make_valid(sfaa_buf_20000_transform)
sfaa_buf_20000_transform_wkt <- sf::st_as_text(sf::st_geometry(sfaa_buf_20000_transform))

##### getting OBIS data #####
# using bounding box variables defined above
library(spocc)
out <- occ(geometry = sfaa_buf_20000_transform_wkt, from = 'obis', limit = 10000)
obisbuffer20000 <- as.data.frame(out$obis$data)

## Convert to sf object using coordinates
obisbuffer20000_sf <- st_as_sf(
  obisbuffer20000,
  coords = c("longitude", "latitude"), # specify coordinate columns
  crs = 4326                         # set CRS to WGS84 (EPSG:4326)
)

## check 
# dim(obisbuffer20000)

```

```{r get_gbif_data, cache=T, echo=F, eval=T} 
##### getting gbif data #####
## first get the original projection back (this CRS is what gbif and GBIF expect)
## ID["EPSG",6326]]
sfaa_transform <- st_transform(sfaa, crs = st_crs(x_geo))
sfaa_transform <- st_make_valid(sfaa_transform)
sfaa_transform_wkt <- sf::st_as_text(sf::st_geometry(sfaa_transform))

## using bounding box variables defined above in another chunk
library(spocc)
out <- occ(geometry = sfaa_transform_wkt, from = 'gbif', limit = 10000)
gbif <- as.data.frame(out$gbif$data)


## Convert to sf object using coordinates
gbif_sf <- st_as_sf(
  gbif,
  coords = c("longitude", "latitude"), # specify coordinate columns
  crs = 4326                         # set CRS to WGS84 (EPSG:4326)
)

## check 
# dim(gbif)
# View(gbif)
# names(gbif)
# table(gbif$catalogNumber)
# table(gbif$id)
# table(gbif$aphiaID)
# plot(gbif_sf)

```

```{r get_gbif_data_in_5000m_buffer, cache=T, echo=F, eval=T}
##### first transform back to the the original projection back (this is what gbif and GBIF expect) #####
sfaa_buf_5000_transform <- st_transform(sfaa_buf_5000, crs = st_crs(x_geo))
sfaa_buf_5000_transform <- st_make_valid(sfaa_buf_5000_transform)
sfaa_buf_5000_transform_wkt <- sf::st_as_text(sf::st_geometry(sfaa_buf_5000_transform))

##### getting gbif data #####
# using bounding box variables defined above
library(spocc)
out <- occ(geometry = sfaa_buf_5000_transform_wkt, from = 'gbif', limit = 10000)
gbifbuffer5000 <- as.data.frame(out$gbif$data)

## Convert to sf object using coordinates
gbifbuffer5000_sf <- st_as_sf(
  gbifbuffer5000,
  coords = c("longitude", "latitude"), # specify coordinate columns
  crs = 4326                         # set CRS to WGS84 (EPSG:4326)
)

## check 
# dim(gbifbuffer5000)
# View(gbifbufrer5000)

```

```{r get_gbif_data_in_20000m_buffer, cache=T, echo=F, eval=F}
##### first transform back to the the original projection back (this is what gbif and GBIF expect) #####
sfaa_buf_20000_transform <- st_transform(sfaa_buf_20000, crs = st_crs(x_geo))
sfaa_buf_20000_transform <- st_make_valid(sfaa_buf_20000_transform)
sfaa_buf_20000_transform_wkt <- sf::st_as_text(sf::st_geometry(sfaa_buf_20000_transform))

##### getting gbif data #####
# using bounding box variables defined above
library(spocc)
out <- occ(geometry = sfaa_buf_20000_transform_wkt, from = 'gbif', limit = 10000)
gbifbuffer20000 <- as.data.frame(out$gbif$data)

## Convert to sf object using coordinates
gbifbuffer20000_sf <- st_as_sf(
  gbifbuffer20000,
  coords = c("longitude", "latitude"), # specify coordinate columns
  crs = 4326                         # set CRS to WGS84 (EPSG:4326)
)

## check 
# dim(gbifbuffer20000)

```

```{r echo=F, eval=F}
# export data files to local folder
##### export Excel file of data records to folder #####
setwd("C:/rworking/deepseatools/indata")
write.csv(as.data.frame(points_in_0_select),'DSCRTP_AOI_no_buffer.csv')
write.csv(as.data.frame(points_in_5000_select),'DSCRTP_AOI_5000m_buffer.csv')
write.xlsx(as.data.frame(obis),'OBIS_AOI_no_buffer.xlsx')
write.xlsx(as.data.frame(obisbuffer5000),'OBIS_AOI_5000m_buffer.xlsx')

```

```{r records_with_images, echo=F}
z <- points_in_5000_select %>% filter(is.na(ImageURL) == F)

```

```{r getting_NOAA_bathy, echo=F, eval=F}
st_bbox(sfaa)
st_crs(sfaa)
st_crs(points)
yo<-st_transform(sfaa, st_crs(points))
st_crs(yo)
st_bbox(yo)
st_bbox(obis_sf)
st_bbox(yo)




```

# Purpose of Report

To create summaries of occurrence data to support fishery resource access decision making.  

# Online Resources

The source RMarkdown code used to generate this report is linked here on GitHub: [LINK](https://github.com/RobertMcGuinn/deepseatools/blob/master/code/20241126_AOI_Analysis_Southeast_RPMcGuinn.Rmd)

# Methods

Coral and sponge occurrences from NOAA National Database for Deep-sea Corals and Sponges and other occurrences from the Ocean Biogeographic Information System (OBIS) and the Global Biodiversity Information Facility (GBIF) are summarized within the SFAA Area of Interest (AOI) and in the SFAA AOI + a 5km buffer.  

Version Notes: OBIS and GBIF database last accessed 2024-12-09.  National database version number 20241022-1.  

* AOI (SFAA)
* AOI (SFAA) + 5km buffer

# Maps

```{r plot, fig.width=6, fig.height=4, dpi=300, echo = F}
## plot to see how they relate
ggplot() +
  # Add layers with legend mappings
  geom_sf(data = oculina, aes(fill = "Oculina Banks HAPC"), alpha = 0.3) +
  # geom_sf(data = sfaa_buf_20000, aes(fill = "SFAA-AOI + 20 km buffer"), alpha = 0.3) +
  geom_sf(data = sfaa_buf_5000, aes(fill = "SFAA-AOI + 5 km buffer"), alpha = 0.3) +
  geom_sf(data = points_in_0_select, aes(color = "National Database points within AOI")) +
  geom_sf(data = sfaa, aes(fill = "SFAA-AOI"), alpha = 1) +
  
  # Custom legend labels and colors
  scale_fill_manual(
    name = "AOIs",
    values = c("Oculina Banks HAPC" = "blue", 
               # "SFAA-AOI + 20 km buffer" = "black", 
               "SFAA-AOI + 5 km buffer" = "green",
               "SFAA-AOI" = "red" )
  ) +
  scale_color_manual(
    name = "Points",
    values = c("National Database points within AOI" = "black")
  ) +
  
  # Minimal theme and optional adjustments
  theme_minimal() +
  theme(
    legend.position = "right",                # Place legend on the right
    axis.text.x = element_text(angle = 90)   # Rotate x-axis labels (if applicable)
  )


```

```{r plot_2, fig.width=6, fig.height=4, dpi=300, echo = F, eval=T}

obisbuffer5000_sf_filter <- obisbuffer5000_sf %>% filter(name %in%  c('Oculina varicosa', 'Oculina'))
# table(obis_sf_filter$phylum)
# table(obis_sf_filter$name)


library(ggplot2)

ggplot() +
  # Basemap data layer with fill mapped for legend
  geom_sf(data = sfaa, aes(fill = "SFAA-AOI"), alpha = 1) +

  # Specific plotted data with color mapped for legend
  geom_sf(data = obisbuffer5000_sf,
          aes(color = "OBIS points within AOI + 5 km buffer"),  # Map color to a label
          shape = 3, alpha = .3, size = 3) +
  geom_sf(data = points_in_5000_select,
          aes(color = "National Database points within AOI + 5 km buffer"),  # Map color to a label
          shape = 4, alpha = .3) +
  geom_sf(data = points_in_0_select,
          aes(color = "National Database points within AOI"),  # Updated label
          shape = 4, size = 3) +
  
  # Legend customization
  scale_fill_manual(
    name = "AOI",
    values = c("SFAA-AOI" = "red")
  ) +
  scale_color_manual(
    name = "Points",
    values = c("OBIS points within AOI + 5 km buffer" = "grey",
               "National Database points within AOI + 5 km buffer" = "red",
               "National Database points within AOI" = "black")  # Match updated label
  ) +
  
  # Theme for better readability
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1),  # Rotate x-axis labels
    legend.position = "right"                          # Position legend
  )


```

```{r plot_3, fig.width=6, fig.height=4, dpi=300, echo = F, eval=T}
# Load required libraries
library(ggplot2)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)

# Load Natural Earth country boundaries
countries <- ne_countries(scale = "medium", returnclass = "sf")

# Define the bounding box for Florida and the nearby East Coast
florida_bbox <- c(xmin = -85, xmax = -75, ymin = 24, ymax = 32)  # Adjust these values as needed

# Create the map
ggplot() +
  # Basemap data layer with fill mapped for legend
  geom_sf(data = sfaa, aes(fill = "SFAA-AOI"), alpha = 1) +

  # Specific plotted data with color mapped for legend
  geom_sf(data = obisbuffer5000_sf,
          aes(color = "OBIS points within AOI + 5 km buffer"),  # Map color to a label
          shape = 3, alpha = .3, size = 3) +
  geom_sf(data = points_in_5000_select,
          aes(color = "National Database points within AOI + 5 km buffer"),  # Map color to a label
          shape = 4, alpha = .3) +
  geom_sf(data = points_in_0_select,
          aes(color = "National Database points within AOI"),  # Updated label
          shape = 4, size = 3) +

  # Add country outlines
  geom_sf(data = countries, fill = NA, color = "black", size = 0.5) +

  # Zoom in to Florida and the nearby East Coast using coord_sf
  coord_sf(xlim = c(florida_bbox["xmin"], florida_bbox["xmax"]),
           ylim = c(florida_bbox["ymin"], florida_bbox["ymax"]),
           expand = FALSE) +

  # Legend customization
  scale_fill_manual(
    name = "AOI",
    values = c("SFAA-AOI" = "red")
  ) +
  scale_color_manual(
    name = "Points",
    values = c("OBIS points within AOI + 5 km buffer" = "grey",
               "National Database points within AOI + 5 km buffer" = "red",
               "National Database points within AOI" = "black")  # Match updated label
  ) +

  # Theme for better readability
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1),  # Rotate x-axis labels
    legend.position = "right"                          # Position legend
  )

```

# Results

## Taxa within the AOI (no buffer) from the DSCRTP NOAA National Database for Deep-sea Corals and Sponges

Number of occurrences within the AOI: `r length(points_in_0_select$CatalogNumber)`

### Table of Taxa Grouped by 'VernacularNameCategory' 

``` {r CoralByVernacularNameCategoryTarget, echo=FALSE }
target <- as.data.frame(points_in_0_select)

sum_tbl <-
  target %>%
  # filter(Phylum == "Cnidaria") %>% 
  group_by(VernacularNameCategory) %>%
  summarize(
    Taxa = toString(unique(ScientificName)),
    Records = n()) %>%
  arrange(desc(Records))
sum_tbl <- kable(sum_tbl, row.names = F, digits = 2)
sum_tbl

```

### Dataset Details from the National Database

``` {r Expedition_details_no_buf, echo=FALSE }

target <- as.data.frame(points_in_0_select)

sum_tbl <-
  target %>%
  group_by(DatasetID) %>%
  summarize(
    SurveyIDs = toString(unique(SurveyID)),
    PIs = toString(unique(PI)),
    Observation_Year = toString(unique(ObservationYear)),
    Records = n()) %>%
  arrange(desc(Records))
sum_tbl <- kable(sum_tbl, row.names = F, digits = 2)
sum_tbl

```

## Taxa within the AOI (no buffer) from OBIS

Number of occurrences within the AOI: `r length(obis$datasetName)`

### Table of Taxa Grouped by Order

``` {r TaxaByOrderTargetOBIS, echo=FALSE }

sum_tbl <-
  obis %>%
  group_by(order) %>%
  summarize(
    Taxa = toString(unique(name)),
    Records = n()) %>%
  arrange(desc(Records))
sum_tbl <- kable(sum_tbl, row.names = F, digits = 2)
sum_tbl

```

### Dataset Details (OBIS)

``` {r Expedition_details_obis, echo=FALSE }
sum_tbl <-
  obis %>%
  group_by(dataset_id) %>%
  summarize(
    Records = n()
  ) %>%
  mutate(
    OBIS_Dataset = paste('https://obis.org/dataset/', dataset_id, sep = '') 
  ) %>%
  dplyr::select(OBIS_Dataset,Records) %>% 
  arrange(desc(Records))

sum_tbl <- kable(sum_tbl, row.names = FALSE, digits = 2)
sum_tbl



```

## Taxa within the AOI (no buffer) from GBIF

Number of occurrences within the AOI: `r length(gbif$datasetName)`

### Table of Taxa Grouped by Order

``` {r TaxaByOrderTargetgbif, echo=FALSE }

sum_tbl <-
  gbif %>%
  group_by(order) %>%
  summarize(
    Taxa = toString(unique(name)),
    Records = n()) %>%
  arrange(desc(Records))
sum_tbl <- kable(sum_tbl, row.names = F, digits = 2)
sum_tbl

```

### Dataset Details (gbif)

``` {r Expedition_details_gbif, echo=FALSE }
sum_tbl <-
  gbif %>%
  group_by(datasetKey) %>%
  summarize(
    Records = n()
  ) %>%
  mutate(
    gbif_Dataset = paste('https://gbif.org/dataset/', datasetKey, sep = '') 
  ) %>%
  dplyr::select(gbif_Dataset,Records) %>% 
  arrange(desc(Records))

sum_tbl <- kable(sum_tbl, row.names = FALSE, digits = 2)
sum_tbl



```

## Taxa within the AOI plus 5000m buffer zone from the National Database

Number of occurrences within the AOI+5000 meter buffer: `r length(points_in_5000_select$CatalogNumber)`

### Table of Taxa Grouped by 'VernacularNameCategory' (National Database)

``` {r CoralByVernacularNameCategoryBuffer, echo=FALSE }
buffer <- as.data.frame(points_in_5000_select)
sum_tbl <-
  buffer %>%
  filter(Phylum == "Cnidaria") %>% 
  group_by(VernacularNameCategory) %>%
  summarize(
    Taxa = toString(unique(ScientificName)),
    Records = n()) %>%
  arrange(desc(Records))
sum_tbl <- kable(sum_tbl, row.names = F, digits = 2)
sum_tbl
```

### Dataset Details (National Database)

``` {r Expedition_details_buf_5000, echo=FALSE }
sum_tbl <-
  buffer %>%
  group_by(DatasetID) %>%
  summarize(
    SurveyIDs = toString(unique(SurveyID)),
    PIs = toString(unique(PI)),
    Observation_Year = toString(unique(ObservationYear)),
    Records = n()) %>%
  arrange(desc(Records))
sum_tbl <- kable(sum_tbl, row.names = F, digits = 2)
sum_tbl

```

## Taxa within the AOI plus 5000m buffer zone (OBIS)

Number of occurrences within the AOI: `r length(obisbuffer5000$datasetName)`

### Table of Taxa Grouped by Order

``` {r TaxaByOrderBufferOBIS, echo=FALSE }

sum_tbl <-
  obisbuffer5000 %>%
  group_by(order) %>%
  summarize(
    Taxa = toString(unique(name)),
    Records = n()) %>%
  arrange(desc(Records))
sum_tbl <- kable(sum_tbl, row.names = F, digits = 2, align = )
sum_tbl

```


### Dataset Details (OBIS)

``` {r Expedition_details_obis_buffer, echo=FALSE }
sum_tbl <-
sum_tbl <-
  obisbuffer5000 %>%
  group_by(dataset_id) %>%
  summarize(
    Records = n()
  ) %>%
  mutate(
    OBIS_Dataset = paste('https://obis.org/dataset/', dataset_id, sep = '') 
  ) %>%
  dplyr::select(OBIS_Dataset,Records) %>% 
  arrange(desc(Records))

sum_tbl <- kable(sum_tbl, row.names = FALSE, digits = 2)
sum_tbl


```


















## Taxa within the AOI plus 5000m buffer zone (GBIF)

Number of occurrences within the AOI: `r length(gbifbuffer5000$datasetName)`

### Table of Taxa Grouped by Order

``` {r TaxaByOrderBuffergbif, echo=FALSE }

sum_tbl <-
  gbifbuffer5000 %>%
  group_by(order) %>%
  summarize(
    Taxa = toString(unique(name)),
    Records = n()) %>%
  arrange(desc(Records))
sum_tbl <- kable(sum_tbl, row.names = F, digits = 2, align = )
sum_tbl

```


### Dataset Details (gbif)

``` {r Expedition_details_gbif_buffer, echo=FALSE }
sum_tbl <-
sum_tbl <-
  gbifbuffer5000 %>%
  group_by(datasetKey) %>%
  summarize(
    Records = n()
  ) %>%
  mutate(
    gbif_Dataset = paste('https://gbif.org/dataset/', datasetKey, sep = '') 
  ) %>%
  dplyr::select(gbif_Dataset,Records) %>% 
  arrange(desc(Records))

sum_tbl <- kable(sum_tbl, row.names = FALSE, digits = 2)
sum_tbl


```

















