---
title: "Area of Interest Report - Northeast Canyons and Seamounts National Monument and Hudson Canyon"
author: "Robert P. McGuinn"
date: "`r Sys.Date()`"
output: 
  word_document:
    reference_docx: "c:/rworking/deepseatools/code/template.docx"
---

```{r packages, echo=F, message=FALSE, warning=FALSE, cache=FALSE, eval=T,}
## load packages
#install.packages(tidyverse)
#install.packages('openxlsx')
library(tidyverse)
library(openxlsx)
library(sf)
library(RColorBrewer)
library(raster)
library(leaflet)
library(extrafont)
loadfonts()
library(RColorBrewer)
library(rmarkdown)
library(knitr)
library(maps)
library(raster)
library(httr)
library(jsonlite)
library(rnaturalearth)
library(rnaturalearthdata)
library(openxlsx)
library(googledrive)
library(spocc)
library(terra)
library(knitr)
library(kableExtra)

```

```{r load_NDB, echo=F, message = FALSE, warning = FALSE, cache = TRUE, eval = F}
## load NDB
source('c:/rworking/deepseatools/code/mod_load_current_NDB.R')

```

```{r set_ndb_version, echo=F, eval=T}
## create needed variables
db_version <- unique(filt$DatabaseVersion)

```

```{r setup, include=FALSE, eval=T}
knitr::opts_chunk$set(echo = FALSE, 
                      fig.height = 4.5,
                      fig.width = 7.5
                      )

```

```{r filter_points_by_geography, echo=F, warning=F, message=F, eval=T}
##### filter occurrences by query or box or both ##### 
## set bounding box variables
minlat <- 37
maxlat <- 41
minlon <- -73
maxlon <- -65

## subset data by coordinates
x <- subset(filt, as.numeric(Latitude) > minlat &
                   as.numeric(Latitude) < maxlat &
                   as.numeric(Longitude) > minlon &
                   as.numeric(Longitude) < maxlon)
                
## also subset data by other variables by other criteria
# x <- x %>% filter(ScientificName == "")

## checking
# dim(x)

```

```{r create_geo_from_points, echo=FALSE, echo = FALSE, message = FALSE, warning=FALSE, eval=T}
## create spatial points data frame from a subset of points
x_geo <- x

## create spdf
coordinates(x_geo) <- c("Longitude", "Latitude")
proj4string(x_geo) <- "+proj=longlat +ellps=WGS84 +datum=WGS84"

## OPTIONAL: write out featureclass to geodatabase from sp object 
# fgdb_path <- "C:/data/aprx/aoi/aoi.gdb"
# arc.write(file.path(fgdb_path, 'x_geo'), data=x_geo, overwrite = TRUE)

```

```{r create_sf_from_points, message = FALSE, warning = FALSE, echo=F, eval=T}
##### transform coral and sponge points to sf #####
points <- st_as_sf(x_geo, wkt = "geom")
```

```{r fish_points_for_interacitive map, eval=F}
##### write points for portal ##### 
fish_points_for_app <- 
  points %>% filter(Phylum == 'Chordata') %>% 
  dplyr::select(CatalogNumber,
                Class, 
                Order,
                Family, 
                ScientificName, Vessel, 
                SurveyID, 
                ObservationDate, 
                PI, 
                DatasetID, 
                DataProvider)
st_write(fish_points_for_app, "c:/rworking/deepseatools/indata/fish_points_for_app.shp", overwrite = TRUE)

fish_points_for_app %>% pull(Family) %>% table(useNA = 'always')

```

```{r import_NE_Canyons_aoi, echo=FALSE, echo = F, message = FALSE, warning=FALSE, results = "hide", eval=F}

##### get the NE Canyons mounument ##### 
shapefile_path <- "C:/rworking/deepseatools/indata/20221104_protected_areas_update_HColeman/20221104_protected_areas.shp"
protected_areas <- st_read(shapefile_path)
ne_canyons <- protected_areas %>% filter(Sitename == 'NE Canyons and Seamounts')

##### reproject to target crs ##### 
ne_canyons_5070 <- st_transform(ne_canyons, crs = 5070)

##### Convert sf to SpatVector ##### 
ne_canyons_5070_spat <- vect(ne_canyons_5070)

##### Write to shapefile #####
# writeVector(ne_canyons_5070_spat, "C:/rworking/deepseatools/indata/ne_canyons_5070.shp", overwrite = TRUE)


##### check ##### 
# st_crs(ne_canyons)

```

```{r import_Hudson_bathy, echo=FALSE, echo = F, message = FALSE, warning=FALSE, results = "hide", eval=F }

# Define the GeoTIFF file path
geotiff_path <- "C:/rworking/deepseatools/indata/hc_rb_bath100/hc_rb_bath100.tif"

# Read the raster data
bathy_hudson <- rast(geotiff_path)

# Reproject the raster to the target_crs
bathy_hudson_5070 <- project(bathy_hudson, "EPSG:5070")

##### write the raster (OPTIONAL) ##### 
writeRaster(bathy_hudson_5070, "c:/rworking/deepseatools/indata/bathy_hudson_5070.tif", overwrite = TRUE)


```

```{r import_Hudson_facies, echo=FALSE, echo = F, message = FALSE, warning=FALSE, results = "hide", eval=F }

##### get the Hudson facies ##### 
shapefile_path <- "C:/rworking/deepseatools/indata/hc_facies/hc_facies.shp"
hc_facies <- st_read(shapefile_path)

##### reproject to target crs ##### 
hc_facies_5070 <- st_transform(hc_facies, crs = 5070)

##### check ##### 
# st_crs(hc_facies)
# plot(hc_facies)
# names(hc_facies)
# table(hc_facies$INTERP)

```

```{r make_Hudson_AOI, echo=FALSE, echo = F, message = FALSE, warning=FALSE, results = "hide", eval=F, }

library(sf)

# Get bounding box and convert it to an sf POLYGON
hudson_aoi_bbox <- st_bbox(hc_facies_5070) %>%
  st_as_sfc() %>%
  st_sf(crs = st_crs(hc_facies_5070))

```

```{r import_noaa_bathy, echo=FALSE, echo = F, message = FALSE, warning=FALSE, results = "hide", eval=F}

library(marmap)

# Define bounding box
lon_min <- -68.68
lon_max <- -65.59
lat_min <- 38.776
lat_max <- 40.766

# Get bathymetry data (default resolution: 4 arc-minutes ~7.4 km)
bathy_data <- getNOAA.bathy(lon1 = lon_min, lon2 = lon_max, lat1 = lat_min, lat2 = lat_max, resolution = 1)  # Use resolution = 1 for higher resolution (~1.8 km)

# Convert to raster and plot
bathy_raster <- as.raster(bathy_data)
plot(bathy_raster)

library(terra)

# Convert raster object to SpatRaster
bathy_ne_canyons <- rast(bathy_raster)

## reproject the raster 
bathy_ne_canyons_5070 <- project(bathy_ne_canyons, "EPSG:5070")

##### reclassify these rasters ##### 
# Define reclassification matrix for negative depth values
rcl_matrix <- matrix(c(
  -9999, -3000, 7,
  -3000, -2500, 6,
  -2500, -2000, 5,
  -2000, -1500, 4,
  -1500, -1000, 3,
  -1000, -500, 2,
  -500, 0, 1
), ncol = 3, byrow = TRUE)

# Reclassify the rasters
bathy_ne_canyons_reclassified <- classify(bathy_ne_canyons_5070, rcl_matrix)


# Plot results
# plot(crm1_reclassified)
# plot(crm2_reclassified)

##### write the raster (OPTIONAL) ##### 
writeRaster(bathy_ne_canyons_reclassified, "c:/rworking/deepseatools/indata/bathy_ne_canyons_reclassified.tif", overwrite = TRUE)
# writeRaster(crm2_reclassified, "c:/rworking/deepseatools/indata/crm2_reclassified_ne_canyons.tif", overwrite = TRUE)

##### check ##### 
# freq(bathy_ne_canyons_reclassified)
# pixel_area <- prod(res(bathy_ne_canyons_reclassified))  # Multiply x and y resolution
# pixel_area


```

```{r import_CRM_subset_reclass_combine, echo=FALSE, echo = F, message = FALSE, warning=FALSE, results = "hide", eval=F, }
##### get data from here ##### 
## https://www.ngdc.noaa.gov/thredds/catalog/crm/cudem/catalog.html
##### load CRM data #####
crm1 <- rast("C:/rworking/deepseatools/indata/crm/crm_vol1_2023.nc")
crm2 <- rast("C:/rworking/deepseatools/indata/crm/crm_vol2_2023.nc")
ext(crm1)

##### crop crm1 to NE Canyons and Seamounts large AOI ##### 
## define custom extent: xmin, ymin, xmax, ymax
xmin <- -73
ymin <- 39
xmax <- -65
ymax <- 41

## create a box polygon from the custom extent
bbox <- st_as_sfc(st_bbox(c(xmin = xmin, ymin = ymin, xmax = xmax, ymax = ymax), crs = 4326))  # Use appropriate CRS

## Convert to sf object
bbox_sf <- st_sf(geometry = bbox)

## Export to shapefile for visualization (OPTIONAL)
# st_write(bbox_sf, "c:/rworking/deepseatools/indata/custom_extent_box.shp", delete_layer = TRUE)

## reproject to make sure the vector crs matcheses the raster 
ne_canyons_proj <- st_transform(bbox_sf, crs(crm1))

## Clip the raster using the vector (crop first, then mask)
cropped_raster <- crop(crm1, ne_canyons_proj)
clipped_raster_ne <- mask(cropped_raster, ne_canyons_proj)

##### crop crm2 to Hudson Canyon southern extension ##### 
## define custom extent: xmin, ymin, xmax, ymax
xmin <- -73
ymin <- 37
xmax <- -68
ymax <- 39

## create a box polygon from the custom extent
bbox <- st_as_sfc(st_bbox(c(xmin = xmin, ymin = ymin, xmax = xmax, ymax = ymax), crs = 4326))  # Use appropriate CRS

## Convert to sf object
bbox_sf <- st_sf(geometry = bbox)

## Export to shapefile for visualization (OPTIONAL)
# st_write(bbox_sf, "c:/rworking/deepseatools/indata/custom_extent_box.shp", delete_layer = TRUE)

## reproject to make sure the vector crs matcheses the raster 
hudson_proj <- st_transform(bbox_sf, crs(crm2))

## Clip the raster using the vector (crop first, then mask)
cropped_raster <- crop(crm2, hudson_proj)
clipped_raster_hudson <- mask(cropped_raster, hudson_proj)

##### reclassify these rasters ##### 
library(terra)

# Define reclassification matrix for negative depth values
rcl_matrix <- matrix(c(
  -9999, -3000, 7,
  -3000, -2500, 6,
  -2500, -2000, 5,
  -2000, -1500, 4,
  -1500, -1000, 3,
  -1000, -500, 2,
  -500, 0, 1, 
  0, 1000, -999
), ncol = 3, byrow = TRUE)

# Perform reclassification
crm_reclass_7_ne <- classify(clipped_raster_ne, rcl_matrix)
crm_reclass_7_hudson <- classify(clipped_raster_hudson, rcl_matrix)

##### combine the rasters in to one ##### 
combined_raster <- mosaic(crm_reclass_7_ne, crm_reclass_7_hudson, fun = "max")

##### write the combined raster for visualization ##### 
writeRaster(
  combined_raster,
  filename = "c:/rworking/deepseatools/indata/crm_reclass_7_combined.tif",
  datatype = "INT2S",        # Ensures integer data
  overwrite = TRUE
)

```

```{r import_GEBCO_2024_data_and_reclass, echo=FALSE, echo = F, message = FALSE, warning=FALSE, results = "hide", eval=T, }
## get data from here: https://download.gebco.net/ (aquired 20250404)
## use this bounding box to subset
##  West	East	South	North
# -73	-65	37	41

gebco_2024_bathy <- rast("C:/rworking/deepseatools/indata/GEBCO_04_Apr_2025_1594a0e93d89/gebco_2024_n41.0_s37.0_w-73.0_e-65.0.tif")

# Define reclassification matrix for negative depth values
rcl_matrix <- matrix(c(
  -9999, -3000, 13,
  -3000, -2750, 12,
  -2750, -2500, 11,
  -2500, -2250, 10,
  -2250, -2000, 9,
  -2000, -1750, 8,
  -1750, -1500, 7,
  -1500, -1250, 6,
  -1250, -1000, 5,
  -1000, -750, 4,
  -750, -500, 3,
  -500, -250, 2,
  -250, 0, 1,
  0, 1000, -999
), ncol = 3, byrow = TRUE)

# rcl_matrix <- matrix(c(
#   -9999, -3000, 7,
#   -3000, -2500, 6,
#   -2500, -2000, 5,
#   -2000, -1500, 4,
#   -750, -500, 3,
#   -500, -250, 2,
#   -250, 0, 1, 
#   0, 1000, -999
# ), ncol = 3, byrow = TRUE)

# Perform reclassification
gebco_2024_bathy_reclass <- classify(gebco_2024_bathy, rcl_matrix)

##### write the combined raster for visualization ##### 
writeRaster(
  gebco_2024_bathy_reclass,
  filename = "c:/rworking/deepseatools/indata/gebco_2024_bathy_reclass_250.tif",
  datatype = "INT2S",        # Ensures integer data
  overwrite = TRUE
)

```

```{r buffering, echo=F, eval=F}
##### buffer operations ##### 
## create 5000 meter buffer
ne_canyons_buf_5000 <- st_buffer(ne_canyons_5070, dist=5000) # map units are in meters

##create 20000 meter buffer
ne_canyons_buf_20000 <- st_buffer(ne_canyons_5070, dist=20000) # map units are in meters

## check
# plot(ne_canyons_buf_20000)

```

```{r intersect_points_with_NECSM, echo=F, eval = F}
##### intersection operations ##### 
## transform CRS for coral and sponge points projection to match polygons
points_transform <- st_transform(points, crs = st_crs(ne_canyons_5070))

## checking
# st_crs(points_transform)

## find points within the polygons
points_in_0 <- st_join(points_transform, ne_canyons_5070, join = st_within) # no buffer
points_in_5000 <- st_join(points_transform, ne_canyons_buf_5000, join = st_within) # 5000 meter buffer
points_in_20000 <- st_join(points_transform, ne_canyons_buf_20000, join = st_within) # 20000 meter buffer

## checking
# dim(points_in_0)
# dim(points_in_5000)
# dim(points_in_20000)
# table(points_in_0$Shape_Area, useNA = 'always')
# table(points_in_5000$Shape_Area, useNA = 'always')
# table(points_in_20000$Shape_Area, useNA = 'always')

## select intersecting points only and create separate data frames
points_in_0_select <- points_in_0 %>% filter(is.na(Shape_Area) == F)
points_in_5000_select <- points_in_5000 %>% filter(is.na(Shape_Area) == F)
points_in_20000_select <- points_in_20000 %>% filter(is.na(Shape_Area) == F)

## checking 
# dim(points_in_0_select)
# dim(points_in_5000_select)
# dim(points_in_20000_select)


```

```{r intersect_points_with_Hudson Canyon, echo=F, eval = F}
##### intersection operations ### 
## transform CRS for coral and sponge points projection to match polygons
points_transform_hudson <- st_transform(points, crs = st_crs(hudson_aoi_bbox))

## checking
# st_crs(points_transform)

## find points within the polygons
points_in_0 <- st_join(points_transform_hudson, hudson_aoi_bbox, join = st_within) # no buffer
points_in_5000 <- st_join(points_transform_hudson, hudson_aoi_bbox, join = st_within) # 5000 meter buffer
points_in_20000 <- st_join(points_transform_hudson, hudson_aoi_bbox, join = st_within) # 20000 meter buffer

## checking
# dim(points_in_0)
# dim(points_in_5000)
# dim(points_in_20000)
# table(points_in_0$Shape_Area, useNA = 'always')
# table(points_in_5000$Shape_Area, useNA = 'always')
# table(points_in_20000$Shape_Area, useNA = 'always')

## select intersecting points only and create separate data frames
points_in_0_select_hudson <- points_in_0 %>% filter(is.na(Shape_Area) == F)
points_in_5000_select_hudson <- points_in_5000 %>% filter(is.na(Shape_Area) == F)
points_in_20000_select_hudson <- points_in_20000 %>% filter(is.na(Shape_Area) == F)

## checking 
# dim(points_in_0_select)
# dim(points_in_5000_select)
# dim(points_in_20000_select)


```

```{r intersect reclass_raster_values to points within NECSMM AOI , echo=F, eval=T}
##### Extract raster values at point locations ##### 

## transform projection
points_in_0_select <- st_transform(points_in_0_select, st_crs(points))

## convert sf object to SpatVector for terra
points_in_0_select_vect <- vect(points_in_0_select)

## extract values

extracted <- extract(gebco_2024_bathy_reclass, points_in_0_select_vect)

## add extracted values to the original point data
points_in_0_select$Classes <- extracted[, 2]  # The second column usually holds the raster values

##### check ##### 
# table(points_in_0_select$Classes, useNA = 'always')

## Save the new point file with class info (POTIONAL)
st_write(points_in_0_select, "c:/rworking/deepseatools/indata/points.gpkg", driver = "GPKG", append = TRUE)



```

```{r intersect Region-wide AOI with GEBCO classified raster, echo=F, eval=T}
##### Extract raster values at point locations ##### 
# Convert sf object to SpatVector for terra
points_vect <- vect(points)

# Extract values
extracted <- extract(gebco_2024_bathy_reclass, points_vect)

# Add extracted values to the original point data
points$Classes <- extracted[, 2]  # The second column usually holds the raster values

# Save the new point file with class info
# st_write(points_for_app, "c:/rworking/deepseatools/indata/points_with_classes2.shp", delete_dsn = TRUE)
# class(points_for_app$Classes)

```

```{r get_obis_data, cache=T, echo=F, eval=F} 
##### getting OBIS data #####
## first get the original projection back (this CRS is what OBIS and GBIF expect)
## ID["EPSG",6326]]
ne_canyons_transform <- st_transform(ne_canyons, crs = st_crs(x_geo))
ne_canyons_transform <- st_make_valid(ne_canyons_transform)
ne_canyons_transform_wkt <- sf::st_as_text(sf::st_geometry(ne_canyons_transform))

## using bounding box variables defined above in another chunk
library(spocc)
out <- occ(geometry = ne_canyons_transform_wkt, from = 'obis', limit = 10000)
obis <- as.data.frame(out$obis$data)


## Convert to sf object using coordinates
obis_sf <- st_as_sf(
  obis,
  coords = c("longitude", "latitude"), # specify coordinate columns
  crs = 4326                         # set CRS to WGS84 (EPSG:4326)
)

## check 
# dim(obis)
# View(obis)
# names(obis)
# table(obis$catalogNumber)
# table(obis$id)
# table(obis$aphiaID)
# plot(obis_sf)

```

```{r export_to_local, echo=F, eval=F}
# export data files to local folder
##### export Excel file of data records to folder #####
setwd("C:/rworking/deepseatools/indata")
write.csv(as.data.frame(points_in_0_select),'DSCRTP_AOI_no_buffer.csv')
write.xlsx(as.data.frame(obis),'OBIS_AOI_no_buffer.xlsx')

```

```{r records_with_images, echo=F, eval=F}
z <- points_in_0_select %>% filter(is.na(ImageURL) == F) %>% pull(ImageURL) %>% unique() %>% length()

```

# Purpose of Report

To create summaries of occurrence data to support decision making in the Northeast Canyons and Seamounts Marine National Monument and Hudson Canyon.   

# Resources

[Interactive map](https://experience.arcgis.com/experience/54c42fae110e4cbbac48c505ddfa0676)

[Code for generating this report](https://github.com/RobertMcGuinn/deepseatools/blob/master/code/20250310-0_NE_Canyons_Monument_143822.Rmd) 

[NOAA's National Database for Deep-sea Corals and Sponges](https://www.ncei.noaa.gov/maps/deep-sea-corals-portal/). 
Note: The current report is generated database version `r db_version`. 

[Google Drive](https://drive.google.com/drive/folders/1g6kNh7mY05i8PvUwrSN_z6blKYEiJM5A?usp=drive_link)

[Redmine tracker](https://vlab.noaa.gov/redmine/issues/143822). For internal use.

# Analysis Notes

Coral, sponge, and fish occurrences from NOAA's National Database for Deep-sea Corals and Sponges are summarized within the Northeast Canyons and Seamounts National Monument and Hudson Canyon. The database version used in this report is `r unique(filt$DatabaseVersion)`.

## Three separate geographic areas of interest (AOI) are defined for the purposes of this report.  

1. Region-wide AOI (minlat = 37,maxlat = 41, minlon = -73, maxlon = -65)
2. NE Canyons and Seamounts NM AOI (official boundary)
3. Hudson Canyon AOI (unofficial boundary)

# Results
## Region-wide AOI

``` {r depth_bin_figure_loop, echo=FALSE, eval=F}
##### Loop to make boxplots in each depth bin #####
# Set color choices
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### Depth Bin Boxplots #####

for(i in 1:(length(depth_bins)-1)) {
  
  mindepth <- depth_bins[i]
  maxdepth <- depth_bins[i+1]
  
  bin_data <- points %>% 
    filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)
  
  if(nrow(bin_data) > 0) {
    
    # Boxplot
    g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                              DepthInMeters,
                              fill = VernacularNameCategory)) +
      geom_boxplot() +
      scale_y_reverse() +
      ylab("Depth (meters)") +
      xlab("Vernacular Name Category") +
      ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
      theme_bw(base_size = 13, base_family = "sans") +
      theme(
        axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
        axis.title.y = element_text(margin = margin(r = 30))
      ) +
      scale_fill_manual(values = my_colors) +
      labs(fill = "Vernacular Name Category")
    
    print(g)  # Plot shows in Word
  }
}



```
\newpage
#### Figure and Table 1: 0-250m
``` {r depth_bin_table_1, echo=FALSE, eval=T}
##### set depth bins #####
i <- 0
mindepth <- i
maxdepth <- i+250
bin_data <- points %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####

bin_data <- data.frame(points) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl



```

\newpage
#### Figure and Table 2: 250-500m
``` {r depth_bin_table_2, echo=FALSE, eval=T}
##### set depth bins #####
i <- 250
mindepth <- i
maxdepth <- i+250
bin_data <- points %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####

bin_data <- data.frame(points) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl



```
\newpage
#### Figure and Table 3: 500-750m
```{r depth_bin_table_3, echo=FALSE, eval=T}
##### set depth bins #####
i <- 500
mindepth <- i
maxdepth <- i+250
bin_data <- points %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####

bin_data <- data.frame(points) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl

```
\newpage
#### Figure and Table 4: 750-1000m
```{r depth_bin_table_4, echo=FALSE, eval=T}
##### set depth bins #####
i <- 750
mindepth <- i
maxdepth <- i+250
bin_data <- points %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####

bin_data <- data.frame(points) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl

```
\newpage
#### Figure and Table 5: 1000-1250m
```{r depth_bin_table_5, echo=FALSE, eval=T}
##### set depth bins #####
i <- 1000
mindepth <- i
maxdepth <- i+250
bin_data <- points %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####

bin_data <- data.frame(points) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl

```
\newpage
#### Figure and Table 6: 1250-1500m
```{r depth_bin_table_6_echo=FALSE, eval=T}
##### set depth bins #####
i <- 1250
mindepth <- i
maxdepth <- i+250
bin_data <- points %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####

bin_data <- data.frame(points) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl

```
\newpage
#### Figure and Table 7: 1500-1750m
```{r depth_bin_table_7, echo=FALSE, eval=T}
##### set depth bins #####
i <- 1500
mindepth <- i
maxdepth <- i+250
bin_data <- points %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####

bin_data <- data.frame(points) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl

```
\newpage
#### Figure and Table 8: 1750-2000m
```{r depth_bin_table_8, echo=FALSE, eval=T}
##### set depth bins #####
i <- 1750
mindepth <- i
maxdepth <- i+250
bin_data <- points %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####

bin_data <- data.frame(points) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl

```
\newpage
#### Figure and Table 9: 2000-2250m
```{r depth_bin_table_9, echo=FALSE, eval=T}
##### set depth bins #####
i <- 2000
mindepth <- i
maxdepth <- i+250
bin_data <- points %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####

bin_data <- data.frame(points) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl

```
\newpage
#### Figure and Table 10: 2250-2500m
```{r depth_bin_table_10, echo=FALSE, eval=T}
##### set depth bins #####
i <- 2250
mindepth <- i
maxdepth <- i+250
bin_data <- points %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####

bin_data <- data.frame(points) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl

```
\newpage
#### Figure and Table 11: 2500-2750m
```{r depth_bin_table_11, echo=FALSE, eval=T}
##### set depth bins #####
i <- 2500
mindepth <- i
maxdepth <- i+250
bin_data <- points %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####

bin_data <- data.frame(points) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl

```
\newpage
#### Figure and Table 12: 2750-3000m
```{r depth_bin_table_12, echo=FALSE, eval=T}
##### set depth bins #####
i <- 2750
mindepth <- i
maxdepth <- i+250
bin_data <- points %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####

bin_data <- data.frame(points) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl

```
\newpage
#### Figure and Table 13: 3000-3250m
```{r depth_bin_table_13, echo=FALSE, eval=T}
##### set depth bins #####
i <- 3000
mindepth <- i
maxdepth <- i+250
bin_data <- points %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####

bin_data <- data.frame(points) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl

```
\newpage
#### Figure and Table 14: 3250-3500m
```{r depth_bin_table_14, echo=FALSE, eval=T}
##### set depth bins #####
i <- 3250
mindepth <- i
maxdepth <- i+250
bin_data <- points %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####

bin_data <- data.frame(points) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl

```
\newpage
#### Figure and Table 15: 3500-3750m
```{r depth_bin_table_15, echo=FALSE, eval=T}
##### set depth bins #####
i <- 3500
mindepth <- i
maxdepth <- i+250
bin_data <- points %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####

bin_data <- data.frame(points) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl

```
\newpage
#### Figure and Table 16: 3750-4500m
```{r depth_bin_table_16, echo=FALSE, eval=T}
##### set depth bins #####
i <- 3750
mindepth <- i
maxdepth <- i+750
bin_data <- points %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####
bin_data <- data.frame(points) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl

```


\newpage
## NE Canyons and Seamounts NM AOI

#### Figure and Table 1: 0-250m
``` {r depth_bin_table__1, echo=FALSE, eval=T}
##### set depth bins #####
i <- 0
mindepth <- i
maxdepth <- i+250
bin_data <- points_in_0_select %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####

bin_data <- data.frame(points_in_0_select) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl



```

\newpage
#### Figure and Table 2: 250-500m
``` {r depth_bin_table__2, echo=FALSE, eval=T}
##### set depth bins #####
i <- 250
mindepth <- i
maxdepth <- i+250
bin_data <- points_in_0_select %>%  
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####

bin_data <- data.frame(points_in_0_select) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl



```
\newpage
#### Figure and Table 3: 500-750m
```{r depth_bin_table__3, echo=FALSE, eval=T}
##### set depth bins #####
i <- 500
mindepth <- i
maxdepth <- i+250
bin_data <- points_in_0_select %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####

bin_data <- data.frame(points_in_0_select) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl

```
\newpage
#### Figure and Table 4: 750-1000m
```{r depth_bin_table__4, echo=FALSE, eval=T}
##### set depth bins #####
i <- 750
mindepth <- i
maxdepth <- i+250
bin_data <- points_in_0_select %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####

bin_data <- data.frame(points_in_0_select) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl

```
\newpage
#### Figure and Table 5: 1000-1250m
```{r depth_bin_table__5, echo=FALSE, eval=T}
##### set depth bins #####
i <- 1000
mindepth <- i
maxdepth <- i+250
bin_data <- points_in_0_select %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####

bin_data <- data.frame(points_in_0_select) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl

```
\newpage
#### Figure and Table 6: 1250-1500m
```{r depth_bin_table__6_echo=FALSE, eval=T}
##### set depth bins #####
i <- 1250
mindepth <- i
maxdepth <- i+250
bin_data <- points_in_0_select %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####

bin_data <- data.frame(points_in_0_select) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl

```
\newpage
#### Figure and Table 7: 1500-1750m
```{r depth_bin_table__7, echo=FALSE, eval=T}
##### set depth bins #####
i <- 1500
mindepth <- i
maxdepth <- i+250
bin_data <- points_in_0_select %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####

bin_data <- data.frame(points_in_0_select) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl

```
\newpage
#### Figure and Table 8: 1750-2000m
```{r depth_bin_table__8, echo=FALSE, eval=T}
##### set depth bins #####
i <- 1750
mindepth <- i
maxdepth <- i+250
bin_data <- points_in_0_select %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####

bin_data <- data.frame(points_in_0_select) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl

```
\newpage
#### Figure and Table 9: 2000-2250m
```{r depth_bin_table__9, echo=FALSE, eval=T}
##### set depth bins #####
i <- 2000
mindepth <- i
maxdepth <- i+250
bin_data <- points_in_0_select %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####

bin_data <- data.frame(points_in_0_select) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl

```
\newpage
#### Figure and Table 10: 2250-2500m
```{r depth_bin_table__10, echo=FALSE, eval=T}
##### set depth bins #####
i <- 2250
mindepth <- i
maxdepth <- i+250
bin_data <- points_in_0_select %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####

bin_data <- data.frame(points_in_0_select) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl

```
\newpage
#### Figure and Table 11: 2500-2750m
```{r depth_bin_table__11, echo=FALSE, eval=T}
##### set depth bins #####
i <- 2500
mindepth <- i
maxdepth <- i+250
bin_data <- points_in_0_select %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####

bin_data <- data.frame(points_in_0_select) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl

```
\newpage
#### Figure and Table 12: 2750-3000m
```{r depth_bin_table__12, echo=FALSE, eval=T}
##### set depth bins #####
i <- 2750
mindepth <- i
maxdepth <- i+250
bin_data <- points_in_0_select %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####

bin_data <- data.frame(points_in_0_select) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl

```
\newpage
#### Figure and Table 13: 3000-3250m
```{r depth_bin_table__13, echo=FALSE, eval=T}
##### set depth bins #####
i <- 3000
mindepth <- i
maxdepth <- i+250
bin_data <- points_in_0_select %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####

bin_data <- data.frame(points_in_0_select) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl

```
\newpage
#### Figure and Table 14: 3250-3500m
```{r depth_bin_table__14, echo=FALSE, eval=T}
##### set depth bins #####
i <- 3250
mindepth <- i
maxdepth <- i+250
bin_data <- points_in_0_select %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####

bin_data <- data.frame(points_in_0_select) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl

```
\newpage
#### Figure and Table 15: 3500-3750m
```{r depth_bin_table__15, echo=FALSE, eval=T}
##### set depth bins #####
i <- 3500
mindepth <- i
maxdepth <- i+250
bin_data <- points_in_0_select %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####

bin_data <- data.frame(points_in_0_select) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl

```
\newpage
#### Figure and Table 16: 3750-4500m
```{r depth_bin_table__16, echo=FALSE, eval=T}
##### set depth bins #####
i <- 3750
mindepth <- i
maxdepth <- i+750
bin_data <- points_in_0_select %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####
bin_data <- data.frame(points_in_0_select) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl

```

\newpage
## Hudson Canyon AOI

#### Figure and Table 1: 0-250m
``` {r depth_bin_table__1, echo=FALSE, eval=T}
##### set depth bins #####
i <- 0
mindepth <- i
maxdepth <- i+250
bin_data <- points_in_0_select %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####

bin_data <- data.frame(points_in_0_select) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl



```

\newpage
#### Figure and Table 2: 250-500m
``` {r depth_bin_table__2, echo=FALSE, eval=T}
##### set depth bins #####
i <- 250
mindepth <- i
maxdepth <- i+250
bin_data <- points_in_0_select %>%  
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####

bin_data <- data.frame(points_in_0_select) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl



```
\newpage
#### Figure and Table 3: 500-750m
```{r depth_bin_table__3, echo=FALSE, eval=T}
##### set depth bins #####
i <- 500
mindepth <- i
maxdepth <- i+250
bin_data <- points_in_0_select %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####

bin_data <- data.frame(points_in_0_select) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl

```
\newpage
#### Figure and Table 4: 750-1000m
```{r depth_bin_table__4, echo=FALSE, eval=T}
##### set depth bins #####
i <- 750
mindepth <- i
maxdepth <- i+250
bin_data <- points_in_0_select %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####

bin_data <- data.frame(points_in_0_select) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl

```
\newpage
#### Figure and Table 5: 1000-1250m
```{r depth_bin_table__5, echo=FALSE, eval=T}
##### set depth bins #####
i <- 1000
mindepth <- i
maxdepth <- i+250
bin_data <- points_in_0_select %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####

bin_data <- data.frame(points_in_0_select) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl

```
\newpage
#### Figure and Table 6: 1250-1500m
```{r depth_bin_table__6_echo=FALSE, eval=T}
##### set depth bins #####
i <- 1250
mindepth <- i
maxdepth <- i+250
bin_data <- points_in_0_select %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####

bin_data <- data.frame(points_in_0_select) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl

```
\newpage
#### Figure and Table 7: 1500-1750m
```{r depth_bin_table__7, echo=FALSE, eval=T}
##### set depth bins #####
i <- 1500
mindepth <- i
maxdepth <- i+250
bin_data <- points_in_0_select %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####

bin_data <- data.frame(points_in_0_select) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl

```
\newpage
#### Figure and Table 8: 1750-2000m
```{r depth_bin_table__8, echo=FALSE, eval=T}
##### set depth bins #####
i <- 1750
mindepth <- i
maxdepth <- i+250
bin_data <- points_in_0_select %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####

bin_data <- data.frame(points_in_0_select) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl

```
\newpage
#### Figure and Table 9: 2000-2250m
```{r depth_bin_table__9, echo=FALSE, eval=T}
##### set depth bins #####
i <- 2000
mindepth <- i
maxdepth <- i+250
bin_data <- points_in_0_select %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####

bin_data <- data.frame(points_in_0_select) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl

```
\newpage
#### Figure and Table 10: 2250-2500m
```{r depth_bin_table__10, echo=FALSE, eval=T}
##### set depth bins #####
i <- 2250
mindepth <- i
maxdepth <- i+250
bin_data <- points_in_0_select %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####

bin_data <- data.frame(points_in_0_select) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl

```
\newpage
#### Figure and Table 11: 2500-2750m
```{r depth_bin_table__11, echo=FALSE, eval=T}
##### set depth bins #####
i <- 2500
mindepth <- i
maxdepth <- i+250
bin_data <- points_in_0_select %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####

bin_data <- data.frame(points_in_0_select) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl

```
\newpage
#### Figure and Table 12: 2750-3000m
```{r depth_bin_table__12, echo=FALSE, eval=T}
##### set depth bins #####
i <- 2750
mindepth <- i
maxdepth <- i+250
bin_data <- points_in_0_select %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####

bin_data <- data.frame(points_in_0_select) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl

```
\newpage
#### Figure and Table 13: 3000-3250m
```{r depth_bin_table__13, echo=FALSE, eval=T}
##### set depth bins #####
i <- 3000
mindepth <- i
maxdepth <- i+250
bin_data <- points_in_0_select %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####

bin_data <- data.frame(points_in_0_select) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl

```
\newpage
#### Figure and Table 14: 3250-3500m
```{r depth_bin_table__14, echo=FALSE, eval=T}
##### set depth bins #####
i <- 3250
mindepth <- i
maxdepth <- i+250
bin_data <- points_in_0_select %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####

bin_data <- data.frame(points_in_0_select) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl

```
\newpage
#### Figure and Table 15: 3500-3750m
```{r depth_bin_table__15, echo=FALSE, eval=T}
##### set depth bins #####
i <- 3500
mindepth <- i
maxdepth <- i+250
bin_data <- points_in_0_select %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####

bin_data <- data.frame(points_in_0_select) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl

```
\newpage
#### Figure and Table 16: 3750-4500m
```{r depth_bin_table__16, echo=FALSE, eval=T}
##### set depth bins #####
i <- 3750
mindepth <- i
maxdepth <- i+750
bin_data <- points_in_0_select %>% 
filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

##### set color choices ##### 
my_colors <- c(
  "stony coral (branching)" =  "#FFFFFF",
  "stony coral (cup coral)" =  "#00E6A9",
  "black coral" =  "#000000",
  "gorgonian coral" = "#FF0000",
  "soft coral" =  "#00734C",
  "sea pen" = "#0000FF",
  "stoloniferan coral" = "#7F7F7F",
  "lace coral" = "#A80084",
  "lithotelestid coral" = "#FF00C5",
  "stony coral (unspecified)" = "#D1FF73",
  "gold coral" =  "#FFFF00",
  "sponge (unspecified)" = "#FF0000",
  "demosponge" = "#00FFC5",
  "glass sponge" = "#C500FF",
  "homoscleromorph sponge" = "#FFFF00",
  "calcareous sponge" = "#55FF00",
  "other coral-like hydrozoan" = "#00FFFF",
  "alcyonacean (unspecified)" = "#D7B09E",
  "fish" = "#7a2d0a"
)

##### make a boxplot of VernacularNameCategory within depth bin ##### 
g <- ggplot(bin_data, aes(reorder(VernacularNameCategory, DepthInMeters, FUN = median),
                          DepthInMeters,
                          fill = VernacularNameCategory)) +
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") +
  xlab("") +
  ggtitle(sprintf("Depth Distribution by Category (%d-%d m)", mindepth, maxdepth)) +
  theme_bw(base_size = 13, base_family = "sans") +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1, face = 'italic'),
    axis.title.y = element_text(margin = margin(r = 30))
  ) +
  scale_fill_manual(values = my_colors) +
  labs(fill = "Vernacular Name Category")

print(g)  # Plot shows in Word

##### Depth Bin Summary Tables #####
bin_data <- data.frame(points_in_0_select) %>% 
  filter(DepthInMeters > mindepth, DepthInMeters <= maxdepth)

# Summary Table
library(flextable)

tab <- bin_data %>%
  group_by(VernacularNameCategory) %>%
  summarise(
    N = n(),
    MinDepth = min(DepthInMeters, na.rm = TRUE),
    MedianDepth = median(DepthInMeters, na.rm = TRUE),
    MaxDepth = max(DepthInMeters, na.rm = TRUE),
    ScientificNames = paste(unique(ScientificName), collapse = " / "),
    .groups = "drop"
  ) %>%
  arrange(MedianDepth)

# Create the flextable
sum_tbl <- flextable(tab) %>%
  fontsize(size = 6) %>%  # Set the font size for content
  compose(j = "N", value = as_paragraph(as_chunk(N))) %>%  # Compose column N
  set_table_properties(layout = "autofit") %>%
  bold(part = "header") %>%  # Optionally, make the header bold
  fontsize(part = "header", size = 7)  # Set a smaller font size for the header

sum_tbl

```



