---
title: "Figure on RecordType"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, echo=F, warning=F, message=F}
#install.packages(tidyverse)
#install.packages('openxlsx')
library(tidyverse)
library(openxlsx)
library(sf)
library(sp)
library(rgdal) # install.packages("rgdal")
library(RColorBrewer)
library(leaflet)# install.packages("leaflet")
library(extrafont)
library(RColorBrewer)
library(rmarkdown)
library(knitr)
library(rgdal)
library(httr)
library(jsonlite)
library(rnaturalearth)# install.packages("rnaturalearth")
library(rnaturalearthdata)# install.packages("rnaturalearthdata")
library(openxlsx)
library(terra) # install.packages("terra")

# install.packages('devtools',dependencies = T)
library(devtools)

```

# Load NDB

```{r load_NDB, eval=T, echo=F, message=FALSE, warning=FALSE, cache=TRUE}
##### input: latest version of NDB #####
setwd("C:/rworking/deepseatools/indata")
indata<-read.csv("DSCRTP_NatDB_20211110-0.csv", header = T)
filt <- indata %>%
  filter(Flag == "0")

```


# Growth of Database Through Time within this Selection

``` {r GrowthThroughTime_charts, echo=FALSE, dpi=300, eval = T}

x <- filt

## set options to remove scientific notation
options(scipen=10000, digits = 0)


subx <- x %>% filter (Phylum == "Cnidaria") %>%
  mutate(EntryDate = replace_na(as.character(EntryDate), '2019-12-18')) %>% 
  mutate(EntryDate = as.factor(EntryDate)) %>% 
  mutate(as.Date(EntryDate))

suby <- x %>% filter (Phylum == "Porifera") %>%
  mutate(EntryDate = replace_na(as.character(EntryDate), '2019-12-18')) %>% 
  mutate(EntryDate = as.factor(EntryDate)) %>% 
  mutate(as.Date(EntryDate))
 
ggplot(subx, aes(x=as.Date(EntryDate), color=Phylum)) +
  stat_bin(data = subx,
           aes(y=cumsum(..count..) / 1000), 
           geom="line", 
           binwidth = 2, 
           size = 1) +
  stat_bin(data = suby,
           aes(y=cumsum(..count..) / 1000),
           geom="line", 
           binwidth = 2, 
           size = 1) +
  ylab("Number of Records (thousands)") + 
  xlab("Year Record Added") + 
  xlim(as.Date('2012-01-01'), as.Date('2020-10-19')) +
  theme(axis.text.x=element_text(hjust= .2)) +
  guides(fill=FALSE) + 
  theme_bw() +
  theme(text=element_text(size=20,  family="Cambria")) +
  #theme(axis.text.x = element_text(angle = 90)) +
  theme(legend.title = element_blank()) +
  theme(axis.title.x = element_text(vjust=-2.5)) +
  theme(axis.title.y = element_text(vjust= 5)) +
  theme(plot.margin = unit(c(1,1,1,1), "cm"))+
  scale_color_manual(values = brewer.pal(12, "Paired")[c(10,9)]) +
  scale_y_continuous(label=scales::comma) +
  scale_y_continuous(labels = scales::number_format(accuracy = 1))+
  geom_vline(xintercept = as.Date("2016-01-01"), color = 'blue')# +
## ggtitle(paste("Database Version:", unique(filt$DatabaseVersion)))

## save the plot
# ggsave(paste("c:/rworking/deepseatools/images/", 
#              "20220130-1_SouthAtlantic_subset", 
#              "NatDB_", 
