---
title: "DSBS16_HI_Analysis"
author: "Robert McGuinn"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true # table of content true
    toc_depth: 3  # upto three depths of headings (specified by #, ## and ###)
    toc_float: true
    number_sections: true  ## if you want number sections at each table header
    theme: lumen  # many options for theme, this one is my favorite.
    highlight: tango  # specifies the syntax highlighting style
   #  css: my.css   # you can add your custom css, should be in same folder
    cold_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, 
                      eval = T, 
                      warning = F,
                      message = F,
                      fig.height = 7 , 
                      fig.width = 8, 
                      dpi = 300,
                      cache = F)
```

```{r packages}
library(tidyr)
library(openxlsx)
library(googlesheets4)
library(googledrive)
library(flextable)
library(dplyr)
library(flextable)
library(vegan)
library(ggplot2)
library(reshape2)
library(extrafont)
library(leaflet)

```

## Purpose of Analysis

To visualize and understand the community structure of high density coral and sponge communities within the Hawaiian Islands.  

## Resources

The main project folder for this analysis is on Google Drive: <a href="https://drive.google.com/drive/folders/1ToTK_v7NF7ZPdwXQwD1hzCwuVZCZRH_d" target="_new">LINK</a>



The R Markdown code used to generate this report is available here: <a href="https://github.com/RobertMcGuinn/deepseatools/blob/master/code/DSBS16_HI_Analysis.Rmd" target="_new">LINK</a>

Comments and code within the R Markdown document further elucidate the procedures being used. Having this code available will facilitate the reanalysis of these data. 

## Input Data

Datasets used in this analysis were acquired on 2021-08-18 from Tom Hourigan via Excel file. The coral and sponge species occurrence records are from a subset of dives in the Hawaiian region (including Johnston Atoll and the Musician Seamounts). This original dataset is shared on Google Drive here: <a href="https://docs.google.com/spreadsheets/d/16izQ_vWsSDgMupbNAEUsVTvMISzmrLaC" target="_new">LINK</a>

Tom Hourigan issued some corrections, via Excel file, to the above dataset on 2021-08-26: <a href="https://docs.google.com/spreadsheets/d/1Rg4fUrvO_0VT2TVNNANPKGbMyxg-jh2p/edit?usp=sharing&ouid=109414727136135095326&rtpof=true&sd=true" target="_new">LINK</a>


```{r main_data_intake, cache=F, eval = T}
##### bringing in dataframe from xlsx stored on drive ####
## creates dataframe called 'sub'
## file path [MANUAL]
filename <- "20210817_Hawaii-HiDensity1500-2500m"

## create a list of files (or single file) that meets title query [MANUAL]
x <- drive_find(q = "name contains '20210817_Hawaii-HiDensity1500-2500m'")

## browse to it
# x %>% drive_browse()

# getting the id as a character string
y <- x$id

# this downloads the file to the specified path
dl <- drive_download(as_id(y),
                     path = paste("C:/rworking/deepseatools/indata/",
                                  filename,
                                  ".xlsx", 
                                  sep = ""),
                     overwrite = TRUE)

## extract just the file of interest from the zip file to a dataframe called 'sub'
sub <- read.xlsx(dl$local_path)

## names 
# names(sub)
```

```{r change_file_intake, cache=T, eval = T}
##### bringing in dataframe from xlsx stored on drive ####
# changes received from THourigan on 20210826
## creates a dataframe called 'change'

## file path [MANUAL]
filename <- "20210826_0_Changes_THourigan"

## create a list of files (or single file) that meets title query [MANUAL]
x <- drive_find(q = "name contains '20210826_0_Changes_THourigan'")

## browse to it
# x %>% drive_browse()

# getting the id as a character string
y <- x$id

# this downloads the file to the specified path
dl <- drive_download(as_id(y),
                     path = paste("C:/rworking/deepseatools/indata/",
                                  filename,
                                  ".xlsx", 
                                  sep = ""),
                     overwrite = TRUE)

## extract just the file of interest from the zip file to a dataframe called 'sub'
change <- read.xlsx(dl$local_path)

## names 
# names(change)

```

```{r data_scrub}
sub <- sub %>% rename(n_km = `#/km`,
                    m_divetrack = `Dive.Track.(m)`,
                    kennedy_class = Kennedy.Class)

sub = subset(sub, select = -c(IdentificationComments,VerbatimSize))


change <- change %>% rename(n_km = `#/km`,
                    m_divetrack = `Dive.Track.(m)`,
                    kennedy_class = Kennedy.Class)

change = subset(change, select = -c(Flag))

```

```{r merging changes}

# this merges the two dataframes by CatalogNumber and then adds changes to the original set where needed and leaves them alone where not needed.
test <- left_join(sub, change, 
                  by = c("CatalogNumber")) %>% 
  mutate(ScientificName.x = ifelse(is.na(ScientificName.y), 
                                   ScientificName.x, 
                                   ScientificName.y)) %>% 
  mutate(VernacularNameCategory.x = ifelse(is.na(VernacularNameCategory.y), 
                                           VernacularNameCategory.x, 
                                           VernacularNameCategory.y)) %>% 
  mutate(TaxonRank.x = ifelse(is.na(TaxonRank.y), 
                              TaxonRank.x, 
                              TaxonRank.y)) %>% 
  mutate(Phylum.x = ifelse(is.na(Phylum.y), 
                           Phylum.x, 
                           Phylum.y)) %>% 
  mutate(Class.x = ifelse(is.na(Class.y), 
                          Class.x, 
                          Class.y)) %>% 
  mutate(Subclass.x = ifelse(is.na(Subclass.y), 
                             Subclass.x, 
                             Subclass.y)) %>% 
  mutate(Order.x = ifelse(is.na(Order.y), 
                          Order.x, 
                          Order.y)) %>% 
  mutate(Suborder.x = ifelse(is.na(Suborder.y), 
                             Suborder.x, 
                             Suborder.y)) %>% 
  mutate(Family.x = ifelse(is.na(Family.y), 
                           Family.x, 
                           Family.y)) %>% 
  mutate(Genus.x = ifelse(is.na(Genus.y), 
                          Genus.x, 
                          Genus.y))


## checking
# test %>% filter(CatalogNumber == "985165") %>% # without "940589" with change"985165"
#   group_by(ScientificName.x, ScientificName.y) %>%
#   summarize(n=n()) %>%
#   View()

## Now get rid of all columns with ".y" in the name since we already harvested the changes and put them the in *.x variables.
test2 <- select(test, -contains(".y"))

# Now get rid of the *.x suffix from all variable names to get back to the corrected file.  
names(test2) <- gsub(".x$","", names(test2))

# Now overwrite 'sub' to continue on down the road 
sub <- test2

## clean up
rm(test)
rm(test2)


```

```{r filter_to_top_ten_2, echo=F, eval=T, warning=F, message=F}
## Data preparation
### Filter to only species appearing in the top ten list  

##### filter d down to just the species appearing in the top ten ##### 
topten <- sub %>%
 group_by(EventID, ScientificName)%>%
 summarize(individual_count_sum = sum(IndividualCount))%>%
 top_n(10, individual_count_sum) %>% 
  arrange(EventID, desc(individual_count_sum))

z <- sub %>% filter(
  ScientificName %in% topten$ScientificName 
  #Phylum == "Porifera"
# EventID %in% c("NA101-H1715","NA101-H1720", "D2-EX1504L2-17")
  )

```

```{r setdepthclass_2, echo=FALSE, cache = FALSE} 
### Setting up depth, temp, and location categories

##### setting depth categories ##### 
# qplot(as.numeric(z$DepthInMeters))
z$DepthCat[as.numeric(z$DepthInMeters) <= 1800] <- "shallower"
z$DepthCat[as.numeric(z$DepthInMeters) > 1800 & 
             as.numeric(z$DepthInMeters) < 2400] <- "medium"
z$DepthCat[as.numeric(z$DepthInMeters) >= 2400] <- "deeper"
z$DepthCat <- factor(z$DepthCat, levels = c("shallower","medium","deeper"))

##### setting location categories ##### 
# qplot(as.numeric(z$Latitude))
z$loc[as.numeric(z$Latitude) >= 20] <- "north"
z$loc[as.numeric(z$Latitude) < 20] <- "south"

##### longitude cat #####
z$longloc[as.numeric(z$Longitude) > -165] <- "east"
z$longloc[as.numeric(z$Longitude) < -165] <- "west"

##### setting temperature categories #####
# qplot(as.numeric(z$Temperature))
z$temp_cat[as.numeric(z$Temperature) > 2.2] <- "warmer"
z$temp_cat[as.numeric(z$Temperature) >= 1.9 & as.numeric(z$Temperature) <= 2.2] <- "cold"
z$temp_cat[as.numeric(z$Temperature) < 1.9] <- "colder"
z$temp_cat <- factor(z$temp_cat, levels = c("warmer", "cold", "colder"))

##### setting oxygen categories #####
# qplot(as.numeric(z$Oxygen))
z$ox_cat[as.numeric(z$Oxygen) >= 2.6] <- "rich"
z$ox_cat[as.numeric(z$Oxygen) < 2.6] <- "poor"

##### filter if necessary ##### 
z <- z %>% 
  filter(
    # temp_cat == "colder"
    # DepthCat == 'deeper'
    # EventID != "D2-EX1605L1-13"
    # Phylum == "Cnidaria"
    #  DensityCategory == ">10000/km" |
    # DensityCategory == "3000-9999/km"
  )

```

``` {r making_matrix, echo=FALSE, cache = FALSE, dpi=300, warning=FALSE, message=FALSE, results='hide'}
### Making the matrix

# First we adjust the abundances by the length of the drive track. 
# This creates a normalized value for 'IndividualCount'. 
# It is calculated as 'IndividualCount' per meter. 
# This value is then entered into the site by species matrix which is used in the NMDS and clustering algorithms.  

##### bring z from above into this function ##### 
x <- z 

## checking 
# length(table(sub$n_km, useNA = 'always'))

## adjust the IndividualCount values 
x$IndividualCount_z <- x$IndividualCount/x$m_divetrack

## create site X species matrix 
library(vegan)
site.sp <- reshape2::dcast(x, EventID ~ ScientificName, value.var = "IndividualCount_z", fun.aggregate = sum)

## set all NA to 0 because the vegan package can't handle the NA
site.sp[is.na(site.sp)] = 0

## checking resulting table
# View(site.sp)

## export to check
# setwd("C:/rworking/deepseatools/indata")
# write.csv(site.sp, 'site.sp.csv')

# creating a site variable
site.sp$site <- site.sp$EventID

# getting rid of non-needed variables
site.sp <- site.sp %>%
  dplyr::select(-EventID)

# # moving the site variable to the beginning
# col_idx <- grep("site", names(site.sp))
# site.sp <- site.sp[, c(col_idx, (1:ncol(site.sp))[-col_idx])]
# # names(site.sp)

# set the site as the row.names
row.names(site.sp) <- site.sp$site

# remove the site variable
site.sp <- site.sp %>%
  dplyr::select(-site)

# making it a matrix
site.sp <- as.matrix(site.sp)

## checking 
# NMDS

## create relative abundances
#site.sp <- decostand(site.sp, method = "total")

``` 

## Tabular Summaries

```{r variables}
n_records <- length(sub$CatalogNumber)

list_imageurls <- sub %>%  filter(is.na(ImageURL)==F) %>% 
  pull(ImageURL)

n_images <- length(unique(list_imageurls))

txt_names <- names(sub)

n_names <- length(names(sub))

n_eventid <- length(unique(sub$EventID))

n_na_eventid <- sub %>% 
  filter(is.na(EventID) == T) %>% 
  pull(EventID) %>% 
  length()

n_densitycat <- length(unique(sub$DensityCategory))

```

The input dataset contains `r n_records` coral and sponge occurrences observed within `r n_eventid` dives ('EventID'). Below are a few summary tables describing the data more fully.    

### Table: Dives organized richness
This table has each dive arranged in descending order by richness.  

```{r richness_per_dive}
### Table: Richness per dive sorted in descending order by richness values.  

x <- z %>% group_by(EventID) %>% 
  summarize(richness = length(unique(ScientificName))) %>% 
  arrange(desc(richness))

## create flextable
myft <- flextable(x)

## modify flextable
myft <- theme_vanilla(myft)
myft<- fontsize(myft, size = 10, part = "body")
myft<- fontsize(myft, size = 10, part = "header")
myft <- align(myft, align = "left", part = "header")
myft <- align(myft, align = "left", part = "body")

myft <- set_table_properties(myft, width = .5, layout = "autofit")
myft


```

### Table: Dives organized by diversity (Simpson)
All taxa (both coral and sponges at all taxonomic ranks included in estimates).

```{r diversity_2, echo=F}

diversity <- diversity(site.sp, index = "simpson") #makes an object
#that the diversity values #are written to.
# summary(plots) #gives summary statistics for the plots
# median(plots) #gives the median
# mean(plots) #gives the mean
# class(plots)
# max(plots[1:30])
# # View(plots)

# creating plots
plots_df <- data.frame(diversity)

# View(plots_df)
plots_df$EventID <- row.names(plots_df)

# names(d)
d <- merge(z, plots_df, all.x = T)
# length(d$EventID)
# length(unique(d$EventID))
x <- d %>% group_by(EventID, diversity) %>% 
  summarize(n=n(), 
            Latitude = mean(as.numeric(Latitude)), 
            Longitude = mean(as.numeric(Longitude))) %>%
  arrange(desc(diversity))

## create flextable
myft <- flextable(x)

## modify flextable
myft <- theme_vanilla(myft)
myft<- fontsize(myft, size = 10, part = "body")
myft<- fontsize(myft, size = 10, part = "header")
myft <- align(myft, align = "left", part = "header")
myft <- align(myft, align = "left", part = "body")

myft <- set_table_properties(myft, width = .5, layout = "autofit")
myft


#View(x)

setwd("C:/rworking/deepseatools/indata")
write.xlsx(x, '20210824-0_Summary_of_Corals_and_Sponges_Diversity_by_EventID.xlsx') 


## Calculating various Diversity Indices

# H <- diversity(site.sp)
# simp <- diversity(site.sp, "simpson")
# invsimp <- diversity(site.sp, "inv")
# ## Unbiased Simpson (Hurlbert 1971, eq. 5) with rarefy:
# unbias.simp <- rarefy(site.sp, 2) - 1
# ## Fisher alpha
# alpha <- fisher.alpha(site.sp)
# ## Plot all
# pairs(cbind(H, simp, invsimp, unbias.simp, alpha, J), pch="+", col="blue")
# ## Species richness (S) and Pielou's evenness (J):
# S <- specnumber(site.sp) ## rowSums(site.sp > 0) does the same...
# J <- H/log(S)

```

### Map of diversity values (Simpson)

```{r mapping_diversity , echo=FALSE, eval=T}

x$Latitude <- as.numeric(x$Latitude)
x$Longitude <- as.numeric(x$Longitude)

pal <- colorNumeric(
  palette = "Reds",
  domain = x$diversity)


m <- leaflet(width = 800, height = 500)
m <- addProviderTiles(m, "Esri.OceanBasemap") #Esri.OceanBasemap, "CartoDB.DarkMatter"
m <- addCircleMarkers(m, data=x,
                      radius=5,
                      weight=0,
                      #fillColor= "green",
                      color = ~pal(diversity),
                      fillOpacity=1,
                      popup = paste0( "EventID: ", 
                                      x$EventID,
                                      "<br>",
                                      "Diversity (Simpson): ", 
                                      x$diversity,
                                      "<br>"))
                                   
m

```

### Table: Dives organized by density category
* *EventID*: Table is grouped by 'EventID'
* *count_per_km*: IndividualCount per linear distance surveyed (km). This field is used for sorting this table in descending order. 
* *DensityCategory*: One of 4 assigned density categories

```{r eventID_table}
## create table
x <- z %>% 
  group_by(EventID) %>% 
  arrange(EventID) %>% 
  summarize(count_per_km = unique(n_km),
            DensityCategory = unique(DensityCategory)) %>% 
  arrange(desc(count_per_km))

## checking
# length(x$EventID)

## create flextable
myft <- flextable(x)

## modify flextable
myft <- theme_vanilla(myft)
myft<- fontsize(myft, size = 10, part = "body")
myft<- fontsize(myft, size = 10, part = "header")
myft <- align(myft, align = "left", part = "header")
myft <- align(myft, align = "left", part = "body")

myft <- set_table_properties(myft, width = .5, layout = "autofit")
myft

# myft <- italic(myft, j = 3)
# myft <- color(myft, ~ drat > 3.5, ~ drat, color = "red")
# myft <- bold(myft, ~ drat > 3.5, ~ drat, bold = TRUE)
```

### Table (Export): All taxa summary by individual taxa grouped by EventID.

This file is loaded as an Excel file to Google Drive at the following location: <a href="https://docs.google.com/spreadsheets/d/1KHBkC3JkJE4rpTiI_buXAiN9ZLcj82S1/edit?usp=sharing&ouid=109414727136135095326&rtpof=true&sd=true" target="_new">LINK</a>

This table is handy for looking at the taxa present in one dive in order of abundance. The table is huge and is much better digested in spreadsheet form. 

```{r table_coral_and_sponge_summary_by_EventID, echo=FALSE} 
yo <- sub %>%
  filter(
    # if needed, filter by TaxonRank
    # TaxonRank == 'genus' |
    #   TaxonRank == 'species' |
    #   TaxonRank == 'subspecies',
    # #Phylum == 'Cnidaria'
  ) %>%
  arrange(ScientificName) %>%
  group_by(EventID, ScientificName) %>%
  summarize(
    IndividualCount_sum = sum(IndividualCount),
    TaxonRank = paste(unique(TaxonRank), collapse= ' | '),
    Phylum = paste(unique(Phylum), collapse= ' | '),
    #Taxa = paste(unique(ScientificName), collapse= ' | '),
    #Densities = paste(unique(Density), collapse = ' | '),
    #Richness = length(unique(ScientificName)),
    Locality = paste(unique(Locality), collapse= ' | '),
    MinimumDepth = min(DepthInMeters),
    MaximumDepth = max(DepthInMeters), 
    NumberRecords = n(),
    ImageURL = paste(unique(ImageURL), collapse= ' | ')
     )

yo <- yo %>% arrange(EventID, desc(IndividualCount_sum))

## create flextable
# myft <- flextable(x)

## modify flextable
# myft <- theme_vanilla(myft)
# myft<- fontsize(myft, size = 10, part = "body")
# myft <- align(myft, align = "left", part = "header")
# myft <- align(myft, align = "left", part = "body")
# 
# myft <- set_table_properties(myft, width = .5, layout = "autofit")
# myft

# myft <- italic(myft, j = 3)
# myft <- color(myft, ~ drat > 3.5, ~ drat, color = "red")
# myft <- bold(myft, ~ drat > 3.5, ~ drat, bold = TRUE)

## write it to excel 
filename = "20210920_Summary_of_all_corals_and_sponges_by_ScientificName_with_groups_Scientific_EventID.xlsx"
setwd("C:/rworking/deepseatools/indata/")
write.xlsx(yo, filename)
```

## NMDS (all data)

### Running NMDS

Here are the statistical results from the NMDS run. The code is also shown below for checking run time parameters.   

```{r NMDS_run_2, echo=F}
##### running NMDS starting with site X species matrix #####
#install.packages("vegan")
x <- z

library(vegan)
NMDS <- metaMDS(site.sp, distance = "bray", binary = T, k=4, trymax = 30)

##### extracting the site and species scores for use with ggplot2 ##### 
#Using the scores function from vegan to extract the site scoresand convert to a data.frame
site.scores <- as.data.frame(scores(NMDS))  
site.scores$site <- rownames(site.scores)  # create a column of site names, from the rownames of data.scores

# head(site.scores)  #look at the data
# merge information from d (to get categories)
site.scores <- merge(site.scores, x, by.x = 'site', by.y = 'EventID')

site.scores <- site.scores %>% group_by(site) %>% summarise(NMDS1 = NMDS1[1],
                                                            NMDS2 = NMDS2[1],
                                                            DepthCat = DepthCat[1],
                                                            temp_cat = temp_cat[1],
                                                            loc = loc[1],
                                                            ox_cat = ox_cat[1], 
                                                            loc2 = Locality[1])

# extracting species scores
species.scores <- as.data.frame(scores(NMDS, "species"))  #Using the scores function from vegan to extract the species scores and convert to a data.frame
species.scores$species <- rownames(species.scores)  # create a column of species, from the rownames of species.scores
# head(species.scores)  #look at the data
species.scores$spec_code <- 1:(length(species.scores$species))

NMDS

```

### Clustering results

```{r, echo=F, fig.height=10, fig.width=10}
##### create clusters #####
## uses 'site.sp' from above 
## creates clustering result ('hclust' object)
sp_bray = vegdist(site.sp, method='bray', binary=T)
fit <- hclust(sp_bray)

##### plot clusters with groups ####
plot(fit, 
     main = "", 
     sub = "", 
     xlab="Sites", 
     axes = T, 
     ylab = "Ecological Distance", 
     hang = -1) # display dendogram
groups <- cutree(fit, k=4) ## cut tree into X clusters
group.order <- groups[fit$order] ## this is not in order by site.
group.in.cluster <- unique(group.order)  ## produces a vector of integers denoting order of clusters
rect.hclust (fit, border = group.in.cluster, k = 4)
legend ('topleft', legend = paste ('Cluster', c(group.in.cluster)), pch = 22, bty = 'n', col = group.in.cluster)


##### take group membership and turn into dataframe with EventID ("site") and cluster group membership. ##### 
df <- data.frame(groups) ## goes to alpha order by site
df <- tibble::rownames_to_column(df, "site") ## adding column header "site"

## checking
# names(df)
# class(site.scores)
# namesite.scores$site)(site.scores)

##### merge groups onto the site.scores dataframe.##### 
test <- merge(site.scores, df, by.x = "site")
site.scores <- test
site.scores$groups <- as.factor(site.scores$groups)  ## create "groups" variable in site.scores

## checking
# site.scores$site
# head(groups)

```

### Interactive map of clusters
Same colors are used as the first NMDS plot below and the clustering results above.

```{r fig.width=5, fig.height=4}

df <- z %>% group_by(EventID) %>% 
  summarise(Latitude = mean(as.numeric(Latitude)), 
            Longitude = mean(as.numeric(Longitude)))


## check
# names(df)
# length(df$EventID)
# table(is.na(df$Latitude))

test <- merge(site.scores, df, by.x = "site", by.y = "EventID")
site.scores_map <- test

## checking

site.scores_map$Latitude <- as.numeric(site.scores_map$Latitude)
site.scores_map$Longitude <- as.numeric(site.scores_map$Longitude)

pal <- colorFactor(
  palette = c("black", "red", "green", "blue"),
  domain = site.scores_map$groups
)


m <- leaflet(width = 800, height = 500)
m <- addProviderTiles(m, "Esri.OceanBasemap") #Esri.OceanBasemap, "CartoDB.DarkMatter"
m <- addCircleMarkers(m, data=site.scores_map,
                      radius=5,
                      weight=0,
                      fill = groups,
                      color = ~pal(groups),
                      fillOpacity=1, 
                      popup = paste0( "EventID: ", 
                                      site.scores_map$site,
                                      "<br>",
                                      "Cluster: ", 
                                      site.scores_map$groups,
                                      "<br>"), 
                      label = site.scores_map$site
                      )
                                   
m
```

### NMDS plot with cluster results

```{r}
plot(NMDS, disp="sites", type="n")
ordispider(NMDS, site.scores$groups, col= "black", label = TRUE)
#ordihull(NMDS, site.scores$groups, col=1:4, lwd=3, label = TRUE)
#ordiellipse(NMDS, groups, col=1:4, kind = "ehull", lwd=3)
ordiellipse(NMDS, groups, col=c("black", "red", "green", "blue"), draw="polygon")
points(NMDS, disp="sites", pch=21, col="red", bg="yellow", cex=1.3)
#text(NMDS, display = "spec", cex=0.7, col="blue")

#points(NMDS, disp="species", pch=21, col="red", bg="red", cex=1.3)

#NMDS$data

```

### NMDS with site names for reference

```{r}
plot(NMDS, disp="sites", type="n")
# ordispider(NMDS, site.scores$loc, col= "black", label = TRUE)
# #ordihull(NMDS, site.scores$groups, col=1:4, lwd=3, label = TRUE)
# #ordiellipse(NMDS, groups, col=1:4, kind = "ehull", lwd=3)
# ordiellipse(NMDS, site.scores$loc, col=c("blue", "green", "red", "black"), draw="polygon")
points(NMDS, disp="sites", pch=21, col="red", bg="yellow", cex=1.3)
text(NMDS, display = "sites", cex=0.7, col="black")
#text(NMDS, display = "sites", cex=0.7, col="black")
#points(NMDS, disp="species", pch=21, col="red", bg="red", cex=1.3)
#NMDS$data

```

### NMDS with species names for reference

```{r}
plot(NMDS, disp="sites", type="n")
# ordispider(NMDS, site.scores$loc, col= "black", label = TRUE)
# #ordihull(NMDS, site.scores$groups, col=1:4, lwd=3, label = TRUE)
# #ordiellipse(NMDS, groups, col=1:4, kind = "ehull", lwd=3)
# ordiellipse(NMDS, site.scores$loc, col=c("blue", "green", "red", "black"), draw="polygon")
points(NMDS, disp="sites", pch=21, col="red", bg="yellow", cex=1.3)
text(NMDS, display = "species", cex=0.7, col="black")
#points(NMDS, disp="species", pch=21, col="red", bg="red", cex=1.3)
#NMDS$data

```


### NMDS plot with depth 

Values: shallower; medium; deeper

```{r}
plot(NMDS, disp="sites", type="n")
ordispider(NMDS, site.scores$DepthCat, col= "black", label = TRUE)
#ordihull(NMDS, site.scores$groups, col=1:4, lwd=3, label = TRUE)
#ordiellipse(NMDS, groups, col=1:4, kind = "ehull", lwd=3)
ordiellipse(NMDS, site.scores$DepthCat, col=c("blue", "green", "red", "black"), draw="polygon")
points(NMDS, disp="sites", pch=21, col="red", bg="yellow", cex=1.3)
#text(NMDS, display = "spec", cex=0.7, col="blue")

#points(NMDS, disp="species", pch=21, col="red", bg="red", cex=1.3)

#NMDS$data

```

### NMDS plot with temperature 

Values: warmer; cold; colder

```{r}
plot(NMDS, disp="sites", type="n")
ordispider(NMDS, site.scores$temp_cat, col= "black", label = TRUE)
#ordihull(NMDS, site.scores$groups, col=1:4, lwd=3, label = TRUE)
#ordiellipse(NMDS, groups, col=1:4, kind = "ehull", lwd=3)
ordiellipse(NMDS, site.scores$temp_cat, col=c("blue", "green", "red", "black"), draw="polygon")
points(NMDS, disp="sites", pch=21, col="red", bg="yellow", cex=1.3)
#text(NMDS, display = "spec", cex=0.7, col="blue")
#points(NMDS, disp="species", pch=21, col="red", bg="red", cex=1.3)
#NMDS$data

```

### NMDS plot with oxygen categories

Values: poor; rich

```{r}
plot(NMDS, disp="sites", type="n")
ordispider(NMDS, site.scores$ox_cat, col= "black", label = TRUE)
#ordihull(NMDS, site.scores$groups, col=1:4, lwd=3, label = TRUE)
#ordiellipse(NMDS, groups, col=1:4, kind = "ehull", lwd=3)
ordiellipse(NMDS, site.scores$ox_cat, col=c("blue", "green", "red", "black"), draw="polygon")
points(NMDS, disp="sites", pch=21, col="red", bg="yellow", cex=1.3)
#text(NMDS, display = "spec", cex=0.7, col="blue")
#points(NMDS, disp="species", pch=21, col="red", bg="red", cex=1.3)
#NMDS$data

```

### NMDS plot with location categories

Values: north; south

```{r}
plot(NMDS, disp="sites", type="n")
ordispider(NMDS, site.scores$loc, col= "black", label = TRUE)
#ordihull(NMDS, site.scores$groups, col=1:4, lwd=3, label = TRUE)
#ordiellipse(NMDS, groups, col=1:4, kind = "ehull", lwd=3)
ordiellipse(NMDS, site.scores$loc, col=c("blue", "green", "red", "black"), draw="polygon")
points(NMDS, disp="sites", pch=21, col="red", bg="yellow", cex=1.3)
#text(NMDS, display = "spec", cex=0.7, col="blue")
#points(NMDS, disp="species", pch=21, col="red", bg="red", cex=1.3)
#NMDS$data

```

