---
title: "DSBS16_HI_Analysis"
author: "Robert McGuinn"
date: "`r Sys.Date()`"
output: 
  word_document:
     reference_docx: template.docx
always_allow_html: yes
editor_options: 
  chunk_output_type: console
---
## ---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, 
                      eval = T, 
                      warning = F,
                      message = F,
                      fig.height = 10, 
                      fig.width = 10, 
                      dpi = 300)


```

```{r packages}
library(tidyr)
library(openxlsx)
library(googlesheets4)
library(googledrive)
library(flextable)
library(dplyr)
library(flextable)
library(vegan)
library(ggplot2)
library(reshape2)
library(extrafont)
library(leaflet)

```

## Resources for this analysis

The main project folder for this analysis is on Google Drive: [LINK](https://drive.google.com/drive/folders/1ToTK_v7NF7ZPdwXQwD1hzCwuVZCZRH_d)

The RMarkdown code used to generate this report is available here: [LINK](https://github.com/RobertMcGuinn/deepseatools/blob/master/code/DSBS16_HI_Analysis.Rmd)

## Input data sets

Datasets used in this analysis were acquired on 2021-08-18 from Tom Hourigan via Excel file. The coral and sponge species occurrence records are from a subset of dives in the Hawaiian region (including Johnston Atoll and the Musician Seamounts). This original dataset is shared on Google Drive here: [LINK](https://docs.google.com/spreadsheets/d/16izQ_vWsSDgMupbNAEUsVTvMISzmrLaC/edit?usp=sharing&ouid=109414727136135095326&rtpof=true&sd=true)


```{r data_intake, cache=T}
##### bringing in dataframe from xlsx stored on drive ####

## file path [MANUAL]
filename <- "20210817_Hawaii-HiDensity1500-2500m"

## create a list of files (or single file) that meets title query [MANUAL]
x <- drive_find(q = "name contains '20210817_Hawaii-HiDensity1500-2500m'")

## browse to it
# x %>% drive_browse()

# getting the id as a character string
y <- x$id

# this downloads the file to the specified path
dl <- drive_download(as_id(y),
                     path = paste("C:/rworking/deepseatools/indata/",
                                  filename,
                                  ".xlsx", 
                                  sep = ""),
                     overwrite = TRUE)

## extract just the file of interest from the zip file to a dataframe called 'sub'
sub <- read.xlsx(dl$local_path)

## names 
# names(sub)
```

```{r change_file_intake}
# received from THourigan on 20210826
##### bringing in dataframe from xlsx stored on drive ####

## file path [MANUAL]
filename <- "20210826_0_Changes_THourigan"

## create a list of files (or single file) that meets title query [MANUAL]
x <- drive_find(q = "name contains '20210826_0_Changes_THourigan'")

## browse to it
# x %>% drive_browse()

# getting the id as a character string
y <- x$id

# this downloads the file to the specified path
dl <- drive_download(as_id(y),
                     path = paste("C:/rworking/deepseatools/indata/",
                                  filename,
                                  ".xlsx", 
                                  sep = ""),
                     overwrite = TRUE)

## extract just the file of interest from the zip file to a dataframe called 'sub'
change <- read.xlsx(dl$local_path)

## names 
# names(change)

```


```{r data_scrub}
sub <- sub %>% rename(n_km = `#/km`,
                    m_divetrack = `Dive.Track.(m)`,
                    kennedy_class = Kennedy.Class)

sub = subset(sub, select = -c(IdentificationComments,VerbatimSize))


change <- change %>% rename(n_km = `#/km`,
                    m_divetrack = `Dive.Track.(m)`,
                    kennedy_class = Kennedy.Class)

change = subset(change, select = -c(Flag))

```

```{r merging changes}

# this merges the two dataframes by CatalogNumber and then adds changes to the original set where needed and leaves them alone where not needed.
test <- left_join(sub, change, 
                  by = c("CatalogNumber")) %>% 
  mutate(ScientificName.x = ifelse(is.na(ScientificName.y), 
                                   ScientificName.x, 
                                   ScientificName.y)) %>% 
  mutate(VernacularNameCategory.x = ifelse(is.na(VernacularNameCategory.y), 
                                           VernacularNameCategory.x, 
                                           VernacularNameCategory.y)) %>% 
  mutate(TaxonRank.x = ifelse(is.na(TaxonRank.y), 
                              TaxonRank.x, 
                              TaxonRank.y)) %>% 
  mutate(Phylum.x = ifelse(is.na(Phylum.y), 
                           Phylum.x, 
                           Phylum.y)) %>% 
  mutate(Class.x = ifelse(is.na(Class.y), 
                          Class.x, 
                          Class.y)) %>% 
  mutate(Subclass.x = ifelse(is.na(Subclass.y), 
                             Subclass.x, 
                             Subclass.y)) %>% 
  mutate(Order.x = ifelse(is.na(Order.y), 
                          Order.x, 
                          Order.y)) %>% 
  mutate(Suborder.x = ifelse(is.na(Suborder.y), 
                             Suborder.x, 
                             Suborder.y)) %>% 
  mutate(Family.x = ifelse(is.na(Family.y), 
                           Family.x, 
                           Family.y)) %>% 
  mutate(Genus.x = ifelse(is.na(Genus.y), 
                          Genus.x, 
                          Genus.y))


## checking
# test %>% filter(CatalogNumber == "985165") %>% # without "940589" with change"985165"
#   group_by(ScientificName.x, ScientificName.y) %>%
#   summarize(n=n()) %>%
#   View()

## Now get rid of all columns with ".y" in the name since we already harvested the changes and put them the in *.x variables.
test2 <- select(test, -contains(".y"))

# Now get rid of the *.x suffix from all variable names to get back to the corrected file.  
names(test2) <- gsub(".x$","", names(test2))

# Now overwrite 'sub' to continue on down the road 
sub <- test2

## clean up
rm(test)
rm(test2)


```

## Summary
```{r variables}
n_records <- length(sub$CatalogNumber)

list_imageurls <- sub %>%  filter(is.na(ImageURL)==F) %>% 
  pull(ImageURL)

n_images <- length(unique(list_imageurls))

txt_names <- names(sub)

n_names <- length(names(sub))

n_eventid <- length(unique(sub$EventID))

n_na_eventid <- sub %>% 
  filter(is.na(EventID) == T) %>% 
  pull(EventID) %>% 
  length()

n_densitycat <- length(unique(sub$DensityCategory))

```

The input dataset contains `r n_records` of coral and sponge occurrences observed within `r n_eventid` dives ('EventID'). 

### Table: Dives organized by Density Category
* *EventID*: Table is grouped by 'EventID'
* *count_per_km*: IndividualCount per linear distance surveyed (km). This field is used for sorting this table in descending order. 
* *DensityCategory*: One of 4 assigned density categories

```{r eventID_table}
## create table
x <- sub %>% 
  group_by(EventID) %>% 
  arrange(EventID) %>% 
  summarize(count_per_km = unique(n_km),
            DensityCategory = unique(DensityCategory)) %>% 
  arrange(desc(count_per_km))

## checking
# length(x$EventID)

## create flextable
myft <- flextable(x)

## modify flextable
myft <- theme_vanilla(myft)
myft<- fontsize(myft, size = 8, part = "body")
myft<- fontsize(myft, size = 8, part = "header")
myft <- align(myft, align = "left", part = "header")
myft <- align(myft, align = "left", part = "body")

myft <- set_table_properties(myft, width = .5, layout = "autofit")
myft

# myft <- italic(myft, j = 3)
# myft <- color(myft, ~ drat > 3.5, ~ drat, color = "red")
# myft <- bold(myft, ~ drat > 3.5, ~ drat, bold = TRUE)
```

### Table: Taxonomic lists (All)
* *EventID*: This table is grouped by dive (EventID)
* *count_per_km*: IndividualCount per linear distance surveyed (km). This field is used for sorting this table in descending order. 
* *number_genera*: The number of genera observed the dive.
* *number_species*: The number of species observed on the dive.
* *genus_list*: List of genera observed on the dive.
* *species_list*: List of species observed on the dive.

```{r taxa_list_table}
## create table
x <- sub %>% arrange(ScientificName) %>% 
  group_by(EventID) %>% 
  summarize(count_per_km = unique(n_km),
            number_species = length(unique(ScientificName)),
            number_genera = length(unique(Genus)),
            genus_list = toString(unique(Genus)),
            species_list = toString(unique(ScientificName))
            ) %>% 
  arrange(desc(count_per_km))
      
## checking
# length(x$EventID)

## create flextable
myft <- flextable(x)

## modify flextable
myft <- theme_vanilla(myft)
myft<- fontsize(myft, size = 8, part = "body")
myft<- fontsize(myft, size = 8, part = "header")
myft <- align(myft, align = "left", part = "header")
myft <- align(myft, align = "left", part = "body")

myft <- set_table_properties(myft, width = .5, layout = "autofit")
myft

# myft <- italic(myft, j = 3)
# myft <- color(myft, ~ drat > 3.5, ~ drat, color = "red")
# myft <- bold(myft, ~ drat > 3.5, ~ drat, bold = TRUE)

## write it out 
setwd("C:/rworking/deepseatools/indata/")
write.xlsx(x, "20210826-0_species_lists_per_EventID.xlsx")

```

### Table: All taxa summary by individual taxa grouped within EventID.

This file is loaded as an Excel file to Google Drive at the following location: [LINK](https://docs.google.com/spreadsheets/d/1KHBkC3JkJE4rpTiI_buXAiN9ZLcj82S1/edit?usp=sharing&ouid=109414727136135095326&rtpof=true&sd=true) 

This table is handy for looking at the taxa present in one dive in order of abundance.  The table is huge and is much better digested in spreadsheet form. 

```{r table_coral_and_sponge_summary_by_EventID, echo=FALSE} 
x <- sub %>%
  filter(
    # if needed, filter by TaxonRank
    # TaxonRank == 'genus' |
    #   TaxonRank == 'species' |
    #   TaxonRank == 'subspecies',
    # #Phylum == 'Cnidaria'
  ) %>%
  arrange(ScientificName) %>%
  group_by(EventID, ScientificName) %>%
  summarize(
    IndividualCount_sum = sum(IndividualCount),
    TaxonRank = paste(unique(TaxonRank), collapse= ' | '),
    Phylum = paste(unique(Phylum), collapse= ' | '),
    #Taxa = paste(unique(ScientificName), collapse= ' | '),
    #Densities = paste(unique(Density), collapse = ' | '),
    #Richness = length(unique(ScientificName)),
    Locality = paste(unique(Locality), collapse= ' | '),
    MinimumDepth = min(DepthInMeters),
    MaximumDepth = max(DepthInMeters), 
    NumberRecords = n(),
    ImageURL = paste(unique(ImageURL), collapse= ' | ')
     )

x <- x %>% arrange(EventID, desc(IndividualCount_sum))

## create flextable
# myft <- flextable(x)

## modify flextable
# myft <- theme_vanilla(myft)
# myft<- fontsize(myft, size = 8, part = "body")
# myft <- align(myft, align = "left", part = "header")
# myft <- align(myft, align = "left", part = "body")
# 
# myft <- set_table_properties(myft, width = .5, layout = "autofit")
# myft

# myft <- italic(myft, j = 3)
# myft <- color(myft, ~ drat > 3.5, ~ drat, color = "red")
# myft <- bold(myft, ~ drat > 3.5, ~ drat, bold = TRUE)

## write it to excel 
filename = "20210826-0_Summary_of_all_corals_and_sponges_by_ScientificName_with_groups_Scientific_EventID.xlsx"
setwd("C:/rworking/deepseatools/indata/")
write.xlsx(x, filename)
```


### Table: Images 
There are `r n_images` unique images within the dataset.  Here is how they break down by each of the `r n_eventid`

```{r images_table}
## create table
x <- sub %>% filter(is.na(ImageURL) == F) %>% 
  group_by(EventID) %>% 
  summarize(n_images = length(unique(ImageURL))
            ) %>% 
  arrange(desc(n_images))

## checking
# length(x$EventID)
sum(x$n_images)

## create flextable
myft <- flextable(x)

## modify flextable
myft <- theme_vanilla(myft)
myft<- fontsize(myft, size = 8, part = "body")
myft<- fontsize(myft, size = 8, part = "header")
myft <- align(myft, align = "left", part = "header")
myft <- align(myft, align = "left", part = "body")

myft <- set_table_properties(myft, width = .5, layout = "autofit")
myft

# myft <- italic(myft, j = 3)
# myft <- color(myft, ~ drat > 3.5, ~ drat, color = "red")
# myft <- bold(myft, ~ drat > 3.5, ~ drat, bold = TRUE)

```

## Images
### Images to local folder

```{r images_from_url, eval=F}
## load any images to a local folder 
## WARNING: This chunk takes a long time.  Go get a coffee
###### download all images and put them in a folder #####
z <- sub %>% filter(is.na(ImageURL) == F)

## [MANUAL]: go create folder and point to it 
setwd("C:/rworking/deepseatools/images/dsbs6/")
for(i in 1:length(z$CatalogNumber)){
  download.file(as.character(z$ImageURL[i]),
                destfile = paste("DSCRTP",
                                 z$CatalogNumber[i],
                                 z$EventID[i],
                                 z$ScientificName[i],
                                 z$DepthInMeters[i],
                                 basename(as.character(z$ImageURL[i])),
                                 sep = '_'),
                mode = "wb")
}

## OPTIONAL: write out featureclass to geodatabase from sf object
fgdb_path <- "C:/data/aprx/aoi/aoi.gdb"
arc.write(file.path(fgdb_path, 'imageset_5000'), data=z, overwrite = TRUE)

```

### Images from local folder to Google Drive
```{r local_images_to_Drive_folder, eval=F}
## load the images to a Google Drive folder
## WARNING: This chunk could take a longish time. Go get a coffee
##### Loading files from a local folder to Google Drive ##### 
## MANUAL CHANGE "folderurl" to the desired drive folder ID
folderurl <- "https://drive.google.com/drive/folders/1oOiCVRWJUc4fQQ1dec9q3APYfk1Ea4nj"

## get the list of files from the local folder
files <- list.files(path="C:/rworking/deepseatools/indata/imageset_5000", full.names=TRUE, recursive=FALSE)

## loop upload images to Google Drive
for(i in files){
drive_upload(i,
             path = as_id(folderurl),
             overwrite = T)
}

```


## NMDS (full set)

### Filter to only species appearing in the top ten list  

```{r filter_to_top_ten_2, echo=F, eval=T, warning=F, message=F}
##### filter d down to just the species appearing in the top ten ##### 
topten <- sub %>%
 group_by(EventID, ScientificName)%>%
 summarize(individual_count_sum = sum(IndividualCount))%>%
 top_n(10, individual_count_sum) %>% 
  arrange(EventID, desc(individual_count_sum))

z <- sub %>% filter(
  ScientificName %in% topten$ScientificName
# EventID %in% c("NA101-H1715","NA101-H1720", "D2-EX1504L2-17")
  )

```

### Setting up depth, temp, and location categories

```{r setdepthclass_2, echo=FALSE, cache = FALSE} 
##### setting depth categories ##### 
# qplot(as.numeric(z$DepthInMeters))
z$DepthCat[as.numeric(z$DepthInMeters) <= 1800] <- "shallower"
z$DepthCat[as.numeric(z$DepthInMeters) > 1800 & 
             as.numeric(z$DepthInMeters) < 2400] <- "medium"
z$DepthCat[as.numeric(z$DepthInMeters) >= 2400] <- "deeper"
z$DepthCat <- factor(z$DepthCat, levels = c("shallower","medium","deeper"))

##### setting location categories ##### 
# qplot(as.numeric(z$Latitude))
z$loc[as.numeric(z$Latitude) >= 20] <- "north"
z$loc[as.numeric(z$Latitude) < 20] <- "south"

##### longitude cat #####
z$longloc[as.numeric(z$Longitude) > -165] <- "east"
z$longloc[as.numeric(z$Longitude) < -165] <- "west"

##### setting temperature categories #####
# qplot(as.numeric(z$Temperature))
z$temp_cat[as.numeric(z$Temperature) > 2.2] <- "warmer"
z$temp_cat[as.numeric(z$Temperature) >= 1.9 & as.numeric(z$Temperature) <= 2.2] <- "cold"
z$temp_cat[as.numeric(z$Temperature) < 1.9] <- "colder"
z$temp_cat <- factor(z$temp_cat, levels = c("warmer", "cold", "colder"))

##### setting oxygen categories #####
# qplot(as.numeric(z$Oxygen))
z$ox_cat[as.numeric(z$Oxygen) >= 2.6] <- "rich"
z$ox_cat[as.numeric(z$Oxygen) < 2.6] <- "poor"

```

### Making the matrix

First we adjust the abundances by the length of the drive track. This creates a normalized value for 'IndividualCount'. It is calculated as 'IndividualCount' per meter. This value is then entered into the site by species matrix which is used in the NMDS and clustering algorithms.  

``` {r NMDS_2, echo=FALSE, cache = FALSE, dpi=300, warning=FALSE, message=FALSE, results='hide', cache=T}
## filter if necessary
x <- z %>% 
  filter(
    #DepthCat == 'deep',
    #EventID != "D2-EX1605L1-13"
    #Phylum == "Cnidaria"
    #DensityCategory == "high and higest"
  )

## checking 
# length(table(sub$n_km, useNA = 'always'))

## adjust the IndividualCount values 
x$IndividualCount_z <- x$IndividualCount/x$m_divetrack

## create site X species matrix 
library(vegan)
site.sp <- reshape2::dcast(x, EventID ~ ScientificName, value.var = "IndividualCount_z", fun.aggregate = sum)

## set all NA to 0 because the vegan package can't handle the NA
site.sp[is.na(site.sp)] = 0

## checking resulting table
# View(site.sp)

## export to check
# setwd("C:/rworking/deepseatools/indata")
# write.csv(site.sp, 'site.sp.csv')

# creating a site variable
site.sp$site <- site.sp$EventID

# getting rid of non-needed variables
site.sp <- site.sp %>%
  dplyr::select(-EventID)

# # moving the site variable to the beginning
# col_idx <- grep("site", names(site.sp))
# site.sp <- site.sp[, c(col_idx, (1:ncol(site.sp))[-col_idx])]
# # names(site.sp)

# set the site as the row.names
row.names(site.sp) <- site.sp$site

# remove the site variable
site.sp <- site.sp %>%
  dplyr::select(-site)

# making it a matrix
site.sp <- as.matrix(site.sp)

## checking 
# NMDS

``` 

### NMDS Run

Here are the statistical results from the NMDS run:   

```{r NMDS_run_2}
##### running NMDS starting with site X species matrix #####
#install.packages("vegan")
library(vegan)
NMDS <- metaMDS(site.sp, distance = "bray", binary = F, k=4, trymax = 30)

##### extracting the site and species scores for use with ggplot2 ##### 
#Using the scores function from vegan to extract the site scoresand convert to a data.frame
site.scores <- as.data.frame(scores(NMDS))  
site.scores$site <- rownames(site.scores)  # create a column of site names, from the rownames of data.scores

# head(site.scores)  #look at the data
# merge information from d (to get categories)
site.scores <- merge(site.scores, x, by.x = 'site', by.y = 'EventID')
site.scores <- site.scores %>% group_by(site) %>% summarise(NMDS1 = NMDS1[1],
                                                            NMDS2 = NMDS2[1],
                                                            DepthCat = DepthCat[1],
                                                            temp_cat = temp_cat[1],
                                                            loc = loc[1],
                                                            ox_cat = ox_cat[1], 
                                                            loc2 = Locality[1])

# extracting species scores
species.scores <- as.data.frame(scores(NMDS, "species"))  #Using the scores function from vegan to extract the species scores and convert to a data.frame
species.scores$species <- rownames(species.scores)  # create a column of species, from the rownames of species.scores
# head(species.scores)  #look at the data
species.scores$spec_code <- 1:(length(species.scores$species))

NMDS

```

### Plot of NMDS with group colors as from depth bin

```{r plotting_NMDS_depth_2}
##### plotting NMDS ##### 
ggplot() +
  geom_point(data=site.scores,aes(x=NMDS2,y=NMDS1, color=DepthCat), size=5) + # add the point markers
  geom_text(data=site.scores,aes(x=NMDS2,y=NMDS1, label = site), size=4, position = position_nudge(y = .0001)) + # add the text
  #geom_point(data=species.scores,aes(x=NMDS1,y=NMDS2), shape = 3, size=4) + 
  coord_equal() +
  theme_bw(base_size = 15, base_family = "Cambria") +
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank())

```

### Plot of NMDS with group colors from temperature

```{r plotting_NMDS_temp_2}

ggplot() +
  geom_point(data=site.scores,aes(x=NMDS2,y=NMDS1, color=temp_cat), size=5) + # add the point markers
  geom_text(data=site.scores,aes(x=NMDS2,y=NMDS1, label = site), size=4, position = position_nudge(y = .0001)) + # add the text
  #geom_point(data=species.scores,aes(x=NMDS1,y=NMDS2), shape = 3, size=4) + 
  coord_equal() +
  theme_bw(base_size = 15, base_family = "Cambria") +
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank())

```

### Plot of NMDS with group colors from location 

Location based on latitude break at X.  

```{r}

ggplot() +
  geom_point(data=site.scores,aes(x=NMDS2,y=NMDS1, color=loc), size=5) + # add the point markers
  geom_text(data=site.scores,aes(x=NMDS2,y=NMDS1, label = site), size=4, position = position_nudge(y = .0001)) + # add the text
  #geom_point(data=species.scores,aes(x=NMDS1,y=NMDS2), shape = 3, size=4) + 
  coord_equal() +
  theme_bw(base_size = 15, base_family = "Cambria") +
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank())

```

### Plot of NMDS with group colors from oxygen values 


```{r}

ggplot() +
  geom_point(data=site.scores,aes(x=NMDS2,y=NMDS1, color=ox_cat), size=5) + # add the point markers
  geom_text(data=site.scores,aes(x=NMDS2,y=NMDS1, label = site), size=4, position = position_nudge(y = .0001)) + # add the text
  #geom_point(data=species.scores,aes(x=NMDS1,y=NMDS2), shape = 3, size=4) + 
  coord_equal() +
  theme_bw(base_size = 15, base_family = "Cambria") +
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank())

```


### Plot of NMDS species scores

Labels are are species code (by number). See table below for species code.   

```{r}

ggplot() +
  #geom_point(data=site.scores,aes(x=NMDS2,y=NMDS1, color=loc), size=5) + # add the point markers
  geom_text(data=species.scores,aes(x=NMDS2,y=NMDS1, label = spec_code), size=4, position = position_nudge(y = .0001)) + # add the text
  geom_point(data=species.scores,aes(x=NMDS2,y=NMDS1), shape = 3, size=4) + 
  coord_equal() +
  theme_bw(base_size = 15, base_family = "Cambria") +
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank())

```

### Table of species codes (key to the lables in the species score NMDS plot above)

```{r spec_codes_2}

x <- species.scores %>% dplyr::select(spec_code)

## convert row names to column
x <- tibble::rownames_to_column(x, "ScientificName")
## create flextable
myft <- flextable(x)

## modify flextable
myft <- theme_vanilla(myft)
myft<- fontsize(myft, size = 8, part = "body")
myft<- fontsize(myft, size = 8, part = "header")
myft <- align(myft, align = "left", part = "header")
myft <- align(myft, align = "left", part = "body")

myft <- set_table_properties(myft, width = .5, layout = "autofit")
myft

```

## Clustering 

```{r}
##### create different distance matrices ##### 
## Bray-Curtis as binary (presence/absence) of species per equation above
# sp_bray_binary = vegdist(site.sp, method='bray', binary=T)

## Bray-Curtis as amount (density) of species
sp_bray_transformed = vegdist(site.sp, method='bray')

## transformed Bray-Curtis as amount (density) of species
#sp_bray_transformed = vegdist(wisconsin(sqrt(site.sp)), method='bray')

##### create different clusters ######

fit <- hclust(sp_bray_transformed)
plot(fit, main = "", sub = "", xlab="Sites", axes = T, ylab = "Ecological Distance (Bray-Curtis - Wisconsin Transformed)", hang = -1) # display dendogram
groups <- cutree(fit, k=5) # cut tree into 5 clusters
# draw dendogram with red borders around the 5 clusters
rect.hclust(fit, k=5, border="red")

# fit <- hclust(sp_bray_binary)
# plot(fit, main = "", sub = "", xlab="Sites", axes = T, ylab = "Ecological Distance (Sorensen)", hang = -1) # display dendogram
# groups <- cutree(fit, k=5) # cut tree into 5 clusters
# # draw dendogram with red borders around the 5 clusters
# rect.hclust(fit, k=5, border="red")
# 
# fit <- hclust(sp_bray)
# plot(fit, main = "", sub = "", xlab="Sites", axes = T, ylab = "Ecological Distance (Bray-Curtis)", hang = -1) # display dendogram
# groups <- cutree(fit, k=5) # cut tree into 5 clusters
# # draw dendogram with red borders around the 5 clusters
# rect.hclust(fit, k=5, border="red")

```

## Diversity estimates.

##### Table: Simpson diversity per EventID
All taxa (both coral and sponges at all taxonomic ranks included in estimates).

```{r diversity_2, echo=F}

diversity <- diversity(site.sp, index = "simpson") #makes an object
#that the diversity values #are written to.
# summary(plots) #gives summary statistics for the plots
# median(plots) #gives the median
# mean(plots) #gives the mean
# class(plots)
# max(plots[1:30])
# # View(plots)

# creating plots
plots_df <- data.frame(diversity)

# View(plots_df)
plots_df$EventID <- row.names(plots_df)

# names(d)
d <- merge(sub, plots_df, all.x = T)
# length(d$EventID)
# length(unique(d$EventID))
x <- d %>% group_by(EventID, diversity) %>% 
  summarize(n=n(), 
            Latitude = mean(as.numeric(Latitude)), 
            Longitude = mean(as.numeric(Longitude))) %>%
  arrange(desc(diversity))

## create flextable
myft <- flextable(x)

## modify flextable
myft <- theme_vanilla(myft)
myft<- fontsize(myft, size = 8, part = "body")
myft<- fontsize(myft, size = 8, part = "header")
myft <- align(myft, align = "left", part = "header")
myft <- align(myft, align = "left", part = "body")

myft <- set_table_properties(myft, width = .5, layout = "autofit")
myft


#View(x)

setwd("C:/rworking/deepseatools/indata")
write.xlsx(x, '20210824-0_Summary_of_Corals_and_Sponges_Diversity_by_EventID.xlsx') 


## Calculating various Diversity Indices

# H <- diversity(site.sp)
# simp <- diversity(site.sp, "simpson")
# invsimp <- diversity(site.sp, "inv")
# ## Unbiased Simpson (Hurlbert 1971, eq. 5) with rarefy:
# unbias.simp <- rarefy(site.sp, 2) - 1
# ## Fisher alpha
# alpha <- fisher.alpha(site.sp)
# ## Plot all
# pairs(cbind(H, simp, invsimp, unbias.simp, alpha, J), pch="+", col="blue")
# ## Species richness (S) and Pielou's evenness (J):
# S <- specnumber(site.sp) ## rowSums(site.sp > 0) does the same...
# J <- H/log(S)

```

```{r mapping_diversity , echo=FALSE, eval=F}

x$Latitude <- as.numeric(x$Latitude)
x$Longitude <- as.numeric(x$Longitude)

pal <- colorNumeric(
  palette = "Reds",
  domain = x$diversity)


m <- leaflet()
m <- addProviderTiles(m, "Esri.OceanBasemap") #Esri.OceanBasemap, "CartoDB.DarkMatter"
m <- addCircleMarkers(m, data=x,
                      radius=5,
                      weight=0,
                      #fillColor= "green",
                      color = ~pal(diversity),
                      fillOpacity=1,
                      popup = paste("EventID:", x$EventID, "<br>"))
                                   
m

```

## NMDS (Two Highest Density Categories)

### Filter to only species appearing in the top ten list  
```{r filter_to_top_ten, echo=F, eval=T, warning=F, message=F}
##### filter d down to just the species appearing in the top ten ##### 
topten <- sub %>%
  filter(DensityCategory == "3000-9999/km" | 
           DensityCategory == "10000/km") %>% 
 group_by(EventID, ScientificName)%>%
 summarize(individual_count_sum = sum(IndividualCount))%>%
 top_n(10, individual_count_sum) %>% 
  arrange(EventID, desc(individual_count_sum))

z <- sub %>% 
  filter(DensityCategory == "3000-9999/km" | 
           DensityCategory == "10000/km") %>% 
  filter(ScientificName %in% topten$ScientificName)

```

### Setting up depth, temp, and location categories

```{r setdepthclass, echo=FALSE, cache = FALSE} 
##### setting depth categories ##### 
# qplot(as.numeric(z$DepthInMeters))
z$DepthCat[as.numeric(z$DepthInMeters) <= 1800] <- "shallower"
z$DepthCat[as.numeric(z$DepthInMeters) > 1800 & 
             as.numeric(z$DepthInMeters) < 2400] <- "medium"
z$DepthCat[as.numeric(z$DepthInMeters) >= 2400] <- "deeper"
z$DepthCat <- factor(z$DepthCat, levels = c("shallower","medium","deeper"))

##### setting location categories ##### 
# qplot(as.numeric(z$Latitude))
z$loc[as.numeric(z$Latitude) >= 20] <- "north"
z$loc[as.numeric(z$Latitude) < 20] <- "south"

##### longitude cat #####
z$longloc[as.numeric(z$Longitude) > -165] <- "east"
z$longloc[as.numeric(z$Longitude) < -165] <- "west"

##### setting temperature categories #####
# qplot(as.numeric(z$Temperature))
z$temp_cat[as.numeric(z$Temperature) > 2.2] <- "warmer"
z$temp_cat[as.numeric(z$Temperature) >= 1.9 & as.numeric(z$Temperature) <= 2.2] <- "cold"
z$temp_cat[as.numeric(z$Temperature) < 1.9] <- "colder"
z$temp_cat <- factor(z$temp_cat, levels = c("warmer", "cold", "colder"))

##### setting oxygen categories #####
# qplot(as.numeric(z$Oxygen))
z$ox_cat[as.numeric(z$Oxygen) >= 2.6] <- "rich"
z$ox_cat[as.numeric(z$Oxygen) < 2.6] <- "poor"

```

### Making the matrix

First we adjust the abundances by the length of the drive track. This creates a normalized value for 'IndividualCount'. It is calculated as 'IndividualCount' per meter. This value is then entered into the site by species matrix which is used in the NMDS and clustering algorithms.  

``` {r NMDS, echo=FALSE, cache = FALSE, dpi=300, warning=FALSE, message=FALSE, results='hide', cache=T}
## filter if necessary
x <- z %>% 
  filter(
    #DepthCat == 'deep',
    #EventID != "D2-EX1605L1-13"
    #Phylum == "Cnidaria"
    #DensityCategory == "high and higest"
  )

## checking 
# length(table(sub$n_km, useNA = 'always'))

## adjust the IndividualCount values 
x$IndividualCount_z <- x$IndividualCount/x$m_divetrack

## create site X species matrix 
library(vegan)
site.sp <- reshape2::dcast(x, EventID ~ ScientificName, value.var = "IndividualCount_z", fun.aggregate = sum)

## set all NA to 0 because the vegan package can't handle the NA
site.sp[is.na(site.sp)] = 0

## checking resulting table
# View(site.sp)

## export to check
# setwd("C:/rworking/deepseatools/indata")
# write.csv(site.sp, 'site.sp.csv')

# creating a site variable
site.sp$site <- site.sp$EventID

# getting rid of non-needed variables
site.sp <- site.sp %>%
  dplyr::select(-EventID)

# # moving the site variable to the beginning
# col_idx <- grep("site", names(site.sp))
# site.sp <- site.sp[, c(col_idx, (1:ncol(site.sp))[-col_idx])]
# # names(site.sp)

# set the site as the row.names
row.names(site.sp) <- site.sp$site

# remove the site variable
site.sp <- site.sp %>%
  dplyr::select(-site)

# making it a matrix
site.sp <- as.matrix(site.sp)

## checking 
# NMDS

``` 

### NMDS Run

Here are the statistical results from the NMDS run:   

```{r NMDS_run}
##### running NMDS starting with site X species matrix #####
#install.packages("vegan")
library(vegan)
NMDS <- metaMDS(site.sp, distance = "bray", binary = F, k=4, trymax = 30)

##### extracting the site and species scores for use with ggplot2 ##### 
#Using the scores function from vegan to extract the site scoresand convert to a data.frame
site.scores <- as.data.frame(scores(NMDS))  
site.scores$site <- rownames(site.scores)  # create a column of site names, from the rownames of data.scores

# head(site.scores)  #look at the data
# merge information from d (to get categories)
site.scores <- merge(site.scores, x, by.x = 'site', by.y = 'EventID')
site.scores <- site.scores %>% group_by(site) %>% summarise(NMDS1 = NMDS1[1],
                                                            NMDS2 = NMDS2[1],
                                                            DepthCat = DepthCat[1],
                                                            temp_cat = temp_cat[1],
                                                            loc = loc[1],
                                                            ox_cat = ox_cat[1], 
                                                            loc2 = Locality[1])

# extracting species scores
species.scores <- as.data.frame(scores(NMDS, "species"))  #Using the scores function from vegan to extract the species scores and convert to a data.frame
species.scores$species <- rownames(species.scores)  # create a column of species, from the rownames of species.scores
# head(species.scores)  #look at the data
species.scores$spec_code <- 1:(length(species.scores$species))

NMDS

```

### Plot of NMDS with group colors as from depth bin

```{r plotting_NMDS_depth}
##### plotting NMDS ##### 
ggplot() +
  geom_point(data=site.scores,aes(x=NMDS2,y=NMDS1, color=DepthCat), size=5) + # add the point markers
  geom_text(data=site.scores,aes(x=NMDS2,y=NMDS1, label = site), size=4, position = position_nudge(y = .0001)) + # add the text
  #geom_point(data=species.scores,aes(x=NMDS1,y=NMDS2), shape = 3, size=4) + 
  coord_equal() +
  theme_bw(base_size = 15, base_family = "Cambria") +
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank())

```

### Plot of NMDS with group colors from temperature

```{r plotting_NMDS_temp}

ggplot() +
  geom_point(data=site.scores,aes(x=NMDS2,y=NMDS1, color=temp_cat), size=5) + # add the point markers
  geom_text(data=site.scores,aes(x=NMDS2,y=NMDS1, label = site), size=4, position = position_nudge(y = .0001)) + # add the text
  #geom_point(data=species.scores,aes(x=NMDS1,y=NMDS2), shape = 3, size=4) + 
  coord_equal() +
  theme_bw(base_size = 15, base_family = "Cambria") +
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank())

```

### Plot of NMDS with group colors from location 

Location based on latitude break at X.  

```{r}

ggplot() +
  geom_point(data=site.scores,aes(x=NMDS2,y=NMDS1, color=loc), size=5) + # add the point markers
  geom_text(data=site.scores,aes(x=NMDS2,y=NMDS1, label = site), size=4, position = position_nudge(y = .0001)) + # add the text
  #geom_point(data=species.scores,aes(x=NMDS1,y=NMDS2), shape = 3, size=4) + 
  coord_equal() +
  theme_bw(base_size = 15, base_family = "Cambria") +
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank())

```

### Plot of NMDS with group colors from oxygen values 


```{r}

ggplot() +
  geom_point(data=site.scores,aes(x=NMDS2,y=NMDS1, color=ox_cat), size=5) + # add the point markers
  geom_text(data=site.scores,aes(x=NMDS2,y=NMDS1, label = site), size=4, position = position_nudge(y = .0001)) + # add the text
  #geom_point(data=species.scores,aes(x=NMDS1,y=NMDS2), shape = 3, size=4) + 
  coord_equal() +
  theme_bw(base_size = 15, base_family = "Cambria") +
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank())

```


### Plot of NMDS species scores

Labels are are species code (by number). See table below for species code.   

```{r}

ggplot() +
  #geom_point(data=site.scores,aes(x=NMDS2,y=NMDS1, color=loc), size=5) + # add the point markers
  geom_text(data=species.scores,aes(x=NMDS2,y=NMDS1, label = spec_code), size=4, position = position_nudge(y = .0001)) + # add the text
  geom_point(data=species.scores,aes(x=NMDS2,y=NMDS1), shape = 3, size=4) + 
  coord_equal() +
  theme_bw(base_size = 15, base_family = "Cambria") +
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank())

```

### Table of species codes (key to the lables in the species score NMDS plot above)

```{r spec_codes}

x <- species.scores %>% dplyr::select(spec_code)

## convert row names to column
x <- tibble::rownames_to_column(x, "ScientificName")
## create flextable
myft <- flextable(x)

## modify flextable
myft <- theme_vanilla(myft)
myft<- fontsize(myft, size = 8, part = "body")
myft<- fontsize(myft, size = 8, part = "header")
myft <- align(myft, align = "left", part = "header")
myft <- align(myft, align = "left", part = "body")

myft <- set_table_properties(myft, width = .5, layout = "autofit")
myft

```

### Clustering 

```{r}
##### create different distance matrices ##### 
## Bray-Curtis as binary (presence/absence) of species per equation above
# sp_bray_binary = vegdist(site.sp, method='bray', binary=T)

## Bray-Curtis as amount (density) of species
sp_bray_transformed = vegdist(site.sp, method='bray')

## transformed Bray-Curtis as amount (density) of species
#sp_bray_transformed = vegdist(wisconsin(sqrt(site.sp)), method='bray')

##### create different clusters ######

fit <- hclust(sp_bray_transformed)
plot(fit, main = "", sub = "", xlab="Sites", axes = T, ylab = "Ecological Distance (Bray-Curtis - Wisconsin Transformed)", hang = -1) # display dendogram
groups <- cutree(fit, k=5) # cut tree into 5 clusters
# draw dendogram with red borders around the 5 clusters
rect.hclust(fit, k=5, border="red")

# fit <- hclust(sp_bray_binary)
# plot(fit, main = "", sub = "", xlab="Sites", axes = T, ylab = "Ecological Distance (Sorensen)", hang = -1) # display dendogram
# groups <- cutree(fit, k=5) # cut tree into 5 clusters
# # draw dendogram with red borders around the 5 clusters
# rect.hclust(fit, k=5, border="red")
# 
# fit <- hclust(sp_bray)
# plot(fit, main = "", sub = "", xlab="Sites", axes = T, ylab = "Ecological Distance (Bray-Curtis)", hang = -1) # display dendogram
# groups <- cutree(fit, k=5) # cut tree into 5 clusters
# # draw dendogram with red borders around the 5 clusters
# rect.hclust(fit, k=5, border="red")

```

### Diversity estimates.

##### Table: Simpson diversity per EventID
All taxa (both coral and sponges at all taxonomic ranks included in estimates).

```{r diversity, echo=F}

diversity <- diversity(site.sp, index = "simpson") #makes an object
#that the diversity values #are written to.
# summary(plots) #gives summary statistics for the plots
# median(plots) #gives the median
# mean(plots) #gives the mean
# class(plots)
# max(plots[1:30])
# # View(plots)

# creating plots
plots_df <- data.frame(diversity)

# View(plots_df)
plots_df$EventID <- row.names(plots_df)

# names(d)
d <- merge(sub, plots_df, all.x = T)
# length(d$EventID)
# length(unique(d$EventID))
x <- d %>% group_by(EventID, diversity) %>% 
  summarize(n=n(), 
            Latitude = mean(as.numeric(Latitude)), 
            Longitude = mean(as.numeric(Longitude))) %>%
  arrange(desc(diversity))

## create flextable
myft <- flextable(x)

## modify flextable
myft <- theme_vanilla(myft)
myft<- fontsize(myft, size = 8, part = "body")
myft<- fontsize(myft, size = 8, part = "header")
myft <- align(myft, align = "left", part = "header")
myft <- align(myft, align = "left", part = "body")

myft <- set_table_properties(myft, width = .5, layout = "autofit")
myft


#View(x)

setwd("C:/rworking/deepseatools/indata")
write.xlsx(x, '20210824-0_Summary_of_Corals_and_Sponges_Diversity_by_EventID.xlsx') 


## Calculating various Diversity Indices

# H <- diversity(site.sp)
# simp <- diversity(site.sp, "simpson")
# invsimp <- diversity(site.sp, "inv")
# ## Unbiased Simpson (Hurlbert 1971, eq. 5) with rarefy:
# unbias.simp <- rarefy(site.sp, 2) - 1
# ## Fisher alpha
# alpha <- fisher.alpha(site.sp)
# ## Plot all
# pairs(cbind(H, simp, invsimp, unbias.simp, alpha, J), pch="+", col="blue")
# ## Species richness (S) and Pielou's evenness (J):
# S <- specnumber(site.sp) ## rowSums(site.sp > 0) does the same...
# J <- H/log(S)

```

```{r mapping_diversity_2 , echo=FALSE, eval=F}

x$Latitude <- as.numeric(x$Latitude)
x$Longitude <- as.numeric(x$Longitude)

pal <- colorNumeric(
  palette = "Reds",
  domain = x$diversity)


m <- leaflet()
m <- addProviderTiles(m, "Esri.OceanBasemap") #Esri.OceanBasemap, "CartoDB.DarkMatter"
m <- addCircleMarkers(m, data=x,
                      radius=5,
                      weight=0,
                      #fillColor= "green",
                      color = ~pal(diversity),
                      fillOpacity=1,
                      popup = paste("EventID:", x$EventID, "<br>"))
                                   
m

```
