---
title: "DSBS16_HI_Analysis"
author: "Robert McGuinn"
date: "`r Sys.Date()`"
output: 
  word_document:
     reference_docx: template.docx
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, 
                      eval = T)
```

```{r packages}
library(tidyr)
library(openxlsx)
library(googlesheets4)
library(googledrive)

```

# Main Project Folder

## Yo

The main project folder for this analysis is on Google Drive: [LINK](https://drive.google.com/drive/folders/1ToTK_v7NF7ZPdwXQwD1hzCwuVZCZRH_d)

## Input data sets

Datasets used in this analysis were acquired on 2021-08-18 from Tom Hourigan. The coral and sponge species occurrence records are from a subset of dives in the Hawaiian region (including Johnston Atoll and the Musician Seamounts). This original dataset is shared on Google Drive here: [LINK]()


```{r data_intake}
##### bringing in dataframe from xlsx stored on drive ####

## file path [MANUAL]
filename <- "20210817_Hawaii-HiDensity1500-2500m"

## create a list of files (or single file) that meets title query [MANUAL]
x <- drive_find(q = "name contains '20210817_Hawaii-HiDensity1500-2500m'")

## browse to it
# x %>% drive_browse()

# getting the id as a character string
y <- x$id

# this downloads the file to the specified path
dl <- drive_download(as_id(y),
                     path = paste("C:/rworking/deepseatools/indata/",
                                  filename,
                                  ".xlsx", 
                                  sep = ""),
                     overwrite = TRUE)

## extract just the file of interest from the zip file to a dataframe called 'sub'
sub <- read.xlsx(dl$local_path)

## names 
names(sub)
```

## Summary 

