---
title: "Area of Interest (AOI) Analysis for the BOEM Virginia Minerals Areas RFI"
author: "NMFS: Deep Sea Coral Research and Technology Program"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: flatly
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
## set global chunk options
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

## load required modern packages
library(sf)           
library(dplyr)        
library(googledrive) 
library(knitr)
library(ggplot2)
library(rnaturalearth)
library(ggspatial) 

```

```{r authenicate}
##### authenticate #####
# Replace with the email tied to your Google Drive
drive_auth(email = "robert.mcguinn@noaa.gov")

```

```{r load_aoi}
## fetch AOI Geodatabase from Google Drive
drive_id <- "10jN0aiw2l8gks7JGmcwqd2w6CXNOT8oA"

## download the zip file to a temporary location
temp_zip <- tempfile(fileext = ".zip")
drive_download(as_id(drive_id), path = temp_zip, overwrite = TRUE)

## unzip to a temporary directory
temp_dir <- tempdir()
unzip(temp_zip, exdir = temp_dir)

## find the .gdb folder
## use list.dirs() because a .gdb is a directory, not a standalone file
all_dirs <- list.dirs(temp_dir, recursive = TRUE)
gdb_path <- all_dirs[grepl("\\.gdb$", all_dirs)][1]

## safety check for GDB presence
if (is.na(gdb_path)) {
  stop("No .gdb folder was found in the downloaded zip. Please check the zip contents.")
}

## read the geodatabase into an sf object
aoi_polygon <- st_read(gdb_path, quiet = TRUE)
```

```{r sf}
## convert 'filt' to a spatial 'sf' object 
filt_sf <- st_as_sf(filt, 
                    coords = c("Longitude", "Latitude"), 
                    crs = 4326,  # WGS84 standard coordinate system
                    remove = FALSE)

## esure the coordinate reference systems (CRS) match before intersection
aoi_polygon <- st_transform(aoi_polygon, crs = st_crs(filt_sf))

## Perform the AOI intersection (keep only occurrences inside the AOI)
filt_aoi <- st_intersection(filt_sf, aoi_polygon)
```
**Point(s) of contact:**     
- Robert McGuinn (robert.mcguinn@noaa.gov)     
- Savannah Goode (goodesavannah@gmail.com)     
- Arvind Shantharam (arvind.shantharam@noaa.gov)     

# Introduction
This analysis was performed by the NMFS Deep-sea Coral Research and Technology Program (DSCRTP) referencing the BOEM critical mineral area off the coast of Virginia. See the following link for more information on the Request for Information (RFI): https://www.boem.gov/marine-minerals/critical-minerals/virginia-mineral-activities. 

# Biological occurrence data summary (presence only)
Summary of presence-only biological occurrences from NOAA's National Database for Deep-sea Corals and Sponges. For more information on the database see: https://deepseacoraldata.noaa.gov/data. **Note: Absence of biological occurrences in this database does not indicate actual absence in the real world.**

**Cite the database as follows:**  
National Oceanic and Atmospheric Administration (NOAA), Deep Sea Coral Research and Technology Program (DSCRTP), 2026. Observations of Deep-Sea Coral and Sponge Occurrences from NOAA's National Database for Deep-sea Corals and Sponges, 1842-Present, version `r unique(filt$DatabaseVersion)`. (NCEI Accession 0145037). NOAA National Centers for Environmental Information. Archived data: https://www.ncei.noaa.gov/archive/accession/0145037. Access portal: https://www.ncei.noaa.gov/maps/deep-sea-corals-portal/.  Accessed [date as YYYY-MM-DD].

Note: Character strings in [bold brackets] should be overwritten with appropriate values and leave out the brackets.  

## Survey summary table (within the AOI)
```{r expedition_summary}
expedition_summary <- filt_aoi %>%
  st_drop_geometry() %>%
  group_by(SurveyID, Vessel, SamplingEquipment, DatasetID, ObservationYear) %>%
  summarize(
    TotalOccurrences = n(),
    .groups = "drop"
  ) %>%
  arrange(ObservationYear, SurveyID)

kable(expedition_summary)
```

## Taxonomic summary table (within the AOI)
```{r taxonomic_summary}
depth_summary <- filt_aoi %>%
  st_drop_geometry() %>% 
  group_by(ScientificName, VernacularNameCategory) %>%
  summarize(
    TotalOccurrences = n(),
    MinDepth_m = min(DepthInMeters, na.rm = TRUE),
    MaxDepth_m = max(DepthInMeters, na.rm = TRUE),
    MeanDepth_m = round(mean(DepthInMeters, na.rm = TRUE), 1),
    .groups = "drop"
  ) %>%
  arrange(desc(TotalOccurrences))

kable(depth_summary)
```

## Map of occurrences (within the AOI)
```{r static-map, fig.width=8, fig.height=6}
##  the US border geometry
us_border <- ne_countries(country = "United States of America", 
                          returnclass = "sf", 
                          scale = "medium")

## ensure the US border CRS matches our AOI
us_border <- st_transform(us_border, crs = st_crs(aoi_polygon))

## extract the bounding box of the AOI and set a wider buffer
aoi_bbox <- st_bbox(aoi_polygon)
zoom_buffer <- 2.0  

## build the map using ggplot2
ggplot() +
  
  geom_sf(data = us_border, fill = "gray95", color = "gray50") +
  geom_sf(data = filt_sf, color = "gray70", size = 1.0, alpha = 0.5) +
  geom_sf(data = aoi_polygon, fill = NA, color = "red", linewidth = 1.2) +
  geom_sf(data = filt_aoi, color = "dodgerblue", size = 1.5, alpha = 0.7) +
  coord_sf(
    xlim = c(aoi_bbox["xmin"] - zoom_buffer, aoi_bbox["xmax"] + zoom_buffer), 
    ylim = c(aoi_bbox["ymin"] - zoom_buffer, aoi_bbox["ymax"] + zoom_buffer),
    expand = FALSE
  ) +
  annotation_north_arrow(
    location = "tr", 
    which_north = "true", 
    pad_x = unit(0.2, "in"), 
    pad_y = unit(0.2, "in"),
    style = north_arrow_fancy_orienteering
  ) +
  annotation_scale(
    location = "bl", 
    width_hint = 0.3,
    text_cex = 0.9
  ) +
  theme_minimal() +
  labs(
    title = "Biological Occurrences: Inside vs. Outside AOI",
    subtitle = "Red-AOI border; Blue- Intersecting AOI; Gray - outside AOI",
    x = "Longitude",
    y = "Latitude"
  ) +
  theme(
    plot.title = element_text(face = "bold"),
    panel.grid.major = element_line(color = "gray90", linetype = "dashed")
  )

```

```{r export_AOI_GeoJSON}

# Export the AOI polygon to a GeoJSON file
st_write(
  obj = aoi_polygon, 
  dsn = "aoi_boundary.geojson", 
  driver = "GeoJSON", 
  delete_dsn = TRUE, # Safely overwrites the file if you run the script again
  quiet = TRUE       # Prevents console messages from printing in your HTML report
)
```

# Habitat suitability model summary 

```{r import-models}
# 1. Define the local path to your zip file
zip_path <- "../indata/DSC_PredictiveModeling_US_Northeast_MidAtlantic.zip"

# Safety check: ensure the file exists before trying to unzip
if (!file.exists(zip_path)) {
  stop("The zip file was not found. Please check the 'indata' folder and path.")
}

# 2. Unzip to a temporary directory
temp_dir <- tempdir()
unzip(zip_path, exdir = temp_dir)

# 3. Find all .shp files inside the extracted folder (including subfolders)
shp_files <- list.files(temp_dir, pattern = "\\.shp$", full.names = TRUE, recursive = TRUE)

if (length(shp_files) == 0) {
  stop("No shapefiles (.shp) were found inside the zip file.")
}

# 4. Read each shapefile, transform its CRS to match the AOI, and store in a list
predictive_models <- lapply(shp_files, function(file) {
  # Read the shapefile
  temp_sf <- st_read(file, quiet = TRUE)
  
  # Transform CRS to match aoi_polygon
  temp_sf <- st_transform(temp_sf, crs = st_crs(aoi_polygon))
  
  return(temp_sf)
})

# 5. Name the items in the list based on their original filenames
names(predictive_models) <- tools::file_path_sans_ext(basename(shp_files))
```

The data used in the following section are based on the habitat suitability models created in 2013 by Matt Poti and Arliss Winship. You can find more documentation about these models at: https://www.ncei.noaa.gov/archive/accession/014592 

**Cite the data as:**
Cite as: Kinlan, Brian; Poti, Matthew; Dorfman, Dan; Caldow, Chris; Drohan, Amy; Packer, David B.; Nizinski, Martha S. (2016). Model output for deep-sea coral habitat suitability in the U.S. North and Mid-Atlantic from 2013 (NCEI Accession 0145923). [indicate subset used]. NOAA National Centers for Environmental Information. Dataset. https://www.ncei.noaa.gov/archive/accession/0145923. Accessed [date].

**The following codes are used to designate the different taxa specific model runs:**
```{r taxon-code-table, echo=FALSE}
# Create the table as a dataframe
taxon_codes <- data.frame(
  Group = c("1", "1a", "1b", "2", "2a", "2b", "3", "3a", "3b"),
  Description = c(
    "Order Alcyonacea",
    "Gorgonian Alcyonacea (Suborders Calcaxonia, Holaxonia, Scleraxonia)",
    "Non-Gorgonian Alcyonacea (Suborders Alcyoniina, Stolonifera)",
    "Order Scleractinia",
    "Family Caryophylliidae",
    "Family Flabellidae",
    "Order Pennatulacea",
    "Suborder Sessiliflorae",
    "Suborder Subsessiliflorae"
  ),
  `Code name` = c(
    "ALCY", "ALCY-GORG", "ALCY-NONGORG", 
    "SCLER", "SCLER-CARYO", "SCLER-FLAB", 
    "PENN", "PENN-SESS", "PENN-SUBSESS"
  ),
  check.names = FALSE # This allows the space in 'Code name'
)

# Print the formatted table using kable
kable(taxon_codes)
```

**Habitat suitability class definitions (GRIDCODE):**
```{r suitability-table, echo=FALSE}
# Create the table as a dataframe
suitability_table <- data.frame(
  `Suitability (GRIDCODE)` = c(0, 1, 2, 5, 10),
  Description = c(
    "Low Suitability", 
    "Medium-Low Suitability", 
    "Medium Suitability", 
    "High Suitability", 
    "Very High Suitability"
  ),
  check.names = FALSE
)

# Print the formatted table with centered numbers
kable(
  suitability_table, 
  align = c("c", "l") # 'c' centers the first column, 'l' left-aligns the second
)
```

## Tabulation of the area of each suitability class within the AOI (in hectares)
```{r calculate-area-intersected}
# Load the units package for easy area conversions
library(units)

# Turn off strict spherical geometry
sf_use_s2(FALSE)

# Loop through each model in the list, intersect with AOI, calculate area, and summarize
area_summaries_aoi <- lapply(names(predictive_models), function(model_name) {
  
  # Extract the individual sf object
  sf_obj <- predictive_models[[model_name]]
  
  # 1. Cookie-cutter the model to the AOI boundaries
  # suppressWarnings prevents console spam about spatial attributes
  sf_intersected <- suppressWarnings(st_intersection(sf_obj, aoi_polygon))
  
  # 2. Safety check: If this specific model doesn't overlap the AOI at all, skip it
  if (nrow(sf_intersected) == 0) {
    return(NULL)
  }
  
  # 3. Calculate the area in hectares for ONLY the intersected geometries
  sf_intersected <- sf_intersected %>%
    mutate(Area_ha = as.numeric(set_units(st_area(geometry), "hectares")))
  
  # 4. Group by the suitability class (GRIDCODE) and sum the areas
  summary_table <- sf_intersected %>%
    st_drop_geometry() %>%
    group_by(GRIDCODE) %>% 
    summarize(
      Total_Area_Hectares = round(sum(Area_ha, na.rm = TRUE), 2),
      .groups = "drop"
    ) %>%
    mutate(Model_Name = model_name) # Add the model name back in
  
  return(summary_table)
})

# Combine the list of summaries into one master dataframe
# bind_rows safely ignores the NULLs from any models that were skipped
master_area_summary_aoi <- bind_rows(area_summaries_aoi)

# Reorder columns logically
master_area_summary_aoi <- master_area_summary_aoi %>%
  select(Model_Name, GRIDCODE, Total_Area_Hectares)

# Output a clean HTML table
kable(master_area_summary_aoi)

```

## Plot of the area of each suitability class within the AOI (in hectares)
```{r plot-area-grid, fig.width=10, fig.height=8}
library(ggplot2)

# Create the faceted bar chart
ggplot(master_area_summary_aoi, aes(x = as.factor(GRIDCODE), y = Total_Area_Hectares, fill = as.factor(GRIDCODE))) +
  # Use geom_col for pre-summarized data (it plots exactly the Y values given)
  geom_col(color = "black", alpha = 0.8) +
  
  # Split into a grid based on Model_Name
  # scales = "free_y" allows each plot to have its own Y-axis range. 
  # If you want them all to share the exact same scale for direct visual comparison, remove that argument!
  facet_wrap(~ Model_Name, scales = "free_y") +
  
  # Clean, modern theme
  theme_minimal() +
  labs(
    x = "Suitability Class (GRIDCODE)",
    y = "Area (Hectares)"
  ) +
  
  # Format the Y-axis to use commas instead of scientific notation
  scale_y_continuous(labels = function(x) format(x, scientific = FALSE, big.mark = ",")) +
  
  # Customize text elements
  theme(
    legend.position = "none", # Hide the legend since the x-axis already shows the GRIDCODE
    strip.text = element_text(face = "bold", size = 10), # Makes the model names bold
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    panel.grid.minor = element_blank()
  )

```

## Plot of the area of each suitability class as a percentage of the total AOI area
Total AOI Area = `r round(aoi_total_area_ha, 0)`
```{r plot-area-grid-percent, fig.width=10, fig.height=8}
library(ggplot2)
library(units)
library(dplyr)

# 1. Calculate the total area of the AOI in hectares
# We use sum() just in case your AOI polygon is made of multiple smaller multipolygon pieces
aoi_total_area_ha <- sum(as.numeric(set_units(st_area(st_make_valid(aoi_polygon)), "hectares")))

# 2. Add the percentage column to your summary dataframe
master_area_summary_aoi <- master_area_summary_aoi %>%
  mutate(Percent_of_AOI = (Total_Area_Hectares / aoi_total_area_ha) * 100)

# 3. Create the faceted bar chart using Percent_of_AOI
ggplot(master_area_summary_aoi, aes(x = as.factor(GRIDCODE), y = Percent_of_AOI, fill = as.factor(GRIDCODE))) +
  
  geom_col(color = "black", alpha = 0.8) +
  
  # facet_wrap splits the charts by Model_Name
  # keeping scales="free_y" so small models are visible, but you can remove it to standardize axes
  facet_wrap(~ Model_Name, scales = "free_y") +
  
  theme_minimal() +
  labs(
    x = "Suitability Class (GRIDCODE)",
    y = "Percent of AOI (%)"
  ) +
  
  # Format the Y-axis to dynamically show the % sign and round to 1 decimal
  scale_y_continuous(labels = function(x) paste0(round(x, 1), "%")) +
  
  theme(
    legend.position = "none", 
    strip.text = element_text(face = "bold", size = 10), 
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    panel.grid.minor = element_blank()
  )
```

## Maps of the suitability classes for each taxa group 
```{r plot-model-maps, fig.width=8, fig.height=6}
# Ensure ggplot2 and viridis are loaded (viridis is great for discrete mapping)
library(ggplot2)
# install.packages('viridis')
library(viridis) 


# Re-establish bounding box and zoom buffer in case they were cleared
aoi_bbox <- st_bbox(aoi_polygon)
zoom_buffer <- 2.0  

# Loop through every model and print a map
for (model_name in names(predictive_models)) {
  
  # Extract the individual model sf object
  model_sf <- predictive_models[[model_name]]
  
  # Build the map
  p <- ggplot() +
    # Add the US Border background
    geom_sf(data = us_border, fill = "gray95", color = "gray50") +
    
    # Add the Model Output (colored by GRIDCODE, borders hidden)
    geom_sf(data = model_sf, aes(fill = as.factor(GRIDCODE)), color = NA, alpha = 0.8) +
    
    # Add the AOI Polygon on top (thick red border, no fill)
    geom_sf(data = aoi_polygon, fill = NA, color = "red", linewidth = 1.2) +
    
    # Zoom the map using the AOI bounding box
    coord_sf(
      xlim = c(aoi_bbox["xmin"] - zoom_buffer, aoi_bbox["xmax"] + zoom_buffer), 
      ylim = c(aoi_bbox["ymin"] - zoom_buffer, aoi_bbox["ymax"] + zoom_buffer),
      expand = FALSE
    ) +
    
    # Apply a modern, colorblind-friendly palette to the suitability classes
    scale_fill_viridis_d(option = "mako", direction = -1, name = "Suitability\n(GRIDCODE)") +
    
    # Add North Arrow
    annotation_north_arrow(
      location = "tr", which_north = "true", 
      pad_x = unit(0.2, "in"), pad_y = unit(0.2, "in"),
      style = north_arrow_fancy_orienteering
    ) +
    
    # Add Scale Bar
    annotation_scale(
      location = "bl", width_hint = 0.3, text_cex = 0.9
    ) +
    
    # Clean theme and dynamic labels
    theme_minimal() +
    labs(
      title = paste("Predictive Model:", model_name),
      subtitle = "Red border indicates the AOI; Colors indicate suitability class",
      x = "Longitude",
      y = "Latitude"
    ) +
    theme(
      plot.title = element_text(face = "bold"),
      panel.grid.major = element_line(color = "gray90", linetype = "dashed"),
      legend.position = "right"
    )
  
  # EXPLICITLY print the plot so it renders in the HTML loop
  print(p)
}
```

