---
title: "highlight_images_eg_edit"
author: "Robert McGuinn"
date: "4/29/2021"
output: html_document
---

```{r}
##### load packages #####
library(tidyverse)
library(knitr)
library(flextable)
library(sp)
library(rgdal)
library(maps)
library(extrafont)
library(googlesheets4)
library(leaflet)
library(RColorBrewer)
library(sf)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

The following imagery is linked to the National Database and was selected by Elizabeth Gugliotti in 2019. It represents the best imagery holdings from the region. 

See the following Google Drive location for all images: 

https://drive.google.com/drive/folders/1LoSALOTqH3OZM6Aml8Hc_2E11fCDqcBH?usp=sharing #####

# Query Google Drive location to get the file list
```{r}

```
# Strip off the CatalogNumber prefixes
# Export the list of CatalogNumbers
# Post to RedMine (manual)
```{r}

```
# Tips and Tricks
# View single image from NDB in a default browser
This code will allow you to browse to a single image within the National Database.  

```{r}
##### best images thing ##### 
## see E. Gugliotti work https://drive.google.com/drive/folders/1LoSALOTqH3OZM6Aml8Hc_2E11fCDqcBH?usp=sharing #####
filt %>% filter(CatalogNumber == '530910') %>% pull(ImageURL) %>% browseURL()

```


# Print image to this page 
