---
title: "plot_db_growth_through_time"
author: "Robert McGuinn"
date: '`r format(Sys.Date(), "%Y-%m-%d")`'
output: html_document
editor_options: 
  chunk_output_type: inline
---
# packages

``` {r install_packages, , echo=T, warning=FALSE, message=FALSE}
library(tidyverse)
library(rmarkdown)
library(knitr)
library(flextable)
library(googlesheets)
library(googledrive)
library(openxlsx)
library(scales)
library(extrafont)
library(RColorBrewer)

```

# load NDB

``` {r data_intake, echo=T, warning=FALSE, message=FALSE, cache=TRUE}
##### input: latest version of NDB #####
setwd("C:/rworking/deepseatools/indata")
indata<-read.csv("DSCRTP_NatDB_20191217-0.csv", header = T)
filt <- indata %>%
  filter(Flag == "0")

```

# clean NDB

```{r clean, echo=T, warning=FALSE, message=FALSE}
##### data improvements (from version DSCRTP_NatDB_20191217-0)#####

filt <- filt %>% mutate(DataProvider =
                          ifelse(DataProvider == "California Academy of Sciences (CAS)",
                                              'California Academy of Sciences',
                                              as.character(DataProvider)))

filt <- filt %>% mutate(DataProvider =
                          ifelse(DataProvider == "NOAA, Office of Exploration and Research | Hawai’i Undersea Research Laboratory, University of Hawai’i", "NOAA, Office of Exploration and Research | Hawai'i Undersea Research Laboratory, University of Hawai'i", 
                                 as.character(DataProvider)))



##### bring in patch file. Raw from David Sallis on 2020-03-10 #####
## edited by RPMcGuinn on 2020-03-10 

yo <- read.xlsx('../indata/20200310-0_entrydate_fix_RPMcGuinn.xlsx')
yo <- yo %>% dplyr::select(-EntryDate)

## check

# setdiff(yo$DatasetID, filt$DatasetID)
# filt %>% filter(grepl('NOAA_CT-13', DatasetID)) %>% pull(DatasetID) %>% unique()

##### mutate step to add missing EntryDate values #####

## merge corrections using DatasetID as a key field
filt <- merge(filt, yo, all.x = TRUE)

## create new variable
filt <- filt %>% 
  mutate(EntryDate2 = if_else(is.na(as.character(EntryDate)), 
                              EntryDate_txt, 
                              as.character(EntryDate)))

## put corrected values into the correct position 
filt$EntryDate <- filt$EntryDate2

## get rid of the extra EntryDate field
filt <- filt %>% select(-EntryDate2, -EntryDate_txt)

```

# merge datasetID key


``` {r merge_key, eval=T}
x <- drive_find(q = "name contains '20191217-0_DatasetID_Key_DSCRTP'")

# # browse to it
# x %>% drive_browse()

# getting the id as a character string
y <- x$id

# this downloads the file to the specified path
dl <- drive_download(as_id(y), path = "C:/rworking/deepseatools/indata/20191217-0_DatasetID_Key_DSCRTP.xlsx", overwrite = TRUE)

# read the file into R as a data frame
key <- read.xlsx(dl$local_path)
filt <- merge(filt, key, all.x = TRUE)
```

# load NDB_trajectory

```{r}

# read the file into R as a data frame from xl
traj <- read.xlsx('../indata/NDB_trajectory.xlsx')

## checking
# names(traj)
# View(traj)
# plot(traj$totrecords)
# plot(traj$corrected)
# plot(traj$quarter)
# plot(traj %>% filter(is.na(totrecords) == F) %>% pull(totrecords))
# plot(traj$totrecord)
# plot(traj$totrecords,as.Date(traj$text_date_posted))
# plot(as.Date(traj$text_date_posted), traj$totrecords)

```

# set up knitr options

``` {r setup_opts, echo=T, warning=FALSE, message=FALSE}

##### knitr options #####
knitr::opts_chunk$set(
  fig.width = 8.0, 
  fig.asp = .5, 
  warning = FALSE,
  message=FALSE,
  cache = TRUE,
  dpi = 300, 
  dev = c('png'), cache.lazy = FALSE, 
  fig.path="C:/rworking/deepseatools/reports/figures/") 

##### setting scientific notation and digits options
options(scipen=10000)
options(digits = 3)

```

# prep the data 

```{r new_and_corrected, echo=F, eval=T, warning=F, message=F}

d <- traj %>%
 pivot_longer(cols = c(totrecords, newrecords, corrected), names_to = "type", values_to = "count")

d2 <- d %>% group_by(fy,type) %>% 
  summarize(sum = sum(count, na.rm = T), max = max(count, na.rm = T)) %>% 
  mutate(merge = ifelse(type == "totrecords", max, sum))

d3 <- d2 %>% filter(type != "totrecords") %>% 
  group_by(fy) %>% 
  summarize(sum2 = sum(sum, na.rm = T))

d4 <- merge(d2,d3, all.x = T)

d5 <- d4 %>%  mutate(sum3 = ifelse(type == "totrecords", 
                                   merge - sum2, 
                                   merge))%>% 
  filter(fy != 2015) %>% 
  filter(fy != 2020)

d5$type <- factor(d5$type)
levels(d5$type)
levels(d5$type) <- c("Improved", "New", "Un-touched")

```

# stacked bar chart

```{r}

cl <- colors(distinct = TRUE)
set.seed(13) # to set random generator seed
mycols <- sample(cl, 3)

ggplot(d5, aes(fill=type, y=sum3, x=fy)) + 
    geom_bar(position="stack", stat="identity") +
scale_colour_manual(values = mycols, 
                    aesthetics = c("colour", "fill"),
                    drop = FALSE, 
                    guide=guide_legend(reverse=F)) + 
  scale_y_continuous(labels = comma) +
  theme_bw() +
  theme(text=element_text(size=14,  family="Cambria")) +
  theme(axis.text.x = element_text(angle = 90)) +
  theme(legend.title = element_blank()) +
  theme(axis.title.x = element_text(vjust=-2.5)) +
  theme(plot.margin = unit(c(1,1,1,1), "cm"))+
  ggtitle("Proportion of new and improved records") + 
  theme(plot.title = element_text(size=15)) +
  theme(axis.title.y = element_text(vjust=5)) +
  xlab('Fiscal Year') +
  ylab('Number of Records')


```

# save the figure

```{r}

ggsave(paste("c:/rworking/deepseatools/images/", "20200311-0_new_improved_RPMcGuinn", ".png", sep = ''), 
         width = 6.5, 
         height = 5.57, 
         units = "in")

```

# change FishCouncilRegion levels for charting purposes

``` {r echo=TRUE, dpi=300}

library(scales)
# ordering factors

filt2 <- filt %>%
  #filter(as.Date(EntryDate) > as.Date('2010-01-01')) %>% 
  dplyr::mutate(FishCouncilRegion = replace_na(as.character(FishCouncilRegion), "International")) %>% 
  dplyr::mutate(Phylum=recode(Phylum, 
                         'Cnidaria' = "Corals",
                         'Porifera' = "Sponges"))
```

# create the growth trajectory figures for each region (saved to file)

``` {r GrowthThroughTime_charts, echo=TRUE, dpi=300, eval = F}
    
## set options to remove scientific notation
options(scipen=10000)

## create a vector of FishCouncilRegions
vect_region <-  c("Caribbean", "Gulf of Mexico", "South Atlantic", "Mid-Atlantic", "New England", "North Pacific", "Pacific","Western Pacific", "International")

for (region in vect_region){
  
x <- filt2 %>% filter (Phylum == "Corals", FishCouncilRegion == region) %>%
  mutate(EntryDate = replace_na(as.character(EntryDate), '2019-12-18')) %>% 
  mutate(EntryDate = as.factor(EntryDate)) %>% 
  mutate(as.Date(EntryDate))

y <- filt2 %>% filter (Phylum == "Sponges", FishCouncilRegion == region) %>%
  mutate(EntryDate = replace_na(as.character(EntryDate), '2019-12-18')) %>% 
  mutate(EntryDate = as.factor(EntryDate)) %>% 
  mutate(as.Date(EntryDate))
 
ggplot(filt2,aes(x=as.Date(EntryDate), color=Phylum)) +
  stat_bin(data = x,
           aes(y=cumsum(..count..)), 
           geom="line", 
           binwidth = 2, 
           size = 1) +
  stat_bin(data = y,
           aes(y=cumsum(..count..)),
           geom="line", 
           binwidth = 2, 
           size = 1) +
  ylab("Cumulative Number of Records") + 
  xlab("Year Record Added") + 
  xlim(as.Date('2012-01-01'), as.Date('2020-01-01')) +
  scale_y_continuous(labels = comma) +
  guides(fill=FALSE) + 
  theme_bw() +
  theme(text=element_text(size=14,  family="Cambria")) +
  theme(axis.text.x = element_text(angle = 90)) +
  theme(legend.title = element_blank()) +
  theme(axis.title.x = element_text(vjust=-2.5)) +
  theme(plot.margin = unit(c(1,1,1,1), "cm"))+
  scale_color_manual(values = brewer.pal(12, "Paired")[c(10,9)])+
  ggtitle(region)

ggsave(paste("c:/rworking/deepseatools/images/", "20200311-1_", region, ".png", sep = ''), 
       width = 6.5, 
       height = 5.57, 
       units = "in")
  
}

```

# summarize EntryDate by Year by Region (save to Excel file) 

```{r}

filt2$EntryDateYear <- format(as.Date(filt$EntryDate),"%Y")

sum <- filt2 %>% group_by(FishCouncilRegion, EntryDateYear) %>% summarize(n = n()) 

sum2 <- sum %>% group_by(FishCouncilRegion) %>% 
  mutate(n_cumulative = cumsum(n))

# View(sum2)
setwd("C:/rworking/deepseatools/indata")
write.xlsx(sum2, '20200310-0_summary_table_growth_by_region.xlsx')

```

# summarize Entrydate by Year by Region, and by Phylum (save to Excel file)

```{r}
filt2$EntryDateYear <- format(as.Date(filt$EntryDate),"%Y") 
sum3 <- filt2 %>% filter(is.na(Phylum) == F) %>% 
  group_by(FishCouncilRegion, EntryDateYear, Phylum) %>% 
  summarize(n = n()) %>% arrange(FishCouncilRegion, FishCouncilRegion, Phylum, EntryDateYear)
   
sum4 <- sum3 %>% group_by(FishCouncilRegion, Phylum) %>% 
  mutate(n_cumulative = cumsum(n)) 

# View(sum2)
setwd("C:/rworking/deepseatools/indata")
write.xlsx(sum4, '20200310-0_summary_table_growth_by_region_by_Phylum_RPMcGuinn.xlsx')
```

# filtering for just Gulf of Mexico latest records

```{r}
## checking for GOMEX records

x <- filt2 %>% filter(
  #as.Date(EntryDate) > '2019-04-18', 
  FishCouncilRegion == 'Gulf of Mexico') %>% 
  group_by(DataProvider,FishCouncilRegion, EntryDate, EntryDateYear) %>% 
  dplyr::summarize(n=n())

# View(x)
# 
# sum(x$n)
# table(filt2$EntryDateYear, useNA = 'always')
# class(filt2$EntryDateYear)
```

# list of institutional contributors in 2018 and 2019

```{r}

dataprovider_summary <- filt2 %>%  
  filter(EntryDateYear == 2018 | EntryDateYear == 2019) %>% 
  group_by(DataProvider) %>% 
  summarize(
    class = paste(unique(class), collapse = ' | '),
    DatasetIDs = paste(unique(DatasetID), collapse = ' | '),
    Repositories = paste(unique(Repository), collapse = ' | '),
    PIs = paste(unique(PI), collapse = ' | '), 
    DataContacts = paste(unique(DataContact), collapse = ' | '),
    Reporter = paste(unique(Reporter), collapse = ' | '))
    
x <- dataprovider_summary$DataProvider
y <- dataprovider_summary$Repositories

institution_list <- c(x,y)
institution_list <- unique(institution_list)
# View(institution_list)

## export
setwd("C:/rworking/deepseatools/indata")
write.xlsx(institution_list, "20200311-0_institution_list_2018_2019_from_NDB_20191207-0_RPMcGuinn.xlsx")

##### final cleaned up list by manual inspection ##### 
## import cleaned up file


```






