---
title: "Data Holdings: MBARI and Pacific FMC"
author: "Robert P. McGuinn"
date: "Report last ran on: `r Sys.Date()`"
output: word_document
---

```{r opts, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(fig.width=6.5, fig.height=4, warning = FALSE, dpi = 300) 
options(scipen=10000)
```

```{r variables, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
version <- '20181005_0'
```

```{r filters, echo=FALSE}
# filter
db <- filt
pac <- db %>% filter(FishCouncilRegion == "Pacific")
mbari <- db %>% filter(DatasetID == "MBARI")
```

### Database Version:
`r version` 

### Number of records in entire National Database:
`r prettyNum(length(db$CatalogNumber), big.mark = ",")` 

### Number of records within the Pacific Fishery Management Council Region:
`r prettyNum(length(pac$CatalogNumber), big.mark = ",")`

### Number of records contributed by MBARI:
`r prettyNum(length(mbari$CatalogNumber), big.mark = ",")`

### Depth distribution by Fishery Management Council Region in entire database:
*Note*: Depths shallower than 50m in the database are flagged (see 'ShallowFlag' variable in the database)

``` {r depth_db, echo=FALSE}
##### Boxplot of depth by FishCouncilRegion #####
filt2 <- filt %>% 
  filter(DepthInMeters != "-999")
  
g <- ggplot(db, aes(factor(FishCouncilRegion), as.numeric(DepthInMeters), color = Phylum)) +   geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") + 
  xlab("FishCouncilRegion") + 
  theme_bw(base_size = 15, base_family = "Cambria") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

#display.brewer.all(colorblindFriendly=TRUE)
g + scale_color_manual(values = brewer.pal(12, "Paired")[c(10,9)])

```


# Whole Database

###Minimum Depth in Database: 
`r min(as.numeric(db$DepthInMeters))` meters

###Median Depth in Database: 
`r median(db$DepthInMeters)` meters

###Maximum Depth in Database: 
`r max(db$DepthInMeters)` meters


# Pacific

###Depth distribution within Pacific Fishery Management Council Region:
``` {r depth_pacific, echo=FALSE}
##### Boxplot of depth by FishCouncilRegion #####

g <- ggplot(pac, aes(factor(FishCouncilRegion), as.numeric(DepthInMeters), color = Phylum)) +   geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") + 
  xlab("FishCouncilRegion") + 
  theme_bw(base_size = 15, base_family = "Cambria") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

#display.brewer.all(colorblindFriendly=TRUE)
g + scale_color_manual(values = brewer.pal(12, "Paired")[c(10,9)])

```

###Minimum Depth in Pacific Fishery Management Council Region: 
`r min(pac$DepthInMeters)` meters

###Median Depth in Pacific Fishery Management Council Region: 
`r median(pac$DepthInMeters)` meters

###Maximum Depth in Pacific Fishery Management Council Region: 
`r max(pac$DepthInMeters)` meters


# MBARI

###Minimum observation date for MBARI records: 
`r min(as.Date(mbari$ObservationDate))`

###Median observation date for MBARI records: 
`r median(as.Date(mbari$ObservationDate))`

###Maximum observation date for MBARI records: 
`r max(as.Date(mbari$ObservationDate))`
