---
title: "Golden Crab EFP Taxonomic Tables"
author: "Robert P. McGuinn"
date: "October 19, 2018"
output: word_document
#html_document
---

``` {r variables_and_paths, echo=FALSE}
version <- "20181005-0"
fgdb_path <- 'C:/data/aprx/GoldenCrab/GoldenCrab.gdb'
```

``` {r creating_aoi_poly, echo = FALSE}
##### _____ Creating aoi polygon from bounding coordinates #####
minLon <- -85.4
maxLon <- -84.2
minLat <- 25
maxLat <- 28

e <- as(raster::extent(minLon, maxLon, minLat, maxLat), "SpatialPolygons")
proj4string(e) <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
data = data.frame(f=99.9)
aoibox = SpatialPolygonsDataFrame(e,data)
# 
# setwd("C:/data/BaseLayers/GoldenCrabAOIp2")
# writeOGR(aoibox, dsn="aoibox", 
#          layer= "aoibox", 
#          driver = "ESRI Shapefile",
#          overwrite_layer = T)

fgdb_path <- 'C:/data/aprx/GoldenCrab/GoldenCrab.gdb'
arc.write(file.path(fgdb_path, 'one'), data=aoibox)

```

``` {r elevation, echo=FALSE}
##### _____ get ETOPO1  and CRM data from NCEI #####

##### set 'aoi' bounding box ##### 
# minLon <- -85.4
# maxLon <- -84.2
# minLat <- 25
# maxLat <- 28
# 
# setwd("C:/data/BaseLayers/GoldenCrabAOIp2")
# url <- paste("http://maps.ngdc.noaa.gov/mapviewer-support/wcs-proxy/",
#                 "wcs.groovy?filename=etopo1_bedrock.tif&",
#                 "request=getcoverage&version=1.0.0&service=wcs&",
#                 "coverage=etopo1_bedrock&CRS=EPSG:4326&format=geotiff&",
#                 "resx=0.000833333333333334&resy=0.000833333333333334&bbox=",
#                 minLon, ",", minLat, ",", maxLon, ",", maxLat, sep="")
# fname <- "etopo_temp.tif"
# download.file(url, fname, mode="wb", cacheOK="false")
# etopo <- raster(fname)
# 
# setwd("C:/data/BaseLayers/GoldenCrabAOIp2")
# writeRaster(etopo,'etopo.tif', overwrite=TRUE)
# 
######_____  get CRM data from NCEI #####
# setwd("C:/data/BaseLayers/GoldenCrabAOIp2")
# url.hi <- paste("http://maps.ngdc.noaa.gov/mapviewer-support/wcs-proxy/",
#                 "wcs.groovy?filename=crm.tif&",
#                 "request=getcoverage&version=1.0.0&service=wcs&",
#                 "coverage=crm&CRS=EPSG:4326&format=geotiff&",
#                 "resx=0.000833333333333334&resy=0.000833333333333334&bbox=",
#                 minLon, ",", minLat, ",", maxLon, ",", maxLat, sep="")
# fname.hi <- "crm.temp.tif"
# 
# download.file(url.hi, fname.hi, mode="wb", cacheOK="false")
# 
# crm <- raster::raster(fname.hi)
# 
# setwd("C:/data/BaseLayers/GoldenCrabAOIp2")
# writeRaster(crm,'crm.tif', overwrite=TRUE)

##### reclassify CRM data to create proper depth zone (see reclass matrix, 'rclmat') #####
# depth zone of interest 548.6-792.4 (class 2 below)
# crm_flip_sign <- crm * -1
# m <- c(0.0, 548.0, 1,  549.0, 792.0, 2,  793.0, 3500.0, 3)
# rclmat <- matrix(m, ncol=3, byrow=TRUE)
# rc <- reclassify(crm_flip_sign, rclmat)
# 
# setwd("C:/data/BaseLayers/GoldenCrabAOIp2")
# writeRaster(rc,'crm_rc.tif', overwrite=TRUE)
```

``` {r subset_db_data, echo=FALSE}
#####_____ setting bounding boxes #####
##### set 'aoi' bounding box ##### 
minLon <- -85.4 
maxLon <- -84.2
minLat <- 25 #25
maxLat <- 28

##### subsetting NDB with 'aoi' extent #####
db_aoi <- subset(filt, as.numeric(Latitude) > minLat & 
                   as.numeric(Latitude) < maxLat & 
                   as.numeric(Longitude) < maxLon &
                   as.numeric(Longitude) > minLon)

##### set 'big' bounding box ##### 
minLon <- -98
maxLon <- -64
minLat <- 15
maxLat <- 37

##### subsetting NDB with 'big' extent #####
db_big <- subset(filt, as.numeric(Latitude) > minLat & 
                   as.numeric(Latitude) < maxLat & 
                   as.numeric(Longitude) < maxLon &
                   as.numeric(Longitude) > minLon)


```

``` {r input_extra_data, echo=FALSE}
##### reading from feature class (after some GIS selection), creates 'arc.feature_impl'object ##### 
fgdb_path <- 'C:/data/aprx/GoldenCrab/GoldenCrab.gdb'
reed <- arc.open(file.path(fgdb_path, 'reed'))
ex1803 <- arc.open(file.path(fgdb_path, 'ex1803'))
ex1711 <- arc.open(file.path(fgdb_path, 'ex1711'))
nf1708 <- arc.open(file.path(fgdb_path, 'nf1708'))

##### checking ##### 
# class(ex1803)
# # [1] "arc.feature_impl"
# # attr(,"package")
# # [1] "arcgisbinding"

##### _____ creating 'sp' spatial polygon dataframes of the polygons ##### 

reed <- arc.select(reed)
##### checking ####
# > class(reed)
# [1] "arc.data"   "data.frame"
# > reed <- arc.data2sp(reed)

reed <- arc.data2sp(reed)
# > class(reed)
# [1] "SpatialPointsDataFrame"
# attr(,"package")
# [1] "sp"

ex1803 <- arc.select(ex1803)
ex1803 <- arc.data2sp(ex1803)

ex1711 <- arc.select(ex1711)
ex1711 <- arc.data2sp(ex1711)

nf1708 <- arc.select(nf1708)
nf1708 <- arc.data2sp(nf1708)

##### checking ######
# proj4string(ex1803)
# proj4string(nf1708)
# proj4string(ex1711)
# proj4string(reed)

```

``` {r rbinding_newdata_to_db, echo = FALSE}
##### ____ rbinding operations for reedbuffer (output is 'yo') #####  

# setdiff(names(reedbuffer), names(aoi))
# setdiff(names(aoi), names(reedbuffer))

##### rename specific column name (#reassign) ####
x <-which(colnames(reed) == 'Order_')
colnames(reed)[x] <- 'Order'

##### output of new reedbuffer1 #####
reedbuffer1 <- reedbuffer[,c(names(aoi))]

##### checking ##### 
# setdiff(names(reedbuffer1), names(aoi))
# setdiff(names(aoi), names(reedbuffer1))

##### bind filt and reedbuffer1 #####
yo<-rbind(aoi, reedbuffer1)

##### _____ rbinding operations for NF1708 (output is 'yo1') #####  
##### checking #####
# setdiff(names(NF1708buffer), names(filt))
# setdiff(names(filt), names(NF1708buffer))

##### rename specific column name (#reassign) ####
x <-which(colnames(NF1708buffer) == 'Order_')
colnames(NF1708buffer)[x] <- 'Order'

##### output of new reedbuffer1 #####
NF1708buffer1 <- NF1708buffer[,c(names(aoi))]

##### bind yo and NF1708buffer1 #####
yo1<-rbind(yo, NF1708buffer1)

##### checking row numbers ##### 
# x <- length(NF1708buffer1$ScientificName)
# y <- length(yo$ScientificName)
# z <- length(yo1$ScientificName)
# x
# y
# x+y
# z

##### _____ rbinding operations for ex1803buffer  (output is 'yo2') #####  
##### checking #####
# setdiff(names(ex1803buffer), names(yo1))
# setdiff(names(yo1), names(ex1803buffer))

##### ______ create a new Latitude and Longitude field ##### 
ex1803buffer$Latitude <- (ex1803buffer$StartLat + ex1803buffer$EndLat)/2
ex1803buffer$Longitude <- (ex1803buffer$StartLon + ex1803buffer$EndLon)/2

##### _____ create a new 'DepthInMeters' field ##### 
ex1803buffer$DepthInMeters <- (ex1803buffer$StartDepth + ex1803buffer$EndDepth)/2

##### _____ create a new 'Habitat' field #####
ex1803buffer$Habitat <- paste(ex1803buffer$PrimSubstrate, ex1803buffer$SecSubstrate, sep = ',')

##### output of new ex1803buffer1) #####
i <- intersect(names(yo1), names(ex1803buffer))
ex1803buffer1 <- ex1803buffer[,c(i)]

##### add all the column names #####
namevector<-setdiff(names(yo1), names(ex1803buffer1))
for(i in namevector)
  ex1803buffer1[,i] <- NA

##### checking #####
# setdiff(names(ex1803buffer1), names(yo1))
# setdiff(names(yo1), names(ex1803buffer1))

##### 
yo2 <-rbind(yo1, ex1803buffer1)

##### checking ##### 
# x <- length(ex1803buffer1$ScientificName)
# y <- length(yo1$ScientificName)
# z <- length(yo2$ScientificName)
# x
# y
# x+y
# z

##### _____ rbinding operations for ex1711buffer  (output is 'yo3') #####  
# setdiff(names(ex1711buffer), names(yo1))
# setdiff(names(yo1), names(ex1711buffer))

##### ______ create a new Latitude and Longitude field ##### 
ex1711buffer$Latitude <- (ex1711buffer$StartLat + ex1711buffer$EndLat)/2
ex1711buffer$Longitude <- (ex1711buffer$StartLon + ex1711buffer$EndLon)/2

##### _____ create a new 'DepthInMeters' field ##### 
ex1711buffer$DepthInMeters <- (ex1711buffer$StartDepth + ex1711buffer$EndDepth)/2

##### output of new ex1711buffer1) #####
i <- intersect(names(yo1), names(ex1711buffer))
ex1711buffer1 <- ex1711buffer[,c(i)]

##### add all the missing column names #####
namevector<-setdiff(names(yo1), names(ex1711buffer1))
for(i in namevector)
  ex1711buffer1[,i] <- NA

# setdiff(names(ex1711buffer1), names(yo1))
# setdiff(names(yo1), names(ex1711buffer1))

##### 
yo3 <- rbind(yo2, ex1711buffer1)

##### checking ##### 
# x <- length(ex1711buffer1$ScientificName)
# y <- length(yo2$ScientificName)
# z <- length(yo3$ScientificName)
# x
# y
# x+y
# z

# names(ex1711buffer)

###### ______ export final rbind result to ArcGIS ###### 

##### exporting a shapefile of 'yo3' ##### 
# coordinates(yo3) <- c("Longitude", "Latitude")
# proj4string(yo3) <- "+proj=longlat +ellps=WGS84 +datum=WGS84"
# 
# setwd("C:/data/BaseLayers/GoldenCrabAOIp2")
# writeOGR(yo3, dsn="yo3",
#          layer= "yo3",
#          driver = "ESRI Shapefile",
#          dataset_options=c("NameField=CatalogNumber"),
#          overwrite_layer = T)

##### checking ##### 
length(ex1803buffer$ScientificName)
length(ex1711buffer)
length(reedbuffer)
length(NF1708buffer)

table(factor(yo3$ScientificName))
length

```

``` {r output_expanded_db_data, echo=FALSE}
##### _____ output routines ##### 
##### _____ output of NDB subsets #####
##### exporting the 'big' and 'aoi' data subsets as a CSV #####
# setwd("C:/rworking/digs/outdata")
# write.csv(db_big,"20181019_0_big_geo_subset_NatDB_20181005-0_RPMcGuinn.csv", row.names = F, quote = T)
# 
# setwd("C:/rworking/digs/outdata")
# write.csv(db_aoi,"20181019_0_aoi_geo_subset_NatDB_20181005-0_RPMcGuinn.csv", row.names = F, quote = T)

##### exporting 
##### exporting db_big and db_aoi as an sp objects through the arcgisbinding to point featureclass #####
coordinates(db_big) <- c("Longitude", "Latitude")
proj4string(db_big) <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"
arc.write(file.path(fgdb_path, 'db_big'), data=db_big, overwrite = TRUE )

coordinates(db_aoi) <- c("Longitude", "Latitude")
proj4string(db_aoi) <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"
arc.write(file.path(fgdb_path, 'db_aoi'), data=db_aoi, overwrite = TRUE)

##### exporting a shapefile of 'db_big' ##### 
# coordinates(db_big) <- c("Longitude", "Latitude")
# proj4string(db_big) <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"
# 
# setwd("C:/data/BaseLayers/GoldenCrabAOIp2")
# writeOGR(db_big, dsn="big", 
#          layer= "big", 
#          driver = "ESRI Shapefile",
#          dataset_options=c("NameField=CatalogNumber"), 
#          overwrite_layer = T)

##### exporting a shapefile of 'db_aoi' ##### 
# coordinates(db_aoi) <- c("Longitude", "Latitude")
# proj4string(db_aoi) <- "+proj=longlat +ellps=WGS84 +datum=WGS84"
# 
# setwd("C:/data/BaseLayers/GoldenCrabAOIp2")
# writeOGR(db_aoi, dsn="db_aoi", 
#          layer= "db_aoi", 
#          driver = "ESRI Shapefile",
#          dataset_options=c("NameField=CatalogNumber"), 
#          overwrite_layer = T)

```

``` {r input_obis, echo = FALSE}
##### _____ getting OBIS data #####
##### setting bounding box for 'aoi' #####
minLon <- -85.4
maxLon <- -84.2
minLat <- 25
maxLat <- 28

##### pinging OBIS for data #####
library(spocc)
bounds <- c(minLon, minLat, maxLon, maxLat)
out <- occ(geometry = bounds, from = 'obis')
obis <- as.data.frame(out$obis$data)

##### check ##### 
# summary(obis$longitude)
# plot(obis$longitude)
# names(obis)
# View(obis)
# table(obis$phylum)
# table(obis$species)

```

``` {r output_obis, echo = FALSE}
##### make a shapefile with the OGR driver #####
# library(rgdal)
# coordinates(obis) <- c("longitude", "latitude")
# proj4string(obis) <- "+proj=longlat +ellps=WGS84 +datum=WGS84"
# 
# setwd("C:/data/BaseLayers/GoldenCrabAOIp2")
# writeOGR(obis, dsn="obis", 
#          layer= "obis", 
#          driver = "ESRI Shapefile",
#          dataset_options=c("NameField=CatalogNumber"), 
#          overwrite_layer = T)
``` 

``` {r read_in_extent_polys, echo=FALSE}
##### _____ bringing in aoi, target, and target_buffer using arc2R bridge #####
##### reading from feature class (after some GIS selection), creates 'arc.feature_impl'object ##### 
fgdb_path <- 'C:/data/aprx/GoldenCrab/GoldenCrab.gdb'
aoi_poly <- arc.open(file.path(fgdb_path, 'aoi'))
target_buffer_poly <- arc.open(file.path(fgdb_path, 'target_Buffer'))
target_poly <- arc.open(file.path(fgdb_path, 'target'))

##### creating 'sp' spatial polygon dataframes of the polygons ##### 
target_poly <- arc.select(target_poly)
target_poly <- arc.data2sp(target_poly)
target_buffer_poly <- arc.select(target_buffer_poly)
target_buffer_poly <- arc.data2sp(target_buffer_poly)
aoi_poly <- arc.select(aoi_poly)
aoi_poly <- arc.data2sp(aoi_poly)

##### checking #####
# plot(aoi_poly)
# plot(target_buffer_poly)
# plot(target_poly)

```

``` {r intersects, echo = FALSE}
##### another way to make target internal to R #####
library(rgdal)
coordinates(db_aoi) <- c("Longitude", "Latitude")
proj4string(db_aoi) <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"

##### intersect db points in aoi with 'target_poly' ##### 
x <-over(db_aoi, target_poly)
# assign the intersection results back to aoio
db_aoi$Id <- x$Id
db_aoi <- as.data.frame(db_aoi)
aoi_target <- db_aoi %>% filter(Id == "1")
coordinates(aoi_target) <- c("Longitude", "Latitude")
proj4string(aoi_target) <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"
fgdb_path <- 'C:/data/aprx/GoldenCrab/GoldenCrab.gdb'
arc.write(file.path(fgdb_path, 'aoi_target'), data=aoi_target)

```

# Analysis details

Target fishing area boundaries:

* Depth zone 548.6 to 792.4 meters
* Bounding box
+ minimum longitude: `r minLon`
+ maximum longitude: `r maxLon`
+ minimum latitude: `r minLat`
+ maximum lataitude: `r maxLat`

Databases used for species occurrences:

* NOAA National Database for Deep Sea Corals and Sponges version: `r version`
* Ocean Biogeographic Informaton System (OBIS) data accessed on 2018-10-19.
* Other data: Reed, NF

# Within the fishing target area

## List of Coral Taxa Grouped by 'VernacularNameCategory' (from National Database)

``` {r CoralByVernacularNameCategoryTarget, echo=FALSE }
target <- as.data.frame(target)
sum_tbl <-
  target %>%
  filter(Phylum == "Cnidaria") %>% 
  group_by(VernacularNameCategory) %>%
  summarize(
    Taxa = toString(unique(ScientificName)),
    Records = n()) %>%
  arrange(desc(Records))
sum_tbl <- kable(sum_tbl, row.names = F, digits = 2)
sum_tbl
```

## List of Sponge Taxa Grouped by 'VernacularNameCategory' (from National Database)

``` {r SpongeByVernacularNameCategoryTarget, echo=FALSE }
target <- as.data.frame(target)
sum_tbl <-
  target %>%
  filter(Phylum == "Porifera") %>% 
  group_by(VernacularNameCategory) %>%
  summarize(
    Taxa = toString(unique(ScientificName)),
    Records = n()) %>%
  arrange(desc(Records))
sum_tbl <- kable(sum_tbl, row.names = F, digits = 2)
sum_tbl
```

## OBIS List of Taxa Grouped by Order

``` {r TaxaByOrderTargetOBIS, echo=FALSE }
sum_tbl <-
  obis %>%
  group_by(order_) %>%
  summarize(
    Taxa = toString(unique(name)),
    Records = n()) %>%
  arrange(desc(Records))
sum_tbl <- kable(sum_tbl, row.names = F, digits = 2)
sum_tbl
```

# Within the fishing target area plus 5000m buffer zone

## List of Coral Taxa Grouped by 'VernacularNameCategory' (From National Database)

``` {r CoralByVernacularNameCategoryBuffer, echo=FALSE }
buffer <- as.data.frame(buffer)
sum_tbl <-
  buffer %>%
  filter(Phylum == "Cnidaria") %>% 
  group_by(VernacularNameCategory) %>%
  summarize(
    Taxa = toString(unique(ScientificName)),
    Records = n()) %>%
  arrange(desc(Records))
sum_tbl <- kable(sum_tbl, row.names = F, digits = 2)
sum_tbl
```

## List of Sponge Taxa Grouped by 'VernacularNameCategory' (From National Database)

``` {r SpongeByVernacularNameCategoryBuffer, echo=FALSE }
sum_tbl <-
  buffer %>%
  filter(Phylum == "Porifera") %>% 
  group_by(VernacularNameCategory) %>%
  summarize(
    Taxa = toString(unique(ScientificName)),
    Records = n()) %>%
  arrange(desc(Records))
sum_tbl <- kable(sum_tbl, row.names = F, digits = 2)
sum_tbl
```

## List of Taxa Grouped by Order (From OBIS)

``` {r TaxaByOrderBufferOBIS, echo=FALSE }
sum_tbl <-
  obisbuffer %>%
  group_by(order_) %>%
  summarize(
    Taxa = toString(unique(name)),
    Records = n()) %>%
  arrange(desc(Records))
sum_tbl <- kable(sum_tbl, row.names = F, digits = 2)
sum_tbl
```
