---
title: "Database Status Update - Figs and Tables"
author: "Robert P. McGuinn"
date: "Report last ran on: `r Sys.Date()`"
output: word_document
---

```{r data_intake, echo=FALSE, warning=FALSE, message=FALSE, cache = TRUE}

##### input: latest version of NDB #####
# setwd("C:/rworking/deepseatools/indata")
# indata<-read.csv("DSCRTP_NatDB_20190920-0.csv", header = T)
# filt <- indata %>%
#   filter(Flag == "0")

```

```{r data_prep, echo=FALSE, warning=FALSE, message=FALSE, cache = FALSE}

##### data improvements (from version DSCRTP_NatDB_20171214-0): Recorded in Redmine #####
# filt <- filt %>% filter(ObservationYear != "-999" | 
#                           is.na(TaxonRank) == F |
#                           DepthInMeters != '-999') 
# 
# filt <- filt %>% mutate(DataProvider = 
#                           ifelse(DataProvider == "Yoklavich, Mary",
#                                               'NOAA, Southwest Fisheries Science Center, Santa Cruz', 
#                                               as.character(DataProvider)))
# filt <- filt %>% mutate(DataProvider = 
#                           ifelse(DataProvider == "Flower Garden Banks National Marine Sanctuary (FGBNMS)",
#                                               'NOAA, Flower Garden Banks National Marine Sanctuary', 
#                                               as.character(DataProvider)))
# 
# filt <- filt %>% mutate(DataProvider = 
#                           ifelse(DataProvider == "Hall-Spencer, Jason",
#                                               'Hall-Spencer, J.', 
#                                               as.character(DataProvider)))
# 
# 
# filt <- filt %>% mutate(DataProvider = 
#                           ifelse(DataProvider == "National Undersea Research Center",
#                                               'National Undersea Research Center, University of Connecticut', 
#                                               as.character(DataProvider)))
# 
# filt <- filt %>% mutate(DataProvider = 
#                           ifelse(DataProvider == "NOAA Center for Coastal Monitoring and Assessment",
#                                               'NOAA, Center for Coastal Monitoring and Assessment', 
#                                               as.character(DataProvider)))
# 
# filt <- filt %>% mutate(LocationAccuracy = 
#                           ifelse(LocationAccuracy == "300 m",
#                                               '500m', 
#           

filt <- filt %>% filter(is.na(Phylum) == F)
                         
##### checker #####

# length(filt1$CatalogNumber)

# sum_tbl <- filt %>%
#    # filter(as.numeric(IndividualCount) > 100,
#    #        grepl('Etnoyer', PI),
#    #        DatasetID == 'NOAA_SJ-10-08'
#    #       ) %>%
#   group_by(AccessionID) %>% 
#   summarize(DataProvider = toString(unique(DataProvider)),
#             Repository = toString(unique(Repository)),
#             Vessel = toString(unique(Vessel)),
#             PI = toString(unique(PI)),
#             SampleID = toString(unique(SampleID)),
#             ScientificName = toString(unique(ScientificName)),
#             ImageFilePath = toString(unique(ImageFilePath)),
#             Records = n()
#   ) %>% 
#   arrange(desc(Records))
# View(sum_tbl)

```

```{r schema, echo=FALSE, echo=FALSE, warning=FALSE, message=FALSE, cache = FALSE}
##### download Google Sheet version of schema for use in R  documents ##### 
# Register and download Google Sheet
s <- gs_title('2019_DSCRTP_National_Database_Schema')
#s <- gs_key('1YDskzxY8OF-34Q8aI04tZvlRbhGZqBSysuie39kYHoI')
s <- gs_read(s)

# s<- s[,-(33:35)]
#names(s)

```

```{r opts, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
##### knitr options #####
knitr::opts_chunk$set(fig.width=6.5, 
                      fig.height=4, 
                      warning = FALSE, 
                      dpi = 300, 
                      dev = c('png'), 
                      fig.path="C:/rworking/deepseatools/reports/2019_status_update_report/figures/") 

options(scipen=10000)

##### setting digit options #####
options(digits = 0)
```

```{r variables, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
version <- '20190920-0'
```

# DRAFT - Status Update for NOAA's National Database for Deep-Sea Corals and Sponges - Figs and Tables (database version `r version`)

Number of genera: `r length(unique(filt$Genus))` 


# Changes and Additions to Database Fields

Two new taxomomy related fields were added.

* **VerbatimScientificName**: `r s$FieldDescription[s$FieldName == "VerbatimScientificName"]`
* **TypeStatus**: `r s$FieldDescription[s$FieldName == "TypeStatus"]`

The previous 'Size' field was changed to 'VerbatimSize' and two new fields were added to capture the minimum and maximum size. 

* **VerbatimSize**: `r s$FieldDescription[s$FieldName == "VerbatimSize"]`
* **MinimumSize**: `r s$FieldDescription[s$FieldName == "MinimumSize"]` 
* **MaximumSize**: `r s$FieldDescription[s$FieldName == "MaximumSize"]`

The following fields were added to capture key habitat characteristics using standardized terminology from the Coastal and Marine Ecological Classification Standard (CMECS).

* **CMECSGeoForm**: `r s$FieldDescription[s$FieldName == "CMECSGeoForm"]`
* **CMECSSubstrate**: `r s$FieldDescription[s$FieldName == "CMECSSubstrate"]`
* **CMECSBiotic**:   `r s$FieldDescription[s$FieldName == "CMECSBiotic"]`

The 'EntryUpdate' field was added to capture the most current date that the Program modified the record. This is in addition to the already added field called 'Modified', which captures the last date that the data provider modifed the record. 

* **EntryUpdate**:  `r s$FieldDescription[s$FieldName == "EntryUpdate"]` 

We added the field 'ShallowFlag' to flag occurrences that are less than 50 meters deep. These occurrences are included in our database because they represent deeper taxa that can also occur at shallower depths than our typical 50 meter cuttoff.

* **ShallowFlag**:  `r s$FieldDescription[s$FieldName == "ShallowFlag"]`

Several new spatially calculated depth fields were added to help ground truth the observed depths within the database.  

* **gisCRMDepth**: `r s$FieldDescription[s$FieldName == "gisCRMDepth"]`
* **gisGEBCODepth**: `r s$FieldDescription[s$FieldName == "gisGEBCODepth"]`
* **gisEtopoDepth**: `r s$FieldDescription[s$FieldName == "gisEtopoDepth"]`

The following location related fields were added to enhance the ability to describe and standarize place names. 

* **gisNGIALocality**:`r s$FieldDescription[s$FieldName == "gisNGIALocality"]`
* **gisNGIADist**:`r s$FieldDescription[s$FieldName == "gisNGIADist"]`
* **gisGEBCOLocality**: `r s$FieldDescription[s$FieldName == "gisGEBCOLocality"]`
* **gisGEBCODist**:`r s$FieldDescription[s$FieldName == "gisGEBCODist"]` 

The following fields have been added to capture the actual spatially explicit boundar of the sample area footprint for each record. 

* **footprintWKT**: `r s$FieldDescription[s$FieldName == "footprintWKT"]`  
* **footprintSRS**: `r s$FieldDescription[s$FieldName == "footprintSRS"]`

Number of 'DataProviders': `r length(unique(factor(filt$DataProvider)))` 

# Schema Metrics

``` {r schema_metrics, echo=FALSE}
#total number of fields in the NDB
a <- length(s$FieldName)

#total number of published fields
b <- s %>%
  filter(InternalUseOnly == '0')
b <- length(b$FieldName)

#total number of unpublished fields
c <- s %>%
  filter(InternalUseOnly == '1')
c <- length(c$FieldName)

#PointHist required
d <- s %>%
  filter(PointHist == 'R')
d <- length(d$FieldName)

#PointNew required
e <- s %>%
  filter(PointNew == 'R')
e <- length(e$FieldName)

#PointProgram required
f <- s %>%
  filter(PointProgram == 'R')
f <- length(f$FieldName)

#TransHist required
g <- s %>%
  filter(TransHist == 'R')
g <- length(g$FieldName)

#TransNew required
h <- s %>%
  filter(TransNew == 'R')
h <- length(h$FieldName)

#TransProgram required
i <- s %>%
  filter(TransProgram == 'R')
i <- length(i$FieldName)

#TrawlHist required
j <- s %>%
  filter(TrawlHist == 'R')
j <- length(j$FieldName)

#TrawlNew required
k <- s %>%
  filter(TrawlNew == 'R')
k <- length(k$FieldName)

#TrawlProgram required
l <- s %>%
  filter(TrawlProgram == 'R')
l <- length(l$FieldName)

```

Total number of fields: `r a` 
Total number of published fields: `r b`
Total number of un-published fields: `r c`
PointHist required: `r d`
PointNew required: `r e`
PointProgram required: `r f`
TransHist required: `r g`
TransNew required: `r h`
TransProgram required: `r i`
TrawlHist required: `r j`
TrawlNew required: `r k` 
TrawlProgram required: `r l`

###### *Table 1*: Alphabetical list of data provider ('DataProvider') with a list of unique entries for data set ('DatasetID'). 

``` {r dataprovider_table, echo=FALSE, caption = 'test caption'}
options(scipen=999)

x <-
  filt %>%
  group_by(DataProvider) %>%
  summarize(
    Number_Datasets = length(unique(DatasetID)),
    Records = n()
            )

colkeys <- names(x[,-1])
ft <- flextable(x) %>% 
  theme_box() %>% #booktabs, vanilla, box, tron, tron_legacy
  set_header_labels(Number_Datasets = "# of Data Sets", Records = "# of Records") %>% 
  colformat_num(col_keys = colkeys, digits = 0, big.mark = ",", na_str = "NA") %>% 
  font(fontname = "Cambria", part = "all") %>% #header, body, all
  autofit() 

ft

read_docx() %>% 
  body_add_flextable(ft) %>% 
  print(target = "c:/rworking/deepseatools/reports/2019_status_update_report/tables/dataprovider_table.docx")



```

###### *Table 2a*: This table displays all of the public fields in the database. They are grouped by 'Field Category'. 

```{r public_fields_table, echo=FALSE, cache = FALSE}

s$DSCRTPCategory <- factor(s$DSCRTPCategory, levels = c("Survey Data","Event Data","Observation Data","TaxonomicID","Environment","Occurrence Detail","GIS Enhancement","Metadata","Record Keeping","Flag"))

x <-
  s %>%
  filter(InternalUseOnly == '0') %>%
  group_by(DSCRTPCategory) %>% 
  summarize(Field_Names = toString(unique(FieldName)),
            n = n())


colkeys <- names(x[,-1])
ft <- flextable(x) %>% 
  theme_booktabs() %>% #booktabs, vanilla, box, tron, tron_legacy
  set_header_labels(Field_Names = "Field Name", n = "# of Records") %>% 
  colformat_num(col_keys = colkeys, digits = 0, big.mark = ",", na_str = "NA") %>% 
  font(fontname = "Cambria", part = "all")  %>% #header, body, all
  autofit() 

ft

read_docx() %>% 
  body_add_flextable(ft) %>% 
  print(target = "c:/rworking/deepseatools/reports/2019_status_update_report/tables/public_fields_table.docx")




# sum_tbl <- kable(sum_tbl, 
#                  row.names = F, 
#                  digits = 2,
#                  col.names = c('Field Category', 'Field Name', 'Number of Fields'))
# 
# options(scipen=10000)
# sum_tbl

#setwd("C:/rworking/digs/outdata")
#write.csv(sum_tbl,"20180216_0_Field_Summary_Schema_20180216-0.csv", row.names = F, quote = T)

```

# Required Fields for Submission

###### *Table 2b*: The following table outlines the required and desired fields for different types of incoming *point-based* data. 'R' means the field is required for data submission and 'D' means that it is desired by the Program. 'P' means that is a value provided by the program.

```{r required_desired_points_table, echo=FALSE, cache = FALSE, warning=FALSE}
x <-
  s %>%
  filter(InternalUseOnly == '0',
         PointHist == 'R' | PointNew == 'R' | PointProgram == 'R'
         ) %>%
  group_by(FieldName) %>% 
  summarize(PointHist = unique(PointHist),
            PointNew = unique(PointNew),
            PointProgram = unique(PointProgram)
                        )
colkeys <- names(x)

ft <- flextable(x) %>% 
  theme_vanilla() %>% #booktabs, vanilla, box, tron, tron_legacy
 # set_header_labels(Field_Names = "Field Name", n = "# of Records") %>% 
  colformat_num(col_keys = colkeys, digits = 0, big.mark = ",", na_str = "NA") %>% 
  font(fontname = "Cambria", part = "all") %>% #header, body, all
  autofit() 

ft

read_docx() %>% 
  body_add_flextable(ft) %>% 
  print(target = "c:/rworking/deepseatools/reports/2019_status_update_report/tables/required_desired_points_table.docx")

```


###### *Table 2c*: The following table outlines the required and desired fields for different types of incoming *transect-based* data. 'R' means the field is required for data submission and 'D' means that it is desired by the Program. 'P' means that is a value provided by the program.

```{r required_desired_transects_table, echo=FALSE, cache = FALSE}
x <-
  s %>%
  filter(InternalUseOnly == '0',
         TransHist == 'R' | TransNew == 'R' | TransProgram == 'R'
         ) %>%
  group_by(DSCRTPCategory, FieldName) %>% 
  summarize(TransHist = unique(TransHist),
            TransNew = unique(TransNew),
            TransProgram = unique(TransProgram)
                        ) %>% 
  arrange(DSCRTPCategory)

colkeys <- names(x[,-1])

ft <- flextable(x) %>% 
  theme_vanilla() %>% #booktabs, vanilla, box, tron, tron_legacy
 # set_header_labels(Field_Names = "Field Name", n = "# of Records") %>% 
  colformat_num(col_keys = colkeys, digits = 0, big.mark = ",", na_str = "NA") %>% 
  font(fontname = "Cambria", part = "all") %>% #header, body, all
  autofit() 

ft

read_docx() %>% 
  body_add_flextable(ft) %>% 
  print(target = "c:/rworking/deepseatools/reports/2019_status_update_report/tables/required_desired_transects_table.docx")


```


###### *Table 2d*: The following table outlines the required and desired fields for different types of incoming *trawl-based* data. 'R' means the field is required for data submission and 'D' means that it is desired by the Program. 'P' means that is a value provided by the program.

```{r required_desired_trawl_table, echo=FALSE, cache = FALSE}
x <-
  s %>%
  filter(InternalUseOnly == '0',
         TrawlHist == 'R' | TrawlNew == 'R' | TrawlProgram == 'R'
         ) %>%
  group_by(DSCRTPCategory, FieldName) %>% 
  summarize(TrawlHist = unique(TrawlHist),
            TrawlNew = unique(TrawlNew),
            TrawlProgram = unique(TrawlProgram)
                        ) 

colkeys <- names(x[,-1])

ft <- flextable(x) %>% 
  theme_vanilla() %>% #booktabs, vanilla, box, tron, tron_legacy
 # set_header_labels(Field_Names = "Field Name", n = "# of Records") %>% 
  colformat_num(col_keys = colkeys, digits = 0, big.mark = ",", na_str = "NA") %>% 
  font(fontname = "Cambria", part = "all") %>% #header, body, all
  autofit() 

ft

read_docx() %>% 
  body_add_flextable(ft) %>% 
  print(target = "c:/rworking/deepseatools/reports/2019_status_update_report/tables/required_desired_trawl_table.docx")


```





###### *Table 3* OBIS Fields  
``` {r obis_table, echo=FALSE}
options(scipen=999)
x <-
  s %>%
  filter(ReleaseToOBIS == '1') %>% 
  group_by(DarwinCoreTerm) %>%
  summarize(
    Field_Name = toString(unique(FieldName))
            )

colkeys <- names(x)

ft <- flextable(x) %>% 
  theme_vanilla() %>% #booktabs, vanilla, box, tron, tron_legacy
 # set_header_labels(Field_Names = "Field Name", n = "# of Records") %>% 
  colformat_num(col_keys = colkeys, digits = 0, big.mark = ",", na_str = "NA") %>% 
  font(fontname = "Cambria", part = "all") %>% #header, body, all
  autofit() 

ft

read_docx() %>% 
  body_add_flextable(ft) %>% 
  print(target = "c:/rworking/deepseatools/reports/2019_status_update_report/tables/obis_table.docx")
```

# Size and Growth of the Database Through Time
```{r calculations_1, echo=FALSE, cache = FALSE}
# length of unflagged records 
a <- round(length(filt$CatalogNumber), -4)

# number of DatasetID
b <- length(unique(filt$DatasetID))

# number of DataProviders
c <- length(unique(filt$DataProvider))

# number of flagged records
d <- length(filter(indata, Flag == '1')$CatalogNumber)

# percent of unflagged records with images
filt2 <- filter(filt, is.na(ImageURL) == F)
e <- round((length(filt2$CatalogNumber)/(length(filt$CatalogNumber))*100),0)

# number of unflagged records with images 
f <- length(filter(filt, is.na(ImageURL) == F)$CatalogNumber)

# stats on number of records per DatasetID
y <- count(filt, DatasetID)
g <- round(median(y$n),0)
h <- min(y$n)
i <- max(y$n)

```

Published records: `r prettyNum(a, big.mark = ",")`  
Number of data sets: `r b` 
Number of data providers: `r c`  
Flagged records `r prettyNum(d, big.mark = ",")` 
Percent of unflagged records with images: `r e`
Number of unflagged records with images: `r prettyNum(f, big.mark = ",")`

## Stats on number of records per DatasetID
* median: `r g`
* minimum: `r h`
* maximum: `r i`

# Growth of Database 
```{r GrowthThroughTime, echo=FALSE, cache=TRUE, dpi=300}

# ordering factors
filt2 <- filt %>%
  filter(as.Date(EntryDate) > as.Date('2010-01-01'))
options(scipen=10000)
g <- ggplot(filt2,aes(x=as.Date(EntryDate), color=Phylum)) +
  stat_bin(data = subset(filt2, Phylum =="Porifera"),aes(y=cumsum(..count..)),geom="line", binwidth = 2, size = 2) +
  stat_bin(data = subset(filt2, Phylum =="Cnidaria"),aes(y=cumsum(..count..)),geom="line", binwidth = 2, size = 2) +
  ylab("Cumulative Number of Records") + 
  xlab("Year Record Added") + 
  guides(fill=FALSE) + 
  theme_bw(base_size = 15, base_family = "Cambria")

g + scale_color_manual(values = brewer.pal(12, "Paired")[c(10,9)])

# write to file

# x <- g + scale_fill_manual(values = brewer.pal(12, "Paired")[c(10,9)])
# 
# png(file = "C:/rworking/deepseatools/reports/2019_status_update_report/growththroughtime.png",
#     width = 1200, height = 800, 
#     units = "px", pointsize = 12, res = 150)
# 
# x
# 
# dev.off()

```

##### *Figure 2*: Growth of the database through time.

# Distribution of Records by U.S. Fishery Management Council
```{r region-counts, echo=FALSE, cache = FALSE}
a <- filt %>% 
  filter(
    is.na(FishCouncilRegion) == T
    ) 
a <- length(a$CatalogNumber)/length(filt$CatalogNumber)
a <- round(a*100, 0)

```

Percent of records without FishCouncilRegion assignments: `r prettyNum(a, big.mark = ",")`

```{r FishCouncilRegion, echo=FALSE, cache = FALSE}
# ordering factors

filt2 <- filt %>% 
  filter(is.na(Phylum) == F , 
         is.na(FishCouncilRegion) == F , 
         FishCouncilRegion != 'NA')
options(scipen=10000)

#table(filt2$FishCouncilRegion, useNA = 'always')

filt2$FishCouncilRegion <- factor(filt2$FishCouncilRegion, levels = c('Pacific', 'North Pacific','Western Pacific','New England','Mid-Atlantic','South Atlantic', 'Gulf of Mexico','Caribbean'))

g <- ggplot(filt2, aes(FishCouncilRegion, fill = Phylum)) +
  geom_bar() + 
  coord_flip() + 
  theme(text = element_text(size=20)) + 
  ylab("Number of Records") + 
  theme_bw(base_size = 15, base_family = "Cambria")

#display.brewer.all(colorblindFriendly=TRUE)
g + scale_fill_manual(values = brewer.pal(12, "Paired")[c(10,9)])

# write to file
# x <- g + scale_fill_manual(values = brewer.pal(12, "Paired")[c(10,9)])
# 
# png(file = "C:/rworking/deepseatools/reports/2019_status_update_report/fishcouncilregion.png",
#     width = 1200, height = 800, 
#     units = "px", pointsize = 12, res = 150)
# 
# x
# 
# dev.off()



```

###### *Figure 3*: Distribution of the coral and sponge occurrences by Fishery Management Council Region ('FishCouncilRegion'). This figure leaves out records that do not fall within a council boundary.

# Images Associated with Records
Per `r e`% or `r prettyNum(f, big.mark = ",")` records have full resolution images available. 

# Depth Distribution
```{r DepthCalc, echo=FALSE, cache = FALSE}
x <- s %>%
  filter(
    DSCRTPGroup == 'Location-Depth Metric',
    InternalUseOnly == "0", 
    DSCRTPCategory != "GIS Enhancement" 
  )
x <- toString(x$FieldName)

y <- s %>%
  filter(DSCRTPGroup == 'Location-Depth Metric', InternalUseOnly == "0", DSCRTPCategory == "GIS Enhancement" )
y <- toString(y$FieldName)

##### checker ##### 
# z <- filt %>% 
#   filter(DepthInMeters == "-999") %>% 
#   group_by(DatasetID) %>% 
#   summarize(
#     DepthInMeters = toString(unique(DepthInMeters)),
#     DataProvider = toString(unique(DataProvider)),
#     n = n()
#     )
# #View(z)
# #length(z$CatalogNumber)
# #[1] 3992
```

Depth fields (not gis-derived): `r x`
Depth fields (gis-derived): `r y`

``` {r depth_vernacularname, echo=FALSE}
##### Boxplot of depth by FishCouncilRegion #####
filt2 <- filt %>% 
  filter(DepthInMeters != "-999")
  
g <- ggplot(filt2, aes(factor(FishCouncilRegion), as.numeric(DepthInMeters), color = Phylum)) +   geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") + 
  xlab("FishCouncilRegion") + 
  theme_bw(base_size = 15, base_family = "Cambria") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

#display.brewer.all(colorblindFriendly=TRUE)
g + scale_color_manual(values = brewer.pal(12, "Paired")[c(10,9)])

```

###### *Figure 5*: Distribution of depth values by Fish Council Region ('FishCouncilRegion'). The 'NA' values are records that are outside of the United States Exclusive Economic Zone.

```{r DepthDensity, echo=FALSE, cache = FALSE, dpi = 300}
filt2 <- filt %>% 
  filter(DepthInMeters != "-999")
options(scipen=10000)

g <- ggplot(filt2, aes(DepthInMeters, fill = Phylum)) +
  geom_density(alpha = 0.5) + 
  xlim(0,7000) +
  ylab("Proportion of Records") + 
  xlab('DepthInMeters') + 
  theme_bw(base_size = 15, base_family = "Cambria")

#display.brewer.all(colorblindFriendly=TRUE)
g + scale_fill_manual(values = brewer.pal(12, "Paired")[c(10,9)])
```

###### *Figure 6*: Density of depth values across the database grouped by Phylum

# ObservationYear 
Range of 'Observation Year': `r min(filt$ObservationYear)` to `r max(filt$ObservationYear)`
Median'Observation Year':`r median(filt$ObservationYear)`

```{r ObservationYear, echo=FALSE, cache = FALSE, dpi = 300}
# ordering factors
filt2 <- filt %>% 
  filter(is.na(ObservationYear) == F, ObservationYear != "-999", is.na(Phylum) == F)
filt2$TaxonRank <- factor(filt2$TaxonRank, levels = c("phylum", "class", "subclass", "order", "suborder","family","subfamily", "genus","subgenus","species","subspecies","forma","variety"))
options(scipen=10000)

g <- ggplot(filt2, aes(ObservationYear, fill = Phylum)) +
  geom_bar() + 
  ylab("Number of Records")  +
  theme_bw(base_size = 15, base_family = "Cambria")

#display.brewer.all(colorblindFriendly=TRUE)
g + scale_fill_manual(values = brewer.pal(12, "Paired")[c(10,9)])

```

###### *Figure 7*: The temporal distribution of coral and sponge observations by 'ObservationYear'

# Individual Count
Records having IndividualCount values:
`r round((length(filter(filt, IndividualCount != '-999')$CatalogNumber)/length(filt$CatalogNumber))*100,2)` % or `r length(filter(filt, IndividualCount != '-999')$CatalogNumber)`
Maximum 'IndividualCount': `r max(filt$IndividualCount)` 
Median 'IndividualCount': `r median(filt$IndividualCount)` 

```{r IndividualCount, echo=FALSE, cache = FALSE, dpi = 300}
# ordering factors
filt2 <- filt %>% 
  filter(is.na(IndividualCount) == F, 
         IndividualCount != '0',
         IndividualCount != '-999', 
         is.na(Phylum) == F, 
         IndividualCount < 50
         )

options(scipen=10000)

g <- ggplot(filt2, aes(IndividualCount, fill = Phylum)) +
  geom_histogram(binwidth = 1) + 
  ylab("Number of Records")  +
  theme_bw(base_size = 15, base_family = "Cambria")

#display.brewer.all(colorblindFriendly=TRUE)
g + scale_fill_manual(values = brewer.pal(12, "Paired")[c(10,9)])

```

###### *Figure 5*: Distribution of 'IndividualCount' values between 0 and 20.  

# Size

``` {r SizeCalc, echo=FALSE, cache = FALSE, warning = FALSE}
yo <- filt %>% 
  filter(
    MinimumSize != "-999",
    MaximumSize != "-999",
    MinimumSize != '0',
    is.na(Phylum) == F
    )

yo$AverageSize <- (yo$MinimumSize + yo$MaximumSize)/2
yo$SizeDiff <- (yo$MaximumSize - yo$MinimumSize)

x <- length(yo$CatalogNumber)
y <- length(filt$CatalogNumber)
```

'MinimumSize' and 'MaximumSize' values reported: `r round((x/y) *100,1)` % of all records (n = `r x`)  

```{r Size, echo=FALSE, cache = FALSE, warning = FALSE, dpi = 300}
yo <- yo %>%
  filter(
   AverageSize < 100  
    )

# ordering factors
options(scipen=10000)
g <- ggplot(yo, aes(as.numeric(AverageSize), fill = Phylum)) +
  geom_histogram(bins = 50) + 
  #coord_cartesian(ylim = c(1, 20000)) +
  ylab("Number of Records") + 
  xlab("Average Size (cm)")  +
  theme_bw(base_size = 15, base_family = "Cambria")

#display.brewer.all(colorblindFriendly=TRUE)
g + scale_fill_manual(values = brewer.pal(12, "Paired")[c(10,9)])

```

###### *Figure 9*: Size distribution. Only records with valid values are shown. The figure has been limited in the x and y axis to focus on the bulk of the data. Limits used: x axis (0-200 cm), y axis (0-5000 records). 

# Condition
```{r ConditionCalc, echo=FALSE, cache = FALSE}

condition <- filt$Condition %in% c('Live','Damaged','Dead')
x <- length(condition[condition == T])
y <- filter(filt, is.na(filt$Condition) == FALSE)
y <- length(y$Condition)
z <- filter(filt, is.na(filt$Condition) == TRUE)
z <- length(z$Condition)
t <- length(filt$Condition)

```

Condition values reported: `r round((y/t) *100,1)` % of all records (n = `r y`)   
Percent valid: `r round((x/y) * 100,1)`   

```{r Condition, echo=FALSE, cache = FALSE, dpi = 300}
yo <- filt %>% 
  filter(
    Condition %in% c('Live','Damaged','Dead') 
    #Condition != 'Live'
         )

# ordering factors
yo$Condition <- factor(yo$Condition, 
                       levels = c('Live','Damaged','Dead'))


options(scipen=10000)
g <- ggplot(yo, aes(Condition, fill = Phylum)) +
  geom_bar() + 
  coord_flip() + 
  ylab("Number of Records") + 
  theme_bw(base_size = 15, base_family = "Cambria")

#display.brewer.all(colorblindFriendly=TRUE)
g + scale_fill_manual(values = brewer.pal(12, "Paired")[c(10,9)])

```

###### *Figure 10*: Distribution of 'Condition' values. This figure *exludes() the 'Live' category. A majority of the the records are in the 'Live' category. 

# Habitat and Substrate
Records with 'Substrate' values: `r (length(filter(filt, is.na(Substrate) == F)$CatalogNumber)/length(filt$CatalogNumber))*100` % or `r length(filter(filt, is.na(Substrate) == F)$CatalogNumber)` 

Records with 'Habitat' values:
`r (length(filter(filt, is.na(Habitat) == F)$CatalogNumber)/length(filt$CatalogNumber))*100` % or `r length(filter(filt, is.na(Habitat) == F)$CatalogNumber)` records have 'Habitat' values.

# Taxonomy 
```{r Location, echo=FALSE, cache = FALSE}
z <- s %>%
  filter(DSCRTPGroup == 'Taxonomy Metadata', InternalUseOnly == "0", FieldName != 'AssociatedTaxa', FieldName != 'Synonyms')
z <-toString(z$FieldName)

```
Taxononmic qualifier and descriptive fields: `r z`. 

```{r taxonranksum, echo=FALSE, cache = FALSE, dpi = 300}
# ordering factors
filt2 <- filt %>% 
  filter(is.na(TaxonRank) == F, is.na(Phylum) == F)

filt2$TaxonRank <- factor(filt2$TaxonRank, levels = c("phylum", "class", "subclass", "order", "suborder","family","subfamily", "genus","subgenus","species","subspecies","forma","variety"))
options(scipen=10000)

g <- ggplot(filt2, aes(TaxonRank, fill = Phylum)) +
  geom_bar() + 
  coord_flip() + 
  ylab("Number of Records") + 
  theme_bw(base_size = 15, base_family = "Cambria")

#display.brewer.all(colorblindFriendly=TRUE)
g + scale_fill_manual(values = brewer.pal(12, "Paired")[c(10,9)])
```

###### *Figure 11*: Distribution of "TaxonRank" across all records. Corals are only included in the database if they are identified at least to Order.

# RecordType

``` {r recordtype, echo = FALSE}
valid <- filt$RecordType %in% c('literature',  'specimen',  'still image', 'video observation', 'video transect','notation', 'catch record')

# number of records with valid values
x <- length(valid[valid == T])

# number of records where the specified value is non-null 
y <- filter(filt, is.na(filt$RecordType) == FALSE)
y <- length(y$RecordType)

# number of records where the specified value is null
z <- filter(filt, is.na(filt$RecordType) == TRUE)
z <- length(z$RecordType)

# total number of unflagged records
t <- length(filt$RecordType)

```

Values reported for 'RecordType': `r round((y/t) *100,1)` % of all records (n = `r y`) have .  
Percent valid values for 'RecordType: `r round((x/y) * 100,1)` %

```{r RecordTypePlotSouth, echo=FALSE, cache = FALSE, dpi = 300}
yo <- filt %>% 
  filter(RecordType %in% c('literature',  'specimen',  'still image', 'video observation', 'notation', 'catch record'))

yo <- within(yo, 
             RecordType <- factor(RecordType, 
                                  levels=names(sort(table(RecordType), decreasing=TRUE))))

options(scipen=10000)
yo <- yo %>% filter(FishCouncilRegion == 'South Atlantic'| FishCouncilRegion == 'Gulf of Mexico')
#table(yo$FishCouncilRegion)

g <- ggplot(yo, aes(RecordType, fill = Phylum)) +
    geom_bar() + 
  coord_flip() + 
  facet_wrap(~FishCouncilRegion) + 
  ylab("Number of Records") + 
  theme_bw(base_size = 15, base_family = "Cambria") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
#display.brewer.all(colorblindFriendly=TRUE)
g + scale_fill_manual(values = brewer.pal(12, "Paired")[c(10,9)])

```

```{r RecordTypePlotPacific, echo=FALSE, cache = FALSE, dpi = 300}
yo <- filt %>% 
  filter(RecordType %in% c('literature',  'specimen',  'still image', 'video observation', 'notation', 'catch record'))

yo <- within(yo, 
             RecordType <- factor(RecordType, 
                                  levels=names(sort(table(RecordType), decreasing=TRUE))))

options(scipen=10000)
yo <- yo %>% filter(FishCouncilRegion == 'Pacific'| FishCouncilRegion == 'North Pacific' | FishCouncilRegion == 'Western Pacific')
#table(yo$FishCouncilRegion)

g <- ggplot(yo, aes(RecordType, fill = Phylum)) +
  geom_bar() + 
  coord_flip() + 
  facet_wrap(~FishCouncilRegion) + 
  ylab("Number of Records") + 
  theme_bw(base_size = 15, base_family = "Cambria") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

#display.brewer.all(colorblindFriendly=TRUE)
g + scale_fill_manual(values = brewer.pal(12, "Paired")[c(10,9)])

```

```{r RecordTypePlotNewEngland, echo=FALSE, cache = FALSE, dpi = 300}
yo <- filt %>% 
  filter(RecordType %in% c('literature',  'specimen',  'still image', 'video observation', 'notation', 'catch record'))

yo <- within(yo, 
             RecordType <- factor(RecordType, 
                                  levels=names(sort(table(RecordType), decreasing=TRUE))))

options(scipen=10000)
yo <- yo %>% filter(FishCouncilRegion == 'New England'| FishCouncilRegion == 'Mid-Atlantic' | FishCouncilRegion == 'Caribbean')
#table(yo$FishCouncilRegion)

g <- ggplot(yo, aes(RecordType, fill = Phylum)) +
   geom_bar() + 
  coord_flip() + 
  facet_wrap(~FishCouncilRegion) + 
  ylab("Number of Records") + 
  theme_bw(base_size = 15, base_family = "Cambria") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

#display.brewer.all(colorblindFriendly=TRUE)
g + scale_fill_manual(values = brewer.pal(12, "Paired")[c(10,9)])

```

###### *Figure 12*: The distribution of 'RecordType' within the database, faceted by 'FishCouncilRegion'

```{r RecordTypePlotbyRegion, echo=FALSE, cache = FALSE, dpi = 300}
yo <- filt %>% 
  filter(RecordType %in% c('literature',  'specimen',  'still image', 'video observation', 'notation', 'catch record'))

yo <- within(yo, 
             RecordType <- factor(RecordType, 
                                  levels=names(sort(table(RecordType), decreasing=TRUE))))

options(scipen=10000)
#yo <- yo %>% filter(FishCouncilRegion == 'South Atlantic'| FishCouncilRegion == 'Caribbean')
#table(yo$FishCouncilRegion)

# ordering factors
options(scipen=10000)
g <- ggplot(yo, aes(RecordType, fill = Phylum)) +
  geom_bar() + 
  coord_flip() + 
  #facet_wrap(~FishCouncilRegion) + 
  ylab("Number of Records") + 
  theme_bw(base_size = 15, base_family = "Cambria") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

#display.brewer.all(colorblindFriendly=TRUE)
g + scale_fill_manual(values = brewer.pal(12, "Paired")[c(10,9)])

```

# SamplingEquipment

``` {r SamplingEquipmentCalcs, echo = FALSE}
valid <- filt$SamplingEquipment %in% c('ROV', 'AUV', 'submersible', 'drop camera', 'towed camera', 'trawl', 'net', 'dredge', 'longline', 'pot', 'hook and line', 'grab', 'corer', 'SCUBA', 'other', 'NA')

#schema %>% filter(FieldName == "SamplingEquipment") %>% dplyr::select(ValidValues)
#View(table(filt$SamplingEquipment))

# number of records with valid values
x <- length(valid[valid == T])

# number of records where the specified value is non-null 
y <- filter(filt, is.na(filt$SamplingEquipment) == FALSE)
y <- length(y$SamplingEquipment)

# number of records where the specified value is null
z <- filter(filt, is.na(filt$SamplingEquipment) == TRUE)
z <- length(z$SamplingEquipment)

# total number of unflagged records
t <- length(filt$SamplingEquipment)

```

`r round((y/t) *100,1)` % of all records (n = `r y`) have 'SamplingEquipment' values reported.  Of those records that have reported values, `r round((x/y) * 100,1)` % comply with the list of valid values specified in the data dictionary. 

```{r SamplingEquipmentPlot, echo=FALSE, cache = FALSE, dpi = 300}
yo <- filt %>% 
  filter(SamplingEquipment %in% c('ROV', 'AUV', 'submersible', 'drop camera', 'towed camera', 'trawl', 'net', 'dredge', 'longline', 'pot', 'hook and line', 'grab', 'corer', 'SCUBA', 'other', 'NA'))

yo <- within(yo, 
             SamplingEquipment <- factor(SamplingEquipment, 
                                  levels=names(sort(table(SamplingEquipment), decreasing=TRUE))))

# ordering factors
options(scipen=10000)
#yo <- yo %>% filter(FishCouncilRegion == 'South Atlantic'| FishCouncilRegion == 'Gulf of Mexico' | FishCouncilRegion == 'Caribbean')
#table(yo$FishCouncilRegion)

g <- ggplot(yo, aes(SamplingEquipment, fill = Phylum)) +
  geom_bar() + 
  coord_flip() +
  #facet_wrap(~ FishCouncilRegion) +
  ylab("Number of Records") + 
  theme_bw(base_size = 15, base_family = "Cambria") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

#display.brewer.all(colorblindFriendly=TRUE)
g + scale_fill_manual(values = brewer.pal(12, "Paired")[c(10,9)])

```

###### *Figure 14*: The distribution of 'SamplingEquipment' valid values across the entire database.

```{r SamplingEquipmentPlotbyRegionPacific, echo=FALSE, cache = FALSE, dpi = 300}
yo <- filt %>% 
  filter(SamplingEquipment %in% c('ROV', 'AUV', 'submersible', 'drop camera', 'towed camera', 'trawl', 'net', 'dredge', 'longline', 'pot', 'hook and line', 'grab', 'corer', 'SCUBA', 'other', 'NA'))

yo <- within(yo, 
             SamplingEquipment <- factor(SamplingEquipment, 
                                  levels=names(sort(table(SamplingEquipment), decreasing=TRUE))))

# ordering factors
options(scipen=10000)
yo <- yo %>% filter(FishCouncilRegion == 'North Pacific'| FishCouncilRegion == 'Pacific' | FishCouncilRegion == 'Western Pacific')
#table(yo$FishCouncilRegion)
g <- ggplot(yo, aes(SamplingEquipment, fill = Phylum)) +
  geom_bar() + 
  coord_flip() + 
  facet_wrap(~FishCouncilRegion) + 
  ylab("Number of Records") + 
  theme_bw(base_size = 15, base_family = "Cambria") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

#display.brewer.all(colorblindFriendly=TRUE)
g + scale_fill_manual(values = brewer.pal(12, "Paired")[c(10,9)])

```
 
```{r SamplingEquipmentPlotbyRegionSouth, echo=FALSE, cache = FALSE, dpi = 300}
yo <- filt %>% 
  filter(SamplingEquipment %in% c('ROV', 'AUV', 'submersible', 'drop camera', 'towed camera', 'trawl', 'net', 'dredge', 'longline', 'pot', 'hook and line', 'grab', 'corer', 'SCUBA', 'other', 'NA'))

yo <- within(yo, 
             SamplingEquipment <- factor(SamplingEquipment, 
                                  levels=names(sort(table(SamplingEquipment), decreasing=TRUE))))

options(scipen=10000)
yo <- yo %>% filter(FishCouncilRegion == 'South Atlantic'| FishCouncilRegion == 'Gulf of Mexico' | FishCouncilRegion == 'Caribbean')
g <- ggplot(yo, aes(SamplingEquipment, fill = Phylum)) +
  geom_bar() + 
  coord_flip() + 
  facet_wrap(~FishCouncilRegion) + 
    ylab("Number of Records") + 
  theme_bw(base_size = 15, base_family = "Cambria") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

#display.brewer.all(colorblindFriendly=TRUE)
g + scale_fill_manual(values = brewer.pal(12, "Paired")[c(10,9)])

```
 
```{r SamplingEquipmentPlotbyRegionNewEngland, echo=FALSE, cache = FALSE, dpi = 300}
yo <- filt %>% 
  filter(SamplingEquipment %in% c('ROV', 'AUV', 'submersible', 'drop camera', 'towed camera', 'trawl', 'net', 'dredge', 'longline', 'pot', 'hook and line', 'grab', 'corer', 'SCUBA', 'other', 'NA'))

yo <- within(yo, 
             SamplingEquipment <- factor(SamplingEquipment, 
                                  levels=names(sort(table(SamplingEquipment), decreasing=TRUE))))

# ordering factors
options(scipen=10000)
yo <- yo %>% filter(FishCouncilRegion == 'New England'| FishCouncilRegion == 'Mid-Atlantic')
#table(yo$FishCouncilRegion)
g <- ggplot(yo, aes(SamplingEquipment, fill = Phylum)) +
  geom_bar() + 
  coord_flip() + 
  facet_wrap(~FishCouncilRegion) + 
  ylab("Number of Records") + 
  theme_bw(base_size = 15, base_family = "Cambria") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

#display.brewer.all(colorblindFriendly=TRUE)
g + scale_fill_manual(values = brewer.pal(12, "Paired")[c(10,9)])

```

###### *Figure 15*: The distribution of 'SamplingEquipment' values, grouped by 'FishCouncilRegion'.

# Appendix 1: Summary of occurrence records by Family. 
The number of genera 'Genus' within in each family is reported in the column named 'Genus_n'. The column heading 'n' represents the number of records within each family. The table is sorted alphabetically to facilitate the location of families of interest.

## Corals

```{r GenusByFamTableCoral, echo=FALSE, cache = FALSE}
sum_tbl <-
 arrange(filt, Genus) %>% filter(Phylum == 'Cnidaria') %>%
  group_by(Family) %>% 
  summarize(Genus_n = length(unique(Genus)),
            n = n(),
            Genera = toString(unique(Genus)))
sum_tbl <- kable(sum_tbl, row.names = F, digits = 2)
sum_tbl
```

## Sponges
```{r GenusByFamTableSponge, echo=FALSE, cache = FALSE}
sum_tbl <-
  arrange(filt, Genus) %>% filter(Phylum == 'Porifera') %>% 
  group_by(Family) %>% 
  summarize(Genus_n = length(unique(Genus)),
            n = n(),
            Genera = toString(unique(Genus)))
sum_tbl <- kable(sum_tbl, row.names = F, digits = 2)
sum_tbl
```

# Appendix 2: Data Dictionary for the National Database
Field names, definitions, and their valid values organized in alphabetical by field name.

```{r DataDictionary, echo=FALSE, cache = FALSE}
sum_tbl <-
  s %>% filter(InternalUseOnly == '0') %>% 
  group_by(FieldName) %>% 
  summarize(Field_Definition = toString(unique(FieldDescription)),
            Valid_Values = toString(unique(ValidValues)))
sum_tbl <- kable(sum_tbl, row.names = F, digits = 2)
sum_tbl
```

# Appendix 3: DatasetID Dashboards
Each of the data sets in the database has unique 'DatasetID'. There are currently `r length(filt$CatalogNumber)` records in the database composed of `r length(unique(filt$DatasetID))` individual data sets from `r length(unique(filt$DataProvider))` different data providers. You will find an an interactive dashboard for each data set at the following location: LINK (not live yet, see reviewer note below) 

*REVEIWER NOTE*:  All of these DatasetID dashboards are now produced and have gone through one round of review from the DSCRTP Data Working Group. These dashboards will ultimately be linked at deepseacoraldata.noaa.gov but for now are available for review at the following Google Drive Location: [LINK](https://drive.google.com/open?id=19JVD2EliPBqQhYR-Iuxvm1lzzjKhKvQW). 

* *Important*: To review these dashboards with appropriate styling, etc. you must download these to your local drive an/d open from there. Clicking directly on the file in Google Drive will not work.  

