---
title: "Database Status Update - Figs and Tables Only"
author: "Robert P. McGuinn"
date: "Report last ran on: `r Sys.Date()`"
output: word_document
---
# Input database

``` {r setup_data_intake, echo=FALSE, message=FALSE, warning=FALSE, cache = TRUE}

# setwd("C:/rworking/deepseatools/indata")
# indata<-read.csv("DSCRTP_NatDB_20190920-0.csv", header = T)
# filt <- indata %>%
#   filter(Flag == "0")

```

# data transformations where needed

``` {r setup_data_prep, echo=FALSE, warning=FALSE, message=FALSE}

# filt <- filt %>% filter(is.na(Phylum) == F)
                         
##### checker #####

# length(filt1$CatalogNumber)

# sum_tbl <- filt %>%
#    # filter(as.numeric(IndividualCount) > 100,
#    #        grepl('Etnoyer', PI),
#    #        DatasetID == 'NOAA_SJ-10-08'
#    #       ) %>%
#   group_by(AccessionID) %>% 
#   summarize(DataProvider = toString(unique(DataProvider)),
#             Repository = toString(unique(Repository)),
#             Vessel = toString(unique(Vessel)),
#             PI = toString(unique(PI)),
#             SampleID = toString(unique(SampleID)),
#             ScientificName = toString(unique(ScientificName)),
#             ImageFilePath = toString(unique(ImageFilePath)),
#             Records = n()
#   ) %>% 
#   arrange(desc(Records))
# View(sum_tbl)

```

# input schema

``` {r setup_schema, echo=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
##### download Google Sheet version of schema for use in R  documents ##### 
# Register and download Google Sheet
s <- gs_title('2019_DSCRTP_National_Database_Schema')
#s <- gs_key('1YDskzxY8OF-34Q8aI04tZvlRbhGZqBSysuie39kYHoI')
s <- gs_read(s)
# s<- s[,-(33:35)]
#names(s)

```

# knitr options

``` {r setup_opts, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
##### knitr options #####
knitr::opts_chunk$set(fig.width=6.5, fig.height=4, warning = FALSE, dpi = 300, dev = c('png'), fig.path="C:/rworking/deepseatools/reports/2019_status_update_report/figures/") 

options(scipen=10000)

##### setting digit options #####
options(digits = 0)
```


``` {r setup_variables,  echo=FALSE, warning=FALSE, message=FALSE}
version <- '20190920-0'
```

# schema metrics variables

``` {r setup_schema_metrics, echo=FALSE}
#total number of fields in the NDB
a <- length(s$FieldName)

#total number of published fields
b <- s %>%
  filter(InternalUseOnly == '0')
b <- length(b$FieldName)

#total number of unpublished fields
c <- s %>%
  filter(InternalUseOnly == '1')
c <- length(c$FieldName)

#PointHist required
d <- s %>%
  filter(PointHist == 'R')
d <- length(d$FieldName)

#PointNew required
e <- s %>%
  filter(PointNew == 'R')
e <- length(e$FieldName)

#PointProgram required
f <- s %>%
  filter(PointProgram == 'R')
f <- length(f$FieldName)

#TransHist required
g <- s %>%
  filter(TransHist == 'R')
g <- length(g$FieldName)

#TransNew required
h <- s %>%
  filter(TransNew == 'R')
h <- length(h$FieldName)

#TransProgram required
i <- s %>%
  filter(TransProgram == 'R')
i <- length(i$FieldName)

#TrawlHist required
j <- s %>%
  filter(TrawlHist == 'R')
j <- length(j$FieldName)

#TrawlNew required
k <- s %>%
  filter(TrawlNew == 'R')
k <- length(k$FieldName)

#TrawlProgram required
l <- s %>%
  filter(TrawlProgram == 'R')
l <- length(l$FieldName)

```

# table: data providers

``` {r dataprovider_table, echo=FALSE, caption = 'test caption'}
options(scipen=999)

x <-
  filt %>%
  group_by(DataProvider) %>%
  summarize(
    Number_Datasets = length(unique(DatasetID)),
    Records = n()
            )

colkeys <- names(x[,-1])
ft <- flextable(x) %>% 
  theme_box() %>% #booktabs, vanilla, box, tron, tron_legacy
  set_header_labels(Number_Datasets = "# of Data Sets", Records = "# of Records") %>% 
  colformat_num(col_keys = colkeys, digits = 0, big.mark = ",", na_str = "NA") %>% 
  font(fontname = "Cambria", part = "all")  %>% #header, body, all
  autofit() %>% 
  width(j = ~ DataProvider, width = 4.2)
  

ft

read_docx() %>% 
  body_add_flextable(ft) %>% 
  print(target = "c:/rworking/deepseatools/reports/2019_status_update_report/tables/dataprovider_table.docx")

```

# table: public fields

``` {r public_fields_table, echo=FALSE}

s$DSCRTPCategory <- factor(s$DSCRTPCategory, levels = c("Survey Data","Event Data","Observation Data","TaxonomicID","Environment","Occurrence Detail","GIS Enhancement","Metadata","Record Keeping","Flag"))

x <-
  s %>% arrange(FieldName) %>% 
  filter(InternalUseOnly == '0') %>%
  group_by(DSCRTPCategory) %>% 
  summarize(Field_Names = toString(unique(FieldName)),
            n = n())


colkeys <- names(x[,-1])
ft <- flextable(x) %>% 
  theme_vanilla() %>% #booktabs, vanilla, box, tron, tron_legacy
  set_header_labels(Field_Names = "Field Name", n = "# of Records", DSCRTPCategory = "DSCRTP Category") %>% 
  colformat_num(col_keys = colkeys, digits = 0, big.mark = ",", na_str = "NA") %>% 
  font(fontname = "Cambria", part = "all")  %>% #header, body, all
  fontsize(size = 10, part = "body") %>% 
  autofit() %>% 
  width(j = ~ n, width = 1.2) %>% 
  width(j = ~ DSCRTPCategory, width = 1.5) %>% 
  width(j = ~ Field_Names, width = 4) %>% 
  align(align = "left") %>% 
  align_text_col(align = "left") %>% 
  align_nottext_col(align = "right")


ft

read_docx() %>% 
  body_add_flextable(ft) %>% 
  print(target = "c:/rworking/deepseatools/reports/2019_status_update_report/tables/public_fields_table.docx")



# sum_tbl <- kable(sum_tbl, 
#                  row.names = F, 
#                  digits = 2,
#                  col.names = c('Field Category', 'Field Name', 'Number of Fields'))
# 
# options(scipen=10000)
# sum_tbl

#setwd("C:/rworking/digs/outdata")
#write.csv(sum_tbl,"20180216_0_Field_Summary_Schema_20180216-0.csv", row.names = F, quote = T)

```

# table: required/desired points

``` {r required_desired_points_table, echo=FALSE,  warning=FALSE}
x <-
  s %>%
  filter(InternalUseOnly == '0',
         PointHist == 'R' | PointNew == 'R' | PointProgram == 'R'
         ) %>%
  group_by(DSCRTPCategory, FieldName) %>% 
  summarize(PointHist = unique(PointHist),
            PointNew = unique(PointNew),
            PointProgram = unique(PointProgram)
                        )
colkeys <- names(x[,-1])

ft <- flextable(x) %>% 
  theme_vanilla() %>% #booktabs, vanilla, box, tron, tron_legacy
 # set_header_labels(Field_Names = "Field Name", n = "# of Records") %>% 
  colformat_num(col_keys = colkeys, digits = 0, big.mark = ",", na_str = "NA") %>% 
  set_header_labels(FieldName = "Field Name",
                    DSCRTPCategory = "DSCRTP Category") %>% 
  font(fontname = "Cambria", part = "all") %>% #header, body, all
  fontsize(size = 10, part = "body") %>% 
  autofit() 

ft

read_docx() %>% 
  body_add_flextable(ft) %>% 
  print(target = "c:/rworking/deepseatools/reports/2019_status_update_report/tables/required_desired_points_table.docx")

```

# table: required/desired transects

``` {r required_desired_transects_table, echo=FALSE}
x <-
  s %>%
  filter(InternalUseOnly == '0',
         TransHist == 'R' | TransNew == 'R' | TransProgram == 'R'
         ) %>%
  group_by(DSCRTPCategory, FieldName) %>% 
  summarize(TransHist = unique(TransHist),
            TransNew = unique(TransNew),
            TransProgram = unique(TransProgram)
                        ) %>% 
  arrange(DSCRTPCategory)

colkeys <- names(x[,-1])

ft <- flextable(x) %>% 
  theme_vanilla() %>% #booktabs, vanilla, box, tron, tron_legacy
 # set_header_labels(Field_Names = "Field Name", n = "# of Records") %>% 
  colformat_num(col_keys = colkeys, digits = 0, big.mark = ",", na_str = "NA") %>% 
  set_header_labels(FieldName = "Field Name",
                    DSCRTPCategory = "DSCRTP Category") %>% 
  font(fontname = "Cambria", part = "all") %>% #header, body, all
  fontsize(size = 10, part = "body") %>% 
  autofit() 

ft

read_docx() %>% 
  body_add_flextable(ft) %>% 
  print(target = "c:/rworking/deepseatools/reports/2019_status_update_report/tables/required_desired_transects_table.docx")


```

# table: required/desired trawl

``` {r required_desired_trawl_table, echo=FALSE}
x <-
  s %>%
  filter(InternalUseOnly == '0',
         TrawlHist == 'R' | TrawlNew == 'R' | TrawlProgram == 'R'
         ) %>%
  group_by(DSCRTPCategory, FieldName) %>% 
  summarize(TrawlHist = unique(TrawlHist),
            TrawlNew = unique(TrawlNew),
            TrawlProgram = unique(TrawlProgram)
                        ) 

colkeys <- names(x[,-1])

ft <- flextable(x) %>% 
  theme_vanilla() %>% #booktabs, vanilla, box, tron, tron_legacy
 # set_header_labels(Field_Names = "Field Name", n = "# of Records") %>% 
  colformat_num(col_keys = colkeys, digits = 0, big.mark = ",", na_str = "NA") %>% 
  set_header_labels(FieldName = "Field Name",
                    DSCRTPCategory = "DSCRTP Category") %>% 
  font(fontname = "Cambria", part = "all") %>% #header, body, all
  fontsize(size = 10, part = "body") %>% 
  autofit() 

ft

read_docx() %>% 
  body_add_flextable(ft) %>% 
  print(target = "c:/rworking/deepseatools/reports/2019_status_update_report/tables/required_desired_trawl_table.docx")


```
  
# table: obis
  
``` {r obis_table, echo=FALSE}
options(scipen=999)
x <-
  s %>% arrange(DarwinCoreTerm) %>% 
  filter(ReleaseToOBIS == '1') %>% 
  group_by(FieldName) %>%
  summarize(
    DarwinCoreTerm = toString(unique(DarwinCoreTerm))
            )

colkeys <- names(x)

ft <- flextable(x) %>% 
  theme_box() %>% #booktabs, vanilla, box, tron, tron_legacy
  set_header_labels(DarwinCoreTerm = "DarwinCore Term",
                    FieldName = "DSCRTP Field Name"
                    ) %>% 
  colformat_num(col_keys = colkeys, digits = 0, big.mark = ",", na_str = "NA") %>% 
  font(fontname = "Cambria", part = "all") %>% #header, body, all
  fontsize(size = 10, part = "body") %>% 
  autofit() %>% 
  width(j = ~ DarwinCoreTerm, width = 2.2) %>% 
  width(j = ~ FieldName, width = 2.2) %>% 
  #width(j = ~ DSCRTPCategory, width = 2) %>% 
  align(align = "right", part = 'all') 
  
ft

read_docx() %>% 
  body_add_flextable(ft) %>% 
  print(target = "c:/rworking/deepseatools/reports/2019_status_update_report/tables/obis_table.docx")
```

# calculatons

``` {r calculations, echo=FALSE}
# length of unflagged records 
a <- round(length(filt$CatalogNumber), -4)

# number of DatasetID
b <- length(unique(filt$DatasetID))

# number of DataProviders
c <- length(unique(filt$DataProvider))

# number of flagged records
d <- length(filter(indata, Flag == '1')$CatalogNumber)

# percent of unflagged records with images
filt2 <- filter(filt, is.na(ImageURL) == F)
e <- round((length(filt2$CatalogNumber)/(length(filt$CatalogNumber))*100),0)

# number of unflagged records with images 
f <- length(filter(filt, is.na(ImageURL) == F)$CatalogNumber)

# stats on number of records per DatasetID
y <- count(filt, DatasetID)
g <- round(median(y$n),0)
h <- min(y$n)
i <- max(y$n)

```

# figure: growth through time

``` {r GrowthThroughTime, echo=FALSE, cache = FALSE, dpi=300}

# ordering factors

filt2 <- filt %>%
  filter(as.Date(EntryDate) > as.Date('2010-01-01'))
options(scipen=10000)
g <- ggplot(filt2,aes(x=as.Date(EntryDate), color=Phylum)) +
  stat_bin(data = subset(filt2, Phylum =="Porifera"),aes(y=cumsum(..count..)),geom="line", binwidth = 2, size = 2) +
  stat_bin(data = subset(filt2, Phylum =="Cnidaria"),aes(y=cumsum(..count..)),geom="line", binwidth = 2, size = 2) +
  ylab("Cumulative Number of Records") + 
  xlab("Year Record Added") + 
  guides(fill=FALSE) + 
  theme_bw(base_size = 15, base_family = "Cambria")

g + scale_color_manual(values = brewer.pal(12, "Paired")[c(10,9)])


```

# calculations: regional

``` {r region-counts, echo=FALSE}
a <- filt %>% 
  filter(
    is.na(FishCouncilRegion) == T
    ) 
a <- length(a$CatalogNumber)/length(filt$CatalogNumber)
a <- round(a*100, 0)

```

# figure: fish council region

``` {r FishCouncilRegion, echo=FALSE}
# ordering factors

filt2 <- filt %>% 
  filter(is.na(Phylum) == F , 
         is.na(FishCouncilRegion) == F , 
         FishCouncilRegion != 'NA')
options(scipen=10000)

#table(filt2$FishCouncilRegion, useNA = 'always')

filt2$FishCouncilRegion <- factor(filt2$FishCouncilRegion, levels = c('Pacific', 'North Pacific','Western Pacific','New England','Mid-Atlantic','South Atlantic', 'Gulf of Mexico','Caribbean'))

g <- ggplot(filt2, aes(FishCouncilRegion, fill = Phylum)) +
  geom_bar() + 
  coord_flip() + 
  theme(text = element_text(size=20)) + 
  ylab("Number of Records") + 
  theme_bw(base_size = 15, base_family = "Cambria")

#display.brewer.all(colorblindFriendly=TRUE)
g + scale_fill_manual(values = brewer.pal(12, "Paired")[c(10,9)])

# write to file
# x <- g + scale_fill_manual(values = brewer.pal(12, "Paired")[c(10,9)])
# 
# png(file = "C:/rworking/deepseatools/reports/2019_status_update_report/fishcouncilregion.png",
#     width = 1200, height = 800, 
#     units = "px", pointsize = 12, res = 150)
# 
# x
# 
# dev.off()



```

# calculations: depth

``` {r DepthCalc, echo=FALSE}
x <- s %>%
  filter(
    DSCRTPGroup == 'Location-Depth Metric',
    InternalUseOnly == "0", 
    DSCRTPCategory != "GIS Enhancement" 
  )
x <- toString(x$FieldName)

y <- s %>%
  filter(DSCRTPGroup == 'Location-Depth Metric', InternalUseOnly == "0", DSCRTPCategory == "GIS Enhancement" )
y <- toString(y$FieldName)

##### checker ##### 
# z <- filt %>% 
#   filter(DepthInMeters == "-999") %>% 
#   group_by(DatasetID) %>% 
#   summarize(
#     DepthInMeters = toString(unique(DepthInMeters)),
#     DataProvider = toString(unique(DataProvider)),
#     n = n()
#     )
# #View(z)
# #length(z$CatalogNumber)
# #[1] 3992
```

# figure: depth distribution

``` {r depth_vernacularname, echo=FALSE}
##### Boxplot of depth by FishCouncilRegion #####
filt2 <- filt %>% 
  filter(DepthInMeters != "-999")
  
g <- ggplot(filt2, aes(factor(FishCouncilRegion), as.numeric(DepthInMeters), color = Phylum)) +   geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") + 
  xlab("FishCouncilRegion") + 
  theme_bw(base_size = 15, base_family = "Cambria") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = -.01))

#display.brewer.all(colorblindFriendly=TRUE)
g + scale_color_manual(values = brewer.pal(12, "Paired")[c(10,9)])

```

# figure: depth distribution density plot

``` {r DepthDensity, echo=FALSE,  dpi = 300}
filt2 <- filt %>% 
  filter(DepthInMeters != "-999")
options(scipen=10000)

g <- ggplot(filt2, aes(DepthInMeters, fill = Phylum)) +
  geom_density(alpha = 0.5) + 
  xlim(0,7000) +
  ylab("Proportion of Records") + 
  xlab('DepthInMeters') + 
  theme_bw(base_size = 15, base_family = "Cambria")

#display.brewer.all(colorblindFriendly=TRUE)
g + scale_fill_manual(values = brewer.pal(12, "Paired")[c(10,9)])
```

# figure: observation year

``` {r ObservationYear, echo=FALSE,  dpi = 300}
# ordering factors
filt2 <- filt %>% 
  filter(is.na(ObservationYear) == F, ObservationYear != "-999", is.na(Phylum) == F)
filt2$TaxonRank <- factor(filt2$TaxonRank, levels = c("phylum", "class", "subclass", "order", "suborder","family","subfamily", "genus","subgenus","species","subspecies","forma","variety"))
options(scipen=10000)

g <- ggplot(filt2, aes(ObservationYear, fill = Phylum)) +
  geom_bar() + 
  ylab("Number of Records")  +
  theme_bw(base_size = 15, base_family = "Cambria")

#display.brewer.all(colorblindFriendly=TRUE)
g + scale_fill_manual(values = brewer.pal(12, "Paired")[c(10,9)])

```

# figure: individual count

``` {r IndividualCount, echo=FALSE,  dpi = 300}
# ordering factors
filt2 <- filt %>% 
  filter(is.na(IndividualCount) == F, 
         IndividualCount != '0',
         IndividualCount != '-999', 
         is.na(Phylum) == F, 
         IndividualCount < 50
         )

options(scipen=10000)

g <- ggplot(filt2, aes(IndividualCount, fill = Phylum)) +
  geom_histogram(binwidth = 1) + 
  ylab("Number of Records")  +
  theme_bw(base_size = 15, base_family = "Cambria")

#display.brewer.all(colorblindFriendly=TRUE)
g + scale_fill_manual(values = brewer.pal(12, "Paired")[c(10,9)])

```

# calculations: size

``` {r SizeCalc, echo=FALSE,  warning = FALSE}
yo <- filt %>% 
  filter(
    MinimumSize != "-999",
    MaximumSize != "-999",
    MinimumSize != '0',
    is.na(Phylum) == F
    )

yo$AverageSize <- (yo$MinimumSize + yo$MaximumSize)/2
yo$SizeDiff <- (yo$MaximumSize - yo$MinimumSize)

x <- length(yo$CatalogNumber)
y <- length(filt$CatalogNumber)
```

# figure: size

``` {r Size, echo=FALSE,  warning = FALSE, dpi = 300}
yo <- yo %>%
  filter(
   AverageSize < 100  
    )

# ordering factors
options(scipen=10000)
g <- ggplot(yo, aes(as.numeric(AverageSize), fill = Phylum)) +
  geom_histogram(bins = 50) + 
  #coord_cartesian(ylim = c(1, 20000)) +
  ylab("Number of Records") + 
  xlab("Average Size (cm)")  +
  theme_bw(base_size = 15, base_family = "Cambria")

#display.brewer.all(colorblindFriendly=TRUE)
g + scale_fill_manual(values = brewer.pal(12, "Paired")[c(10,9)])

```

# calculations: condition

``` {r ConditionCalc, echo=FALSE}

condition <- filt$Condition %in% c('Live','Damaged','Dead')
x <- length(condition[condition == T])
y <- filter(filt, is.na(filt$Condition) == FALSE)
y <- length(y$Condition)
z <- filter(filt, is.na(filt$Condition) == TRUE)
z <- length(z$Condition)
t <- length(filt$Condition)

```

# figure: condition

``` {r Condition, echo=FALSE,  dpi = 300}
yo <- filt %>% 
  filter(
    Condition %in% c('Live','Damaged','Dead') 
    #Condition != 'Live'
         )

# ordering factors
yo$Condition <- factor(yo$Condition, 
                       levels = c('Live','Damaged','Dead'))


options(scipen=10000)
g <- ggplot(yo, aes(Condition, fill = Phylum)) +
  geom_bar() + 
  coord_flip() + 
  ylab("Number of Records") + 
  theme_bw(base_size = 15, base_family = "Cambria")

#display.brewer.all(colorblindFriendly=TRUE)
g + scale_fill_manual(values = brewer.pal(12, "Paired")[c(10,9)])

```

# calculations: 

``` {r Location, echo=FALSE}
z <- s %>%
  filter(DSCRTPGroup == 'Taxonomy Metadata', InternalUseOnly == "0", FieldName != 'AssociatedTaxa', FieldName != 'Synonyms')
z <-toString(z$FieldName)

```



``` {r taxonranksum, echo=FALSE,  dpi = 300}
# ordering factors
filt2 <- filt %>% 
  filter(is.na(TaxonRank) == F, is.na(Phylum) == F)

filt2$TaxonRank <- factor(filt2$TaxonRank, levels = c("phylum", "class", "subclass", "order", "suborder","family","subfamily", "genus","subgenus","species","subspecies","forma","variety"))
options(scipen=10000)

g <- ggplot(filt2, aes(TaxonRank, fill = Phylum)) +
  geom_bar() + 
  coord_flip() + 
  ylab("Number of Records") + 
  theme_bw(base_size = 15, base_family = "Cambria")

#display.brewer.all(colorblindFriendly=TRUE)
g + scale_fill_manual(values = brewer.pal(12, "Paired")[c(10,9)])
```

# figure: recordtype

``` {r recordtype, echo = FALSE}
valid <- filt$RecordType %in% c('literature',  'specimen',  'still image', 'video observation', 'video transect','notation', 'catch record')

# number of records with valid values
x <- length(valid[valid == T])

# number of records where the specified value is non-null 
y <- filter(filt, is.na(filt$RecordType) == FALSE)
y <- length(y$RecordType)

# number of records where the specified value is null
z <- filter(filt, is.na(filt$RecordType) == TRUE)
z <- length(z$RecordType)

# total number of unflagged records
t <- length(filt$RecordType)

```

``` {r RecordTypePlotSouth, echo=FALSE,  dpi = 300}
yo <- filt %>% 
  filter(RecordType %in% c('literature',  'specimen',  'still image', 'video observation', 'notation', 'catch record'))

yo <- within(yo, 
             RecordType <- factor(RecordType, 
                                  levels=names(sort(table(RecordType), decreasing=TRUE))))

options(scipen=10000)
yo <- yo %>% filter(FishCouncilRegion == 'South Atlantic'| FishCouncilRegion == 'Gulf of Mexico')
#table(yo$FishCouncilRegion)

g <- ggplot(yo, aes(RecordType, fill = Phylum)) +
    geom_bar() + 
  coord_flip() + 
  facet_wrap(~FishCouncilRegion) + 
  ylab("Number of Records") + 
  theme_bw(base_size = 15, base_family = "Cambria") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
#display.brewer.all(colorblindFriendly=TRUE)
g + scale_fill_manual(values = brewer.pal(12, "Paired")[c(10,9)])

```

``` {r RecordTypePlotPacific, echo=FALSE,  dpi = 300}
yo <- filt %>% 
  filter(RecordType %in% c('literature',  'specimen',  'still image', 'video observation', 'notation', 'catch record'))

yo <- within(yo, 
             RecordType <- factor(RecordType, 
                                  levels=names(sort(table(RecordType), decreasing=TRUE))))

options(scipen=10000)
yo <- yo %>% filter(FishCouncilRegion == 'Pacific'| FishCouncilRegion == 'North Pacific' | FishCouncilRegion == 'Western Pacific')
#table(yo$FishCouncilRegion)

g <- ggplot(yo, aes(RecordType, fill = Phylum)) +
  geom_bar() + 
  coord_flip() + 
  facet_wrap(~FishCouncilRegion) + 
  ylab("Number of Records") + 
  theme_bw(base_size = 15, base_family = "Cambria") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

#display.brewer.all(colorblindFriendly=TRUE)
g + scale_fill_manual(values = brewer.pal(12, "Paired")[c(10,9)])

```

``` {r RecordTypePlotNewEngland, echo=FALSE,  dpi = 300}
yo <- filt %>% 
  filter(RecordType %in% c('literature',  'specimen',  'still image', 'video observation', 'notation', 'catch record'))

yo <- within(yo, 
             RecordType <- factor(RecordType, 
                                  levels=names(sort(table(RecordType), decreasing=TRUE))))

options(scipen=10000)
yo <- yo %>% filter(FishCouncilRegion == 'New England'| FishCouncilRegion == 'Mid-Atlantic' | FishCouncilRegion == 'Caribbean')
#table(yo$FishCouncilRegion)

g <- ggplot(yo, aes(RecordType, fill = Phylum)) +
   geom_bar() + 
  coord_flip() + 
  facet_wrap(~FishCouncilRegion) + 
  ylab("Number of Records") + 
  theme_bw(base_size = 15, base_family = "Cambria") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

#display.brewer.all(colorblindFriendly=TRUE)
g + scale_fill_manual(values = brewer.pal(12, "Paired")[c(10,9)])

```

``` {r RecordTypePlotbyRegion, echo=FALSE,  dpi = 300}
yo <- filt %>% 
  filter(RecordType %in% c('literature',  'specimen',  'still image', 'video observation', 'notation', 'catch record'))

yo <- within(yo, 
             RecordType <- factor(RecordType, 
                                  levels=names(sort(table(RecordType), decreasing=TRUE))))

options(scipen=10000)
#yo <- yo %>% filter(FishCouncilRegion == 'South Atlantic'| FishCouncilRegion == 'Caribbean')
#table(yo$FishCouncilRegion)

# ordering factors
options(scipen=10000)
g <- ggplot(yo, aes(RecordType, fill = Phylum)) +
  geom_bar() + 
  coord_flip() + 
  #facet_wrap(~FishCouncilRegion) + 
  ylab("Number of Records") + 
  theme_bw(base_size = 15, base_family = "Cambria") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

#display.brewer.all(colorblindFriendly=TRUE)
g + scale_fill_manual(values = brewer.pal(12, "Paired")[c(10,9)])

```

``` {r SamplingEquipmentCalcs, echo = FALSE}
valid <- filt$SamplingEquipment %in% c('ROV', 'AUV', 'submersible', 'drop camera', 'towed camera', 'trawl', 'net', 'dredge', 'longline', 'pot', 'hook and line', 'grab', 'corer', 'SCUBA', 'other', 'NA')

#schema %>% filter(FieldName == "SamplingEquipment") %>% dplyr::select(ValidValues)
#View(table(filt$SamplingEquipment))

# number of records with valid values
x <- length(valid[valid == T])

# number of records where the specified value is non-null 
y <- filter(filt, is.na(filt$SamplingEquipment) == FALSE)
y <- length(y$SamplingEquipment)

# number of records where the specified value is null
z <- filter(filt, is.na(filt$SamplingEquipment) == TRUE)
z <- length(z$SamplingEquipment)

# total number of unflagged records
t <- length(filt$SamplingEquipment)

```

``` {r SamplingEquipmentPlot, echo=FALSE,  dpi = 300}
yo <- filt %>% 
  filter(SamplingEquipment %in% c('ROV', 'AUV', 'submersible', 'drop camera', 'towed camera', 'trawl', 'net', 'dredge', 'longline', 'pot', 'hook and line', 'grab', 'corer', 'SCUBA', 'other', 'NA'))

yo <- within(yo, 
             SamplingEquipment <- factor(SamplingEquipment, 
                                  levels=names(sort(table(SamplingEquipment), decreasing=TRUE))))

# ordering factors
options(scipen=10000)
#yo <- yo %>% filter(FishCouncilRegion == 'South Atlantic'| FishCouncilRegion == 'Gulf of Mexico' | FishCouncilRegion == 'Caribbean')
#table(yo$FishCouncilRegion)

g <- ggplot(yo, aes(SamplingEquipment, fill = Phylum)) +
  geom_bar() + 
  coord_flip() +
  #facet_wrap(~ FishCouncilRegion) +
  ylab("Number of Records") + 
  theme_bw(base_size = 15, base_family = "Cambria") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

#display.brewer.all(colorblindFriendly=TRUE)
g + scale_fill_manual(values = brewer.pal(12, "Paired")[c(10,9)])

```

``` {r SamplingEquipmentPlotbyRegionPacific, echo=FALSE,  dpi = 300}
yo <- filt %>% 
  filter(SamplingEquipment %in% c('ROV', 'AUV', 'submersible', 'drop camera', 'towed camera', 'trawl', 'net', 'dredge', 'longline', 'pot', 'hook and line', 'grab', 'corer', 'SCUBA', 'other', 'NA'))

yo <- within(yo, 
             SamplingEquipment <- factor(SamplingEquipment, 
                                  levels=names(sort(table(SamplingEquipment), decreasing=TRUE))))

# ordering factors
options(scipen=10000)
yo <- yo %>% filter(FishCouncilRegion == 'North Pacific'| FishCouncilRegion == 'Pacific' | FishCouncilRegion == 'Western Pacific')
#table(yo$FishCouncilRegion)
g <- ggplot(yo, aes(SamplingEquipment, fill = Phylum)) +
  geom_bar() + 
  coord_flip() + 
  facet_wrap(~FishCouncilRegion) + 
  ylab("Number of Records") + 
  theme_bw(base_size = 15, base_family = "Cambria") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

#display.brewer.all(colorblindFriendly=TRUE)
g + scale_fill_manual(values = brewer.pal(12, "Paired")[c(10,9)])

```
 
``` {r SamplingEquipmentPlotbyRegionSouth, echo=FALSE,  dpi = 300}
yo <- filt %>% 
  filter(SamplingEquipment %in% c('ROV', 'AUV', 'submersible', 'drop camera', 'towed camera', 'trawl', 'net', 'dredge', 'longline', 'pot', 'hook and line', 'grab', 'corer', 'SCUBA', 'other', 'NA'))

yo <- within(yo, 
             SamplingEquipment <- factor(SamplingEquipment, 
                                  levels=names(sort(table(SamplingEquipment), decreasing=TRUE))))

options(scipen=10000)
yo <- yo %>% filter(FishCouncilRegion == 'South Atlantic'| FishCouncilRegion == 'Gulf of Mexico' | FishCouncilRegion == 'Caribbean')
g <- ggplot(yo, aes(SamplingEquipment, fill = Phylum)) +
  geom_bar() + 
  coord_flip() + 
  facet_wrap(~FishCouncilRegion) + 
    ylab("Number of Records") + 
  theme_bw(base_size = 15, base_family = "Cambria") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

#display.brewer.all(colorblindFriendly=TRUE)
g + scale_fill_manual(values = brewer.pal(12, "Paired")[c(10,9)])

```
 
``` {r SamplingEquipmentPlotbyRegionNewEngland, echo=FALSE,  dpi = 300}
yo <- filt %>% 
  filter(SamplingEquipment %in% c('ROV', 'AUV', 'submersible', 'drop camera', 'towed camera', 'trawl', 'net', 'dredge', 'longline', 'pot', 'hook and line', 'grab', 'corer', 'SCUBA', 'other', 'NA'))

yo <- within(yo, 
             SamplingEquipment <- factor(SamplingEquipment, 
                                  levels=names(sort(table(SamplingEquipment), decreasing=TRUE))))

# ordering factors
options(scipen=10000)
yo <- yo %>% filter(FishCouncilRegion == 'New England'| FishCouncilRegion == 'Mid-Atlantic')
#table(yo$FishCouncilRegion)
g <- ggplot(yo, aes(SamplingEquipment, fill = Phylum)) +
  geom_bar() + 
  coord_flip() + 
  facet_wrap(~FishCouncilRegion) + 
  ylab("Number of Records") + 
  theme_bw(base_size = 15, base_family = "Cambria") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

#display.brewer.all(colorblindFriendly=TRUE)
g + scale_fill_manual(values = brewer.pal(12, "Paired")[c(10,9)])

```

# table: coral

``` {r genus_by_family_coral_table, echo=FALSE}
x <-
 arrange(filt, Genus) %>% filter(Phylum == 'Cnidaria') %>%
  group_by(Family) %>% 
  summarize(Genus_n = length(unique(Genus)),
            n = n(),
            Genera = toString(unique(Genus)))

colkeys <- names(x[,-1])
ft <- flextable(x) %>% 
  theme_zebra() %>% #booktabs, vanilla, box, tron, tron_legacy, vader
  set_header_labels(Genus_n = "# of Genera", n = "# of Records") %>% 
  colformat_num(col_keys = colkeys, digits = 0, big.mark = ",", na_str = "NA") %>% 
  font(fontname = "Cambria", part = "all")  %>% #header, body, all
  autofit() %>% 
  width(j = ~ Genera, width = 3.5) %>% 
  width(j = ~ Genus_n + n + Family, width = 1.2) %>% 
  width(j = ~ Family, width = 1.5)

ft

read_docx() %>% 
  body_add_flextable(ft) %>% 
  print(target = "c:/rworking/deepseatools/reports/2019_status_update_report/tables/genus_by_family_coral_table.docx")
```

## table:  sponges

``` {r genus_by_family_sponge_table, echo=FALSE}
x <-
  arrange(filt, Genus) %>% filter(Phylum == 'Porifera') %>% 
  group_by(Family) %>% 
  summarize(Genus_n = length(unique(Genus)),
            n = n(),
            Genera = toString(unique(Genus)))

colkeys <- names(x[,-1])
ft <- flextable(x) %>% 
  theme_zebra() %>% #booktabs, vanilla, box, tron, tron_legacy, zebra
  set_header_labels(Genus_n = "# of Genera", n = "# of Records") %>% 
  colformat_num(col_keys = colkeys, digits = 0, big.mark = ",", na_str = "NA") %>% 
  font(fontname = "Cambria", part = "all")  %>% #header, body, all
  autofit() %>% 
  width(j = ~ Genera, width = 3.5) %>% 
  width(j = ~ Genus_n + n + Family, width = 1.2) %>% 
  width(j = ~ Family, width = 1.5)

  

ft

read_docx() %>% 
  body_add_flextable(ft) %>% 
  print(target = "c:/rworking/deepseatools/reports/2019_status_update_report/tables/genus_by_family_sponge_table.docx")
```

## table: data dictionary

``` {r data_dictionary_table, echo=FALSE}
x <-
  s %>% filter(InternalUseOnly == '0') %>% 
  group_by(FieldName) %>% 
  summarize(Field_Definition = toString(unique(FieldDescription)),
            Valid_Values = toString(unique(ValidValues)))

colkeys <- names(x[,-1])
ft <- flextable(x) %>% 
  theme_box() %>% #booktabs, vanilla, box, tron, tron_legacy, zebra
  set_header_labels(Field_Definition = "Field Definition", 
                  Valid_Values = "Valid Values",
                    FieldName = "Field Name") %>% 
  colformat_num(col_keys = colkeys, digits = 0, big.mark = ",", na_str = "NA") %>% 
  font(fontname = "Cambria", part = "all")  %>% #header, body, all
  fontsize(size = 10, part = "body") %>% 
  autofit() %>% 
  width(j = ~ FieldName , width = 1.7) %>%
  width(j = ~ Field_Definition, width = 3.5) %>% 
  width(j = ~ Valid_Values, width = 2) %>%
  align(align = "left", part = 'all')
  


  #width(j = ~ DSCRTPCategory, width = 2) %>% 
  

ft

read_docx() %>% 
  body_add_flextable(ft) %>% 
  print(target = "c:/rworking/deepseatools/reports/2019_status_update_report/tables/data_dictionary_table.docx")

```




