---
title: "scratch_NMDS_plots"
author: "Robert McGuinn"
date: "9/17/2021"
output: html_document
---

### NMDS plot with depth bins (alternate)

```{r plotting_NMDS_depth_2}
##### plotting NMDS ##### 
ggplot() +
  geom_point(data=site.scores,aes(x=NMDS1,y=NMDS2, color=DepthCat), size=5) + # add the point markers
  geom_text(data=site.scores,aes(x=NMDS1,y=NMDS2, label = site), size=4, position = position_nudge(y = .0001)) + # add the text
  #geom_point(data=species.scores,aes(x=NMDS1,y=NMDS2), shape = 3, size=4) + 
  coord_equal() +
  theme_bw(base_size = 15, base_family = "Cambria") +
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank())

```

### Plot of NMDS with group colors from temperature

```{r plotting_NMDS_temp_2}

ggplot() +
  geom_point(data=site.scores,aes(x=NMDS2,y=NMDS1, color=temp_cat), size=5) + # add the point markers
  geom_text(data=site.scores,aes(x=NMDS2,y=NMDS1, label = site), size=4, position = position_nudge(y = .0001)) + # add the text
  #geom_point(data=species.scores,aes(x=NMDS1,y=NMDS2), shape = 3, size=4) + 
  coord_equal() +
  theme_bw(base_size = 15, base_family = "Cambria") +
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank())

```

### Plot of NMDS with group colors from location 

Location based on latitude break at X.  

```{r}

ggplot() +
  geom_point(data=site.scores,aes(x=NMDS2,y=NMDS1, color=loc), size=5) + # add the point markers
  geom_text(data=site.scores,aes(x=NMDS2,y=NMDS1, label = site), size=4, position = position_nudge(y = .0001)) + # add the text
  #geom_point(data=species.scores,aes(x=NMDS1,y=NMDS2), shape = 3, size=4) + 
  coord_equal() +
  theme_bw(base_size = 15, base_family = "Cambria") +
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank())

```

### Plot of NMDS with group colors from oxygen values 


```{r}

ggplot() +
  geom_point(data=site.scores,aes(x=NMDS2,y=NMDS1, color=ox_cat), size=5) + # add the point markers
  geom_text(data=site.scores,aes(x=NMDS2,y=NMDS1, label = site), size=4, position = position_nudge(y = .0001)) + # add the text
  #geom_point(data=species.scores,aes(x=NMDS1,y=NMDS2), shape = 3, size=4) + 
  coord_equal() +
  theme_bw(base_size = 15, base_family = "Cambria") +
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank())

```

### Plot of NMDS species scores

Labels are are species code (by number). See table below for species code.   

```{r}

ggplot() +
  #geom_point(data=site.scores,aes(x=NMDS2,y=NMDS1, color=loc), size=5) + # add the point markers
  geom_text(data=species.scores,aes(x=NMDS2,y=NMDS1, label = spec_code), size=5, position = position_nudge(y = .01)) + # add the text
  geom_point(data=species.scores,aes(x=NMDS2,y=NMDS1), shape = 3, size=4) + 
  coord_equal() +
  theme_bw(base_size = 15, base_family = "Cambria") +
  theme(axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=18), # remove x-axis labels
        axis.title.y = element_text(size=18), # remove y-axis labels
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank())

```

### Key: Species codes (key to the lables in the species score NMDS plot above)

```{r spec_codes_2}

x <- species.scores %>% dplyr::select(spec_code)

## convert row names to column
x <- tibble::rownames_to_column(x, "ScientificName")
## create flextable
myft <- flextable(x)

## modify flextable
myft <- theme_vanilla(myft)
myft<- fontsize(myft, size = 12, part = "body")
myft<- fontsize(myft, size = 12, part = "header")
myft <- align(myft, align = "left", part = "header")
myft <- align(myft, align = "left", part = "body")

myft <- set_table_properties(myft, width = .5, layout = "autofit")
myft

```



<!-- ## Clustering -->

<!-- ```{r} -->
<!-- ##### create different distance matrices ##### -->
<!-- ## Bray-Curtis as binary (presence/absence) of species per equation above -->
<!-- sp_bray_binary = vegdist(site.sp, method='bray', binary=T) -->

<!-- ## Bray-Curtis as amount (density) of species -->
<!-- # sp_bray = vegdist(site.sp, method='bray', binary = F) -->

<!-- ## transformed Bray-Curtis as amount (density) of species -->
<!-- # sp_bray_transformed = vegdist(wisconsin(sqrt(site.sp)), method='bray') -->

<!-- ##### create different clusters ###### -->

<!-- fit <- hclust(sp_bray_binary) -->
<!-- plot(fit, main = "", sub = "", xlab="Sites", axes = T, ylab = "Ecological Distance (Sorensen)", hang = -1) # display dendogram -->
<!-- groups <- cutree(fit, k=4) # cut tree into 5 clusters -->

<!-- # draw dendogram with red borders around the 5 clusters -->
<!-- rect.hclust(fit, k=4, border="red") -->

<!-- # fit <- hclust(sp_bray_binary) -->
<!-- # plot(fit, main = "", sub = "", xlab="Sites", axes = T, ylab = "Ecological Distance (Sorensen)", hang = -1) # display dendogram -->
<!-- # groups <- cutree(fit, k=5) # cut tree into 5 clusters -->
<!-- # # draw dendogram with red borders around the 5 clusters -->
<!-- # rect.hclust(fit, k=5, border="red") -->
<!-- # -->
<!-- # fit <- hclust(sp_bray) -->
<!-- # plot(fit, main = "", sub = "", xlab="Sites", axes = T, ylab = "Ecological Distance (Bray-Curtis)", hang = -1) # display dendogram -->
<!-- # groups <- cutree(fit, k=5) # cut tree into 5 clusters -->
<!-- # # draw dendogram with red borders around the 5 clusters -->
<!-- # rect.hclust(fit, k=5, border="red") -->

<!-- ``` -->

<!-- ## NMDS (Two Highest Density Categories) -->

<!-- ### Filter to only species appearing in the top ten list -->
<!-- ```{r filter_to_top_ten, echo=F, eval=T, warning=F, message=F} -->
<!-- ##### filter d down to just the species appearing in the top ten ##### -->
<!-- topten <- sub %>% -->
<!--   filter(DensityCategory == "3000-9999/km" | -->
<!--            DensityCategory == "10000/km") %>% -->
<!--  group_by(EventID, ScientificName)%>% -->
<!--  summarize(individual_count_sum = sum(IndividualCount))%>% -->
<!--  top_n(10, individual_count_sum) %>% -->
<!--   arrange(EventID, desc(individual_count_sum)) -->

<!-- z <- sub %>% -->
<!--   filter(DensityCategory == "3000-9999/km" | -->
<!--            DensityCategory == "10000/km") %>% -->
<!--   filter(ScientificName %in% topten$ScientificName) -->

<!-- ``` -->

<!-- ### Setting up depth, temp, and location categories -->

<!-- ```{r setdepthclass, echo=FALSE, cache = FALSE} -->
<!-- ##### setting depth categories ##### -->
<!-- # qplot(as.numeric(z$DepthInMeters)) -->
<!-- z$DepthCat[as.numeric(z$DepthInMeters) <= 1800] <- "shallower" -->
<!-- z$DepthCat[as.numeric(z$DepthInMeters) > 1800 & -->
<!--              as.numeric(z$DepthInMeters) < 2400] <- "medium" -->
<!-- z$DepthCat[as.numeric(z$DepthInMeters) >= 2400] <- "deeper" -->
<!-- z$DepthCat <- factor(z$DepthCat, levels = c("shallower","medium","deeper")) -->

<!-- ##### setting location categories ##### -->
<!-- # qplot(as.numeric(z$Latitude)) -->
<!-- z$loc[as.numeric(z$Latitude) >= 20] <- "north" -->
<!-- z$loc[as.numeric(z$Latitude) < 20] <- "south" -->

<!-- ##### longitude cat ##### -->
<!-- z$longloc[as.numeric(z$Longitude) > -165] <- "east" -->
<!-- z$longloc[as.numeric(z$Longitude) < -165] <- "west" -->

<!-- ##### setting temperature categories ##### -->
<!-- # qplot(as.numeric(z$Temperature)) -->
<!-- z$temp_cat[as.numeric(z$Temperature) > 2.2] <- "warmer" -->
<!-- z$temp_cat[as.numeric(z$Temperature) >= 1.9 & as.numeric(z$Temperature) <= 2.2] <- "cold" -->
<!-- z$temp_cat[as.numeric(z$Temperature) < 1.9] <- "colder" -->
<!-- z$temp_cat <- factor(z$temp_cat, levels = c("warmer", "cold", "colder")) -->

<!-- ##### setting oxygen categories ##### -->
<!-- # qplot(as.numeric(z$Oxygen)) -->
<!-- z$ox_cat[as.numeric(z$Oxygen) >= 2.6] <- "rich" -->
<!-- z$ox_cat[as.numeric(z$Oxygen) < 2.6] <- "poor" -->

<!-- ``` -->

<!-- ### Making the matrix -->

<!-- First we adjust the abundances by the length of the drive track. This creates a normalized value for 'IndividualCount'. It is calculated as 'IndividualCount' per meter. This value is then entered into the site by species matrix which is used in the NMDS and clustering algorithms. -->

<!-- ``` {r NMDS, echo=FALSE, cache = FALSE, dpi=300, warning=FALSE, message=FALSE, results='hide', cache=T} -->
<!-- ## filter if necessary -->
<!-- x <- z %>% -->
<!--   filter( -->
<!--     #DepthCat == 'deep', -->
<!--     #EventID != "D2-EX1605L1-13" -->
<!--     #Phylum == "Cnidaria" -->
<!--     #DensityCategory == "high and higest" -->
<!--   ) -->

<!-- ## checking -->
<!-- # length(table(sub$n_km, useNA = 'always')) -->

<!-- ## adjust the IndividualCount values -->
<!-- x$IndividualCount_z <- x$IndividualCount/x$m_divetrack -->

<!-- ## create site X species matrix -->
<!-- library(vegan) -->
<!-- site.sp <- reshape2::dcast(x, EventID ~ ScientificName, value.var = "IndividualCount_z", fun.aggregate = sum) -->

<!-- ## set all NA to 0 because the vegan package can't handle the NA -->
<!-- site.sp[is.na(site.sp)] = 0 -->

<!-- ## checking resulting table -->
<!-- # View(site.sp) -->

<!-- ## export to check -->
<!-- # setwd("C:/rworking/deepseatools/indata") -->
<!-- # write.csv(site.sp, 'site.sp.csv') -->

<!-- # creating a site variable -->
<!-- site.sp$site <- site.sp$EventID -->

<!-- # getting rid of non-needed variables -->
<!-- site.sp <- site.sp %>% -->
<!--   dplyr::select(-EventID) -->

<!-- # # moving the site variable to the beginning -->
<!-- # col_idx <- grep("site", names(site.sp)) -->
<!-- # site.sp <- site.sp[, c(col_idx, (1:ncol(site.sp))[-col_idx])] -->
<!-- # # names(site.sp) -->

<!-- # set the site as the row.names -->
<!-- row.names(site.sp) <- site.sp$site -->

<!-- # remove the site variable -->
<!-- site.sp <- site.sp %>% -->
<!--   dplyr::select(-site) -->

<!-- # making it a matrix -->
<!-- site.sp <- as.matrix(site.sp) -->

<!-- ## checking -->
<!-- # NMDS -->

<!-- ``` -->

<!-- ### NMDS Run -->

<!-- Here are the statistical results from the NMDS run: -->

<!-- ```{r NMDS_run} -->
<!-- ##### running NMDS starting with site X species matrix ##### -->
<!-- #install.packages("vegan") -->
<!-- library(vegan) -->
<!-- NMDS <- metaMDS(site.sp, distance = "bray", binary = T, k=4, trymax = 30) -->

<!-- ##### extracting the site and species scores for use with ggplot2 ##### -->
<!-- #Using the scores function from vegan to extract the site scoresand convert to a data.frame -->
<!-- site.scores <- as.data.frame(scores(NMDS)) -->
<!-- site.scores$site <- rownames(site.scores)  # create a column of site names, from the rownames of data.scores -->

<!-- # head(site.scores)  #look at the data -->
<!-- # merge information from d (to get categories) -->
<!-- site.scores <- merge(site.scores, x, by.x = 'site', by.y = 'EventID') -->
<!-- site.scores <- site.scores %>% group_by(site) %>% summarise(NMDS1 = NMDS1[1], -->
<!--                                                             NMDS2 = NMDS2[1], -->
<!--                                                             DepthCat = DepthCat[1], -->
<!--                                                             temp_cat = temp_cat[1], -->
<!--                                                             loc = loc[1], -->
<!--                                                             ox_cat = ox_cat[1], -->
<!--                                                             loc2 = Locality[1]) -->

<!-- # extracting species scores -->
<!-- species.scores <- as.data.frame(scores(NMDS, "species"))  #Using the scores function from vegan to extract the species scores and convert to a data.frame -->
<!-- species.scores$species <- rownames(species.scores)  # create a column of species, from the rownames of species.scores -->
<!-- # head(species.scores)  #look at the data -->
<!-- species.scores$spec_code <- 1:(length(species.scores$species)) -->

<!-- NMDS -->

<!-- ``` -->

<!-- ### Plot of NMDS with group colors as from depth bin -->

<!-- ```{r plotting_NMDS_depth} -->
<!-- ##### plotting NMDS ##### -->
<!-- ggplot() + -->
<!--   geom_point(data=site.scores,aes(x=NMDS2,y=NMDS1, color=DepthCat), size=5) + # add the point markers -->
<!--   geom_text(data=site.scores,aes(x=NMDS2,y=NMDS1, label = site), size=4, position = position_nudge(y = .0001)) + # add the text -->
<!--   #geom_point(data=species.scores,aes(x=NMDS1,y=NMDS2), shape = 3, size=4) + -->
<!--   coord_equal() + -->
<!--   theme_bw(base_size = 15, base_family = "Cambria") + -->
<!--   theme(axis.text.x = element_blank(),  # remove x-axis text -->
<!--         axis.text.y = element_blank(), # remove y-axis text -->
<!--         axis.ticks = element_blank(),  # remove axis ticks -->
<!--         axis.title.x = element_text(size=18), # remove x-axis labels -->
<!--         axis.title.y = element_text(size=18), # remove y-axis labels -->
<!--         panel.background = element_blank(), -->
<!--         panel.grid.major = element_blank(),  #remove major-grid labels -->
<!--         panel.grid.minor = element_blank(),  #remove minor-grid labels -->
<!--         plot.background = element_blank()) -->

<!-- ``` -->

<!-- ### Plot of NMDS with group colors from temperature -->

<!-- ```{r plotting_NMDS_temp} -->

<!-- ggplot() + -->
<!--   geom_point(data=site.scores,aes(x=NMDS2,y=NMDS1, color=temp_cat), size=5) + # add the point markers -->
<!--   geom_text(data=site.scores,aes(x=NMDS2,y=NMDS1, label = site), size=4, position = position_nudge(y = .0001)) + # add the text -->
<!--   #geom_point(data=species.scores,aes(x=NMDS1,y=NMDS2), shape = 3, size=4) + -->
<!--   coord_equal() + -->
<!--   theme_bw(base_size = 15, base_family = "Cambria") + -->
<!--   theme(axis.text.x = element_blank(),  # remove x-axis text -->
<!--         axis.text.y = element_blank(), # remove y-axis text -->
<!--         axis.ticks = element_blank(),  # remove axis ticks -->
<!--         axis.title.x = element_text(size=18), # remove x-axis labels -->
<!--         axis.title.y = element_text(size=18), # remove y-axis labels -->
<!--         panel.background = element_blank(), -->
<!--         panel.grid.major = element_blank(),  #remove major-grid labels -->
<!--         panel.grid.minor = element_blank(),  #remove minor-grid labels -->
<!--         plot.background = element_blank()) -->

<!-- ``` -->

<!-- ### Plot of NMDS with group colors from location -->

<!-- Location based on latitude break at X. -->

<!-- ```{r} -->

<!-- ggplot() + -->
<!--   geom_point(data=site.scores,aes(x=NMDS2,y=NMDS1, color=loc), size=5) + # add the point markers -->
<!--   geom_text(data=site.scores,aes(x=NMDS2,y=NMDS1, label = site), size=4, position = position_nudge(y = .0001)) + # add the text -->
<!--   #geom_point(data=species.scores,aes(x=NMDS1,y=NMDS2), shape = 3, size=4) + -->
<!--   coord_equal() + -->
<!--   theme_bw(base_size = 15, base_family = "Cambria") + -->
<!--   theme(axis.text.x = element_blank(),  # remove x-axis text -->
<!--         axis.text.y = element_blank(), # remove y-axis text -->
<!--         axis.ticks = element_blank(),  # remove axis ticks -->
<!--         axis.title.x = element_text(size=18), # remove x-axis labels -->
<!--         axis.title.y = element_text(size=18), # remove y-axis labels -->
<!--         panel.background = element_blank(), -->
<!--         panel.grid.major = element_blank(),  #remove major-grid labels -->
<!--         panel.grid.minor = element_blank(),  #remove minor-grid labels -->
<!--         plot.background = element_blank()) -->

<!-- ``` -->

<!-- ### Plot of NMDS with group colors from oxygen values -->


<!-- ```{r} -->

<!-- ggplot() + -->
<!--   geom_point(data=site.scores,aes(x=NMDS2,y=NMDS1, color=ox_cat), size=5) + # add the point markers -->
<!--   geom_text(data=site.scores,aes(x=NMDS2,y=NMDS1, label = site), size=4, position = position_nudge(y = .0001)) + # add the text -->
<!--   #geom_point(data=species.scores,aes(x=NMDS1,y=NMDS2), shape = 3, size=4) + -->
<!--   coord_equal() + -->
<!--   theme_bw(base_size = 15, base_family = "Cambria") + -->
<!--   theme(axis.text.x = element_blank(),  # remove x-axis text -->
<!--         axis.text.y = element_blank(), # remove y-axis text -->
<!--         axis.ticks = element_blank(),  # remove axis ticks -->
<!--         axis.title.x = element_text(size=18), # remove x-axis labels -->
<!--         axis.title.y = element_text(size=18), # remove y-axis labels -->
<!--         panel.background = element_blank(), -->
<!--         panel.grid.major = element_blank(),  #remove major-grid labels -->
<!--         panel.grid.minor = element_blank(),  #remove minor-grid labels -->
<!--         plot.background = element_blank()) -->

<!-- ``` -->


<!-- ### Plot of NMDS species scores -->

<!-- Labels are are species code (by number). See table below for species code. -->

<!-- ```{r} -->

<!-- ggplot() + -->
<!--   #geom_point(data=site.scores,aes(x=NMDS2,y=NMDS1, color=loc), size=5) + # add the point markers -->
<!--   geom_text(data=species.scores,aes(x=NMDS2,y=NMDS1, label = spec_code), size=4, position = position_nudge(y = .0001)) + # add the text -->
<!--   geom_point(data=species.scores,aes(x=NMDS2,y=NMDS1), shape = 3, size=4) + -->
<!--   coord_equal() + -->
<!--   theme_bw(base_size = 15, base_family = "Cambria") + -->
<!--   theme(axis.text.x = element_blank(),  # remove x-axis text -->
<!--         axis.text.y = element_blank(), # remove y-axis text -->
<!--         axis.ticks = element_blank(),  # remove axis ticks -->
<!--         axis.title.x = element_text(size=18), # remove x-axis labels -->
<!--         axis.title.y = element_text(size=18), # remove y-axis labels -->
<!--         panel.background = element_blank(), -->
<!--         panel.grid.major = element_blank(),  #remove major-grid labels -->
<!--         panel.grid.minor = element_blank(),  #remove minor-grid labels -->
<!--         plot.background = element_blank()) -->

<!-- ``` -->

<!-- ### Table of species codes (key to the lables in the species score NMDS plot above) -->

<!-- ```{r spec_codes} -->

<!-- x <- species.scores %>% dplyr::select(spec_code) -->

<!-- ## convert row names to column -->
<!-- x <- tibble::rownames_to_column(x, "ScientificName") -->
<!-- ## create flextable -->
<!-- myft <- flextable(x) -->

<!-- ## modify flextable -->
<!-- myft <- theme_vanilla(myft) -->
<!-- myft<- fontsize(myft, size = 10, part = "body") -->
<!-- myft<- fontsize(myft, size = 10, part = "header") -->
<!-- myft <- align(myft, align = "left", part = "header") -->
<!-- myft <- align(myft, align = "left", part = "body") -->

<!-- myft <- set_table_properties(myft, width = .5, layout = "autofit") -->
<!-- myft -->

<!-- ``` -->

<!-- ### Clustering -->

<!-- ```{r} -->
<!-- ##### create different distance matrices ##### -->
<!-- ## Bray-Curtis as binary (presence/absence) of species per equation above -->
<!-- sp_bray_binary = vegdist(site.sp, method='bray', binary=T) -->

<!-- ## Bray-Curtis as amount (density) of species -->
<!-- # sp_bray = vegdist(site.sp, method='bray') -->

<!-- ## transformed Bray-Curtis as amount (density) of species -->
<!-- #sp_bray_transformed = vegdist(wisconsin(sqrt(site.sp)), method='bray') -->

<!-- ##### create different clusters ###### -->

<!-- fit <- hclust(sp_bray_binary) -->
<!-- plot(fit, main = "", sub = "", xlab="Sites", axes = T, ylab = "Ecological Distance (Sorensen)", hang = -1) # display dendogram -->
<!-- groups <- cutree(fit, k=4) # cut tree into 5 clusters -->
<!-- # draw dendogram with red borders around the 5 clusters -->
<!-- rect.hclust(fit, k=4, border="red") -->

<!-- # fit <- hclust(sp_bray_binary) -->
<!-- # plot(fit, main = "", sub = "", xlab="Sites", axes = T, ylab = "Ecological Distance (Sorensen)", hang = -1) # display dendogram -->
<!-- # groups <- cutree(fit, k=5) # cut tree into 5 clusters -->
<!-- # # draw dendogram with red borders around the 5 clusters -->
<!-- # rect.hclust(fit, k=5, border="red") -->
<!-- # -->
<!-- # fit <- hclust(sp_bray) -->
<!-- # plot(fit, main = "", sub = "", xlab="Sites", axes = T, ylab = "Ecological Distance (Bray-Curtis)", hang = -1) # display dendogram -->
<!-- # groups <- cutree(fit, k=5) # cut tree into 5 clusters -->
<!-- # # draw dendogram with red borders around the 5 clusters -->
<!-- # rect.hclust(fit, k=5, border="red") -->

<!-- ``` -->

<!-- ### Diversity estimates. -->

<!-- ##### Table: Simpson diversity per EventID -->
<!-- All taxa (both coral and sponges at all taxonomic ranks included in estimates). -->

<!-- ```{r diversity, echo=F} -->

<!-- diversity <- diversity(site.sp, index = "simpson") #makes an object -->
<!-- #that the diversity values #are written to. -->
<!-- # summary(plots) #gives summary statistics for the plots -->
<!-- # median(plots) #gives the median -->
<!-- # mean(plots) #gives the mean -->
<!-- # class(plots) -->
<!-- # max(plots[1:30]) -->
<!-- # # View(plots) -->

<!-- # creating plots -->
<!-- plots_df <- data.frame(diversity) -->

<!-- # View(plots_df) -->
<!-- plots_df$EventID <- row.names(plots_df) -->

<!-- # names(d) -->
<!-- d <- merge(sub, plots_df, all.x = T) -->
<!-- # length(d$EventID) -->
<!-- # length(unique(d$EventID)) -->
<!-- x <- d %>% group_by(EventID, diversity) %>% -->
<!--   summarize(n=n(), -->
<!--             Latitude = mean(as.numeric(Latitude)), -->
<!--             Longitude = mean(as.numeric(Longitude))) %>% -->
<!--   arrange(desc(diversity)) -->

<!-- ## create flextable -->
<!-- myft <- flextable(x) -->

<!-- ## modify flextable -->
<!-- myft <- theme_vanilla(myft) -->
<!-- myft<- fontsize(myft, size = 10, part = "body") -->
<!-- myft<- fontsize(myft, size = 10, part = "header") -->
<!-- myft <- align(myft, align = "left", part = "header") -->
<!-- myft <- align(myft, align = "left", part = "body") -->

<!-- myft <- set_table_properties(myft, width = .5, layout = "autofit") -->
<!-- myft -->


<!-- #View(x) -->

<!-- setwd("C:/rworking/deepseatools/indata") -->
<!-- write.xlsx(x, '20210824-0_Summary_of_Corals_and_Sponges_Diversity_by_EventID.xlsx') -->


<!-- ## Calculating various Diversity Indices -->

<!-- # H <- diversity(site.sp) -->
<!-- # simp <- diversity(site.sp, "simpson") -->
<!-- # invsimp <- diversity(site.sp, "inv") -->
<!-- # ## Unbiased Simpson (Hurlbert 1971, eq. 5) with rarefy: -->
<!-- # unbias.simp <- rarefy(site.sp, 2) - 1 -->
<!-- # ## Fisher alpha -->
<!-- # alpha <- fisher.alpha(site.sp) -->
<!-- # ## Plot all -->
<!-- # pairs(cbind(H, simp, invsimp, unbias.simp, alpha, J), pch="+", col="blue") -->
<!-- # ## Species richness (S) and Pielou's evenness (J): -->
<!-- # S <- specnumber(site.sp) ## rowSums(site.sp > 0) does the same... -->
<!-- # J <- H/log(S) -->

<!-- ``` -->

<!-- ## Diversity -->

<!-- ```{r mapping_diversity_2 , echo=FALSE, eval=F} -->

<!-- x$Latitude <- as.numeric(x$Latitude) -->
<!-- x$Longitude <- as.numeric(x$Longitude) -->

<!-- pal <- colorNumeric( -->
<!--   palette = "Reds", -->
<!--   domain = x$diversity) -->


<!-- m <- leaflet() -->
<!-- m <- addProviderTiles(m, "Esri.OceanBasemap") #Esri.OceanBasemap, "CartoDB.DarkMatter" -->
<!-- m <- addCircleMarkers(m, data=x, -->
<!--                       radius=5, -->
<!--                       weight=0, -->
<!--                       #fillColor= "green", -->
<!--                       color = ~pal(diversity), -->
<!--                       fillOpacity=1, -->
<!--                       popup = paste("EventID:", x$EventID, "<br>")) -->

<!-- m -->

<!-- ``` -->




