---
title: "NOAA National Database for Deep-sea Corals and Sponges: Area of Interest / Query Summary"
author: "NOAA-NFMS-OHC-DSCRTP_RPMcGuinn"
date: 'Robert.McGuinn@NOAA.gov; 843-460-9696. Report last ran: `r Sys.Date()`'
output: word_document
---

# --

```{r packages, echo=F, warning=F, message=F}
#install.packages(tidyverse)
#install.packages('openxlsx')
library(tidyverse)
# library(openxlsx)
# library(sf)
# library(sp)
# library(rgdal) # install.packages("rgdal")
# library(RColorBrewer)
# library(leaflet)# install.packages("leaflet")
# library(extrafont)
# library(RColorBrewer)
# library(rmarkdown)
# library(knitr)
# library(rgdal)
# library(httr)
# library(jsonlite)
# library(rnaturalearth)# install.packages("rnaturalearth")
# library(rnaturalearthdata)# install.packages("rnaturalearthdata")
# library(openxlsx)
# library(terra) # install.packages("terra")


```

# Load NDB

```{r load_NDB, eval=T, echo=F, message=FALSE, warning=FALSE, cache=TRUE}
##### input: latest version of NDB #####
setwd("C:/rworking/deepseatools/indata")
indata<-read.csv("DSCRTP_NatDB_20201021-0.csv", header = T)
filt <- indata %>%
  filter(Flag == "0")
rm(indata)
```

# Filter by Box (OR)

```{r filter_geo_OR, echo= T, warning=F, message=F, eval=F}
## set bounding box variables
minlat <- 14.87
maxlat <- 21.89
minlon <- -68.48
maxlon <- -63.86

## subset data by coordinates
x <- subset(filt, as.numeric(Latitude) > minlat &
                   as.numeric(Latitude) < maxlat &
                   as.numeric(Longitude) > minlon &
                   as.numeric(Longitude) < maxlon &
                   Flag == "0")
```

# Query used in this Report (OR) 
Language: R

```{r filter_query_OR, echo=T, warning=F, message=F, eval=T}

ecogig <- c("NOAA_EX-12-02-L2",
"BOEM_Lophelia_II",
"WHOI_AT-18",
"SOI_FK006B",
"NOAA_HC-11-10",
"OET_NA028",
"OET_NA043",
"OET_NA057",
"OET_NA058",
"ECOGIG_OI-16-10",
"ECOGIG_O2-17-06")

ecogig <- unique(ecogig)

## DatasetIDs in the SEDCI rescue project (as of 2020-10-19)
sedci_rescue <- c("HBOI_SJ-10-07", "HBOI_SJ-10-07", "NOAA_SJ-09-08", "NOAA_SJ-09-08","HBOI_SJ-08-09","NOAA_SJ-07-05", "NOAA_SJ-07-05", "NOAA_SJ-07-06", "HBOI_SJ-06-02", "HBOI_SJ-06-05-L1", "HBOI_SJ-06-05-L3", "HBOI_SJ-06-02", "NOAA_SJ-05-04", "NOAA_SJ-05-08", "NOAA_SJ-05-10", "NOAA_SJ-05-11", "NOAA_SJ-04-05", "NOAA_SJ-02-08", "HBOI_EL-99-08", "HBOI_SJ-88-11", "USGS_TM-97-10", "NOAA_RB-03-07", "NOAA_SJ-05-11",
"NOAA_NF-09-01", "NOAA_NF-08-04", "NOAA_NF-07-06", "NOAA_NF-05-05", "NOAA_SJ-07-06", "HBOI_SJ-08-09", "NOAA_SJ-09-08", "NOAA_PC-10-02", 
"NOAA_NF-14-01", "NOAA_NF-13-02", "NOAA_NF-12-01", "NOAA_NF-11-09", "NOAA_NF-11-09-L3", "NOAA_PC-11-05-L1", "NOAA_NF-11-09", "NOAA_NF-11-09-L3", "Sharuga_SM_2014", "NOAA_PC-12-03", "NOAA_NF-13-02", "NOAA_PC-13-03", "NOAA_NF-14-08","NOAA_PC-15-02", "NOAA_PC-16-02", "OET_NA058", "OET_NA028", "NOAA_PC-17-02", "HBOI_WS-17-125")

## unique DatasetIDs
sedci_rescue<- unique(sedci_rescue)

## sedci_rescue add ecogig
sedci_rescue <- union(sedci_rescue, ecogig)

## DatasetIDs in the SEDCI rescue project (as of 2020-10-19)
sedci_new <- c("NOAA_DFH-30", "NOAA_NF-17-08", "NOAA_NF-17-08", "NOAA_DFH-32-33", "NOAA_DFH-32-33", "NOAA_EX-17-11", "NOAA_EX-18-03", "NOAA_EX-18-06","NOAA_NF-18-04","NOAA_FGBNMS_DFH-35", "NOAA_FGBNMS_DFH-37", "NOAA_RESTORE_MT18")

sedci_new <- unique(sedci_new)

## create geographic filter for just the SEDCI region.
x <- filt %>% filter(FishCouncilRegion == "Gulf of Mexico" |
                       FishCouncilRegion == "South Atlantic" |  
                       FishCouncilRegion == "Caribbean")

## define not in
`%notin%` <- Negate(`%in%`)

x$sedci <- 
  case_when(
  x$DatasetID %in% sedci_rescue ~ "rescue" ,
  x$DatasetID %in% sedci_new ~ "new",
)

length(x$Flag)
table(x$sedci, useNA = 'always')

x$sedci2 <- 
  case_when(
  x$DatasetID %in% sedci_rescue ~ "Post SEDCI" ,
  x$DatasetID %in% sedci_new ~ "Post SEDCI",
  x$DatasetID %notin% sedci_rescue ~ "Pre SEDCI",
  x$DatasetID %notin% sedci_new ~ "Pre SEDCI"
)

x$sedci2 <- factor(x$sedci2, levels = c('Post SEDCI', 'Pre SEDCI'))


## check
# x %>% pull(CatalogNumber) %>% length()

# 
# filt %>%
#   filter(grepl("EX", DatasetID),
#          is.na(ImageURL) == F#,
#          #ObservationDate == '-999'
#          ) %>%
#   pull(DatasetID) %>% 
#   unique() %>% 
#   length()

```

# Get Bounding Box from Query

```{r bbox_from_query, echo=F, warning=F, message=F, eval=T}
## set bounding box variables with query results.
minlat <- min(x$Latitude)
maxlat <- max(x$Latitude)
minlon <- min(x$Longitude)
maxlon <- max(x$Longitude)
```

# Create SPDF from subset of points

```{r export_GIS, echo=FALSE, echo = FALSE, message = FALSE, warning=FALSE, eval=T}

##### load packages #####
library(arcgisbinding)
arc.check_product()

x_geo <- x

##### create spdf #####

coordinates(x_geo) <- c("Longitude", "Latitude")
proj4string(x_geo) <- "+proj=longlat +ellps=WGS84 +datum=WGS84"

##### create feature-class #####

# fgdb_path <- 'C:/rworking/sf/sf.gdb'
# arc.write(file.path(fgdb_path, 'x_geo'), data=x_geo, overwrite = TRUE)

```

# Assignments for Protected Status 

```{r gis_assignment}
##### bring in data #####
setwd("C:/data/geoindata.gdb/")
mpas_sp <- readOGR(".", "mpas_RPMcGuinn_20201019_0")
oculina_sp <- readOGR(".", "oculina_bank_po_RPMcGuinn_20201019_0")
oculinaexp_sp <- readOGR(".", "oculina_bank_exp_po_RPMcGuinn_20201019_0")
hapc_sp <- readOGR(".", "coral_hapc_RPMGuinn_20191019_0")

##### transform to sf #####
points <- st_as_sf(x_geo, wkt = "geom")
mpas <- st_as_sf(mpas_sp, wkt = "geom")
oculina <- st_as_sf(oculina_sp, wkt = "geom")
oculinaexp <- st_as_sf(oculinaexp_sp, wkt = "geom")
hapc <- st_as_sf(hapc_sp, wkt = "geom")

##### get everything into the same projection #####
mpas <- st_transform(mpas, crs = proj4string(x_geo))
oculina <- st_transform(oculina, crs = proj4string(x_geo))
oculinaexp <- st_transform(oculinaexp, crs = proj4string(x_geo))
hapc <- st_transform(hapc, crs = proj4string(x_geo))

##### spatial joins with points #####
points_mpas <- st_join(points, mpas) #NAME
points_oculina <- st_join(points, oculina) #AREA_NAME
points_oculinaexp <- st_join(points, oculinaexp) #AreaName
points_hapc <- st_join(points, hapc) #NAME

##### stripping off CatalogNumbers from spatial joins #####
mpas_cats <- points_mpas %>% filter(is.na(NAME) == F)
oculina_cats <- points_oculina %>%  filter(is.na(AREA_NAME) == F)
oculinaexp_cats <- points_oculinaexp %>% filter(is.na(AreaName) == F)
hapc_cats<- points_hapc %>% filter(is.na(NAME) == F)

## binding all of the CatalogNumbers together 
yo <- c(mpas_cats$CatalogNumber, oculina_cats$CatalogNumber, oculinaexp_cats$CatalogNumber, hapc_cats$CatalogNumber)

## define not in
`%notin%` <- Negate(`%in%`)

## assign management status ## 
x$mngstatus <- 
  case_when(
  x$CatalogNumber %in% yo ~ "Inside Managed Area",
  x$CatalogNumber %notin% yo ~ "Outside Managed Area"
)

## checking
# length(mpas_cats$CatalogNumber)
# length(oculina_cats$CatalogNumber)
# length(oculinaexp_cats$CatalogNumber)
# length(hapc_cats$CatalogNumber)
# names(points_oculina)
# z <- points_oculina %>% filter(is.na(AREA_NAME) == F)
# plot(z["AREA_NAME"])
# names(points_mpas)
# z <- points_mpas %>% filter(is.na(NAME) == F)
# plot(z["NAME"])
# names(points_mpas)
# z <- points_oculinaexp %>% filter(is.na(AREA_NAME) == F)
# plot(z["AREA_NAME"])

```

# Assigning Depth Bins

``` {r DepthClassTable, echo=FALSE, cache = FALSE} 
hi <- 600
y <- 300
z <- 150

x$DepthCat[x$DepthInMeters > hi] <- "very deep"
x$DepthCat[x$DepthInMeters > y & x$DepthInMeters <= hi] <- "deep"
x$DepthCat[x$DepthInMeters >= z & x$DepthInMeters <= y] <- "mesophotic"
x$DepthCat[x$DepthInMeters < z] <- "very shallow"
x$DepthCat <- factor(x$DepthCat, levels = c('very shallow','mesophotic', 'deep', 'very deep'))

```

# Assigning EcoRegions

```{r}
x <- x %>%
  mutate(
    EcoRegion = case_when(
      Latitude >= 32 ~ "Northern Carolinian",
      Latitude < 32 & Latitude >= 28.489 ~ "Southern Carolinian", 
      Latitude < 28.489 ~ "Floridian"
    )
  )

x$EcoRegion <- factor(x$EcoRegion, levels = c('Northern Carolinian', 'Southern Carolinian', 'Floridian'))

class(x$EcoRegion)
table(x$EcoRegion)

```

# Setting Genera

``` {r genera_list, echo=FALSE}
##### creating list of specific genera from Tom #####
# shallow genera 
genera_1 <- c('Phyllangia',
              'Renilla',
              'Leptogorgia',
              'Titanideum',
              'Balanophyllia',
              'Stichopathes',
              'Ellisella',
              'Tanacetipathes',
              'Diodogorgia', 
              'Nidalia',
              'Telesto',
              'Oculina',
              'Cladocora',
              'Thesea',
              'Paracyathus',
              'Dasmosmilia',
              'Polymyces')

# deeper genera
genera_2 <- c('Bathypathes',
              'Stylaster',
              'Paramuricea',
              'Plumarella',
              'Thecopsammia',
              'Leiopathes',
              'Deltocyathus',
              'Madrepora',
              'Lophelia',
              'Enallopsammia',
              'Bathypsammia',
              'Swiftia',
              'Javania',
              'Pseudodrifa',
              'Clavularia',
              'Anthomastus',
              'Acanella',
              'Keratoisis',
              'Eunicella'
)

#all genera
genera <- c('Phyllangia',
            'Renilla',
            'Leptogorgia',
            'Titanideum',
            'Balanophyllia',
            'Stichopathes',
            'Ellisella',
            'Tanacetipathes',
            'Diodogorgia', 
            'Nidalia',
            'Telesto',
            'Oculina',
            'Cladocora',
            'Thesea',
            'Paracyathus',
            'Dasmosmilia',
            'Polymyces',
            'Bathypathes',
            'Stylaster',
            'Paramuricea',
            'Plumarella',
            'Thecopsammia',
            'Leiopathes',
            'Deltocyathus',
            'Madrepora',
            'Lophelia',
            'Enallopsammia',
            'Bathypsammia',
            'Swiftia',
            'Javania',
            'Pseudodrifa',
            'Clavularia',
            'Anthomastus',
            'Acanella',
            'Keratoisis',
            'Eunicella'
)

```

# Figure: Protection by DepthCat

```{r}
##### building big pallette #####
# to get the colors from a palette:
palette1 <- brewer.pal(9,"Set1")
palette2 <- brewer.pal(8,"Set2")
palette3 <- brewer.pal(9,"Set3")
palette4 <- brewer.pal(8,"Dark2")
palette5 <- brewer.pal(8,"Accent")

# We can just stick a few color palettes together
big_palette <- c(palette1,palette2,palette3, palette4, palette5)

##### making the figure #####
options(scipen=10000)

g <- x %>% filter(
 # sedci == ""
  ) %>% 
  ggplot(aes(DepthCat, fill = sedci2)) +
  geom_bar() + 
  coord_flip() + 
  facet_wrap(~ mngstatus) +
  theme(text = element_text(size=20)) + 
  ylab("Number of Records") + 
    xlab("Depth Zone") + 
    scale_x_discrete(limits = rev(levels(x$DepthCat))) + 
  theme_bw(base_size = 15, base_family = "Cambria") +
  theme(legend.title = element_blank())

#display.brewer.all(colorblindFriendly=TRUE)
set.seed(10)
g + scale_fill_manual(values = sample(big_palette)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

## save the plot
ggsave(paste("c:/rworking/deepseatools/images/", 
             "20201020-0_SAFMC_Protection_by_Depth_SEDCI", 
             "NatDB_", 
             unique(x$DatabaseVersion), 
             ".png", 
             sep = ''),
       width = 8,
       height = 5.57,
       units = "in")


```

# Figure: Pie Charts

```{r}

# Pie charts
#####
set.seed(3)
# build a color matching data frame
colormatchdf <- data.frame(Genus = genera, color = sample(big_palette[1:36], length(genera)))
#get a summary vector of Genus Counts
GenusCounts <- summary(x$Genus)
# limit by count
GenusCounts <- GenusCounts[GenusCounts>50]
# make count into data frame
GenusCounts <- data.frame(Genus=names(GenusCounts), value=GenusCounts, row.names=NULL)
# merge to get color field into cound summary dataframe
GenusCountsColor <- merge(GenusCounts, colormatchdf, by = "Genus")
# sort by value to make prettyer chart
GenusCountsColor <- GenusCountsColor[order(as.numeric(GenusCountsColor$value)),]
# make color vector
colors <- as.vector(GenusCountsColor$color)
# set pallete
palette(colors)
# make pie chart
pie(GenusCountsColor$value,labels = GenusCountsColor$Genus, col = c(1:length(GenusCountsColor$value)), cex = 1.5)

```

# Write data subset to file

```{r write_data, echo=F, warning=F, message=F, eval=F}
## write data to Excel
setwd("C:/rworking/deepseatools/indata")
write.xlsx(x,"20201019-2_subset_SEDCI_from_DSCRTP_NatDBNatDB_20200710-2_RPMcGuinn.xlsx",
           row.names = F, quote = T)
```

```{r erddap_link, echo = FALSE}
x$erddap <- paste("https://www.ncei.noaa.gov/erddap/tabledap/deep_sea_corals.csv?ShallowFlag%2CDatabaseVersion%2CDatasetID%2CCatalogNumber%2CSampleID%2CTrackingID%2CImageURL%2CHighlightImageURL%2CCitation%2CRepository%2CScientificName%2CVerbatimScientificName%2CVernacularNameCategory%2CVernacularName%2CTaxonRank%2CAphiaID%2CLifeScienceIdentifier%2CPhylum%2CClass%2CSubclass%2COrder%2CSuborder%2CFamily%2CSubfamily%2CGenus%2CSubgenus%2CSpecies%2CSubspecies%2CScientificNameAuthorship%2CTypeStatus%2COperationalTaxonomicUnit%2CMorphospecies%2CCombinedNameID%2CSynonyms%2CIdentificationComments%2CIdentifiedBy%2CIdentificationDate%2CIdentificationQualifier%2CIdentificationVerificationStatus%2CAssociatedSequences%2COcean%2CLargeMarineEcosystem%2CCountry%2CFishCouncilRegion%2CLocality%2Clatitude%2Clongitude%2CDepthInMeters%2CDepthMethod%2CMinimumDepthInMeters%2CMaximumDepthInMeters%2CLocationComments%2CObservationDate%2CObservationYear%2CObservationTime%2CSurveyID%2CVessel%2CPI%2CPIAffiliation%2CPurpose%2CSurveyComments%2CStation%2CEventID%2CSamplingEquipment%2CVehicleName%2CSampleAreaInSquareMeters%2CfootprintWKT%2CfootprintSRS%2CIndividualCount%2CCategoricalAbundance%2CDensity%2CCover%2CVerbatimSize%2CMinimumSize%2CMaximumSize%2CWeightInKg%2CCondition%2CAssociatedTaxa%2COccurrenceComments%2CStartLatitude%2CStartLongitude%2CEndLatitude%2CEndLongitude%2CVerbatimLatitude%2CVerbatimLongitude%2CLocationAccuracy%2CNavType%2COtherData%2CHabitat%2CSubstrate%2CCMECSGeoForm%2CCMECSSubstrate%2CCMECSBiotic%2CTemperature%2CSalinity%2COxygen%2CpH%2CpHscale%2CpCO2%2CTA%2CDIC%2CRecordType%2CDataProvider%2CDataContact%2CModified%2CWebSite%2CEntryDate%2CReporter%2CReporterEmail%2CReporterComments%2CgisCRMDepth%2CgisGEBCODepth%2CgisEtopoDepth%2CgisMEOW%2CSynonymSearchProxy", "&latitude%3E=",
                  minlat, "&latitude%3C=", 
                  maxlat, "&longitude%3E=", 
                  minlon, "&longitude%3C=",
                  maxlon, sep='')

y <- unique(x$erddap)

```

# Download data
## Download comma delimited (CSV) file from ERDDAP: 
[LINK](`r y`)

Download data dictionary: [LINK](https://deepseacoraldata.noaa.gov/internal-documents/program-guidance/science-team-guidance-for-data-management/20170707.xlsx/at_download/file)

**Note:** When importing CSV files into Excel, use the data import wizard (under the 'data' tab) and set all column types to 'text' to avoid data tranformation issues. This step will prevent the automated alteration of dates in Excel.

## Database Version
`r paste('DSCRTP_NatDB_', x$DatabaseVersion[1], sep = '')`

# Area of Interest (AOI) Map

```{r zoom, echo=F}

# create a spatial points data frame
spdf<-x
coordinates(spdf) <- c("Longitude", "Latitude", "DepthInMeters")
proj4string(spdf) <- "+proj=longlat +ellps=WGS84 +datum=WGS84"
bbox<-bbox(spdf)

library(rnaturalearth)
library(rnaturalearthdata)

world <- ne_countries(scale = "medium", returnclass = "sf")
#class(world)

zoom <- 5
g <- ggplot(world) +
    geom_sf() +
    coord_sf(xlim = c(minlon-zoom, maxlon+zoom), ylim = c(minlat-zoom, maxlat+zoom), expand = FALSE)

# add sampling locations

g <- g + geom_point(aes(x=Longitude, y=Latitude), data=x, alpha=0.5, color = 'red', size = 2)
g

```

# 3D Bounding Box

```{r bounding_box, echo=FALSE, echo = FALSE, message = FALSE, warning=FALSE, dpi=300, fig.width=8, fig.height=4.1, eval=TRUE, fig.align='center'}
options(digits = 5)
spdf<-x
coordinates(spdf) <- c("Longitude", "Latitude", "DepthInMeters")
proj4string(spdf) <- "+proj=longlat +ellps=WGS84 +datum=WGS84"
bbox<-bbox(spdf)
bbox
```




# Summary of this Selection by the Numbers

*Note:* The numbers below only reflect published records. Additional records may be retained by the Program for further review.

* **Number of records:** `r prettyNum(length(x$CatalogNumber), big.mark = ",")`
* **Number of coral records:** `r prettyNum(length(x$CatalogNumber[x$Phylum == "Cnidaria"]), big.mark = ",")` 
* **Number of sponge records:** `r prettyNum(length(x$CatalogNumber[x$Phylum == "Porifera"]), big.mark = ",")`
* **Records with images:** `r length(x$ImageURL[is.na(x$ImageURL) == FALSE])`
* **Record type(s):** `r toString(unique(x$RecordType))`
* **Minimum depth (meters):** `r prettyNum(min(x[x$DepthInMeters != "-999",]$DepthInMeters), big.mark = ",")`
* **Maximum depth (meters):** `r prettyNum(max(x[x$DepthInMeters != "-999",]$DepthInMeters), big.mark = ",")`
* **Time frame:** `r min(x[x$ObservationYear != "-999",]$ObservationYear)` to `r max(x[x$ObservationYear != "-999",]$ObservationYear)`

# Where?

``` {r Ocean, echo=FALSE}
x <- x %>% mutate(Country = ifelse(is.na(Country) == T, 'International Waters', as.character(Country)))

sum_tbl <-
  x %>%
  #filter(Phylum == "Porifera") %>%
  group_by(Country) %>%
  summarize(Ocean = toString(unique(Ocean)), 
            FishCouncilRegion = toString(unique(FishCouncilRegion)),
            n = n()) %>%
  arrange(desc(n))
sum_tbl <- kable(sum_tbl, row.names = F, digits = 2, format.args = list(big.mark = ","))
sum_tbl

```

# Corals vs. sponges

```{r corals_sponges, echo=FALSE, cache = FALSE, dpi=300, fig.height=3, fig.width=8}
#library(extrafont)
#loadfonts(device = "win")

g <- ggplot(x, aes(Phylum, fill = Phylum)) +
  geom_bar() + 
 # coord_flip() + 
  theme(text = element_text(size=20)) + 
  ylab("Number of Records") + 
  theme_bw(base_size = 15, base_family = "Cambria")

#display.brewer.all(colorblindFriendly=TRUE)
g + scale_fill_manual(values = brewer.pal(12, "Paired")[c(8,7)])

```

# Distribution of Taxa by Order

```{r orders, echo=FALSE, cache = FALSE, dpi=300, fig.height=6, fig.width=8, message=FALSE, warning=FALSE}
##### building big pallette #####
# to get the colors from a palette:
palette1 <- brewer.pal(9,"Set1")
palette2 <- brewer.pal(8,"Set2")
palette3 <- brewer.pal(9,"Set3")
palette4 <- brewer.pal(8,"Dark2")
palette5 <- brewer.pal(8,"Accent")

# We can just stick a few color palettes together
big_palette <- c(palette1,palette2,palette3, palette4, palette5)

x <- within(x, 
             Order <- factor(Order, 
                                  levels=names(sort(table(Order), decreasing=TRUE))))

##### graphing the orders #####
g <- ggplot(x, aes(Order)) +
  geom_bar() + 
 # coord_flip() + 
  #theme(text = element_text(size=10)) + 
  ylab("Number of Records") + 
  facet_wrap(~Phylum, scales="free") +
  theme_bw(base_size = 12, base_family = "Cambria")

set.seed(7)
g + scale_fill_manual(values = sample(big_palette)) + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.05))

```

# Who Contributed to this Selection?

``` {r DataProvider, echo=FALSE}
# # Data Provider corrections
# x$DataProvider <- 
#   plyr::revalue(x$DataProvider,
#                 c('Kelley, Christopher (CKelley@hawaii.edu)' = 'NOAA, Deep Sea Coral Research & Technology Program and Office of Ocean Exploration and Research'))
# 
# x$DataContact <- 
#   plyr::revalue(x$DataContact, 
#                 c('2017-12-18' =  'Kelley, Chris: ckelley@hawaii.edu','Pante, Eric (eric.pante@univ-lr.fr)' = 'Pante, Eric: eric.pante@univ-lr.fr', "Battista, Tim; Tim.Battista@noaa.gov" = "Battista, Tim: Tim.Battista@noaa.gov", "Thoma, Jana" = "Thoma, Jana: jana.thoma@louisiana.edu"))
                  
# table(factor(x$DataProvider))
# table(factor(x$PI))
# table(factor(x$DataContact))

x$url <- paste0('https://deepseacoraldata.noaa.gov/Dataset%20Summaries/',
                      x$DatasetID, 
                      '.html', 
                      sep = '')

x$DashBoard <- paste0("[", x$DatasetID, "](", x$url, ")")

sum_tbl <-
  x %>%
  group_by(DataProvider) %>%
  summarize(DatasetID = toString(unique(DashBoard)), 
            PI = toString(unique(unlist(unique(strsplit(as.character(PI), '; '))))),
            DataContact = toString(unique(DataContact)),
            Reporter = toString(unique(Reporter)),
            n = n()) %>%
  arrange(desc(DatasetID))
sum_tbl <- kable(sum_tbl, row.names = F, digits = 2, format.args = list(big.mark = ","))
sum_tbl
``` 

# Expedition Details

``` {r exped_details, echo=FALSE}
sum_tbl <-
  x %>%
  group_by(DatasetID, SurveyID) %>% 
  summarize(Vessel = toString(unique(Vessel)),
            VehicleName = toString(unique(VehicleName)),
            PI = toString(unique(PI)),
            SamplingEquipment = toString(unique(SamplingEquipment)),
            RecordType = toString(unique(RecordType)),
            #BeginYear= min(as.numeric(ObservationYear)),
            EndYear= max(as.numeric(ObservationYear)),
            n = prettyNum(n(),big.mark = ',')) %>% 
  arrange(desc(EndYear))

View(sum_tbl)

sum_tbl$url <- paste0('https://deepseacoraldata.noaa.gov/Dataset%20Summaries/',
                      sum_tbl$DatasetID, 
                      '.html', 
                      sep = '')

sum_tbl$DashBoard <- paste0("[", "Dashboard Link", "](", sum_tbl$url, ")")
sum_tbl <- sum_tbl %>% dplyr::select(-url)
sum_tbl <- kable(sum_tbl, row.names = F, digits = 2)
sum_tbl
```

# Growth of Database Through Time within this Selection

``` {r GrowthThroughTime_charts, echo=FALSE, dpi=300, eval = T}
## set options to remove scientific notation
options(scipen=10000, digits = 0)


subx <- x %>% filter (Phylum == "Cnidaria") %>%
  mutate(EntryDate = replace_na(as.character(EntryDate), '2019-12-18')) %>% 
  mutate(EntryDate = as.factor(EntryDate)) %>% 
  mutate(as.Date(EntryDate))

suby <- x %>% filter (Phylum == "Porifera") %>%
  mutate(EntryDate = replace_na(as.character(EntryDate), '2019-12-18')) %>% 
  mutate(EntryDate = as.factor(EntryDate)) %>% 
  mutate(as.Date(EntryDate))
 
ggplot(subx, aes(x=as.Date(EntryDate), color=Phylum)) +
  stat_bin(data = subx,
           aes(y=cumsum(..count..) / 1000), 
           geom="line", 
           binwidth = 2, 
           size = 1) +
  stat_bin(data = suby,
           aes(y=cumsum(..count..) / 1000),
           geom="line", 
           binwidth = 2, 
           size = 1) +
  ylab("Number of Records (thousands)") + 
  xlab("Year Record Added") + 
  xlim(as.Date('2012-01-01'), as.Date('2020-10-19')) +
  theme(axis.text.x=element_text(hjust= .2)) +
  guides(fill=FALSE) + 
  theme_bw() +
  theme(text=element_text(size=20,  family="Cambria")) +
  #theme(axis.text.x = element_text(angle = 90)) +
  theme(legend.title = element_blank()) +
  theme(axis.title.x = element_text(vjust=-2.5)) +
  theme(axis.title.y = element_text(vjust= 5)) +
  theme(plot.margin = unit(c(1,1,1,1), "cm"))+
  scale_color_manual(values = brewer.pal(12, "Paired")[c(10,9)]) +
  scale_y_continuous(label=scales::comma) +
  scale_y_continuous(labels = scales::number_format(accuracy = 1))+
  geom_vline(xintercept = as.Date("2016-01-01"), color = 'blue')# +
## ggtitle(paste("Database Version:", unique(filt$DatabaseVersion)))

## save the plot
ggsave(paste("c:/rworking/deepseatools/images/", 
             "20220130-1_SouthAtlantic_subset", 
             "NatDB_", 
             unique(x$DatabaseVersion), 
             ".png", 
             sep = ''),
       width = 8,
       height = 5.57,
       units = "in")

```

# Subset data to just South Atlantic FMC 

```{r meow, cache = FALSE, echo=FALSE}
sa <- x %>% filter(FishCouncilRegion == 'South Atlantic')
table(sa$gisMEOW, useNA = 'always')

```

# Creating a Depth Class Table

``` {r DepthClassTable, echo=FALSE, cache = FALSE} 
x <- 600
y <- 300
z <- 150

sa$DepthCat[sa$DepthInMeters > x] <- "very deep"
sa$DepthCat[sa$DepthInMeters > y & sa$DepthInMeters <= x] <- "deep"
sa$DepthCat[sa$DepthInMeters >= z & sa$DepthInMeters <= y] <- "mesophotic"
sa$DepthCat[sa$DepthInMeters < z] <- "very shallow"
sa$DepthCat <- factor(sa$DepthCat, levels = c('very deep', 'deep', 'mesophotic', 'very shallow'))

```


