---
title: "Untitled"
author: "Robert McGuinn"
date: "8/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Table: Taxonomic lists (All)
* *EventID*: This table is grouped by dive (EventID)
* *count_per_km*: IndividualCount per linear distance surveyed (km). This field is used for sorting this table in descending order. 
* *number_genera*: The number of genera observed the dive.
* *number_species*: The number of species observed on the dive.
* *genus_list*: List of genera observed on the dive.
* *species_list*: List of species observed on the dive.

```{r taxa_list_table}
## create table
x <- sub %>% arrange(ScientificName) %>% 
  group_by(EventID) %>% 
  summarize(count_per_km = unique(n_km),
            number_species = length(unique(ScientificName)),
            number_genera = length(unique(Genus)),
            genus_list = toString(unique(Genus)),
            species_list = toString(unique(ScientificName))
            ) %>% 
  arrange(desc(count_per_km))
      
## checking
# length(x$EventID)

## create flextable
myft <- flextable(x)

## modify flextable
myft <- theme_vanilla(myft)
myft<- fontsize(myft, size = 8, part = "body")
myft<- fontsize(myft, size = 8, part = "header")
myft <- align(myft, align = "left", part = "header")
myft <- align(myft, align = "left", part = "body")

myft <- set_table_properties(myft, width = .5, layout = "autofit")
myft

# myft <- italic(myft, j = 3)
# myft <- color(myft, ~ drat > 3.5, ~ drat, color = "red")
# myft <- bold(myft, ~ drat > 3.5, ~ drat, bold = TRUE)

## write it out 
setwd("C:/rworking/deepseatools/indata/")
write.xlsx(x, "20210826-0_species_lists_per_EventID.xlsx")

```
