---
title: "taxonomy_checker"
author: "Robert McGuinn"
date: "3/24/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---
# opts

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# packages

```{r}
library(tidyverse)
library(googledrive)
library(googlesheets4)
# install.packages("worrms")
library(worrms)
library(rmarkdown)
library(knitr)
library(openxlsx)
```

# load taxonomy tables

```{r}
##### load the most current taxonomy from Google Sheets #####
# https://drive.google.com/open?id=0B9c2c_XdhpFBT29NQmxIeUQ4Tlk
## manual: make sure the IDs below are pointing at the correct sheets


tax <- read_sheet('12PofxQbhE0nwxX4tFznuI17wPhSugRMST32YM_3N4D0')
taxfl <- read_sheet('1bGee86NQTHGfMMWt83nA1zCHnuNrzBOSkJJy5q29FbI')
taxch <- read_sheet('1S4e_RKgeMl5YoRDgDz9UU0HpnD3NgVpTGBysqt03SFw')


```

# load dataset of interest 

```{r}
##### get dataset from CSV #####
x <- '20201001-4_NOAA_OER_EX1811_Oceano_Profundo_2018_2018'
setwd("C:/rworking/deepseatools/indata")
sub <- read.csv(paste(x,'.csv', sep = ''), header = T)

##### --OR-- get dataset via Excel #####
# setwd("C:/rworking/deepseatools/indata")
# sub <- read.xlsx('20200313-0_NMNH-NewRecords_11-2019a_THourigan.xlsx', sheet = 1)

```

# create taxa list 4 possible ways

```{r}
##### create a character vector of non-matching taxa #####
# taxa <- as.character(
#   setdiff(unique(sub$ScientificName),
#           tax$ScientificName)
#   )

## checking
# taxa
# length(taxa)

##### -OR- just bring in a list of mismatches from csv #####

setwd("C:/rworking/deepseatools/indata")
taxa <- read.csv('taxa.csv', header = F)
taxa$V1 <- gsub("'", '', taxa$V1)
taxa <- taxa$V1


##### -OR- bring in a single taxa #####

#taxa <- as.character("Chrysogorgia binata")

##### -OR- just bring in all of the taxa from the sub #####
# taxa <- unique(sub$ScientificName)

##### create a raw version of taxa list #####
# taxaraw <- taxa

##### note: taxa list will go on for cleaning #####
## 'taxaraw' will be used for stripping taxonomy string
## directly from the 'sub' when it exists. This
## condition happens often with incoming Smithsonian records. 
## Warning: the incoming taxonomy is sometimes problematic and
## accepting it, whole cloth into the master taxonomy table 
## is probably risky no matter how expedient. Getting the 
## match directly from WoRMS is the preferred method.  
## Addtional code development time is needed to operationalize that.
## This would also mean the retirement of the master taxonomy list 
## as a customized entity. The manual assignment

```

# clean taxa list 

```{r}
# taxa <- taxa %>% str_replace(' sp.', '')
# taxa <- taxa %>% str_replace('cf. ', '')
taxa <- setdiff(taxa, tax$ScientificName)
taxa <- trimws(taxa)
taxa <- str_sort(taxa)

```

# check taxa list 

```{r}
length(taxa)
length(setdiff(sub$AphiaID, tax$AphiaID))
```

# match taxa (nonmatching taxa will not show)

```{r}
##### run the loop to match World Registry of Marine Species (WoRMS) for matching #####
## this loop and the subsequent rbind()step is helpful in getting around the
## WoRMS 50 taxa limit within their API.

for (val in taxa)
{
  tryCatch({
   x <- wm_records_taxamatch(name = val,
                            ids = TRUE,
                            verbose = TRUE,
                            marine_only = TRUE,
                            sleep_btw_chunks_in_sec = 0.2
  )

  x <- x[[1]]
  x$VerbatimScientificName <- val
  assign(paste('spec', str_replace(val, ' ', '_'), sep='_'), x)
  }, error=function(e){})
}

##### bind all of the resulting matches together #####
z <- mget(ls(pattern='spec')) %>%
              bind_rows()

## checking
# z %>% select(VerbatimScientificName,
#              scientificname,
#              valid_name) %>% View()
# 
# x <- tax %>%  filter(tax$ScientificName %in% z$scientificname) %>% 
#   select(ScientificName, AphiaID)
# View(x)
# 
# setdiff(x$AphiaID, z$AphiaID)
# setdiff(z$AphiaID, x$AphiaID)
# val

## cleanup
# rm(list = ls(pattern = 'spec'))
# rm(z)

```

# creating a new taxonomy empty table

```{r}

##### creating an empty taxonomy table (using information from accepted names) to populate (output new object) #####
newtax <- tax[0,]

#adding enough empty rows
newtax[1:length(z$scientificname),] <- NA

```

# bringing info from taxonomic match table (z)

```{r}
##### adding information from (match from Worms) to taxonomy table #####
newtax$ScientificName <- z$valid_name
newtax$AphiaID <- z$valid_AphiaID
newtax$ScientificNameAuthorship <- z$valid_authority
newtax$Phylum <- z$phylum
newtax$Class <- z$class
newtax$Order <- z$order
newtax$Family <- z$family
newtax$Genus <- z$genus
newtax$TaxonRank <- tolower(z$rank)
newtax$SynonymAphiaID <- "-999"
newtax$HigherTaxonNameAuthorship <- NA

## checking
# length(newtax$VernacularNameCategory)

## OPTIONAL: getting just the ones that aren't in the flag list 
x <- setdiff(newtax$ScientificName,taxfl$ScientificName)
newtax <- newtax %>% filter(ScientificName %in% x)

## checking
# View(newtax)


```

# joining new records with current taxonomy table 

```{r}
##### binding new taxonomy table with existing #####
newtax <- rbind(tax, newtax)
## checking
length(tax$VernacularNameCategory)
length(newtax$VernacularNameCategory)

## creating just data.frame
newtax <- as.data.frame(newtax)

## checking
# View(newtax)
```

# write updated taxonomy table to file

```{r}
##### write out new file #####
setwd("C:/rworking/deepseatools/indata")
newtax %>%
  write_csv('newtax.csv')
```

# --OR-- : extract a taxonomic table from the 'sub'

```{r}
##### prepare the column names list to use for select statment #####
taxstring <- names(tax) %>% 
  str_remove('HigherTaxonNameAuthorship') %>% 
  str_remove('SynonymAphiaID')
## getting rid of NA
taxstring <- taxstring[taxstring != ""]

##### select columns and create distinct record list #####
x <- sub %>% filter(sub$ScientificName %in% taxaraw) %>% 
  select(taxstring) %>% distinct()
## checking
# setdiff(x$ScientificName, sub$ScientificName)

##### adding back missing variables #####
x$HigherTaxonNameAuthorship <- NA
x$SynonymAphiaID <- -999
##checking
setdiff(names(tax), names(x))
setdiff(names(x), names(tax))
```

# joining extracted records with current taxonomy table 

```{r}
##### binding new taxonomy table with existing #####
newtax <- rbind(tax, x)
## checking
length(tax$VernacularNameCategory)
length(newtax$VernacularNameCategory)
```


