---
title: "30 Day Map Challenge"
author: "Robert McGuinn"
date: "5/16/2020"
output: html_document
---
# purpose

The writer here is Robert P. McGuinn on 2020-05-16 at 1648 in Charleston, South Carolina, USA.  

I will be working through the 30 day map challenge from https://rud.is/books/30-day-map-challenge/index.html written by Bob Rudis. Please see his original for more detail. This document is my copy and includes experimentation outside of the realm of original work by Rudis.  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# packages

```{r}
library(sf)
library(hrbrthemes)
library(googlesheets4)
library(ggthemes)
library(tidyverse)
```

# load data

* accessing the sheet via the sheet ID
* making a year_public variable so we can facet by year
* thinning out the data frame to just the year and address

```{r}
xdf <-read_sheet("https://docs.google.com/spreadsheets/d/1p-_GRo4YPW7m4QnjvErKD4U67t8-O6aDBlRjy9V8g8Y/edit?usp=sharing") %>% 
  mutate(
    year_public = ifelse(
      year_public < 2000, lubridate::year(date_added), year_public
    ) %>% factor()) %>% 
  select(year_public, city_st, lat, lng) 

## checking
glimpse(xdf)
```

# make it spatial

* get rid of missing lat/lng
* make it an sf object

```{r}
incidents <- xdf %>% filter(is.na(lat) == F)
incidents <- incidents %>% filter(is.na(lng) == F)
incidents <- incidents %>% st_as_sf(coords = c("lng", "lat"), crs =  '+proj=laea +lat_0=45 +lon_0=-100 +x_0=0 +y_0=0 +a=6370997 +b=6370997 +units=m +no_defs')
```


```{r}
ggplot() +
  # geom_sf( # plot the base map
  #   data = usa, fill = "#3B454A", size = 0.125, color = "#b2b2b277"
  # ) +
  geom_sf( # plot the points
    data = incidents, aes(fill = lab), 
    color = "white", size = 1.5, alpha = 2/3, shape = 21,
    show.legend = FALSE # no legend since we're using facets and it'd just be redundant
  ) + 
  scale_color_tableau() +
  coord_sf(datum = NA) + # using datum = NA turns off the graticule 
  facet_wrap(~lab) +
  labs(
    title = "Locations of K-12 Ransomware Incidents 2016-2019 (Sept)",
    caption = "2019 #30DayMapChallenge"
  ) +
  theme_ft_rc(grid="", strip_text_face = "bold") +
  theme(axis.text = element_blank()) +
  theme(strip.text = element_text(color = "white"))
```

