---
title: "Southeast US Deep Sea Community Analysis"
author: "Robert P. McGuinn"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: word_document
---

```{r installpackages, echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE}

#install.packages("beanplot")
library(beanplot)
#install.packages("stringr")
library(stringr)
#install.packages("knitr")
library(knitr)
#install.packages("tidyr")
library(tidyr)
#install.packages("sp")
library(sp)
#install.packages("maptools")
library(maptools)
#install.packages("maps")
library(maps)
#install.packages("reshape")
library(reshape)
#install.packages("reshape2")
library(reshape2)
#install.packages("psych")
library(psych)
#install.packages("ggplot2")
library(ggplot2)
#install.packages("data.table")
library(data.table)
#install.packages("dplyr")
library(dplyr)
#install.packages("car")
library(car)
#install.packages("gdata")
library(gdata)
#install.packages("digest")
library(digest)
#install.packages("rgdal")
library(rgdal)
#install.packages("ggmap")
library(ggmap)
#install.packages("rerddap")
library(rerddap)
#install.packages("raster")
library(raster)
#install.packages("rworldxtra")
library(rworldxtra)
#install.packages("ggrepel")
library(ggrepel)
#install.packages("xtable")
library(xtable)
library(taxize)
library(rgdal)
library(dplyr)
#install.packages("tidyverse")
library(tidyverse)
library(lattice)
library(extrafont)
library(RColorBrewer)
library(googlesheets)
```

```{r data, cache = FALSE, echo=FALSE}

setwd("C:/rworking/digs/indata")
#indata<-read.csv("DSCRTP_NatDB_20171214-0.csv", header = T)
indata<-read.csv('20180130_0_Genus_NoFlag_TH.csv')

filt <- indata %>%
  filter(Flag == '0' | Flag == '2', 
         ObservationYear != "-999",
         DepthInMeters != "-999", 
         is.na(Phylum)== F)
options(digits = 1)

# changing levels of 'Ecoregion'
filt2 <- filt %>% mutate(Ecoregion = ifelse(Ecoregion == 'N_Carolinean', 'North Carolinean', as.character(Ecoregion)))

minLon <- -83
maxLon <- -74 
minLat <- 23
maxLat <- 36

geofilt <- 
  filt2 %>% filter(as.numeric(Latitude) > minLat, 
                       as.numeric(Latitude) < maxLat, 
                       as.numeric(Longitude) < maxLon,
                       as.numeric(Longitude) > minLon)

coordinates(geofilt) <- c("Longitude", "Latitude")
proj4string(geofilt) <- "+proj=longlat +ellps=WGS84 +datum=WGS84"

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.height = 4, fig.width = 7.5)
##### building big pallette #####
# to get the colors from a palette:
palette1 <- brewer.pal(9,"Set1")
palette2 <- brewer.pal(8,"Set2")
palette3 <- brewer.pal(9,"Set3")
palette4 <- brewer.pal(8,"Dark2")
palette5 <- brewer.pal(8,"Accent")

# We can just stick a few color palettes together
big_palette <- c(palette1,palette2,palette3, palette4, palette5)
```

``` {r DepthClassTable, echo=FALSE, cache = FALSE} 

x <- 600
y <- 300
z <- 150

geofilt$DepthCat[geofilt$DepthInMeters > x] <- "very deep"
geofilt$DepthCat[geofilt$DepthInMeters > y & geofilt$DepthInMeters <= x] <- "deep"
geofilt$DepthCat[geofilt$DepthInMeters >= z & geofilt$DepthInMeters <= y] <- "mesophotic"
geofilt$DepthCat[geofilt$DepthInMeters < z] <- "very shallow"
geofilt$DepthCat <- factor(geofilt$DepthCat, levels = c('very deep', 'deep', 'mesophotic', 'very shallow'))

```

######Figure 1: North Carolinean ecoregion - all depths. These data are filtered for genera that have more than 30 occurrence. 'very deep' = >`r x` meters. 'deep' is <= `r x` and >= `r y`. 'mesophotic' is <= `r y` and >= `r z`. 'very shallow' is < `r z` 

``` {r ordered histograms1, echo=FALSE, cache=FALSE, dpi=300, warnings = FALSE}
sum_tbl <-
  as.data.frame(geofilt) %>%
  group_by(Genus) %>%
  summarize(
    Phylum = toString(unique(Phylum)),
    Taxon_Rank = toString(unique(TaxonRank)),
    ScientificNames = toString(unique(ScientificName)),
    Records = n(),
    MinDepth = min(DepthInMeters),
    MedianDepth = median(DepthInMeters),
    MaxDepth = max(DepthInMeters)) %>%
  arrange(Records)
options(scipen=10000)
#View(sum_tbl)

sum_tbl <- sum_tbl %>%
  filter(Records > 30)

geofilt2 <- as.data.frame(geofilt) %>%
  filter(Genus %in% sum_tbl$Genus, 
         Phylum == "Cnidaria", 
         Ecoregion == 'North Carolinean',
         as.numeric(DepthInMeters) < 1250 )

g <- ggplot(geofilt2, aes(reorder(Genus, DepthInMeters, FUN=median), as.numeric(DepthInMeters))) +   
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") + 
  xlab("Genus") +
  theme_bw(base_size = 14, base_family = "Cambria") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0))  +
  geom_hline(yintercept = 150, col = 'grey') +
  geom_hline(yintercept = 300, col = 'grey') +
  geom_hline(yintercept = 600, col = 'grey')

g + scale_color_manual(values = brewer.pal(12, "Paired")[c(10,9,8,7,6,5,4,3,2,1)])

# Pie charts
GenusCounts = summary(geofilt2$Genus)
set.seed(8)
pie(GenusCounts[GenusCounts>5],col=sample(big_palette), cex = .7)

```

######Figure 2: South Carolinean ecoregion - all depths. These data are filtered for genera that have more than 30 occurrences. 'very deep' = >`r x` meters. 'deep' is <= `r x` and >= `r y`. 'mesophotic' is <= `r y` and >= `r z`. 'very shallow' is < `r z` 

``` {r ordered histograms2, echo=FALSE, cache=FALSE, dpi=300, warnings = FALSE}
sum_tbl <-
  as.data.frame(geofilt) %>%
  group_by(Genus) %>%
  summarize(
    Phylum = toString(unique(Phylum)),
    Taxon_Rank = toString(unique(TaxonRank)),
    ScientificNames = toString(unique(ScientificName)),
    Records = n(),
    MinDepth = min(DepthInMeters),
    MedianDepth = median(DepthInMeters),
    MaxDepth = max(DepthInMeters)) %>%
  arrange(Records)
options(scipen=10000)
#View(sum_tbl)

sum_tbl <- sum_tbl %>%
  filter(Records > 30)

geofilt2 <- as.data.frame(geofilt) %>%
  filter(Genus %in% sum_tbl$Genus, 
         Phylum == "Cnidaria", 
         Ecoregion == 'South Carolinean',
         #DepthCat == 'very shallow'
         as.numeric(DepthInMeters) < 1250 )
g <- ggplot(geofilt2, aes(reorder(Genus, DepthInMeters, FUN=median), as.numeric(DepthInMeters))) +   
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") + 
  xlab("Genus") +
  theme_bw(base_size = 14, base_family = "Cambria") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0))  + 
  geom_hline(yintercept = 150, col = 'grey') +
  geom_hline(yintercept = 300, col = 'grey') +
  geom_hline(yintercept = 600, col = 'grey')

g + scale_color_manual(values = brewer.pal(12, "Paired")[c(10,9,8,7,6,5,4,3,2,1)])

# Pie charts
GenusCounts = summary(geofilt2$Genus)
set.seed(8)
pie(GenusCounts[GenusCounts > 30],col=sample(big_palette), cex = .7)

```

######Figure 3: Floridian ecoregion - all depths. These data are filtered for genera that have more than 30 occurrences. 'very deep' = >`r x` meters. 'deep' is <= `r x` and >= `r y`. 'mesophotic' is <= `r y` and >= `r z`. 'very shallow' is < `r z` 

``` {r ordered histograms3, echo=FALSE, cache=FALSE, dpi=300, warnings = FALSE}
sum_tbl <-
  as.data.frame(geofilt) %>%
  group_by(Genus) %>%
  summarize(
    Phylum = toString(unique(Phylum)),
    Taxon_Rank = toString(unique(TaxonRank)),
    ScientificNames = toString(unique(ScientificName)),
    Records = n(),
    MinDepth = min(DepthInMeters),
    MedianDepth = median(DepthInMeters),
    MaxDepth = max(DepthInMeters)) %>%
  arrange(Records)
options(scipen=10000)
#View(sum_tbl)

sum_tbl <- sum_tbl %>%
  filter(Records > 30)

geofilt2 <- as.data.frame(geofilt) %>%
  filter(Genus %in% sum_tbl$Genus, 
         Phylum == "Cnidaria", 
         Ecoregion == 'Floridian',#,
         #DepthCat == 'very shallow'
         as.numeric(DepthInMeters) < 1250 )
g <- ggplot(geofilt2, aes(reorder(Genus, DepthInMeters, FUN=median), as.numeric(DepthInMeters))) +   
  geom_boxplot() +
  scale_y_reverse() +
  ylab("Depth (meters)") + 
  xlab("Genus") +
  theme_bw(base_size = 14, base_family = "Cambria") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0))  +
  geom_hline(yintercept = 150, col = 'grey') +
  geom_hline(yintercept = 300, col = 'grey') +
  geom_hline(yintercept = 600, col = 'grey')

g + scale_color_manual(values = brewer.pal(12, "Paired")[c(10,9,8,7,6,5,4,3,2,1)])

# Pie charts
GenusCounts = summary(geofilt2$Genus)
set.seed(8)
pie(GenusCounts[GenusCounts>30],col=sample(big_palette),cex = .7)

```

######Figure 4: Bean plots of 'DepthInMeters' by 'Genus'. This depicts the genera ranked 1-6 by number of records within the study region. 

```{r Genus1, echo=FALSE, cache = FALSE, dpi=300}
sum_tbl <-
  as.data.frame(geofilt) %>%
  group_by(Genus) %>%
  summarize(
    Phylum = toString(unique(Phylum)),
    Taxon_Rank = toString(unique(TaxonRank)),
    Records = n()) %>%
  arrange(desc(Records))

geofilt2 <- as.data.frame(geofilt) %>%
  filter(Genus %in% sum_tbl$Genus[2:7])
g <- ggplot(geofilt2, aes(factor(Genus), as.numeric(DepthInMeters))) +   
  geom_violin() +
  scale_y_reverse() +
  ylab("Depth (meters)") + 
  xlab("Genus") +
  facet_wrap(~Order) +
  theme_bw(base_size = 15, base_family = "Cambria") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

g + scale_color_manual(values = brewer.pal(12, "Paired")[c(10,9,8,7,6,5)])

```

######Figure 5: Bean plots of 'DepthInMeters' by 'Genus'. This depicts the genera ranked 7-12 by number of records within the study region.  

```{r Genus2, echo=FALSE, cache = FALSE, dpi=300}
geofilt2 <- as.data.frame(geofilt) %>%
  filter(Genus %in% sum_tbl$Genus[8:13])

g <- ggplot(geofilt2, aes(factor(Genus), as.numeric(DepthInMeters))) +   geom_violin() +
  scale_y_reverse() +
  ylab("Depth (meters)") + 
  xlab("Genus") +
  facet_wrap(~Order) +
  theme_bw(base_size = 15, base_family = "Cambria") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

g + scale_color_manual(values = brewer.pal(12, "Paired")[c(10,9,8,7,6,5)])
```

######Figure 6: Bean plots of 'DepthInMeters' by 'Genus'. This depicts the genera ranked 13-18 by number of records within the study region.  

```{r Genus3, echo=FALSE, cache = FALSE, dpi=300}
geofilt2 <- as.data.frame(geofilt) %>%
  filter(Genus %in% sum_tbl$Genus[14:19])
g <- ggplot(geofilt2, aes(factor(Genus), as.numeric(DepthInMeters))) +   
  geom_violin() +
  scale_y_reverse() +
  ylab("Depth (meters)") + 
  xlab("Genus") +
  facet_wrap(~Order) +
  theme_bw(base_size = 15, base_family = "Cambria") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

g + scale_color_manual(values = brewer.pal(12, "Paired")[c(10,9,8,7,6,5)])
```

######Figure 7: Bean plots of 'DepthInMeters' by 'Genus'. This depicts the genera ranked 19-24 by number of records within the study region.  

```{r Genus4, echo=FALSE, cache = FALSE, dpi=300}

geofilt2 <- as.data.frame(geofilt) %>%
  filter(Genus %in% sum_tbl$Genus[20:25])
g <- ggplot(geofilt2, aes(factor(Genus), as.numeric(DepthInMeters))) +   geom_violin() +
  scale_y_reverse() +
  ylab("Depth (meters)") + 
  xlab("Genus") +
  facet_wrap(~Order) +
  theme_bw(base_size = 15, base_family = "Cambria") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))


g + scale_color_manual(values = brewer.pal(12, "Paired")[c(10,9,8,7,6,5)])
```

######Figure 8: Shows the number of records within each Depth Category regardless of subregion. 'very deep' = >`r x` meters. 'deep' is <= `r x` and >= `r y`. 'mesophotic' is <= `r y` and >= `r z`. 'very shallow' is < `r z`

``` {r DepthClassPlotFacet, echo=FALSE, cache = FALSE, dpi=300} 

options(scipen=10000)
geofilt <- as.data.frame(geofilt) %>%
  filter(Ecoregion != 'Virginian' , is.na(Ecoregion) == F)
g <- ggplot(as.data.frame(geofilt), aes(DepthCat, fill = Order)) +
  geom_bar() + 
  coord_flip() + 
  theme(text = element_text(size=20)) + 
  ylab("Number of Records") + 
  theme_bw(base_size = 15, base_family = "Cambria")

#display.brewer.all(colorblindFriendly=TRUE)
set.seed(8)
g + scale_fill_manual(values = sample(big_palette))

```

######Figure 9: shows the number of records within each Depth Category, faceted by Ecoregion. 'very deep' = >`r x` meters. 'deep' is <= `r x` and >= `r y`. 'mesophotic' is <= `r y` and >= `r z`. 'very shallow' is < `r z`
 
``` {r DepthClassPlot, echo=FALSE, cache = FALSE, dpi=300} 

options(scipen=10000)
geofilt <- as.data.frame(geofilt) %>%
  filter(Ecoregion != 'Virginian' , is.na(Ecoregion) == F)
g <- ggplot(as.data.frame(geofilt), aes(DepthCat, fill = Order)) +
  geom_bar() + 
  coord_flip() +
 #scale_x_reverse() +
  facet_wrap(~Ecoregion) +
  ylab("Number of Records") + 
  theme_bw(base_size = 15, base_family = "Cambria") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0))

#display.brewer.all(colorblindFriendly=TRUE)
set.seed(8)
g + scale_fill_manual(values = sample(big_palette))

```

######Table 1: Table of records by depth category grouped by Ecoregion.  The 'Genera' column is simply the number of unique genera and 'n' is the number of records. 'very deep' = >`r x` meters. 'deep' is <= `r x` and >= `r y`. 'mesophotic' is <= `r y` and >= `r z`. 'very shallow' is < `r z`.

``` {r DepthClassPlotFacetByMEOW, echo=FALSE, cache = FALSE, dpi=300}
sum_tbl <-
  as.data.frame(geofilt) %>%
  filter(is.na(DepthInMeters) != '-999')%>%
  group_by(Ecoregion, DepthCat) %>%
  summarize(
    Genera_n = length(unique(na.omit(Genus))),
    n = n()
  )
kable <- kable(sum_tbl, row.names = F, digits = 2)
kable

```

######Table 2: Frequency table of the number of records within each genera within the study area.  Table is descending order by the *frequency of occurrence* within each genera.

``` {r ByGenusCount, echo=FALSE }
sum_tbl <-
  as.data.frame(geofilt) %>%
  group_by(Genus) %>%
  summarize(
    Phylum = toString(unique(Phylum)),
    Taxon_Rank = toString(unique(TaxonRank)),
    ScientificNames = toString(unique(ScientificName)),
    Records = n(),
    MinDepth = min(DepthInMeters),
    MedianDepth = median(DepthInMeters),
    MaxDepth = max(DepthInMeters)) %>%
  arrange(desc(Records))
kable <- kable(sum_tbl, row.names = F, digits = 2)
kable

```

######Table 3:  Frequency table grouped by 'Genus'. Arranged in *ascending order by median depth* of each genera. 

``` {r ByGenusMedianDepth, echo=FALSE }
sum_tbl <-
  as.data.frame(geofilt) %>%
  group_by(Genus) %>%
  summarize(
    Phylum = toString(unique(Phylum)),
    Taxon_Rank = toString(unique(TaxonRank)),
    ScientificNames = toString(unique(ScientificName)),
    Records = n(),
    MinDepth = min(DepthInMeters),
    MedianDepth = median(DepthInMeters),
    MaxDepth = max(DepthInMeters)) %>%
  arrange(MedianDepth)
kable <- kable(sum_tbl, row.names = F, digits = 2)
kable
```

